### 2019-11-06 paper.tex
the inital commit

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}
%
\title{Contribution Title\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\input{marco}
\maketitle              % typeset the header of the contribution
%
\input{sections/abstract.tex}
\input{sections/introduction.tex}
\input{sections/preliminaries.tex}
\input{sections/i_uip.tex}
\input{sections/evaluation.tex}
%
\begin{thebibliography}{8}

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2019-11-06 sections/i_uip.tex
the inital commit

\section{Find i-uip Clause}
Classical CDCL SAT solvers analyze each conflict with 1-UIP resolution learning scheme to derive an asserting clause $\oneUIPClause$. Even though prior studies shows that further resolutions on $\oneUIPClause$ against the assertion trail $\assertionTrail$ cannot reduce the clause's literal block distance ($\LBD$), the resolutions could potentially reduce the clause's size. Clause size is an important quality measurement because a smaller clause 1) consumes less memory, 2) requires less steps to force a literal and 3) decreases the size of future conflict clauses.

Inspired by both $\LBD$ and clause size, i-UIP resolution learning attempts to resolve away literals in $\oneUIPClause$ against the assertion trail $\assertionTrail$ to minimize the clause size without increasing the clause's $\LBD$. The goal of i-UIP learning is to find a clause $\iUIPClause$ whose literals are either the unique implication points in their respective decision levels or are directly implied by literals from a foreign decision level. Alg. ~\ref{alg:i-uip} is a pseudo-code implementation of i-UIP learning.

\begin{algorithm}[t]
\caption{\IUIP}\label{alg:i-uip}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid and minimized 1-UIP clause
\Require  $\assertionTrail$ is a valid assertion trail
\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$} 
    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause} \label{ln:init}
    \State \code{DecisionLvs} $\gets$ Decision levels in $\oneUIPClause$ in descending order
\For{$i \in \code{DecisionLvs}$} \label{ln:iterDecisionLvls}
    \State $L_{i} \gets \Set{l}{\code{level}(l) = i \wedge l \in \iUIPClause}$
    \While{$|\code{unmarked}(L_i)| > 1$} \label{ln:uip}
        \State $p \gets \text{lit with the highest trail position in } L_{i} $ \label{ln:trailaccess}
        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs}} \label{ln:resolvable}
            \If{$\IUIPPURE$}
                \State Unresolve literals at level $i$  \label{ln:unresolve}
                \State Go to the next decision level $i$ \label{ln:abondon}
            \ElsIf{$\IUIPMIN$}
                \State $\code{Mark}(p)$ \label{ln:mark}
            \EndIf
        \Else 
            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(\neg{p}, \assertionTrail)$
           \State  $\code{Update}(L_{i})$
        \EndIf
    \EndWhile
\EndFor \label{ln:endCompleteResolution}
\If{$|\iUIPClause| < |\oneUIPClause|$} \label{ln:checkSize}
    \State \textbf{return} $\iUIPClause$
\Else
    \State \textbf{return} $\oneUIPClause$
\EndIf  \label{ln:cSizeCheck}
\EndProcedure
\end{algorithmic}
\end{algorithm}

The algorithm $\IUIP$ computes $\iUIPClause$ from (line~\ref{ln:init} to \ref{ln:endCompleteResolution}), and then returns the smaller clause between 
$\oneUIPClause$ and $\iUIPClause$ (line~\ref{ln:checkSize} to \ref{ln:cSizeCheck}). 
The algorithm first initializes $\iUIPClause$ with the minimized\cite{} $\oneUIPClause$. Next, the algorithm iterate through the decision levels of $\iUIPClause$ in descending order (line~\ref{ln:iterDecisionLvls}) and tries to find the unique implication point in each level (line~\ref{ln:uip}). Since $\IUIP$ needs to preserve $\iUIPClause$'s $\LBD$ during resolutions, before resolving away a target literal $p$, the algorithm preemptively checks whether the reasoning clause of $\neg{p}$ contains any literal $q$ from an unseen decision level (line~\ref{ln:resolvable}). If such a literal $q$ exists, then the algorithm cannot find the unique implication point (UIP) at the current level $i$ because resolving away $p$ will introduce $q$ into the clause and increases $\LBD$. One solution is to abandon level $i$ by reverting back to the state before any literal at level $i$ is resolved away (line~\ref{ln:unresolve}), and then move on to the next level (line~\ref{ln:abondon}). We denote $\IUIP$ learning with this solution as $\IUIPPURE$. The final $\iUIPClause$ produced by $\IUIPPURE$ will contains exactly one literal for levels where the LBD-persevering UIP exist. However, $\IUIPPURE$ does not minimize the number of literals in the decision levels where the LBD-preserving UIP does not exist.

  When $\IUIP$ cannot resolve away a literal $p$ without increasing $\iUIPClause$'s LBD, instead of skipping the entire decision level $i$, a more practical solution $\IUIPMIN$ will mark and keep $p$ in $\iUIPClause$ (line~\ref{ln:mark}), and continue resolutions at level $i$. $\IUIPMIN$ will ignore all the unsolvable literals and find the "local" unique implication point at every decision level.  The algorithm terminates when there is exactly one unmarked literal left for each decision level of $\iUIPClause$, representing the local unique implication points.

The core algorithm of $\IUIP$ as a clause reduction technique is simple. However, to make the algorithm a practical learning scheme that is compatible with modern SAT solvers, a number of issues need to be addressed. The rest of the section details the key optimizations and augmentations of $\IUIP$ as a practical clause learning scheme.

\subsection{Control i-UIP Learning}
This simple and greedy $\IUIP$ learning scheme can produce significantly smaller clause. However, when $\IUIP$ does not reduce the clause size, the cost of the additional resolution steps will hurt the solver's performance. Since resolution cannot reduce a clause's $\LBD$, The maximum size reduction from $\IUIP$ is the difference between the clause's size and $\LBD$, denote as the clause's gap value ($\GAP(\oneUIPClause) = |\oneUIPClause| - \LBD(\oneUIPClause)$). For an 1-UIP clause with a small $\GAP$, applying $\IUIP$ is unlikely to achieve cost effective results. Therefore, we propose a heuristic based approach to to enable and disable $\IUIP$ learning based on input clause's $\GAP$.

\begin{algorithm}[t]
\caption{Control-$\IUIP$}\label{alg:enableCondition}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Require  $ t_{gap} \ge 0$ is a dynamically calculated gap threshold
\Procedure{Control-$\IUIP$}{$\oneUIPClause,  t_{gap}$} 
    \State $Gap \gets |\oneUIPClause| - \textit{LBD}(\oneUIPClause)$
     \If{$Gap > t_{gap}$} \label{ln:compareGap}
        \State $\iUIPClause \gets \IUIP(\oneUIPClause, \phi)$
        \If {$|\iUIPClause| < |\oneUIPClause|$} 
            \State $\text{Succeed} \gets \text{Succeed}+1$ \label{ln:updateS}
        \EndIf
        \State $\text{Attempted} \gets \text{Attempted}+1$ \label{ln:updateA}
     \EndIf
\EndProcedure
\end{algorithmic}
\end{algorithm}

Alg.~\ref{alg:enableCondition} compares the input $\oneUIPClause$'s $\GAP$ against a floating target threshold $t_{gap}$ (line~\ref{ln:compareGap}). The threshold $t_{gap}$ represent the expected minimal $\GAP$ required for $\oneUIPClause$ to achieve a predetermined success rate (80\%) from performing $\IUIP$ learning.

The gap threshold $t_{gap}$ is initialized to 0, and is updated at every restart based on $\IUIP$'s success rate from the previous restart interval. More specifically, the algorithm collects the statistics of the number of $\IUIP$ learning attempted (line~\ref{ln:updateA} in alg.~\ref{alg:enableCondition}) and the number of attempts succeed (line~\ref{ln:updateS}) for each restart interval, and use them to calculate the success rate. If the success rate is below 80, the threshold $t_{gap}$ is increased to restrict $\IUIP$for the next restart interval. On the other hand, the threshold is decreased to encourage more aggressive $\IUIP$ learning for the next restart interval.

\[
    t_{gap}=
\begin{cases}
    t_{gap} + 1& \text{if } \frac{\text{Succeed}}{\text{Attempted}} < 0.8\\
    max(t_{gap} - 1, 0),              & \text{otherwise}
\end{cases}
\]

\subsection{Early stop i-UIP Learning}
At any point of $\IUIP$ learning, if the number of marked literals in $\iUIPClause$ (literals which are forced into the clause to preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the final $\iUIPClause$ is at least the size of $\oneUIPClause$. The early stopping rule prevents solver from wasting time on traversing a large implication graph when $\IUIP$ has already failed.

\subsection{Greedy Active Clause Selection}
Even though $\IUIP$ learning can reduce the size of the learnt clause $\iUIPClause$, but it may introduce inactive literals into $\iUIPClause$.  Inactive literals prevents the clause from being asserted to force literal implication. Therefore, a practical clause learning scheme should consider both size and variable activity. We propose an optional extension $\IUIPGreedy$ to filter out inactive $\oneUIPClause$.

\begin{algorithm}[t]
\caption{$\IUIPGreedy$}\label{alg:greedySelection}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Require  $\assertionTrail$ is a valid assertion trail
\Procedure{$\IUIPGreedy$}{$\oneUIPClause,  \assertionTrail$} 
   \State $\iUIPClause \gets \IUIP(\oneUIPClause, \assertionTrail)$
   \If{$|\iUIPClause| < |\oneUIPClause| \wedge (\code{AvgVarAct}(\iUIPClause) > \code{AvgVarAct}(\oneUIPClause)) $} \label{ln:checkSizeAndActivity}
    \State \textbf{return} $\iUIPClause$
\Else
    \State \textbf{return} $\oneUIPClause$
\EndIf  \label{ln:cSizeCheck}
\EndProcedure
\end{algorithmic}
\end{algorithm}

After computing the $\iUIPClause$, $\IUIPGreedy$ compares both the size and the average variable activity for $\oneUIPClause$ and $\iUIPClause$ (alg.~\ref{alg:greedySelection} at line~\ref{ln:checkSizeAndActivity}). The algorithm learns $\iUIPClause$ if the clause has smaller size and higher average variable activity.

\subsection{Adjust Variable Activity}
Two popular branching heuristics for modern SAT solvers are VSIDS and LBR. Both heuristics increase 
the variable activity for all literals involved in resolutions during 1-UIP learning. Since $\IUIP$ extends 1-UIP with deeper resolutions against the trail, the variables activities for the fresh literals involved in the additional resolution steps need to be adjusted as well. We purpose two optional schemes for adjusting variable activities, $\IUIPActive$ and $\IUIPDist$.

After learning $\oneUIPClause$ from 1-UIP scheme, $\IUIPActive$ collects all literals appear  in the the $\IUIP$ clause $\iUIPClause$, and increase their variable activity uniformly according to the current branching heuristic if the literals' variable activity have not yet been bumped during 1-UIP. The scheme does not bump variable activities for transient literals that are resolved away at non-conflicting level because, unlike transient literals at conflicting level, these literals cannot be re-asserted after the immediate backtracking. Notice that other post-analyze extensions such as Reason Side Rate (RSR) and Locality\cite{} can be applied after applying $\IUIPActive$ on $\iUIPClause$ so that no literal's variable activity is double bumped.

\[ \forall l \in  \iUIPClause \cdot NotBumped(var(l)) \implies  bumpActivity(var(l)) \]

$\IUIPDist$ collects literals appear exclusively in $\iUIPClause$ and bump their variable activity uniformly. It also find all the literals in $\oneUIPClause$ that are resolved away during $\IUIP$, and unbump their variable activity if they have been bumped during 1-UIP. The unbumped literals are no longer in the learned clause $\iUIPClause$, keep their variable activity bumped does not help solver to use $\iUIPClause$.

\[ \forall l_{i} \in  \iUIPClause \setminus \oneUIPClause \cdot bumpActivity(var(l_{i})) 
\]
\[ \forall l_{1} \in  \oneUIPClause \setminus \iUIPClause \cdot unbumpActivity(var(l_{1})) 
\]

\subsection{Integrate with Chronological Backtracking}
In SAT solver with chronological backtracking, literals on the assertion trail $\assertionTrail$ are not always sorted by decision levels. This change imposes a challenge to $\IUIP$ learning since the previous implementation relies on solver's ability to efficiently access all literals from any decision level in descending trail order (alg.~\ref{alg:i-uip} line~\ref{ln:trailaccess}).

To mitigate this challenge, we modify the solver to track the precise trail position of all asserted literals with a single vector. We then build a priority queue $lit\_Order$ to manage literal's resolution order. The queue $lit\_Order$  prioritizes literals with higher decision level, and it favors literal with higher trail position when decision levels are tied. The order of $lit\_Order$  represents the correct resolution order of $\IUIP$ because 1) an asserted literal's decision level is the maximum level among all of its reasoning literals, and 2) an asserted literal always appears higher in the trail than its reasoning literal.

\begin{algorithm}[t]
\caption{\IUIP-CB}\label{alg:i-uip-CB}
\begin{algorithmic}[1]
\Require  $\oneUIPClause$ is a valid 1-UIP clause
\Require  $\assertionTrail$ is a valid assertion trail
\Require  $lit\_Order$ is a priority queue
\Procedure{\IUIP-CB}{$\oneUIPClause, \assertionTrail, lit\_Order $} 
    \State ...
    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause}
    \State forall $l \in \oneUIPClause \cdot \code{Enqueu}(lit\_Order, l)$  \label{ln:enqueue}
    \State \code{DecisionLvs} $\gets \Set{\code{level}(l)}{l \in \oneUIPClause}$
    \While{$lit\_Order \neq \emptyset$} \label{ln:pop}
        \State $p \gets \code{dequeu}(lit\_Order, l) $ \label{ln:dequeue}
        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs} \\ $\vee p$ is the last reminaing lit in its decision level } \label{ln:resolvable}
          \State Pass
        \Else 
            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(q, \assertionTrail)$
            \State   forall $l \in \code{Reason}(q, \assertionTrail) \cdot \code{Enqueu}(lit\_Order, l)$ \label{ln:reequeue}
        \EndIf
    \EndWhile
\State ...
\EndProcedure
\end{algorithmic}
\end{algorithm}

Alg.~\ref{alg:i-uip-CB} is the pesudo-code implementation of the augmented $\IUIP$ for chronological backtracking with the priority queue $lit\_Order$. The algorithm first populates $lit\_Order$ with all literals in $\oneUIPClause$ (line~\ref{ln:enqueue}), and then continuously pops literals until the queue is empty (line~\ref{ln:pop}). When a literal is resolved away, all of its reasoning literals are added into $lit\_Order$ if they are not already in the queue (line~\ref{ln:reequeue}). The algorithm will always terminate because a literal cannot entered the queue twice (guaranteed by the properties of the trail order) and there are finite amount of literals on the trail.

### 2019-11-06 sections/evaluation.tex
describe first part of the experiment

\section{Implementation and Experiments}

\subsection{Reduce Clause Size with $\IUIP$}
To evaluate the effectiveness of $\IUIP$ as a clause reduction technique, we implement  $\IUIPPURE$ and $\IUIPMIN$  on top of \text{\MapleBase} \cite{}, the winner of SAT Race 2015 application track.  We than compare the performance of the baseline $\MapleBase$ with $\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from SAT RACE 2019 main track.

The benchmark contains 400 instances divided into two groups of 200, new and old, representing historical instances and fresh instances in the 2019 race, respectively. I partition the old group instances into six partitions of size 30 and one partition of size 20. Each partition is then assigned to a Intel CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory. The new group is partitioned based on their contributor (e.g. Heule contributed 22 matrix multiplication instances), and each partition is assigned to a aforementioned CPU node. To speed up the experiment, we allow a CPU node to solver at most seven instances concurrently.

Beside solved instances and PAR-2 score, we additionally measure the average clause length and memory consumption for each instances. For $\MapleIUIPPURE$  and $\MapleIUIMIN$, we also captures the $\IUIP$ learning attempted rate and success rate.

Inserted a table here, and graphs and analysis

\begin{center}
\begin{tabular}{ | m{4cm} | m{2cm}| m{2cm} | m{2cm} | m{2cm} | } 
\hline
Solver & \# solved & PAR-2 & Clause Size & Memory \\ 
\hline
$\MapleBase$ & cell5 & cell6 & cell 7 & cell 8 \\ 
\hline
$\MapleIUIPPURE$ & cell8 & cell9 & cell10 & cell11\\ 
\hline
$\MapleIUIMIN$ & cell8 & cell9 & cell10 & cell11\\ 
\hline
\end{tabular}
\end{center}

A solver produce smaller clauses can construct smaller proofs as well. For UNSAT instances, we additionally measure their DRUP\cite{} proof checking time as well as the size of the optimized DRUP proof. We used DART-trim \cite{} with 10000 timeout to check and optimize DRUP proofs.

Insert a comparison graphs here.

### 2019-11-07 sections/evaluation.tex
added marco's for comments.
make change to optimization 1

@@
 
-\subsection{Reduce Clause Size with $\IUIP$}
-To evaluate the effectiveness of $\IUIP$ as a clause reduction technique, we implement  $\IUIPPURE$ and $\IUIPMIN$  on top of \text{\MapleBase} \cite{}, the winner of SAT Race 2015 application track.
+\subsection{Clause Reduction with $\IUIP$}
+To evaluate $\IUIP$'s effectiveness as a clause reduction technique, we implement  $\IUIPPURE$ and $\IUIPMIN$  on top of \text{\MapleBase} \cite{}, the winner of SAT Race 2015 application track.
 We than compare the performance of the baseline $\MapleBase$ with $\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from SAT RACE 2019 main track.
@@
 
-Inserted a table here, and graphs and analysis
+\nf{Inserted a table here, and graphs and analysis}
 
 \begin{center}
-\begin{tabular}{ | m{4cm} | m{2cm}| m{2cm} | m{2cm} | m{2cm} | } 
+\begin{tabular}{ | m{3.5cm} | m{2cm}| m{2cm} | m{2cm} | m{2cm} | } 
 \hline
@@
 For UNSAT instances, we additionally measure their DRUP\cite{} proof checking time as well as the size of the optimized DRUP proof.
-We used DART-trim \cite{} with 10000 timeout to check and optimize DRUP proofs.
+We used DART-trim \cite{} with 5000 timeout to check and optimize DRUP proofs.
 
-Insert a comparison graphs here.
+\nf{Insert a comparison graphs here. (X and Y comparisons graph, showing both time and optimized proof size).}
+
+\subsection{$\IUIP$ as a Practical Learning Scheme}
+To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with $\IUIPGreedy$ (sec~\ref{sec:greedy}) and two variable activity adjustment heuristics,  $\IUIPActive$ and $\IUIPDist$ (sec~\ref{sec: varajust}).

### 2019-11-07 sections/i_uip.tex
make change to optimization 1

@@
         \State $\iUIPClause \gets \IUIP(\oneUIPClause, \phi)$
+        \State \textcolor{gray}{\textbf{I-UIP-Greedy}($\iUIPClause$ , $\oneUIPClause$)}  \Comment{additional clause selection policy, see sec~\ref{sec:greedy}}
         \If {$|\iUIPClause| < |\oneUIPClause|$} 
@@
 
-\subsection{Greedy Active Clause Selection}
-Even though $\IUIP$ learning can reduce the size of the learnt clause $\iUIPClause$, but it may introduce inactive literals into $\iUIPClause$.
+\subsection{Greedy Active Clause Selection} \label{sec:greedy}
+Even though $\IUIP$ learning can reduce the size of the learnt clause $\iUIPClause$, but it may introduce literals with low variable activity into $\iUIPClause$.
 Inactive literals prevents the clause from being asserted to force literal implication.
@@
 \begin{algorithmic}[1]
+\Require  $\iUIPClause$ is a valid i-UIP clause
 \Require  $\oneUIPClause$ is a valid 1-UIP clause
-\Require  $\assertionTrail$ is a valid assertion trail
-\Procedure{$\IUIPGreedy$}{$\oneUIPClause,  \assertionTrail$} 
-   \State $\iUIPClause \gets \IUIP(\oneUIPClause, \assertionTrail)$
+\Procedure{$\IUIPGreedy$}{$\iUIPClause, \oneUIPClause$} 
    \If{$|\iUIPClause| < |\oneUIPClause| \wedge (\code{AvgVarAct}(\iUIPClause) > \code{AvgVarAct}(\oneUIPClause)) $} \label{ln:checkSizeAndActivity}
@@
 
-\subsection{Adjust Variable Activity}
+\subsection{Adjust Variable Activity} \label{sec: varajust}
 Two popular branching heuristics for modern SAT solvers are VSIDS and LBR. Both heuristics increase 

### 2019-11-07 sections/evaluation.tex
add evaluation section template

@@
 \subsection{$\IUIP$ as a Practical Learning Scheme}
-To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with $\IUIPGreedy$ (sec~\ref{sec:greedy}) and two variable activity adjustment heuristics,  $\IUIPActive$ and $\IUIPDist$ (sec~\ref{sec: varajust}).
+To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
+We evaluated four different configurations (i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances and PAR-2 score.
+
+\nf{Inserted a table here, and graphs and analysis}
+
+\subsection{$\IUIP$ on Modern SAT solvers}
+To validate $\IUIP$ as a generalizable learning scheme on modern SAT solvers, we re-implement $\IUIP$ on $\MapleSeven$ \cite{},  $\MapleNine$ \cite{} and $\expSAT$.  
+The first two solvers are the winner of 2017 and 2019 SAT race, respectively.
+$\expSAT$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branch.
+For each solver, we compare the base 1-UIP learning scheme against top two $\IUIP$ configurations, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
+We report solved instances and PAR-2 score.
+
+\nf{Inserted a table here, and graphs and analysis}

### 2019-11-09 sections/evaluation.tex
add len comparsion plot

@@
 
-Beside solved instances and PAR-2 score, we additionally measure the average clause length and memory consumption for each instances.
+Beside solved instances and PAR-2 score, we additionally measure the average clause length and clause reduction ratio for each instances.
 For $\MapleIUIPPURE$  and $\MapleIUIMIN$, we also captures the $\IUIP$ learning attempted rate and success rate.
@@
 \begin{center}
-\begin{tabular}{ | m{3.5cm} | m{2cm}| m{2cm} | m{2cm} | m{2cm} | } 
+\begin{tabular}{ | m{3.5cm} | m{2cm}| m{2cm} | m{2cm} | m{2.5cm} | } 
 \hline
-Solver & \# solved & PAR-2 & Clause Size & Memory \\ 
+Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
 \hline
-$\MapleBase$ & cell5 & cell6 & cell 7 & cell 8 \\ 
+$\MapleBase$ & cell5 & cell6 & 59.3 & 37.53\% \\ 
 \hline
-$\MapleIUIPPURE$ & cell8 & cell9 & cell10 & cell11\\ 
+$\MapleIUIPPURE$ & 221 & cell9 & cell10 & cell11\\ 
 \hline
-$\MapleIUIMIN$ & cell8 & cell9 & cell10 & cell11\\ 
+$\MapleIUIMIN$ & 226 & cell9 & 41.5 & 49.5\%\\ 
 \hline

### 2019-11-10 sections/evaluation.tex
add length comparsion and proof size graphs

@@
 \hline
-$\MapleBase$ & cell5 & cell6 & 59.3 & 37.53\% \\ 
+$\MapleBase$ & 221 & 5005.2 & 62.6 & 36.53\% \\ 
 \hline
-$\MapleIUIPPURE$ & 221 & cell9 & cell10 & cell11\\ 
+$\MapleIUIPPURE$ & cell9 & cell9 & cell10 & cell11\\ 
 \hline
-$\MapleIUIMIN$ & 226 & cell9 & 41.5 & 49.5\%\\ 
+$\MapleIUIMIN$ & 226 & 4933.6 & 45.2 & 47.5\%\\ 
 \hline
@@
 \end{center}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth, natwidth=610,natheight=642]{clause_length_PDF.png}
+    \caption{Average clause (relative to 1-UIP clauses) size distribution. X axis indicates the relative size difference, and Y axis indicates the PDF.}
+     \label{fig:len_pdf}
+\end{figure}
+\begin{figure} \label{fig:len_compare}
+    \centering
+    \includegraphics[width=0.8\textwidth, natwidth=610,natheight=642 ]{i-uip-sizes-compare.png}
+    \caption{Average clause size comparison plot. Each point in the plot represents an benchmark instance. X and Y axis shows the clause length from $\MapleIUIMIN$ and $\MapleBase$, respectively  }
+    \label{fig:len_compare}
+\end{figure}
 

### 2019-11-16 sections/evaluation.tex
continue evaluation

@@
 
-Beside solved instances and PAR-2 score, we additionally measure the average clause length and clause reduction ratio for each instances.
+Beside solved instances count and PAR-2 score, we additionally measure the average clause length and clause reduction ratio for each instances.
 For $\MapleIUIPPURE$  and $\MapleIUIMIN$, we also captures the $\IUIP$ learning attempted rate and success rate.
 
-\nf{Inserted a table here, and graphs and analysis}
-
+\begin{figure} 
 \begin{center}
@@
 \hline
-$\MapleBase$ & 221 & 5005.2 & 62.6 & 36.53\% \\ 
+$\MapleBase$ & 221 & 5018.89 & 62.6 & 36.53\% \\ 
 \hline
-$\MapleIUIPPURE$ & cell9 & cell9 & cell10 & cell11\\ 
+$\MapleIUIPPURE$ & 226 & 4920.04 & 49.6 & 41.6\% \\ 
 \hline
-$\MapleIUIMIN$ & 226 & 4933.6 & 45.2 & 47.5\%\\ 
+$\MapleIUIMIN$ & 226 & 4890.67 & 45.2 & 47.8\%\\ 
 \hline
@@
 \end{center}
+\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.}
+\label{fig:t1}
+\end{figure}
+
+Fig.~\ref{fig:t1} shows that both version of $\IUIP$ solved five more instances than the baseline solver with lower PAR-2 scores.
+$\IUIPMIN$ has marginally lower PAR-2 score than $\IUIPPURE$.
+Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with significantly smaller size than 1-UIP by 27.7\% and 20.7\%, respectively.
+Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the average clause length from $\IUIPMIN$ relative to 1-UIP.
+$\IUIPMIN$ learning produces shorter clauses for 88.25\% instances, and average relative reduction from 1-UIP is 18.685\%.Fig.~\ref{fig:len_compare} compares the absolute average clause size from $\IUIPMIN$ and 1-UIP, and it shows that $\IUIPMIN$ in general produces smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
+
+We also looked at the 14 instances solved by $\IUIPMIN$ but not by 1-UIP.
+$\IUIPMIN$ produces smaller clauses for all of them with average relative reduction of 22\% and maximum 77\% (30 vs 135).
+Seven out of 14 instances has size relative reduction over 30\%.
+For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause by 33\% and with average relative reduction of 3.3\$.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth, natwidth=610,natheight=642]{clause_length_PDF.png}
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{clause_length_PDF.png}
     \caption{Average clause (relative to 1-UIP clauses) size distribution. X axis indicates the relative size difference, and Y axis indicates the PDF.}
@@
     \centering
-    \includegraphics[width=0.8\textwidth, natwidth=610,natheight=642 ]{i-uip-sizes-compare.png}
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{i-uip-sizes-compare.png}
     \caption{Average clause size comparison plot. Each point in the plot represents an benchmark instance. X and Y axis shows the clause length from $\MapleIUIMIN$ and $\MapleBase$, respectively  }
@@
 
-A solver produce smaller clauses can construct smaller proofs as well.
+$\IUIPMIN$ outperformed $\IUIPPURE$ in both PAR-2 score and clause size.
+This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.
+Remark that the success of $\IUIP$ learning is determined by the size of the learned i-UIP clause $\iUIPClause$, and the $\IUIP$ learning frequency is also indirectly controlled by $\IUIP$'s success rate from the previous restart interval.
+The results indicates  $\IUIPMIN$ shortened $\iUIPClause$'s size through further minimization of $\iUIPClause$ at non-unique implication decision level.
+
+\begin{figure} 
+\begin{center} 
+\begin{tabular}{ | m{3.5cm} | m{5cm}| m{5cm} | } 
+\hline
+Solver & $\IUIP$ attempt rate & $\IUIP$ success rate  \\ 
+\hline
+$\MapleIUIPPURE$ & 16.1\% & 43.4\% \\ 
+\hline
+$\MapleIUIMIN$ & 28.8\% & 59.3\% \\ 
+\hline
+\end{tabular}
+\end{center}
+\caption{Compare $\IUIPPURE$ and $\IUIPMIN$ i-uip attempt rate and success rate.
+$\IUIPMIN$ scheme attempted $\IUIP$ more frequently, and it is more likely to successfully produce smaller $\iUIPClause$ clause .}
+\label{fig:t2}
+\end{figure}
+
+A solver produce smaller clauses can construct smaller proofs.
 For UNSAT instances, we additionally measure their DRUP\cite{} proof checking time as well as the size of the optimized DRUP proof.
@@
 
-\nf{Insert a comparison graphs here. (X and Y comparisons graph, showing both time and optimized proof size).}
+Fig.~\ref{fig:t3} shows that the optimized proof construct by $\IUIPMIN$ and $\IUIPPURE$ are significantly smaller than 1-UIP proofs.
+The relative proof size reduction rougly correlates to the average clause size reduction.
+Fig.~\ref{fig:proof_compare} shows the absolute proof size comparison results.
+
+\begin{figure} 
+\begin{center} 
+\begin{tabular}{ | m{3.5cm} | m{5cm}| m{5cm} | } 
+\hline
+Solver & optimized proof size (MB) & relative reduction size  \\ 
+\hline
+$\MapleBase$ & 613.9 & 0  \\ 
+\hline
+$\MapleIUIPPURE$ & 487.2 & 6.90\% \\ 
+\hline
+$\MapleIUIMIN$ & 413.2 & 17.18\% \\ 
+\hline
+\end{tabular}
+\end{center}
+\caption{Optimized UNSAT proof comparison for 1-UIP $\IUIPPURE$ and $\IUIPMIN$.
+Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
+\label{fig:t3}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{proof_size_compare.png}
+    \caption{Average optimized proof size between 1-uip and $\IUIPMIN$.}
+    \label{fig:proof_compare}
+\end{figure}
 
@@
 To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
-We evaluated four different configurations (i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances and PAR-2 score.
+We evaluated four different configurations (i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances, PAR-2 score and average clause size.
+
+Fig.~\ref{fig:t4} summarizes the result of the experiment.
+Learning scheme i-UIP-greedy solves the same amount of instance (226) as i-UIP with less PAR-2 score.
+The inclusive activity adjustment solves the most SAT instances (138) and the least UNSAT instances (87).
+The exclusive activity adjustment scheme produces the shortest average clause size, but solved the second least instances, one more instance than the baseline.
+
+\begin{figure} 
+\begin{center}
+\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
+\hline
+Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
+\hline
+1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
+\hline
+i-UIP & \textbf{226} (135, \textbf{91}) & 4890.67 & 45.2 \\ 
+\hline
+i-UIP-greedy & \textbf{226} (135, \textbf{91})  & \textbf{4866.94} & 47.7 \\
+\hline
+i-UIP-active-Inclusive & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
+\hline
+i-UIP-active-Exclusive & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of 1-UIP ($\MapleBase$), i-UIP($\IUIPMIN$), i-UIP-Greedy,
+i-UIP-Inclusive, and i-UIP-Exclusive on SAT2019 race main track.}
+\label{fig:t4}
+\end{figure}
 
@@
 The first two solvers are the winner of 2017 and 2019 SAT race, respectively.
-$\expSAT$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branch.
+$\expSAT$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
 For each solver, we compare the base 1-UIP learning scheme against top two $\IUIP$ configurations, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.

### 2019-11-24 sections/evaluation.tex
change evalation formmating, add existing benchmark results for the multi-solver tests

@@
 Seven out of 14 instances has size relative reduction over 30\%.
-For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause by 33\% and with average relative reduction of 3.3\$.
+For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for three instances and with average relative reduction of 3.3\%.
 
@@
     \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{i-uip-sizes-compare.png}
-    \caption{Average clause size comparison plot. Each point in the plot represents an benchmark instance. X and Y axis shows the clause length from $\MapleIUIMIN$ and $\MapleBase$, respectively  }
+    \caption{Average clause size comparison plot.
+Each point in the plot represents an benchmark instance.
+X and Y axis shows the clause length from $\IUIP$ and 1-UIP, respectively.
+Each green (red) dot represents an compared instance between $\MapleBase$ and $\MapleIUIMIN$ ($\IUIPPURE$).
+}
     \label{fig:len_compare}
@@
 Fig.~\ref{fig:t3} shows that the optimized proof construct by $\IUIPMIN$ and $\IUIPPURE$ are significantly smaller than 1-UIP proofs.
-The relative proof size reduction rougly correlates to the average clause size reduction.
+The relative proof size reduction roughly correlates to the average clause size reduction.
 Fig.~\ref{fig:proof_compare} shows the absolute proof size comparison results.
@@
 \subsection{$\IUIP$ on Modern SAT solvers}
-To validate $\IUIP$ as a generalizable learning scheme on modern SAT solvers, we re-implement $\IUIP$ on $\MapleSeven$ \cite{},  $\MapleNine$ \cite{} and $\expSAT$.  
+To validate $\IUIP$ as a generalizable learning scheme on modern SAT solvers, we re-implement $\IUIP$ on $\MapleSeven$ \cite{},  $\MapleNine$ \cite{}($\MapleNineShort$) and $\expSAT$\cite{} ($\expSATShort$).
 The first two solvers are the winner of 2017 and 2019 SAT race, respectively.
-$\expSAT$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
-For each solver, we compare the base 1-UIP learning scheme against top two $\IUIP$ configurations, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
-We report solved instances and PAR-2 score.
+$\expSATShort$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
+For each solver, we compare the base 1-UIP learning scheme against $\IUIP$ learning and the top two $\IUIP$ variants, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
+We report solved instances, PAR-2 score and the average clause size.
 
-\nf{Inserted a table here, and graphs and analysis}
+\begin{figure} 
+\begin{center}
+\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
+\hline
+Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
+\hline
+$\MapleSeven$ & 232 (135, 97)  & 4755.96 & 61.9  \\ 
+\hline
+$\MapleSeven$-i-uip & TBA & TBA & TBA \\ 
+\hline
+$\MapleSeven$-i-greedy & \textbf{237} (\textbf{140}, 97) & \textbf{4678.434} & \textbf{43.62} \\ 
+\hline
+$\MapleSeven$-i-inclusive & TBA & TBA & TBA \\ 
+\hline
+\hline
+$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
+\hline
+$\expSATShort$-i-uip & TBA & TBA & TBA \\ 
+\hline
+$\expSATShort$-i-greedy & 244 (143, \textbf{101}) & \textbf{4460.92} & 47.25 \\
+\hline
+$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) & 4475.76 & \textbf{45.33} \\
+\hline
+\hline
+$\MapleNineShort$ & 243 (145, \textbf{98}) & 4450.24 & 61.00 \\
+\hline
+$\MapleNineShort$-i-uip & 244 (148, 96) & \textbf{4409.88} & \textbf{37.43} \\
+\hline
+$\MapleNineShort$-i-greedy & 243 (146, 97) & 4476.73 & 40.65 \\
+\hline
+$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) & 4425.99 & 37.60 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of 1-UIP, i-UIP($\IUIPMIN$), i-UIP-Greedy,
+i-UIP-Inclusive on SAT2019 race main track.}
+\label{fig:t5}
+\end{figure}

### 2019-11-25 sections/evaluation.tex
add a non-cumulative i-uip cl reduction data

@@
 
-Beside solved instances count and PAR-2 score, we additionally measure the average clause length and clause reduction ratio for each instances.
+Beside solved instances count and PAR-2 score, we additionally measure the average clause length and clause reduction ratio (both cumulative and non-cumulative)\footnote{The cumulative reduction ratio is obtained through learning all clauses with the target learning scheme; Therefore, the reduction is cumulative.
+The non-cumulative reduction ratio is obtained by running the target scheme for measurement only (the minimized 1-UIP clause is learned); Therefore, the reduction is not cumulative.
+}  for each instances.
 For $\MapleIUIPPURE$  and $\MapleIUIMIN$, we also captures the $\IUIP$ learning attempted rate and success rate.
@@
 \hline
-$\MapleIUIMIN$ & 226 & 4890.67 & 45.2 & 47.8\%\\ 
+$\MapleIUIMIN$ & 226 & 4890.67 & 45.2 & 47.8\%  (51.19\%)\\ 
 \hline
@@
 \end{center}
-\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.}
+\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.
+CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
 \label{fig:t1}

### 2019-11-25 sections/evaluation.tex
add new benchmark results

@@
 I partition the old group instances into six partitions of size 30 and one partition of size 20.
-Each partition is then assigned to a Intel CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory.
+Each partition is then assigned to a XeonE5-2 CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory.
 The new group is partitioned based on their contributor (e.g.
@@
 $\expSATShort$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
+We chose $\expSATShort$ because 1) it is the best solver in the 2019 SAT Race without using chronological backtracking; 2) the random walk simulation branching heuristic allows our learning schemes to partially sidestep the problem of variable activity.
 For each solver, we compare the base 1-UIP learning scheme against $\IUIP$ learning and the top two $\IUIP$ variants, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
@@
 \hline
-$\MapleSeven$-i-uip & TBA & TBA & TBA \\ 
+$\MapleSeven$-i-uip & \textbf{240} (\textbf{144}, 96) & \textbf{4601.25} & \textbf{36.97} \\ 
 \hline
-$\MapleSeven$-i-greedy & \textbf{237} (\textbf{140}, 97) & \textbf{4678.434} & \textbf{43.62} \\ 
+$\MapleSeven$-i-greedy & 237 (140, \textbf{97}) & 4678.434 & 43.62 \\ 
 \hline
-$\MapleSeven$-i-inclusive & TBA & TBA & TBA \\ 
+$\MapleSeven$-i-inclusive & 234 (137, 97) & 4718.03 & 37.96 \\ 
 \hline
@@
 \end{figure}
+
+Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers.
+All three configurations of $\IUIP$ outperformed 1-UIP on $\MapleSeven$.
+i-UIP, i-UIP-Greedy and i-UIP-inclusive solved 8, 5 and 2 more instances, respectively, whiling producing smaller clauses.
+The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction.
+This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because: 1) i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through unit propagation.
+2) using $\IUIP$ instead of 1-UIP scheme to analyze conflicts during vivification could derive shorter clauses.
+
+Similar to the observation on $\MapleSeven$, all three configurations of $\IUIP$ significantly outperformed 1-UIP on $\expSATShort$.
+i-UIP, i-UIP-Greedy and i-UIP-inclusive solved x, 7 and 8 more instances, respectively, while producing smaller clauses.
+We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because $\expSATShort$ allows the learning schemes to partially sidestep the activity problem.
+The learning adjustment for variable activities should have less impact on the solver's performance, and we observed the expected performance from the benchmark results.
+$\IUIP$ improved $\expSATShort$ the most across all compared solvers for both solved instances (+8) and PAR-2 score (-168), and this is might be due to the partially mitigation of the side-effect from variable activity.
+
+$\IUIP$ learning schemes show moderate improvement on $\MapleNineShort$ as i-UIP, i-UIP-greedy and i-UIP-inclusive solved 1, 0 and 2 more instances, respectively, while producing smaller clauses.
+Our learning schemes didn't improve $\MapleNineShort$ significantly because the side-effects from chronological backtracking (CB).
+More specifically, We believe CB prevents decision levels from being compressed through the process of backtracking and re-asserting literals.
+Comparing to 1-uip clause, the shorter i-uip clauses can bring related literals tighter in a decision level, and consequently produces smaller decision levels with  backtracking.
+However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.
+We believe we have observed an interesting interaction effect that shows the limitations of both CB and $\IUIP$ learning for future research.

### 2019-11-26 sections/evaluation.tex
add the final piece result

@@
 $\expSATShort$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
-We chose $\expSATShort$ because 1) it is the best solver in the 2019 SAT Race without using chronological backtracking; 2) the random walk simulation branching heuristic allows our learning schemes to partially sidestep the problem of variable activity.
+We chose $\expSATShort$ because 1) it is the best solver in the 2019 SAT Race without using chronological backtracking; 2) the combination of random walk simulation and variable activity branching heuristic allows our learning schemes to partially sidestep the problem of variable activity.
 For each solver, we compare the base 1-UIP learning scheme against $\IUIP$ learning and the top two $\IUIP$ variants, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
@@
 \hline
-$\expSATShort$-i-uip & TBA & TBA & TBA \\ 
+$\expSATShort$-i-uip & 241 (143, 98) & 4524.28 & 46.29 \\ 
 \hline
@@
 Similar to the observation on $\MapleSeven$, all three configurations of $\IUIP$ significantly outperformed 1-UIP on $\expSATShort$.
-i-UIP, i-UIP-Greedy and i-UIP-inclusive solved x, 7 and 8 more instances, respectively, while producing smaller clauses.
-We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because $\expSATShort$ allows the learning schemes to partially sidestep the activity problem.
-The learning adjustment for variable activities should have less impact on the solver's performance, and we observed the expected performance from the benchmark results.
-$\IUIP$ improved $\expSATShort$ the most across all compared solvers for both solved instances (+8) and PAR-2 score (-168), and this is might be due to the partially mitigation of the side-effect from variable activity.
+i-UIP, i-UIP-Greedy and i-UIP-inclusive solved 4, 7 and 8 more instances, respectively, while producing smaller clauses.
+We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance.
+However, we observed that both i-UIP-greedy and i-UIP-inclusive outperformed the default i-UIP learning scheme.
+One possible explanation is that i-UIP-greedy and i-UIP-inclusive schemes can easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration can use future search information to mintage the effect of the overcompensation.
+The i-UIP-greedy and i-UIP-inclusive schemes improved $\expSATShort$ the most among all the compared solvers for both solved instances (+7, +8) and PAR-2 score (-168, -153) possibly due to the aforementioned reasons.
 
@@
 More specifically, We believe CB prevents decision levels from being compressed through the process of backtracking and re-asserting literals.
-Comparing to 1-uip clause, the shorter i-uip clauses can bring related literals tighter in a decision level, and consequently produces smaller decision levels with  backtracking.
+Comparing to 1-uip clause, the shorter i-uip clauses can bring related literals tighter in a decision level, and consequently produces less and smaller decision levels with  backtracking.
 However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.

### 2019-11-26 sections/evaluation.tex
fix typos

@@
 Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with significantly smaller size than 1-UIP by 27.7\% and 20.7\%, respectively.
-Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the average clause length from $\IUIPMIN$ relative to 1-UIP.
-$\IUIPMIN$ learning produces shorter clauses for 88.25\% instances, and average relative reduction from 1-UIP is 18.685\%.Fig.~\ref{fig:len_compare} compares the absolute average clause size from $\IUIPMIN$ and 1-UIP, and it shows that $\IUIPMIN$ in general produces smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
+Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the average clause length reduction of $\IUIPMIN$ relative to 1-UIP.
+$\IUIPMIN$ learning produces shorter clauses for 88.25\% instances, and average relative reduction from 1-UIP is 18.685\%.
+Fig.~\ref{fig:len_compare} compares the absolute average clause size from $\IUIPMIN$ and 1-UIP, and it shows that $\IUIPMIN$ in general produces smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
 
@@
 However, we observed that both i-UIP-greedy and i-UIP-inclusive outperformed the default i-UIP learning scheme.
-One possible explanation is that i-UIP-greedy and i-UIP-inclusive schemes can easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration can use future search information to mintage the effect of the overcompensation.
-The i-UIP-greedy and i-UIP-inclusive schemes improved $\expSATShort$ the most among all the compared solvers for both solved instances (+7, +8) and PAR-2 score (-168, -153) possibly due to the aforementioned reasons.
+One possible explanation is that i-UIP-greedy and i-UIP-inclusive schemes can easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration can use future search information to mitigate the negative effects of our overcompensation.
+The i-UIP-greedy and i-UIP-inclusive schemes improved $\expSATShort$ the most among all the compared solvers for both solved instances (+7, +8) and PAR-2 score (-168, -153) possibly due to the aforementioned reason.
 
 $\IUIP$ learning schemes show moderate improvement on $\MapleNineShort$ as i-UIP, i-UIP-greedy and i-UIP-inclusive solved 1, 0 and 2 more instances, respectively, while producing smaller clauses.
-Our learning schemes didn't improve $\MapleNineShort$ significantly because the side-effects from chronological backtracking (CB).
+Our learning schemes didn't improve $\MapleNineShort$ significantly possibly due to the side-effects of chronological backtracking (CB).
 More specifically, We believe CB prevents decision levels from being compressed through the process of backtracking and re-asserting literals.
-Comparing to 1-uip clause, the shorter i-uip clauses can bring related literals tighter in a decision level, and consequently produces less and smaller decision levels with  backtracking.
+Comparing to 1-UIP clause, the shorter i-UIP clauses can bring related literals closer with shorter implications, and consequently produces less and smaller decision levels through  backtracking.
 However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.

### 2019-12-03 sections/evaluation.tex
address a reproducibility issue in comment

@@
 
+\nf{Benchmark results from running $\expSAT$ are harder to produce due to the randomness caused by the solver's strategy for selecting branching heuristics: the solver runs 2500s with LBR and the remaining 2500s with VSIDS.
+This is problematic for reproducing results because two runs of the same solver can be in different states at 2500s mark, and switching heuristics at different states will cause further divergence.
+The divergence can be quiet impactful (+- 2 instances at most for i-uip-active).
+For reproducible benchmark results, should we turn off the 2500s heuristic switching?
+Should we add a new section for the reproducible result?
+or remove the existing ones?}
+
 Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers.

### 2019-12-09 sections/i_uip.tex
i-uip pure clause can be minimized, add a step of minimization in the algo

@@
-\section{Find i-uip Clause}
+\section{Find i-uip Clause} \label{sec:i-uip}
 Classical CDCL SAT solvers analyze each conflict with 1-UIP resolution learning scheme to derive an asserting clause $\oneUIPClause$.
@@
 \EndFor \label{ln:endCompleteResolution}
+\\
+\If{$\IUIPPURE$}\label{ln: 2ndMinStart}   \Comment{$\iUIPClause$ from $\IUIPPURE$ can be minimized}
+    \State $\iUIPClause  \gets \textbf{Minimize}(\iUIPClause)$ 
+\EndIf\label{ln: 2ndMinEnd}
+\\
 \If{$|\iUIPClause| < |\oneUIPClause|$} \label{ln:checkSize}
@@
 The final $\iUIPClause$ produced by $\IUIPPURE$ will contains exactly one literal for levels where the LBD-persevering UIP exist.
-However, $\IUIPPURE$ does not minimize the number of literals in the decision levels where the LBD-preserving UIP does not exist.
+However, $\IUIPPURE$ does not minimize the number of literals in the decision levels where the LBD-preserving UIP does not exist, and clause minimization techniques\cite{} can be used to further reduce the clause size (line~\ref{ln: 2ndMinStart} to \ref{ln: 2ndMinEnd}).
 
-  When $\IUIP$ cannot resolve away a literal $p$ without increasing $\iUIPClause$'s LBD, instead of skipping the entire decision level $i$, a more practical solution $\IUIPMIN$ will mark and keep $p$ in $\iUIPClause$ (line~\ref{ln:mark}), and continue resolutions at level $i$.
+  When $\IUIP$ cannot resolve away a literal $p$ without increasing $\iUIPClause$'s LBD, instead of skipping the entire decision level $i$, a more practical solution $\IUIPMIN$ will mark and keep $p$ in $\iUIPClause$ (line~\ref{ln:mark}), and continue the resolutions at level $i$.
 $\IUIPMIN$ will ignore all the unsolvable literals and find the "local" unique implication point at every decision level.
 The algorithm terminates when there is exactly one unmarked literal left for each decision level of $\iUIPClause$, representing the local unique implication points.
+The clause $\iUIPClause$ learnt by
+  $\IUIPMIN$ is minimized\footnote{clause size cannot be reduced via minimization techniques\cite{}.}, we defer the minimization proof to Appendix~\ref{}.
 

### 2019-12-10 sections/evaluation.tex
add some new results from i-uip-pure, it did almost too well

@@
 \begin{center}
-\begin{tabular}{ | m{3.5cm} | m{2cm}| m{2cm} | m{2cm} | m{2.5cm} | } 
+\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
 \hline
@@
 \hline
-$\MapleBase$ & 221 & 5018.89 & 62.6 & 36.53\% \\ 
+$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
 \hline
-$\MapleIUIPPURE$ & 226 & 4920.04 & 49.6 & 41.6\% \\ 
+$\MapleIUIPPURE$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
 \hline
-$\MapleIUIMIN$ & 226 & 4890.67 & 45.2 & 47.8\%  (51.19\%)\\ 
+$\MapleIUIMIN$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
 \hline
@@
 Fig.~\ref{fig:len_compare} compares the absolute average clause size from $\IUIPMIN$ and 1-UIP, and it shows that $\IUIPMIN$ in general produces smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
-
-We also looked at the 14 instances solved by $\IUIPMIN$ but not by 1-UIP.
-$\IUIPMIN$ produces smaller clauses for all of them with average relative reduction of 22\% and maximum 77\% (30 vs 135).
-Seven out of 14 instances has size relative reduction over 30\%.
-For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for three instances and with average relative reduction of 3.3\%.
-
 \begin{figure}
     \centering
-    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{clause_length_PDF.png}
-    \caption{Average clause (relative to 1-UIP clauses) size distribution. X axis indicates the relative size difference, and Y axis indicates the PDF.}
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/factor_clause_length_PDF.png}
+    \caption{ Relative Clause size distribution. X axis indicates the relative size of i-uip clause (with respect to 1-uip clause) for each instance, and Y axis indicates the probability density.}
      \label{fig:len_pdf}
@@
     \caption{Average clause size comparison plot.
-Each point in the plot represents an benchmark instance.
+Each point in the plot represents an instance.
 X and Y axis shows the clause length from $\IUIP$ and 1-UIP, respectively.
@@
 \end{figure}
+
+We also looked at the 14 instances solved by $\IUIPMIN$ but not by 1-UIP.
+$\IUIPMIN$ produces smaller clauses for all of them with average relative reduction of 22\% and maximum 77\% (30 vs 135).
+Seven out of 14 instances has size relative reduction over 30\%.
+For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for three instances and with average relative reduction of 3.3\%.
 
@@
 \hline
-i-UIP & \textbf{226} (135, \textbf{91}) & 4890.67 & 45.2 \\ 
+\nf{i-UIP-PURE} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
 \hline
-i-UIP-greedy & \textbf{226} (135, \textbf{91})  & \textbf{4866.94} & 47.7 \\
+i-UIP & 226 (135, 91) & 4890.67 & 45.2 \\ 
+\hline
+i-UIP-greedy & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of 1-UIP ($\MapleBase$), i-UIP($\IUIPMIN$), i-UIP-Greedy,
+\caption{Benchmark results of 1-UIP ($\MapleBase$),i-UIP-PURE, i-UIP($\IUIPMIN$), i-UIP-Greedy,
 i-UIP-Inclusive, and i-UIP-Exclusive on SAT2019 race main track.}
@@
 \hline
-$\MapleSeven$-i-uip & \textbf{240} (\textbf{144}, 96) & \textbf{4601.25} & \textbf{36.97} \\ 
+$\MapleSeven$-i-uip & 240 (144, 96) & 4601.25 & \textbf{36.97} \\ 
 \hline
-$\MapleSeven$-i-greedy & 237 (140, \textbf{97}) & 4678.434 & 43.62 \\ 
+$\MapleSeven$-i-greedy & 237 (140, 97) & 4678.434 & 43.62 \\ 
 \hline
 $\MapleSeven$-i-inclusive & 234 (137, 97) & 4718.03 & 37.96 \\
+\hline
+$\MapleSeven$-i-pure & 245 (146, \nf{99}) & \nf{4488.30} & 43.76 \\
 \hline
@@
 This is problematic for reproducing results because two runs of the same solver can be in different states at 2500s mark, and switching heuristics at different states will cause further divergence.
-The divergence can be quiet impactful (+- 2 instances at most for i-uip-active).
+The result can be quiet impactful (+- 2 instances at most for i-uip-active).
 For reproducible benchmark results, should we turn off the 2500s heuristic switching?
@@
 The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction.
-This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because: 1) i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through unit propagation.
-2) using $\IUIP$ instead of 1-UIP scheme to analyze conflicts during vivification could derive shorter clauses.
+This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through more unit propagation.
 

### 2019-12-10 sections/evaluation.tex
revision and reformatting

@@
 \subsection{Clause Reduction with $\IUIP$}
-To evaluate $\IUIP$'s effectiveness as a clause reduction technique, we implement  $\IUIPPURE$ and $\IUIPMIN$  on top of \text{\MapleBase} \cite{}, the winner of SAT Race 2015 application track.
+To evaluate $\IUIP$'s effectiveness as a clause reduction technique, we implement  $\IUIPPURE$ and $\IUIPMIN$ with Control-I-UIP and Early Stopping rules in Sec~\ref{sec:i-uip} on top of \text{\MapleBase} \cite{}, the winner of SAT Race 2015 application track.
 We than compare the performance of the baseline $\MapleBase$ with $\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from SAT RACE 2019 main track.
@@
 
-Fig.~\ref{fig:t1} shows that both version of $\IUIP$ solved five more instances than the baseline solver with lower PAR-2 scores.
-$\IUIPMIN$ has marginally lower PAR-2 score than $\IUIPPURE$.
-Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with significantly smaller size than 1-UIP by 27.7\% and 20.7\%, respectively.
-Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the average clause length reduction of $\IUIPMIN$ relative to 1-UIP.
-$\IUIPMIN$ learning produces shorter clauses for 88.25\% instances, and average relative reduction from 1-UIP is 18.685\%.
-Fig.~\ref{fig:len_compare} compares the absolute average clause size from $\IUIPMIN$ and 1-UIP, and it shows that $\IUIPMIN$ in general produces smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
+Fig.~\ref{fig:t1} shows that $\IUIPPURE$ ($\IUIPPURE$) solved seven (five) more instances than the baseline solver with lower PAR-2 scores.
+$\IUIPPURE$ has marginally lower PAR-2 score than $\IUIPPURE$.
+Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with significantly smaller size than 1-UIP by  20.4\% and 27.7\%, respectively.
+Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the relative clause size of both i-uip learning schemes ($\IUIPPURE$ in red and $\IUIPMIN$ in green) for each instance.
+$\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 76.5\% (88.25\%) of instances, and average relative reduction from 1-UIP is 16.7\% (18.6\%).
+Fig.~\ref{fig:len_compare} compares the absolute clause size of $\IUIPMIN$, $\IUIPPURE$ and 1-UIP, and it shows that both i-UIP learning schemes in general produce the smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
+Moreover, $\IUIPMIN$ clauses (indicated in green) is consistently smaller than $\IUIPPURE$ clauses.
+
 \begin{figure}
@@
     \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/factor_clause_length_PDF.png}
-    \caption{ Relative Clause size distribution. X axis indicates the relative size of i-uip clause (with respect to 1-uip clause) for each instance, and Y axis indicates the probability density.}
+    \caption{ Relative Clause size distribution.
+X axis indicates the relative size of i-uip clause (with respect to 1-uip clause) for each instance, and Y axis indicates the probability density.
+$\IUIPMIN$ and $\IUIPPURE$ are illustrated with green and red, respectively}
      \label{fig:len_pdf}
@@
 \end{figure}
-
-We also looked at the 14 instances solved by $\IUIPMIN$ but not by 1-UIP.
+\nf{ Do we need this?
+We additionally looked at the 14 instances solved by $\IUIPMIN$ but not by 1-UIP.
 $\IUIPMIN$ produces smaller clauses for all of them with average relative reduction of 22\% and maximum 77\% (30 vs 135).
 Seven out of 14 instances has size relative reduction over 30\%.
-For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for three instances and with average relative reduction of 3.3\%.
+For the 9 instances solved by 1-UIP but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for three instances and with average relative reduction of 3.3\%.}
 
-$\IUIPMIN$ outperformed $\IUIPPURE$ in both PAR-2 score and clause size.
+$\IUIPMIN$ outperformed $\IUIPPURE$ in clause size.
 This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.
 Remark that the success of $\IUIP$ learning is determined by the size of the learned i-UIP clause $\iUIPClause$, and the $\IUIP$ learning frequency is also indirectly controlled by $\IUIP$'s success rate from the previous restart interval.
-The results indicates  $\IUIPMIN$ shortened $\iUIPClause$'s size through further minimization of $\iUIPClause$ at non-unique implication decision level.
+The results indicates  $\IUIPMIN$ shortened $\iUIPClause$'s size through further resolutions of $\iUIPClause$ at non-unique implication decision level.
 
@@
 A solver produce smaller clauses can construct smaller proofs.
-For UNSAT instances, we additionally measure their DRUP\cite{} proof checking time as well as the size of the optimized DRUP proof.
+For UNSAT instances, we measure their DRUP\cite{} proof checking time as well as the size of the optimized DRUP proof.
 We used DART-trim \cite{} with 5000 timeout to check and optimize DRUP proofs.
@@
 
-\subsection{$\IUIP$ as a Practical Learning Scheme}
-To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
-We evaluated four different configurations (i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances, PAR-2 score and average clause size.
-
-Fig.~\ref{fig:t4} summarizes the result of the experiment.
-Learning scheme i-UIP-greedy solves the same amount of instance (226) as i-UIP with less PAR-2 score.
-The inclusive activity adjustment solves the most SAT instances (138) and the least UNSAT instances (87).
-The exclusive activity adjustment scheme produces the shortest average clause size, but solved the second least instances, one more instance than the baseline.
-
 \begin{figure} 
@@
 \end{figure}
+
+\subsection{$\IUIP$ as a Practical Learning Scheme}
+To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
+We evaluated five configurations (i-UIP-PURE, i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances, PAR-2 score and average clause size.
+
+Fig.~\ref{fig:t4} summarizes the result of the experiment.
+Learning scheme i-UIP-PURE the most overall instances (228) and the most UNSAT instances (93).
+The i-uip-greedy scheme has the lowest PAR-2 score.
+The inclusive activity adjustment solves the most SAT instances (138).
+The exclusive activity adjustment scheme produces the shortest average clause size, but solved the second least instances, one more instance than the baseline.
+All configurations of $\IUIP$ outperform the baseline 1-UIP scheme in solved instances, PAR-2 score and average clause size.
 

### 2019-12-10 sections/i_uip.tex
revision and reformatting

@@
 
-  When $\IUIP$ cannot resolve away a literal $p$ without increasing $\iUIPClause$'s LBD, instead of skipping the entire decision level $i$, a more practical solution $\IUIPMIN$ will mark and keep $p$ in $\iUIPClause$ (line~\ref{ln:mark}), and continue the resolutions at level $i$.
+  When $\IUIP$ cannot resolve away a literal $p$ without increasing $\iUIPClause$'s LBD, instead of skipping the entire decision level $i$, another solution $\IUIPMIN$  simply marks literal $p$ and keep it  in $\iUIPClause$ (line~\ref{ln:mark}).
+The learning scheme then continue the resolutions at level $i$.
 $\IUIPMIN$ will ignore all the unsolvable literals and find the "local" unique implication point at every decision level.
@@
 The clause $\iUIPClause$ learnt by
-  $\IUIPMIN$ is minimized\footnote{clause size cannot be reduced via minimization techniques\cite{}.}, we defer the minimization proof to Appendix~\ref{}.
+  $\IUIPMIN$ is minimized\footnote{clause size cannot be reduced via minimization techniques\cite{}.}, we defer the proof of minimization to Appendix~\ref{}.
 
@@
 \subsection{Early stop i-UIP Learning}
-At any point of $\IUIP$ learning, if the number of marked literals in $\iUIPClause$ (literals which are forced into the clause to preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the final $\iUIPClause$ is at least the size of $\oneUIPClause$.
+At any point of $\IUIPMIN$ learning, if the number of marked literals in $\iUIPClause$ (literals which are forced into the clause to preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the final $\iUIPClause$ is at least the size of $\oneUIPClause$.
 The early stopping rule prevents solver from wasting time on traversing a large implication graph when $\IUIP$ has already failed.
+
+The early stopping rule may prematurely stops $\IUIPPURE$ learning which would otherwise succeed because $\iUIPClause$ can be further minimized.
+However, the experiments indicates such false negative cases are rare, and the early stopping rule generally helps the learning scheme's performance.
 
@@
 The scheme does not bump variable activities for transient literals that are resolved away at non-conflicting level because, unlike transient literals at conflicting level, these literals cannot be re-asserted after the immediate backtracking.
-Notice that other post-analyze extensions such as Reason Side Rate (RSR) and Locality\cite{} can be applied after applying $\IUIPActive$ on $\iUIPClause$ so that no literal's variable activity is double bumped.
+Notice that other post-analyze extensions such as Reason Side Rate (RSR) and Locality\cite{} can be applied after $\IUIPActive$.
 

### 2019-12-15 sections/evaluation.tex
Add 2018 SAT winner's data

@@
 \hline
-\nf{i-UIP-PURE} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
+\nf{$\IUIPPURE$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
 \hline
-i-UIP & 226 (135, 91) & 4890.67 & 45.2 \\ 
+$\IUIPMIN$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
 \hline
-i-UIP-greedy & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
+$\IUIPGreedy$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
 \hline
-i-UIP-active-Inclusive & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
+$\IUIPActive$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
 \hline
-i-UIP-active-Exclusive & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
+$\IUIPDist$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of 1-UIP ($\MapleBase$),i-UIP-PURE, i-UIP($\IUIPMIN$), i-UIP-Greedy,
-i-UIP-Inclusive, and i-UIP-Exclusive on SAT2019 race main track.}
+\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$,
+$\IUIPActive$, and $\IUIPDist$ on SAT2019 race main track.}
 \label{fig:t4}
@@
 \subsection{$\IUIP$ as a Practical Learning Scheme}
-To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
-We evaluated five configurations (i-UIP-PURE, i-UIP, i-UIP-Greedy, i-UIP-Inclusive, and i-UIP-Exclusive) of $\IUIP$ and 1-UIP learning on the SAT Race 2019 main track benchmark and report each configuration's solved instances, PAR-2 score and average clause size.
+To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we re-implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned in section~\ref{sec:i-uip}.
+We evaluated 1-UIP learning and five $\IUIP$ configurations ( $\IUIPMIN$, $\IUIPPURE$, $\IUIPGreedy$, $\IUIPActive$, and $\IUIPDist$) on the SAT Race 2019 main track benchmark and reported solved instances, PAR-2 score and average clause size.
 
 Fig.~\ref{fig:t4} summarizes the result of the experiment.
-Learning scheme i-UIP-PURE the most overall instances (228) and the most UNSAT instances (93).
-The i-uip-greedy scheme has the lowest PAR-2 score.
-The inclusive activity adjustment solves the most SAT instances (138).
-The exclusive activity adjustment scheme produces the shortest average clause size, but solved the second least instances, one more instance than the baseline.
-All configurations of $\IUIP$ outperform the baseline 1-UIP scheme in solved instances, PAR-2 score and average clause size.
-
-\nf{Inserted a table here, and graphs and analysis}
+Learning scheme $\IUIPPURE$ solved the most overall instances (228) and the most UNSAT instances (93).
+$\IUIPGreedy$ had the lowest PAR-2 score.
+$\IUIPActive$ solved the most SAT instances (138).
+$\IUIPDist$ produced the shortest average clause size, but solved the second least instances, one more instance than the baseline 1-UIP learning.
+All configurations of $\IUIP$ outperformed the baseline 1-UIP scheme in solved instances, PAR-2 score and average clause size.
 
 \subsection{$\IUIP$ on Modern SAT solvers}
-To validate $\IUIP$ as a generalizable learning scheme on modern SAT solvers, we re-implement $\IUIP$ on $\MapleSeven$ \cite{},  $\MapleNine$ \cite{}($\MapleNineShort$) and $\expSAT$\cite{} ($\expSATShort$).
-The first two solvers are the winner of 2017 and 2019 SAT race, respectively.
+To validate $\IUIP$ as a generalizable learning scheme on modern SAT solvers, we re-implement $\IUIP$ on the winners of 2017, 2018 and 2019 SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$).
 $\expSATShort$ is a top ten solver from 2019 SAT race which uses random walk simulation to help branching.
-We chose $\expSATShort$ because 1) it is the best solver in the 2019 SAT Race without using chronological backtracking; 2) the combination of random walk simulation and variable activity branching heuristic allows our learning schemes to partially sidestep the problem of variable activity.
-For each solver, we compare the base 1-UIP learning scheme against $\IUIP$ learning and the top two $\IUIP$ variants, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
+We chose $\expSATShort$ because 1) it is a top solver in the 2019 SAT Race without using chronological backtracking; 2) the combination of random walk simulation and variable activity branching heuristic allows our learning schemes to partially sidestep the problem of variable activity.
+For each solver, we compare the base 1-UIP learning scheme against $\IUIPPURE$, $\IUIPMIN$ and the top two $\IUIP$ variants, $\IUIPGreedy$ and $\IUIPActive$, on the SAT Race 2019 main track benchmark.
 We report solved instances, PAR-2 score and the average clause size.
@@
 \hline
+SAT 2017 Winner & & & \\
 $\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
 \hline
-$\MapleSeven$-i-uip & 240 (144, 96) & 4601.25 & \textbf{36.97} \\ 
+$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
 \hline
-$\MapleSeven$-i-greedy & 237 (140, 97) & 4678.434 & 43.62 \\ 
+$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
 \hline
-$\MapleSeven$-i-inclusive & 234 (137, 97) & 4718.03 & 37.96 \\
+$\MapleSeven$-i-greedy & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
 \hline
-$\MapleSeven$-i-pure & 245 (146, \nf{99}) & \nf{4488.30} & 43.76 \\
+$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
 \hline
 \hline
+SAT 2018 Winner & & & \\
+$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
+\hline
+$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
+\hline
+$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
+\hline
+$\MapleEightShort$-i-greedy & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
+\hline
+$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
+\hline
+\hline
+SAT 2019 Winner & & & \\
+$\MapleNineShort$ & 243 (145, \textbf{98}) & 4450.24 & 61.00 \\
+\hline
+$\MapleNineShort$-i-pure & 244 (148, 95) +0 & 4435.88 &  43.27\\
+\hline
+$\MapleNineShort$-i-min & 244 (148, 96) +1 & \textbf{4409.88} & \textbf{37.43} \\
+\hline
+$\MapleNineShort$-i-greedy & 243 (146, 97) + 0 & 4476.73 & 40.65 \\
+\hline
+$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) +2 & 4425.99 & 37.60 \\
+\hline
+\hline
+SAT 2019 Competitor & & &\\
 $\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
 \hline
-$\expSATShort$-i-uip & 241 (143, 98) & 4524.28 & 46.29 \\ 
+$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
 \hline
-$\expSATShort$-i-greedy & 244 (143, \textbf{101}) & \textbf{4460.92} & 47.25 \\
+$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
 \hline
-$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) & 4475.76 & \textbf{45.33} \\
+$\expSATShort$-i-greedy & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
 \hline
-\hline
-$\MapleNineShort$ & 243 (145, \textbf{98}) & 4450.24 & 61.00 \\
-\hline
-$\MapleNineShort$-i-uip & 244 (148, 96) & \textbf{4409.88} & \textbf{37.43} \\
-\hline
-$\MapleNineShort$-i-greedy & 243 (146, 97) & 4476.73 & 40.65 \\
-\hline
-$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) & 4425.99 & 37.60 \\
+$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of 1-UIP, i-UIP($\IUIPMIN$), i-UIP-Greedy,
-i-UIP-Inclusive on SAT2019 race main track.}
+\caption{Benchmark results of 1-UIP, $\IUIPPURE$. $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ on SAT2019 race main track.}
 \label{fig:t5}
@@
 Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers.
-All three configurations of $\IUIP$ outperformed 1-UIP on $\MapleSeven$.
-i-UIP, i-UIP-Greedy and i-UIP-inclusive solved 8, 5 and 2 more instances, respectively, whiling producing smaller clauses.
+All four configurations of $\IUIP$ outperformed 1-UIP learning for the SAT 2017 race winner, $\MapleSeven$.
+More specifically, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ solved 12, 8, 5 and 2 more instances, respectively, whiling producing smaller clauses.
+$\IUIPPURE$ solved more UNSAT and SAT instances while other configurations improved on solving SAT instances.
 The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction.
@@
 
-Similar to the observation on $\MapleSeven$, all three configurations of $\IUIP$ significantly outperformed 1-UIP on $\expSATShort$.
-i-UIP, i-UIP-Greedy and i-UIP-inclusive solved 4, 7 and 8 more instances, respectively, while producing smaller clauses.
-We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance.
-However, we observed that both i-UIP-greedy and i-UIP-inclusive outperformed the default i-UIP learning scheme.
-One possible explanation is that i-UIP-greedy and i-UIP-inclusive schemes can easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration can use future search information to mitigate the negative effects of our overcompensation.
-The i-UIP-greedy and i-UIP-inclusive schemes improved $\expSATShort$ the most among all the compared solvers for both solved instances (+7, +8) and PAR-2 score (-168, -153) possibly due to the aforementioned reason.
-
-$\IUIP$ learning schemes show moderate improvement on $\MapleNineShort$ as i-UIP, i-UIP-greedy and i-UIP-inclusive solved 1, 0 and 2 more instances, respectively, while producing smaller clauses.
-Our learning schemes didn't improve $\MapleNineShort$ significantly possibly due to the side-effects of chronological backtracking (CB).
-More specifically, We believe CB prevents decision levels from being compressed through the process of backtracking and re-asserting literals.
+We observed significant improvement of $\IUIP$ for the SAT 2018 winner, $\MapleEightShort$.
+Three out of four $\IUIP$ configurations outperformed 1-UIP by 5, 4 an 4 instances, respectively.
+$\MapleEightShort$ improved from $\MapleSeven$ by using chronological backtracking (CB) for long distance backtracks.
+Therefore, we used $\IUIP$-CB extension for all $\IUIP$ configurations.
+The results indicates that the improvement of $\IUIP$ is slightly shadowed by the adoption of CB.
+More specifically, we believe CB prevents decision levels from being compressed through the process of backtracking and literal assertion.
 Comparing to 1-UIP clause, the shorter i-UIP clauses can bring related literals closer with shorter implications, and consequently produces less and smaller decision levels through  backtracking.
@@
 We believe we have observed an interesting interaction effect that shows the limitations of both CB and $\IUIP$ learning for future research.
+
+$\IUIP$ learning schemes show marginal improvement on $\MapleNineShort$.
+Despite significantly reducing clause size,  only two out of four configurations improved solved instances by 1 and 2 instances, respectively.
+$\MapleNineShort$ remembers previously learned clauses, and prioritize the clause if the same clause (denote as duplicate clause) is learned again later.
+$\IUIP$ and duplicate learning (DL) didn't synergies well possibly because multiple 1-UIP clause can be derived to the same i-UIP clause through the $\IUIP$; therefore, duplicate clauses could appear too frequently for duplicate learning to prioritize the right clauses.
+
+We observed significant improvement of $\IUIP$ for $\MapleSeven$, three out of four configurations of $\IUIP$ significantly outperformed 1-UIP learning by 4, 7 and 8 more instances, respectively, while producing smaller clauses.
+We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance.
+However, we observed that both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP learning scheme.
+One possible explanation is that $\IUIPGreedy$ and $\IUIPActive$ schemes could easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration could use future search information to mitigate the negative effects of our overcompensation.

### 2019-12-16 sections/evaluation.tex
fix typo, add a follow up question

@@
 
-$\IUIP$ learning schemes show marginal improvement on $\MapleNineShort$.
+$\IUIP$ learning schemes show marginal improvement for $\MapleNineShort$.
 Despite significantly reducing clause size,  only two out of four configurations improved solved instances by 1 and 2 instances, respectively.
-$\MapleNineShort$ remembers previously learned clauses, and prioritize the clause if the same clause (denote as duplicate clause) is learned again later.
-$\IUIP$ and duplicate learning (DL) didn't synergies well possibly because multiple 1-UIP clause can be derived to the same i-UIP clause through the $\IUIP$; therefore, duplicate clauses could appear too frequently for duplicate learning to prioritize the right clauses.
+$\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
+$\IUIP$ and DL didn't synergies well possibly because different 1-UIP clause can be reduced to the same i-UIP clause through the $\IUIP$ learning; therefore, DL could encounter too many duplicate i-uip clauses to correctly prioritize the important clauses.
+\nf{should I run some experiment to validate my guess?, if the guess is valid, should I adjust DL settings to make it compatible with $\IUIP$}
 

### 2019-12-17 sections/evaluation.tex
rerun experiments with the same machine settings to measure DL rate, update results

@@
 SAT 2019 Winner & & & \\
-$\MapleNineShort$ & 243 (145, \textbf{98}) & 4450.24 & 61.00 \\
+$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
 \hline
-$\MapleNineShort$-i-pure & 244 (148, 95) +0 & 4435.88 &  43.27\\
+$\MapleNineShort$-i-pure & 238 (140, 98) +0 & 4519.08 &  43.32\\
 \hline
-$\MapleNineShort$-i-min & 244 (148, 96) +1 & \textbf{4409.88} & \textbf{37.43} \\
+$\MapleNineShort$-i-min & 244 (148, 96) +6 & \textbf{4409.88} & \textbf{37.43} \\
 \hline
-$\MapleNineShort$-i-greedy & 243 (146, 97) + 0 & 4476.73 & 40.65 \\
+$\MapleNineShort$-i-greedy & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
 \hline
-$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) +2 & 4425.99 & 37.60 \\
+$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) +7 & 4425.99 & 37.60 \\
 \hline
@@
 
-$\IUIP$ learning schemes show marginal improvement for $\MapleNineShort$.
-Despite significantly reducing clause size,  only two out of four configurations improved solved instances by 1 and 2 instances, respectively.
+$\IUIP$ learning schemes showed significant improvement for $\MapleNineShort$.
+Three out of four configurations improved solved instances by 6, 5 and 7 instances, respectively.
 $\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
-$\IUIP$ and DL didn't synergies well possibly because different 1-UIP clause can be reduced to the same i-UIP clause through the $\IUIP$ learning; therefore, DL could encounter too many duplicate i-uip clauses to correctly prioritize the important clauses.
+$\IUIP$ and DL didn't synergies well possibly because different 1-UIP clause can be reduced to the same i-UIP clause through $\IUIP$ learning; therefore, DL could encounter too many duplicate i-uip clauses to correctly prioritize the important clauses.
 \nf{should I run some experiment to validate my guess?, if the guess is valid, should I adjust DL settings to make it compatible with $\IUIP$}

### 2019-12-17 sections/evaluation.tex
summarize DL experiment observation

@@
 \hline
-$\MapleNineShort$-i-pure & 238 (140, 98) +0 & 4519.08 &  43.32\\
+$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
 \hline
@@
 $\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
-$\IUIP$ and DL didn't synergies well possibly because different 1-UIP clause can be reduced to the same i-UIP clause through $\IUIP$ learning; therefore, DL could encounter too many duplicate i-uip clauses to correctly prioritize the important clauses.
-\nf{should I run some experiment to validate my guess?, if the guess is valid, should I adjust DL settings to make it compatible with $\IUIP$}
+$\IUIP$ and DL didn't synergies well possibly because i-uip clauses are less likely to be duplicated.
+We observed that $\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the core clause database.
+One possible explanation is that the conflict blocked by a single 1-uip clauses can be alternatively blocked by one of many i-uip clauses.
+Therefore, the same 1-uip clause can be reduced to different i-uip clauses depending on the solver's assertion trail.
 
-We observed significant improvement of $\IUIP$ for $\MapleSeven$, three out of four configurations of $\IUIP$ significantly outperformed 1-UIP learning by 4, 7 and 8 more instances, respectively, while producing smaller clauses.
+We observed significant improvement of $\IUIP$ for $\expSATShort$, three out of four configurations of $\IUIP$ significantly outperformed 1-UIP learning by 4, 7 and 8 more instances, respectively, while producing smaller clauses.
 We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance.

### 2019-12-17 sections/evaluation.tex
update clause size reduction representation and description

@@
 Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the relative clause size of both i-uip learning schemes ($\IUIPPURE$ in red and $\IUIPMIN$ in green) for each instance.
-$\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 76.5\% (88.25\%) of instances, and average relative reduction from 1-UIP is 16.7\% (18.6\%).
+$\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 77.7\% (88.5\%) of instances, and average relative reduction from 1-UIP is 16.7\% (18.6\%).
 Fig.~\ref{fig:len_compare} compares the absolute clause size of $\IUIPMIN$, $\IUIPPURE$ and 1-UIP, and it shows that both i-UIP learning schemes in general produce the smaller clauses, and the size reduction is more significant for instances with large average 1-UIP clause size.
@@
     \centering
-    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/factor_clause_length_PDF.png}
-    \caption{ Relative Clause size distribution.
-X axis indicates the relative size of i-uip clause (with respect to 1-uip clause) for each instance, and Y axis indicates the probability density.
-$\IUIPMIN$ and $\IUIPPURE$ are illustrated with green and red, respectively}
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/clause_reduction_PDF.png}
+     \caption{ Relative clause size reduction distribution.
+X axis indicates the relative size of difference between i-uip and 1-uip clauses (calculated as $\dfrac{|\oneUIPClause|-|\iUIPClause|}{|\oneUIPClause|}$ ) for each instance, and Y axis shows the probability density.}
      \label{fig:len_pdf}

### 2020-01-27 sections/evaluation.tex
rerun experiment for data validation
sync up once again

@@
 The benchmark contains 400 instances divided into two groups of 200, new and old, representing historical instances and fresh instances in the 2019 race, respectively.
-I partition the old group instances into six partitions of size 30 and one partition of size 20.
+We randomly partition the old group instances into six partitions of size 30 and one partition of size 20.
 Each partition is then assigned to a XeonE5-2 CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory.
@@
 This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.
-Remark that the success of $\IUIP$ learning is determined by the size of the learned i-UIP clause $\iUIPClause$, and the $\IUIP$ learning frequency is also indirectly controlled by $\IUIP$'s success rate from the previous restart interval.
-The results indicates  $\IUIPMIN$ shortened $\iUIPClause$'s size through further resolutions of $\iUIPClause$ at non-unique implication decision level.
 
@@
 \begin{center} 
-\begin{tabular}{ | m{3.5cm} | m{5cm}| m{5cm} | } 
+\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
 \hline
@@
 \hline
-$\MapleNineShort$-i-min & 244 (148, 96) +6 & \textbf{4409.88} & \textbf{37.43} \\
+$\MapleNineShort$-i-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
 \hline
@@
 \hline
-$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) +7 & 4425.99 & 37.60 \\
+$\MapleNineShort$-i-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
 \hline
@@
 
-\nf{Benchmark results from running $\expSAT$ are harder to produce due to the randomness caused by the solver's strategy for selecting branching heuristics: the solver runs 2500s with LBR and the remaining 2500s with VSIDS.
-This is problematic for reproducing results because two runs of the same solver can be in different states at 2500s mark, and switching heuristics at different states will cause further divergence.
-The result can be quiet impactful (+- 2 instances at most for i-uip-active).
-For reproducible benchmark results, should we turn off the 2500s heuristic switching?
-Should we add a new section for the reproducible result?
-or remove the existing ones?}
-
 Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers.
@@
 The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction.
-This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through more unit propagation.
+This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through unit propagation.
 
@@
 More specifically, we believe CB prevents decision levels from being compressed through the process of backtracking and literal assertion.
-Comparing to 1-UIP clause, the shorter i-UIP clauses can bring related literals closer with shorter implications, and consequently produces less and smaller decision levels through  backtracking.
+The shorter i-UIP clauses can bring related literals closer, which in turns compresses the assertion stack and the decision levels.
 However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.
@@
 $\IUIP$ learning schemes showed significant improvement for $\MapleNineShort$.
-Three out of four configurations improved solved instances by 6, 5 and 7 instances, respectively.
+Three out of four configurations improved solved instances by 6, 5 and 5 instances, respectively.
 $\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
-$\IUIP$ and DL didn't synergies well possibly because i-uip clauses are less likely to be duplicated.
-We observed that $\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the core clause database.
-One possible explanation is that the conflict blocked by a single 1-uip clauses can be alternatively blocked by one of many i-uip clauses.
-Therefore, the same 1-uip clause can be reduced to different i-uip clauses depending on the solver's assertion trail.
+$\IUIP$ and DL didn't synergies well possibly because i-UIP clauses are less likely to be duplicated.
+We observed that $\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the core clause database than 1-UIP.
+One possible explanation is that $\IUIP$ learning can reduce an 1-UIP clause to different i-UIP clauses for different assertion trails.
+Therefore, the solver is less likely to learn the same i-UIP clauses.
 
@@
 We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance.
-However, we observed that both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP learning scheme.
+However, we instead, observed that both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP learning scheme.
 One possible explanation is that $\IUIPGreedy$ and $\IUIPActive$ schemes could easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration could use future search information to mitigate the negative effects of our overcompensation.

### 2020-02-02 sections/i_uip.tex
sync up once again
add an example to illstruate i-uip, i-uip-pure and i-uip-min

@@
   $\IUIPMIN$ is minimized\footnote{clause size cannot be reduced via minimization techniques\cite{}.}, we defer the proof of minimization to Appendix~\ref{}.
+
+\begin{example}
+Consider the assertion trail $\assertionTrail = {..a_2,b_2,c_2,d_2..e_5,f_5,g_5,h_6,i_6,j_6,k_6..}$ where literals $a_2 ...
+k_{6}$ are asserted in order with decision level shown as the underscript.
+Let $L_{ext}$ denote the set of literals on $\assertionTrail$ outside of decision level 2,5,6 and 10, and $l^{*}_{ext}$ be some literal in $L_{ext}$.
+The reasoning clauses involved in building $\assertionTrail$ are:
+\begin{enumerate}
+  \item[] $C_k = \{\neg{f_5} \vee \neg{j_6} \vee k_6\}$,
+  $C_j = \{\neg{f_5} \vee \neg{i_6} \vee j_6\}$, 
+  $C_i = \{\neg{e_5} \vee \neg{h_6} \vee i_6\}$,
+  $C_g = \{\neg{a_2} \vee \neg{f_5} \vee g_5\}$
+  $C_f = \{\neg{l^{*}_{ext}} \vee \neg{e_5} \vee f_5\}$,
+  $C_d = \{\neg{b_2} \vee \neg{c_2} \vee d_2\}$,
+  $C_c = \{\neg{a_2} \vee \neg{b_2} \vee c_2 \}$,
+  $C_b = \{\neg{l^{*}_{ext}} \vee \neg{a_2} \vee b_2\}$
+\end{enumerate}
+Suppose 1-UIP learning produces clause 
+$\oneUIPClause = \{\neg{m_{10}} \vee \neg{k_6} \vee \neg{j_6} \vee \neg{i_6} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{d_2} \vee \neg{c_2}\}$ where $\neg{k_{10}}$ is the UIP from the conflicting level.
+$\IUIP$ first tries to find the UIP for level 6 by resolving $\oneUIPClause$ with $C_k$, $C_j$ and, $C_i$ in order and produces clause
+$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee  \neg{c_2} \}$ where $\neg{h_6}$ is the UIP.
+
+$\IUIP$ attempts to find the UIP for level 5 by resolving $C_i$ with 
+$C_g$ and $C_f$ in order. However, resolving with $C_f$ will introduce $\neg{l^{*}_{ext}}$ into 
+$C_i$ and increase the clause's LBD.
+$\IUIPPURE$ undoes all resolutions at level 5 and skips the entire level, whereas $\IUIPMIN$ only skips the resolution with $C_f$ and tries to find the UIP by ignoring $\neg{f_5}$.
+After level 5, $\IUIPPURE$ preserves the same clause and $\IUIPMIN$ produces $\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \vee \neg{a_2}\}$ where $\neg{e_5}$ is the UIP when $\neg{f_5}$ is ignored.
+The same $\IUIP$ procedure is then performed on level 2 to produce the final $\iUIPClause$.
+
+$\IUIPPURE$ learned $\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee  \neg{c_2} \}$ and reduces the learnt clause size by 1.
+$\IUIPMIN$ learned $\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee \neg{e_5} \vee \neg{b_2} \vee \neg{a_2}\}$ and reduces the learnt clause size by 2.
+\end{example}
 

### 2020-02-06 paper.tex
describe first part of the experiment
update marco and package user
added marco's for comments.
make change to optimization 1
add length comparsion and proof size graphs
added author name
first fb changes

@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
 \documentclass[runningheads]{llncs}
-%
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
+\usepackage{times}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{algorithm}
 \usepackage{algpseudocode}
+%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
 \usepackage{xspace}
 \usepackage{mathtools}
-\usepackage{amsmath}
 \usepackage{graphicx}
+\usepackage{array}
+\usepackage{xcolor}
+\usepackage{theorem}
+\usepackage[T1,hyphens]{url}
 
 \begin{document}
-%
-\title{Contribution Title\thanks{Supported by organization x.}}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\input{marco}
+\title{$i$-UIP Clause Learning}
+\author{Nick Feng \and Fahiem Bacchus}
+\authorrunning{N. Feng and F. Bacchus}
+\institute{Department of Computer Science, University of Toronto, Canada, 
+  \email{\{nfeng,fbacchus\}@cs.toronto.edu}}
+
+\newcommand{\sat}{SAT\xspace}
+\newcommand{\code}[1]{\text{#1}}
+\newcommand{\inputF}{\ensuremath{F}}
+\newcommand{\oneUIPClause}{\ensuremath{C_{1}}}
+\newcommand{\iUIPClause}{\ensuremath{C_{i}}}
+\newcommand{\assertionTrail}{\ensuremath{\phi}}
+\newcommand{\LBD}{\text{LBD}}
+\newcommand{\GAP}{\text{Gap}}
+\newcommand{\IUIP}{\textbf{$i$-UIP}}
+\newcommand{\IUIPPURE}{\textbf{Pure-i-UIP}}
+\newcommand{\IUIPMIN}{\textbf{Min-i-UIP}}
+\newcommand{\IUIPGreedy}{\textbf{i-UIP-Greedy}}
+\newcommand{\IUIPActive}{\textbf{i-UIP-Inclusive}}
+\newcommand{\IUIPDist}{\textbf{i-UIP-Exclusive}}
+\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
+\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
+\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
+\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
+\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
+\newcommand{\expSAT}{\textit{expMaple\_CM\_GCBumpOnlyLRB} }
+\newcommand{\expSATShort}{\textit{expMaple} }
+\newcommand{\MapleIUIPPURE}{\text{Maple-\IUIPPURE}}
+\newcommand{\MapleIUIMIN}{\text{Maple-\IUIPMIN}}
+\newcommand{\MapleEightShort}{\textit{MapleCB}}
+\newcommand\Set[2]{\{\,#1\mid#2\,\}}
+\newcommand\Array[2]{[\,#1\mid#2\,]}
+\newcommand\target{\textbf{Succ_t}}
+\newcommand\trail{\ensuremath{\mathcal{T}}}
+\newcommand\trailIdx[1]{\ensuremath{\iota(#1)}}
+\newcommand\var{\text{var}}
+\newcommand\true{\textsc{true}\xspace}
+\newcommand\false{\textsc{false}\xspace}
+\newcommand\reason{\textit{reason}}
+\setlength{\theorempreskipamount}{3pt}
+\setlength{\theorempostskipamount}{3pt}
+\newtheorem{Obs}{Observation}
+\newtheorem{Cor}{Corollary}
+\newtheorem{defn}{Definition}
+\newtheorem{thm}{Theorem}
+
+\newcommand{\whitebox}{\raisebox{.5ex}{\fbox{\hspace*{.2ex}}}}
+
+%Comments
+\newcommand{\nf}[1]{\textcolor{red}{#1}}
+\newcommand{\fb}[1]{\textcolor{blue}{#1}}
+
 \maketitle              % typeset the header of the contribution
 % 
-\input{sections/abstract.tex}
-\input{sections/introduction.tex}
-\input{sections/preliminaries.tex}
-\input{sections/i_uip.tex}
-\input{sections/evaluation.tex}
-%
-\begin{thebibliography}{8}
+\begin{abstract}
+\end{abstract}
 
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
+\section{Introduction}
+Clause learning is an essential technique in \sat solvers. And there
+is good empirical evidence that clause learning is, in fact, the most
+important technique used in modern \sat solvers
+\cite{DBLP:conf/sat/KatebiSS11}. In early \sat research a number of
+different clause learning techniques were proposed
+\cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
+However, following the revolutionary performance improvements achieved
+by the Chaff \sat solver, the field has converged on using the 1-UIP
+(first Unique Implication Point) scheme
+\cite{DBLP:conf/iccad/ZhangMMM01} employed in Chaff
+\cite{DBLP:conf/dac/MoskewiczMZZM01} (as well as other techniques
+pioneered in the Chaff solver).\footnote{The idea of UIP clauses was
+  first mentioned in \cite{DBLP:journals/tc/Marques-SilvaS99}, and
+  1-UIP clauses along with other UIP clauses were learnt and used in
+  the earlier GRASP \sat solver.} Since then almost all \sat solvers
+have employed the 1-UIP clause learning scheme, along with clause
+minimization \cite{DBLP:conf/sat/SorenssonB09}, as their primary
+method for learning new clauses.
+
+However, as this early work showed, other clause learning schemes can
+be used in SAT solvers without changes to their main data
+structures. Furthermore, advances in our understanding allow us to
+better understand the potential advantages and disadvantages of these
+alternate schemes. In this paper we reexamine some of these previous
+clause learning schemes, with a focus on the $i$-UIP schme described
+in \cite{DBLP:conf/iccad/ZhangMMM01}. With current ideas we can see
+that in its original form $i$-UIP clause learning suffers a
+significant disadvantages over 1-UIP clause learning. We present a
+number of different methods to impove $i$-UIP clause learning so as to
+over come these disadvantages. We show that these different methods
+for $i$-UIP clause learning can provide useful performance gains over
+the standard 1-UIP scheme.
+
+\section{Background}
+We assume the reader has some familiarity with Conjunctive Normal Form
+(CNF) formulas, conflict-driven clause learning (CDCL) \sat solvers
+\cite{DBLP:series/faia/SilvaLM09}, unit-propagation, and conflict
+analysis.
+
+\begin{enumerate}
+\item Trail; Reasons and their properties; decision levels, trail order.
+\item conflict clause. Trail has the property that conflict exists
+    only at the deepest level.
+\item Clauses derived by trail resolution. Identified by variables
+    appearing on the trail.
+\item No point in doing this in non-trail order.
+\item a specific case of trivial resolutions.
+\item Clause derived by trail resolution iff it is a cut in the
+    implication graph.
+\item Note that all implemented clause learning techniques use
+    sequences of trail resolutions. In particular, they all start with
+    a specific conflict and resolve only with reason clauses.
+    This includes clause minimization.
+\item Asserting clause---one lit at deepest level of trail. Only
+    asserting clauses are gauranteed to be new.
+\item 1-UIP is asserting clause generated with fewest trail resolution
+    steps. Each trail resolution step can only increase the number of
+    levels in the clause. So LBD is minimized by 1-UIP clauses.
+\item Description of i-uip, all-uip, all decisions (useful in some cases)
+\end{enumerate}
+
+CDCL \sat solvers maintain a \textbf{trail} which is a sequence of
+literals that have been assigned \true by the solver. Literals newly
+assigned \true are placed at the end of the trail, and literals are
+always unassigned and removed from the end of the trail first;
+i.e. the trail acts as a stack.
+
+Let $\trail$ be the trail. For any literal $\ell$ we let
+$\trailIdx{\ell}$ be the index of $\ell$ in $\trail$. That is,
+$\trail[\trailIdx{\ell}] = \ell$. If $\ell$ is not in the trail, then
+$\trailIdx{\ell} = \varnothing$, and we use the convention that
+$\trail[\varnothing] = \varnothing$. The literals on \trail are
+sequenced in the order they were assigned \true. Hence
+$\trailIdx{\ell} < \trailIdx{\ell'}$ implies that $\ell$ was assigned
+\true before $\ell'$. When $\trailIdx{\ell} < \trailIdx{\ell'}$ we say
+that $\ell$ \textit{appears above} $\ell'$ on the trail. We use the
+convention that $\trailIdx{\ell} < \trailIdx{\ell'}$ whenever $\ell$
+is in the trail and $\ell'$ is not; i.e., $i < \varnothing$ for any
+index $i$.
+
+The literals on $\trail$ have been assigned \true by the solver by
+either (a) having been set as a decisions or (b) having forced by
+unit-propagation. If $\ell\in \trail$ was was set by unit-propgation
+it will have an associated reason clause $\reason(\ell)$. This clause
+has two properties (1) $\ell\in \reason(\ell)$ and (2) for every
+$x\in \reason(\ell)$ we have that
+$\trailIdx{\lnot x} < \trailIdx{\ell}$. That is, $\reason(\ell)$ is a
+clause that implies that $\ell$ must assigned \true because every
+other literal in it has been assigned \false above $\ell$ in
+$\trail$. Literals $\ell$ set as decsions do not have reason clauses
+and we let $\reason(\ell) = \varnothing$.
+
+CDCL \sat solvers perform clause learning when the literals on the
+trail falsify a clause $c$. '
+
+\section{Find i-uip Clause}
+\label{sec:i-uip}
+Classical CDCL SAT solvers analyze each conflict with 1-UIP resolution
+learning scheme to derive an asserting clause $\oneUIPClause$. Even
+though prior studies shows that further resolutions on $\oneUIPClause$
+against the assertion trail $\assertionTrail$ cannot reduce the
+clause's literal block distance ($\LBD$), the resolutions could
+potentially reduce the clause's size. Clause size is an important
+quality measurement because a smaller clause 1) consumes less memory,
+2) requires less steps to force a literal and 3) decreases the size of
+future conflict clauses.
+
+Inspired by both $\LBD$ and clause size, i-UIP resolution learning
+attempts to resolve away literals in $\oneUIPClause$ against the
+assertion trail $\assertionTrail$ to minimize the clause size without
+increasing the clause's $\LBD$. The goal of i-UIP learning is to find
+a clause $\iUIPClause$ whose literals are either the unique
+implication points in their respective decision levels or are directly
+implied by literals from a foreign decision
+level. Alg. ~\ref{alg:i-uip} is a pseudo-code implementation of i-UIP
+learning.
+
+\begin{algorithm}[t]
+\caption{\IUIP}\label{alg:i-uip}
+\begin{algorithmic}[1]
+\Require  $\oneUIPClause$ is a valid and minimized 1-UIP clause
+\Require  $\assertionTrail$ is a valid assertion trail
+\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$} 
+    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause} \label{ln:init}
+    \State \code{DecisionLvs} $\gets$ Decision levels in $\oneUIPClause$ in descending order
+\For{$i \in \code{DecisionLvs}$} \label{ln:iterDecisionLvls}
+    \State $L_{i} \gets \Set{l}{\code{level}(l) = i \wedge l \in \iUIPClause}$
+    \While{$|\code{unmarked}(L_i)| > 1$} \label{ln:uip}
+        \State $p \gets \text{lit with the highest trail position in } L_{i} $ \label{ln:trailaccess}
+        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs}} \label{ln:resolvable}
+            \If{$\IUIPPURE$}
+                \State Unresolve literals at level $i$  \label{ln:unresolve}
+                \State Go to the next decision level $i$ \label{ln:abondon}
+            \ElsIf{$\IUIPMIN$}
+                \State $\code{Mark}(p)$ \label{ln:mark}
+            \EndIf
+        \Else 
+            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(\neg{p}, \assertionTrail)$
+           \State  $\code{Update}(L_{i})$
+        \EndIf
+    \EndWhile
+\EndFor \label{ln:endCompleteResolution}
+\\
+\If{$\IUIPPURE$}\label{ln: 2ndMinStart}   \Comment{$\iUIPClause$ from $\IUIPPURE$ can be minimized}
+    \State $\iUIPClause  \gets \textbf{Minimize}(\iUIPClause)$ 
+\EndIf\label{ln: 2ndMinEnd}
+\\
+\If{$|\iUIPClause| < |\oneUIPClause|$} \label{ln:checkSize}
+    \State \textbf{return} $\iUIPClause$
+\Else
+    \State \textbf{return} $\oneUIPClause$
+\EndIf  \label{ln:cSizeCheck}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+The algorithm $\IUIP$ computes $\iUIPClause$ from (line~\ref{ln:init}
+to \ref{ln:endCompleteResolution}), and then returns the smaller
+clause between $\oneUIPClause$ and $\iUIPClause$
+(line~\ref{ln:checkSize} to \ref{ln:cSizeCheck}).  The algorithm first
+initializes $\iUIPClause$ with the minimized\cite{}
+$\oneUIPClause$. Next, the algorithm iterate through the decision
+levels of $\iUIPClause$ in descending order
+(line~\ref{ln:iterDecisionLvls}) and tries to find the unique
+implication point in each level (line~\ref{ln:uip}). Since $\IUIP$
+needs to preserve $\iUIPClause$'s $\LBD$ during resolutions, before
+resolving away a target literal $p$, the algorithm preemptively checks
+whether the reasoning clause of $\neg{p}$ contains any literal $q$
+from an unseen decision level (line~\ref{ln:resolvable}). If such a
+literal $q$ exists, then the algorithm cannot find the unique
+implication point (UIP) at the current level $i$ because resolving
+away $p$ will introduce $q$ into the clause and increases $\LBD$. One
+solution is to abandon level $i$ by reverting back to the state before
+any literal at level $i$ is resolved away (line~\ref{ln:unresolve}),
+and then move on to the next level (line~\ref{ln:abondon}). We denote
+$\IUIP$ learning with this solution as $\IUIPPURE$. The final
+$\iUIPClause$ produced by $\IUIPPURE$ will contains exactly one
+literal for levels where the LBD-persevering UIP exist. However,
+$\IUIPPURE$ does not minimize the number of literals in the decision
+levels where the LBD-preserving UIP does not exist, and clause
+minimization techniques\cite{} can be used to further reduce the
+clause size (line~\ref{ln: 2ndMinStart} to \ref{ln: 2ndMinEnd}).
+
+When $\IUIP$ cannot resolve away a literal $p$ without increasing
+$\iUIPClause$'s LBD, instead of skipping the entire decision level
+$i$, another solution $\IUIPMIN$ simply marks literal $p$ and keep it
+in $\iUIPClause$ (line~\ref{ln:mark}). The learning scheme then
+continue the resolutions at level $i$. $\IUIPMIN$ will ignore all the
+unsolvable literals and find the "local" unique implication point at
+every decision level.  The algorithm terminates when there is exactly
+one unmarked literal left for each decision level of $\iUIPClause$,
+representing the local unique implication points. The clause
+$\iUIPClause$ learnt by $\IUIPMIN$ is minimized\footnote{clause size
+  cannot be reduced via minimization techniques\cite{}.}, we defer the
+proof of minimization to Appendix~\ref{}.
+
+\begin{example}
+    Consider the assertion trail
+    $\assertionTrail =
+    {..a_2,b_2,c_2,d_2..e_5,f_5,g_5,h_6,i_6,j_6,k_6..}$ where literals
+    $a_2 ... k_{6}$ are asserted in order with decision level shown as
+    the underscript. Let $L_{ext}$ denote the set of literals on
+    $\assertionTrail$ outside of decision level 2,5,6 and 10, and
+    $l^{*}_{ext}$ be some literal in $L_{ext}$. The reasoning clauses
+    involved in building $\assertionTrail$ are:
+\begin{enumerate}
+  \item[] $C_k = \{\neg{f_5} \vee \neg{j_6} \vee k_6\}$,
+  $C_j = \{\neg{f_5} \vee \neg{i_6} \vee j_6\}$, 
+  $C_i = \{\neg{e_5} \vee \neg{h_6} \vee i_6\}$,
+  $C_g = \{\neg{a_2} \vee \neg{f_5} \vee g_5\}$
+  $C_f = \{\neg{l^{*}_{ext}} \vee \neg{e_5} \vee f_5\}$,
+  $C_d = \{\neg{b_2} \vee \neg{c_2} \vee d_2\}$,
+  $C_c = \{\neg{a_2} \vee \neg{b_2} \vee c_2 \}$,
+  $C_b = \{\neg{l^{*}_{ext}} \vee \neg{a_2} \vee b_2\}$
+\end{enumerate}
+Suppose 1-UIP learning produces clause
+$\oneUIPClause = \{\neg{m_{10}} \vee \neg{k_6} \vee \neg{j_6} \vee
+\neg{i_6} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{d_2} \vee
+\neg{c_2}\}$ where $\neg{k_{10}}$ is the UIP from the conflicting
+level. $\IUIP$ first tries to find the UIP for level 6 by resolving
+$\oneUIPClause$ with $C_k$, $C_j$ and, $C_i$ in order and produces
+clause
+$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
+\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ where
+$\neg{h_6}$ is the UIP.
+
+$\IUIP$ attempts to find the UIP for level 5 by resolving $C_i$ with
+$C_g$ and $C_f$ in order. However, resolving with $C_f$ will introduce
+$\neg{l^{*}_{ext}}$ into $C_i$ and increase the clause's
+LBD. $\IUIPPURE$ undoes all resolutions at level 5 and skips the
+entire level, whereas $\IUIPMIN$ only skips the resolution with $C_f$
+and tries to find the UIP by ignoring $\neg{f_5}$.  After level 5,
+$\IUIPPURE$ preserves the same clause and $\IUIPMIN$ produces
+$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
+\neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \vee \neg{a_2}\}$ where
+$\neg{e_5}$ is the UIP when $\neg{f_5}$ is ignored. The same $\IUIP$
+procedure is then performed on level 2 to produce the final
+$\iUIPClause$.
+
+$\IUIPPURE$ learned
+$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
+\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ and reduces
+the learnt clause size by 1. $\IUIPMIN$ learned
+$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
+\neg{e_5} \vee \neg{b_2} \vee \neg{a_2}\}$ and reduces the learnt
+clause size by 2.
+\end{example}
+
+The core algorithm of $\IUIP$ as a clause reduction technique is
+simple. However, to make the algorithm a practical learning scheme
+that is compatible with modern SAT solvers, a number of issues need to
+be addressed. The rest of the section details the key optimizations
+and augmentations of $\IUIP$ as a practical clause learning scheme.
+
+\subsection{Control i-UIP Learning}
+This simple and greedy $\IUIP$ learning scheme can produce
+significantly smaller clause. However, when $\IUIP$ does not reduce
+the clause size, the cost of the additional resolution steps will hurt
+the solver's performance. Since resolution cannot reduce a clause's
+$\LBD$, The maximum size reduction from $\IUIP$ is the difference
+between the clause's size and $\LBD$, denote as the clause's gap value
+($\GAP(\oneUIPClause) = |\oneUIPClause| - \LBD(\oneUIPClause)$). For
+an 1-UIP clause with a small $\GAP$, applying $\IUIP$ is unlikely to
+achieve cost effective results. Therefore, we propose a heuristic
+based approach to to enable and disable $\IUIP$ learning based on
+input clause's $\GAP$.
+
+\begin{algorithm}[t]
+\caption{Control-$\IUIP$}\label{alg:enableCondition}
+\begin{algorithmic}[1]
+\Require  $\oneUIPClause$ is a valid 1-UIP clause
+\Require  $ t_{gap} \ge 0$ is a dynamically calculated gap threshold
+\Procedure{Control-$\IUIP$}{$\oneUIPClause,  t_{gap}$} 
+    \State $Gap \gets |\oneUIPClause| - \textit{LBD}(\oneUIPClause)$
+     \If{$Gap > t_{gap}$} \label{ln:compareGap}
+        \State $\iUIPClause \gets \IUIP(\oneUIPClause, \phi)$
+        \State \textcolor{gray}{\textbf{I-UIP-Greedy}($\iUIPClause$ , $\oneUIPClause$)}
+               \Comment{additional clause selection policy, see sec~\ref{sec:greedy}}
+        \If {$|\iUIPClause| < |\oneUIPClause|$} 
+            \State $\text{Succeed} \gets \text{Succeed}+1$ \label{ln:updateS}
+        \EndIf
+        \State $\text{Attempted} \gets \text{Attempted}+1$ \label{ln:updateA}
+     \EndIf
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+Alg.~\ref{alg:enableCondition} compares the input $\oneUIPClause$'s
+$\GAP$ against a floating target threshold $t_{gap}$
+(line~\ref{ln:compareGap}). The threshold $t_{gap}$ represent the
+expected minimal $\GAP$ required for $\oneUIPClause$ to achieve a
+predetermined success rate (80\%) from performing $\IUIP$ learning.
+
+The gap threshold $t_{gap}$ is initialized to 0, and is updated at
+every restart based on $\IUIP$'s success rate from the previous
+restart interval. More specifically, the algorithm collects the
+statistics of the number of $\IUIP$ learning attempted
+(line~\ref{ln:updateA} in alg.~\ref{alg:enableCondition}) and the
+number of attempts succeed (line~\ref{ln:updateS}) for each restart
+interval, and use them to calculate the success rate. If the success
+rate is below 80, the threshold $t_{gap}$ is increased to restrict
+$\IUIP$for the next restart interval. On the other hand, the threshold
+is decreased to encourage more aggressive $\IUIP$ learning for the
+next restart interval.
+
+\[
+    t_{gap}=
+\begin{cases}
+    t_{gap} + 1& \text{if } \frac{\text{Succeed}}{\text{Attempted}} < 0.8\\
+    max(t_{gap} - 1, 0),              & \text{otherwise}
+\end{cases}
+\]
+
+\subsection{Early stop i-UIP Learning}
+At any point of $\IUIPMIN$ learning, if the number of marked literals
+in $\iUIPClause$ (literals which are forced into the clause to
+preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we
+can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the
+final $\iUIPClause$ is at least the size of $\oneUIPClause$.  The
+early stopping rule prevents solver from wasting time on traversing a
+large implication graph when $\IUIP$ has already failed.
+
+The early stopping rule may prematurely stops $\IUIPPURE$ learning
+which would otherwise succeed because $\iUIPClause$ can be further
+minimized. However, the experiments indicates such false negative
+cases are rare, and the early stopping rule generally helps the
+learning scheme's performance.
+
+\subsection{Greedy Active Clause Selection} \label{sec:greedy} Even
+though $\IUIP$ learning can reduce the size of the learnt clause
+$\iUIPClause$, but it may introduce literals with low variable
+activity into $\iUIPClause$.  Inactive literals prevents the clause
+from being asserted to force literal implication. Therefore, a
+practical clause learning scheme should consider both size and
+variable activity. We propose an optional extension $\IUIPGreedy$ to
+filter out inactive $\oneUIPClause$.
+
+\begin{algorithm}[t]
+\caption{$\IUIPGreedy$}\label{alg:greedySelection}
+\begin{algorithmic}[1]
+\Require  $\iUIPClause$ is a valid i-UIP clause
+\Require  $\oneUIPClause$ is a valid 1-UIP clause
+\Procedure{$\IUIPGreedy$}{$\iUIPClause, \oneUIPClause$} 
+   \If{$|\iUIPClause| < |\oneUIPClause| \wedge (\code{AvgVarAct}(\iUIPClause) > \code{AvgVarAct}(\oneUIPClause)) $} \label{ln:checkSizeAndActivity}
+    \State \textbf{return} $\iUIPClause$
+\Else
+    \State \textbf{return} $\oneUIPClause$
+\EndIf  \label{ln:cSizeCheck}
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+After computing the $\iUIPClause$, $\IUIPGreedy$ compares both the
+size and the average variable activity for $\oneUIPClause$ and
+$\iUIPClause$ (alg.~\ref{alg:greedySelection} at
+line~\ref{ln:checkSizeAndActivity}). The algorithm learns
+$\iUIPClause$ if the clause has smaller size and higher average
+variable activity.
+
+\subsection{Adjust Variable Activity} \label{sec: varajust} Two
+popular branching heuristics for modern SAT solvers are VSIDS and
+LBR. Both heuristics increase the variable activity for all literals
+involved in resolutions during 1-UIP learning. Since $\IUIP$ extends
+1-UIP with deeper resolutions against the trail, the variables
+activities for the fresh literals involved in the additional
+resolution steps need to be adjusted as well. We purpose two optional
+schemes for adjusting variable activities, $\IUIPActive$ and
+$\IUIPDist$.
+
+After learning $\oneUIPClause$ from 1-UIP scheme, $\IUIPActive$
+collects all literals appear in the the $\IUIP$ clause $\iUIPClause$,
+and increase their variable activity uniformly according to the
+current branching heuristic if the literals' variable activity have
+not yet been bumped during 1-UIP. The scheme does not bump variable
+activities for transient literals that are resolved away at
+non-conflicting level because, unlike transient literals at
+conflicting level, these literals cannot be re-asserted after the
+immediate backtracking. Notice that other post-analyze extensions such
+as Reason Side Rate (RSR) and Locality\cite{} can be applied after
+$\IUIPActive$.
+
+\[ \forall l \in  \iUIPClause \cdot NotBumped(var(l)) \implies  bumpActivity(var(l)) \]
+
+$\IUIPDist$ collects literals appear exclusively in $\iUIPClause$ and
+bump their variable activity uniformly. It also find all the literals
+in $\oneUIPClause$ that are resolved away during $\IUIP$, and unbump
+their variable activity if they have been bumped during 1-UIP. The
+unbumped literals are no longer in the learned clause $\iUIPClause$,
+keep their variable activity bumped does not help solver to use
+$\iUIPClause$.
+
+\[ \forall l_{i} \in  \iUIPClause \setminus \oneUIPClause \cdot bumpActivity(var(l_{i})) 
+\]
+\[ \forall l_{1} \in  \oneUIPClause \setminus \iUIPClause \cdot unbumpActivity(var(l_{1})) 
+\]
+
+\subsection{Integrate with Chronological Backtracking}
+In SAT solver with chronological backtracking, literals on the
+assertion trail $\assertionTrail$ are not always sorted by decision
+levels. This change imposes a challenge to $\IUIP$ learning since the
+previous implementation relies on solver's ability to efficiently
+access all literals from any decision level in descending trail order
+(alg.~\ref{alg:i-uip} line~\ref{ln:trailaccess}).
+
+To mitigate this challenge, we modify the solver to track the precise
+trail position of all asserted literals with a single vector. We then
+build a priority queue $lit\_Order$ to manage literal's resolution
+order. The queue $lit\_Order$ prioritizes literals with higher
+decision level, and it favors literal with higher trail position when
+decision levels are tied. The order of $lit\_Order$ represents the
+correct resolution order of $\IUIP$ because 1) an asserted literal's
+decision level is the maximum level among all of its reasoning
+literals, and 2) an asserted literal always appears higher in the
+trail than its reasoning literal.
+
+\begin{algorithm}[t]
+\caption{\IUIP-CB}\label{alg:i-uip-CB}
+\begin{algorithmic}[1]
+\Require  $\oneUIPClause$ is a valid 1-UIP clause
+\Require  $\assertionTrail$ is a valid assertion trail
+\Require  $lit\_Order$ is a priority queue
+\Procedure{\IUIP-CB}{$\oneUIPClause, \assertionTrail, lit\_Order $} 
+    \State ...
+    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause}
+    \State forall $l \in \oneUIPClause \cdot \code{Enqueu}(lit\_Order, l)$  \label{ln:enqueue}
+    \State \code{DecisionLvs} $\gets \Set{\code{level}(l)}{l \in \oneUIPClause}$
+    \While{$lit\_Order \neq \emptyset$} \label{ln:pop}
+        \State $p \gets \code{dequeu}(lit\_Order, l) $ \label{ln:dequeue}
+        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs} \\ $\vee p$ is the last reminaing lit in its decision level } \label{ln:resolvable}
+          \State Pass
+        \Else 
+            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(q, \assertionTrail)$
+            \State   forall $l \in \code{Reason}(q, \assertionTrail) \cdot \code{Enqueu}(lit\_Order, l)$ \label{ln:reequeue}
+        \EndIf
+    \EndWhile
+\State ...
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+Alg.~\ref{alg:i-uip-CB} is the pesudo-code implementation of the
+augmented $\IUIP$ for chronological backtracking with the priority
+queue $lit\_Order$. The algorithm first populates $lit\_Order$ with
+all literals in $\oneUIPClause$ (line~\ref{ln:enqueue}), and then
+continuously pops literals until the queue is empty
+(line~\ref{ln:pop}). When a literal is resolved away, all of its
+reasoning literals are added into $lit\_Order$ if they are not already
+in the queue (line~\ref{ln:reequeue}). The algorithm will always
+terminate because a literal cannot entered the queue twice (guaranteed
+by the properties of the trail order) and there are finite amount of
+literals on the trail.
+
+\section{Implementation and Experiments}
+
+\subsection{Clause Reduction with $\IUIP$}
+To evaluate $\IUIP$'s effectiveness as a clause reduction technique,
+we implement $\IUIPPURE$ and $\IUIPMIN$ with Control-I-UIP and Early
+Stopping rules in Sec~\ref{sec:i-uip} on top of \text{\MapleBase}
+\cite{}, the winner of SAT Race 2015 application track.  We than
+compare the performance of the baseline $\MapleBase$ with
+$\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from
+SAT RACE 2019 main track.
+
+The benchmark contains 400 instances divided into two groups of 200,
+new and old, representing historical instances and fresh instances in
+the 2019 race, respectively. We randomly partition the old group
+instances into six partitions of size 30 and one partition of size
+20. Each partition is then assigned to a XeonE5-2 CPU node with 16
+cores (2 sockets 8 cores and 1 thread) and 96649 MB memory. The new
+group is partitioned based on their contributor (e.g. Heule
+contributed 22 matrix multiplication instances), and each partition is
+assigned to a aforementioned CPU node. To speed up the experiment, we
+allow a CPU node to solver at most seven instances concurrently.
+
+Beside solved instances count and PAR-2 score, we additionally measure
+the average clause length and clause reduction ratio (both cumulative
+and non-cumulative)\footnote{The cumulative reduction ratio is
+  obtained through learning all clauses with the target learning
+  scheme; Therefore, the reduction is cumulative. The non-cumulative
+  reduction ratio is obtained by running the target scheme for
+  measurement only (the minimized 1-UIP clause is learned); Therefore,
+  the reduction is not cumulative. }  for each instances. For
+$\MapleIUIPPURE$ and $\MapleIUIMIN$, we also captures the $\IUIP$
+learning attempted rate and success rate.
+
+\begin{figure} 
+\begin{center}
+\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
+\hline
+Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
+\hline
+$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
+\hline
+$\MapleIUIPPURE$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
+\hline
+$\MapleIUIMIN$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.
+CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
+\label{fig:t1}
+\end{figure}
+
+Fig.~\ref{fig:t1} shows that $\IUIPPURE$ ($\IUIPPURE$) solved seven
+(five) more instances than the baseline solver with lower PAR-2
+scores. $\IUIPPURE$ has marginally lower PAR-2 score than
+$\IUIPPURE$. Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with
+significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
+respectively. Fig.~\ref{fig:len_pdf} shows the probability density
+distribution (PDF) of the relative clause size of both i-uip learning
+schemes ($\IUIPPURE$ in red and $\IUIPMIN$ in green) for each
+instance. $\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 77.7\%
+(88.5\%) of instances, and average relative reduction from 1-UIP is
+16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
+clause size of $\IUIPMIN$, $\IUIPPURE$ and 1-UIP, and it shows that
+both i-UIP learning schemes in general produce the smaller clauses,
+and the size reduction is more significant for instances with large
+average 1-UIP clause size. Moreover, $\IUIPMIN$ clauses (indicated in
+green) is consistently smaller than $\IUIPPURE$ clauses.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/clause_reduction_PDF.png}
+    \caption{ Relative clause size reduction distribution.
+X axis indicates the relative size of difference between i-uip and 1-uip clauses (calculated as $\dfrac{|\oneUIPClause|-|\iUIPClause|}{|\oneUIPClause|}$ ) for each instance, and Y axis shows the probability density.}
+     \label{fig:len_pdf}
+\end{figure}
+\begin{figure} \label{fig:len_compare}
+    \centering
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{i-uip-sizes-compare.png}
+    \caption{Average clause size comparison plot.
+Each point in the plot represents an instance.
+X and Y axis shows the clause length from $\IUIP$ and 1-UIP, respectively.
+Each green (red) dot represents an compared instance between $\MapleBase$ and $\MapleIUIMIN$ ($\IUIPPURE$).
+}
+    \label{fig:len_compare}
+\end{figure}
+\nf{ Do we need this?  We additionally looked at the 14 instances
+  solved by $\IUIPMIN$ but not by 1-UIP. $\IUIPMIN$ produces smaller
+  clauses for all of them with average relative reduction of 22\% and
+  maximum 77\% (30 vs 135). Seven out of 14 instances has size
+  relative reduction over 30\%. For the 9 instances solved by 1-UIP
+  but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for
+  three instances and with average relative reduction of 3.3\%.}
+
+$\IUIPMIN$ outperformed $\IUIPPURE$ in clause size.
+This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.
+
+\begin{figure} 
+\begin{center} 
+\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
+\hline
+Solver & $\IUIP$ attempt rate & $\IUIP$ success rate  \\ 
+\hline
+$\MapleIUIPPURE$ & 16.1\% & 43.4\% \\ 
+\hline
+$\MapleIUIMIN$ & 28.8\% & 59.3\% \\ 
+\hline
+\end{tabular}
+\end{center}
+\caption{Compare $\IUIPPURE$ and $\IUIPMIN$ i-uip attempt rate and success rate.
+$\IUIPMIN$ scheme attempted $\IUIP$ more frequently, and it is more likely to successfully produce smaller $\iUIPClause$ clause .}
+\label{fig:t2}
+\end{figure}
+
+A solver produce smaller clauses can construct smaller proofs. For
+UNSAT instances, we measure their DRUP\cite{} proof checking time as
+well as the size of the optimized DRUP proof. We used DART-trim
+\cite{} with 5000 timeout to check and optimize DRUP proofs.
+
+Fig.~\ref{fig:t3} shows that the optimized proof construct by
+$\IUIPMIN$ and $\IUIPPURE$ are significantly smaller than 1-UIP
+proofs. The relative proof size reduction roughly correlates to the
+average clause size reduction. Fig.~\ref{fig:proof_compare} shows the
+absolute proof size comparison results.
+
+\begin{figure} 
+\begin{center} 
+\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
+\hline
+Solver & optimized proof size (MB) & relative reduction size  \\ 
+\hline
+$\MapleBase$ & 613.9 & 0  \\ 
+\hline
+$\MapleIUIPPURE$ & 487.2 & 6.90\% \\ 
+\hline
+$\MapleIUIMIN$ & 413.2 & 17.18\% \\ 
+\hline
+\end{tabular}
+\end{center}
+\caption{Optimized UNSAT proof comparison for 1-UIP $\IUIPPURE$ and $\IUIPMIN$.
+Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
+\label{fig:t3}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{proof_size_compare.png}
+    \caption{Average optimized proof size between 1-uip and $\IUIPMIN$.}
+    \label{fig:proof_compare}
+\end{figure}
+
+\begin{figure} 
+\begin{center}
+\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
+\hline
+Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
+\hline
+1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
+\hline
+\nf{$\IUIPPURE$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
+\hline
+$\IUIPMIN$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
+\hline
+$\IUIPGreedy$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
+\hline
+$\IUIPActive$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
+\hline
+$\IUIPDist$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$,
+$\IUIPActive$, and $\IUIPDist$ on SAT2019 race main track.}
+\label{fig:t4}
+\end{figure}
+
+\subsection{$\IUIP$ as a Practical Learning Scheme}
+To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we
+re-implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned
+in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
+$\IUIP$ configurations ( $\IUIPMIN$, $\IUIPPURE$, $\IUIPGreedy$,
+$\IUIPActive$, and $\IUIPDist$) on the SAT Race 2019 main track
+benchmark and reported solved instances, PAR-2 score and average
+clause size.
+
+Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
+scheme $\IUIPPURE$ solved the most overall instances (228) and the
+most UNSAT instances (93). $\IUIPGreedy$ had the lowest PAR-2
+score. $\IUIPActive$ solved the most SAT instances (138). $\IUIPDist$
+produced the shortest average clause size, but solved the second least
+instances, one more instance than the baseline 1-UIP learning. All
+configurations of $\IUIP$ outperformed the baseline 1-UIP scheme in
+solved instances, PAR-2 score and average clause size.
+
+\subsection{$\IUIP$ on Modern SAT solvers}
+To validate $\IUIP$ as a generalizable learning scheme on modern SAT
+solvers, we re-implement $\IUIP$ on the winners of 2017, 2018 and 2019
+SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$). $\expSATShort$
+is a top ten solver from 2019 SAT race which uses random walk
+simulation to help branching. We chose $\expSATShort$ because 1) it is
+a top solver in the 2019 SAT Race without using chronological
+backtracking; 2) the combination of random walk simulation and
+variable activity branching heuristic allows our learning schemes to
+partially sidestep the problem of variable activity. For each solver,
+we compare the base 1-UIP learning scheme against $\IUIPPURE$,
+$\IUIPMIN$ and the top two $\IUIP$ variants, $\IUIPGreedy$ and
+$\IUIPActive$, on the SAT Race 2019 main track benchmark. We report
+solved instances, PAR-2 score and the average clause size.
+
+\begin{figure} 
+\begin{center}
+\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
+\hline
+Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
+\hline
+SAT 2017 Winner & & & \\
+$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
+\hline
+$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
+\hline
+$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
+\hline
+$\MapleSeven$-i-greedy & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
+\hline
+$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
+\hline
+\hline
+SAT 2018 Winner & & & \\
+$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
+\hline
+$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
+\hline
+$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
+\hline
+$\MapleEightShort$-i-greedy & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
+\hline
+$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
+\hline
+\hline
+SAT 2019 Winner & & & \\
+$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
+\hline
+$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
+\hline
+$\MapleNineShort$-i-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
+\hline
+$\MapleNineShort$-i-greedy & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
+\hline
+$\MapleNineShort$-i-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
+\hline
+\hline
+SAT 2019 Competitor & & &\\
+$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
+\hline
+$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
+\hline
+$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
+\hline
+$\expSATShort$-i-greedy & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
+\hline
+$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of 1-UIP, $\IUIPPURE$. $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ on SAT2019 race main track.}
+\label{fig:t5}
+\end{figure}
+
+Table~\ref{fig:t5} shows the benchmark result of $\IUIP$
+configurations on different solvers. All four configurations of
+$\IUIP$ outperformed 1-UIP learning for the SAT 2017 race winner,
+$\MapleSeven$. More specifically, $\IUIPPURE$, $\IUIPMIN$,
+$\IUIPGreedy$ and $\IUIPActive$ solved 12, 8, 5 and 2 more instances,
+respectively, whiling producing smaller clauses. $\IUIPPURE$ solved
+more UNSAT and SAT instances while other configurations improved on
+solving SAT instances.  The improvement of $\IUIP$ is more significant
+on $\MapleSeven$ than on $\MapleBase$ for both solved instances and
+clause size reduction. This may suggests that $\IUIP$ and the recent
+learnt clause minimization approach~\cite{} synergies well because
+i-UIP clauses are shorter with more common literals which allows
+vivification~\cite{} to prune literals more aggressively through unit
+propagation.
+
+We observed significant improvement of $\IUIP$ for the SAT 2018
+winner, $\MapleEightShort$. Three out of four $\IUIP$ configurations
+outperformed 1-UIP by 5, 4 an 4 instances,
+respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using
+chronological backtracking (CB) for long distance
+backtracks. Therefore, we used $\IUIP$-CB extension for all $\IUIP$
+configurations. The results indicates that the improvement of $\IUIP$
+is slightly shadowed by the adoption of CB. More specifically, we
+believe CB prevents decision levels from being compressed through the
+process of backtracking and literal assertion. The shorter i-UIP
+clauses can bring related literals closer, which in turns compresses
+the assertion stack and the decision levels. However, since CB
+discourages long distance backtracking, the effect of learning shorter
+clauses is shadowed until a full restart.  We believe we have observed
+an interesting interaction effect that shows the limitations of both
+CB and $\IUIP$ learning for future research.
+
+$\IUIP$ learning schemes showed significant improvement for
+$\MapleNineShort$. Three out of four configurations improved solved
+instances by 6, 5 and 5 instances, respectively. $\MapleNineShort$
+uses duplicate learning (DL) to prioritize clauses that are learned
+multiple times. $\IUIP$ and DL didn't synergies well possibly because
+i-UIP clauses are less likely to be duplicated. We observed that
+$\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the
+core clause database than 1-UIP. One possible explanation is that
+$\IUIP$ learning can reduce an 1-UIP clause to different i-UIP clauses
+for different assertion trails. Therefore, the solver is less likely
+to learn the same i-UIP clauses.
+
+We observed significant improvement of $\IUIP$ for $\expSATShort$,
+three out of four configurations of $\IUIP$ significantly outperformed
+1-UIP learning by 4, 7 and 8 more instances, respectively, while
+producing smaller clauses. We expect all three configurations of
+$\IUIP$ to solve the similar amount of instances with close PAR-2
+scores because the additional random walk exploration allows the
+learning schemes to partially sidestep the activity problem; hence,
+the learning adjustment for variable activities should have less
+impact on the solver's performance. However, we instead, observed that
+both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP
+learning scheme. One possible explanation is that $\IUIPGreedy$ and
+$\IUIPActive$ schemes could easily overcompensate variable activities
+for i-UIP clauses, and $\expSATShort$'s random walk exploration could
+use future search information to mitigate the negative effects of our
+overcompensation.
+
+\bibliography{sat}{}
+\bibliographystyle{splncs04}
+
 \end{document}

### 2020-02-06 paper.tex
change figure link to PDF

@@
     \centering
-    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{figures/clause_reduction_PDF.png}
+    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF.pdf}
     \caption{ Relative clause size reduction distribution.
@@
     \centering
-    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{i-uip-sizes-compare.png}
+    \includegraphics[width=0.8\textwidth]{figures/i-uip-sizes-compare.pdf}
     \caption{Average clause size comparison plot.
@@
     \centering
-    \includegraphics[width=0.8\textwidth,natwidth=610,natheight=642]{proof_size_compare.png}
+    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
     \caption{Average optimized proof size between 1-uip and $\IUIPMIN$.}

### 2020-02-12 paper.tex
more fb changes

@@
 \usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[noend]{algpseudocode}
 %\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
@@
 \usepackage{theorem}
-\usepackage[T1,hyphens]{url}
-
-\begin{document}
-\title{$i$-UIP Clause Learning}
+\usepackage[T1]{fontenc}
+% \usepackage[T1,hyphens]{url}
+\usepackage{hyperref}
+%\PassOptionsToPackage{hypnens}{url}
+\title{Clause size reduction with $i$-UIP Learning}
 \author{Nick Feng \and Fahiem Bacchus}
@@
 
+\algrenewcommand\algorithmicprocedure{}
+\algrenewcommand\algorithmicthen{}
+\algrenewcommand\algorithmicdo{}
+
+\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
+\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
 \newcommand{\sat}{SAT\xspace}
@@
 \newcommand{\iUIPClause}{\ensuremath{C_{i}}}
-\newcommand{\assertionTrail}{\ensuremath{\phi}}
+\newcommand{\assertionTrail}{\trail}
 \newcommand{\LBD}{\text{LBD}}
 \newcommand{\GAP}{\text{Gap}}
-\newcommand{\IUIP}{\textbf{$i$-UIP}}
-\newcommand{\IUIPPURE}{\textbf{Pure-i-UIP}}
-\newcommand{\IUIPMIN}{\textbf{Min-i-UIP}}
+\newcommand{\IUIP}{\textbf{LBDstable-$i$UIP}}
+\newcommand{\tryuiplevel}{\textbf{try-uip-level}\xspace}
+\newcommand{\IUIPPURE}{\textbf{Pure-$i$UIP}}
+\newcommand{\IUIPMIN}{\textbf{Min-$i$UIP}}
 \newcommand{\IUIPGreedy}{\textbf{i-UIP-Greedy}}
@@
 \newcommand{\MapleEightShort}{\textit{MapleCB}}
-\newcommand\Set[2]{\{\,#1\mid#2\,\}}
-\newcommand\Array[2]{[\,#1\mid#2\,]}
-\newcommand\target{\textbf{Succ_t}}
-\newcommand\trail{\ensuremath{\mathcal{T}}}
-\newcommand\trailIdx[1]{\ensuremath{\iota(#1)}}
-\newcommand\var{\text{var}}
-\newcommand\true{\textsc{true}\xspace}
-\newcommand\false{\textsc{false}\xspace}
-\newcommand\reason{\textit{reason}}
+\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
+\newcommand{\Array}[2]{[\,#1\mid#2\,]}
+\newcommand{\target}{\textbf{Succ_t}}
+\newcommand{\trail}{\ensuremath{\mathcal{T}}}
+\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
+\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
+\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
+\newcommand{\var}{\text{var}}
+\newcommand{\true}{\textsc{true}\xspace}
+\newcommand{\false}{\textsc{false}\xspace}
+\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
+\newcommand{\resolve}{\bowtie}
+\renewcommand{\implies}{\rightarrow}
+\newcommand{\ctry}{C_{\mathit{try}}}
+\newcommand{\type}{\textit{type}}
+\newcommand{\unmarked}{\textbf{unmarked}}
+\newcommand{\mrk}{\textbf{mark}}
 \setlength{\theorempreskipamount}{3pt}
 \setlength{\theorempostskipamount}{3pt}
+
 \newtheorem{Obs}{Observation}
@@
 
+\begin{document}
 \maketitle              % typeset the header of the contribution
@@
 \section{Introduction}
-Clause learning is an essential technique in \sat solvers. And there
-is good empirical evidence that clause learning is, in fact, the most
-important technique used in modern \sat solvers
-\cite{DBLP:conf/sat/KatebiSS11}. In early \sat research a number of
-different clause learning techniques were proposed
+Clause learning is an essential technique in \sat solvers. There is good
+evidence to indicate that it is, in fact, the most important technique
+used in modern \sat solvers \cite{DBLP:conf/sat/KatebiSS11}. In early
+\sat research a number of different clause learning techniques were
+proposed
 \cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
@@
 
-However, as this early work showed, other clause learning schemes can
-be used in SAT solvers without changes to their main data
-structures. Furthermore, advances in our understanding allow us to
-better understand the potential advantages and disadvantages of these
-alternate schemes. In this paper we reexamine some of these previous
-clause learning schemes, with a focus on the $i$-UIP schme described
-in \cite{DBLP:conf/iccad/ZhangMMM01}. With current ideas we can see
-that in its original form $i$-UIP clause learning suffers a
-significant disadvantages over 1-UIP clause learning. We present a
-number of different methods to impove $i$-UIP clause learning so as to
-over come these disadvantages. We show that these different methods
-for $i$-UIP clause learning can provide useful performance gains over
-the standard 1-UIP scheme.
-
-\section{Background}
-We assume the reader has some familiarity with Conjunctive Normal Form
-(CNF) formulas, conflict-driven clause learning (CDCL) \sat solvers
-\cite{DBLP:series/faia/SilvaLM09}, unit-propagation, and conflict
-analysis.
-
-\begin{enumerate}
-\item Trail; Reasons and their properties; decision levels, trail order.
-\item conflict clause. Trail has the property that conflict exists
-    only at the deepest level.
-\item Clauses derived by trail resolution. Identified by variables
-    appearing on the trail.
-\item No point in doing this in non-trail order.
-\item a specific case of trivial resolutions.
-\item Clause derived by trail resolution iff it is a cut in the
-    implication graph.
-\item Note that all implemented clause learning techniques use
-    sequences of trail resolutions. In particular, they all start with
-    a specific conflict and resolve only with reason clauses.
-    This includes clause minimization.
-\item Asserting clause---one lit at deepest level of trail. Only
-    asserting clauses are gauranteed to be new.
-\item 1-UIP is asserting clause generated with fewest trail resolution
-    steps. Each trail resolution step can only increase the number of
-    levels in the clause. So LBD is minimized by 1-UIP clauses.
-\item Description of i-uip, all-uip, all decisions (useful in some cases)
-\end{enumerate}
-
-CDCL \sat solvers maintain a \textbf{trail} which is a sequence of
-literals that have been assigned \true by the solver. Literals newly
-assigned \true are placed at the end of the trail, and literals are
-always unassigned and removed from the end of the trail first;
-i.e. the trail acts as a stack.
-
-Let $\trail$ be the trail. For any literal $\ell$ we let
-$\trailIdx{\ell}$ be the index of $\ell$ in $\trail$. That is,
-$\trail[\trailIdx{\ell}] = \ell$. If $\ell$ is not in the trail, then
-$\trailIdx{\ell} = \varnothing$, and we use the convention that
-$\trail[\varnothing] = \varnothing$. The literals on \trail are
-sequenced in the order they were assigned \true. Hence
-$\trailIdx{\ell} < \trailIdx{\ell'}$ implies that $\ell$ was assigned
-\true before $\ell'$. When $\trailIdx{\ell} < \trailIdx{\ell'}$ we say
-that $\ell$ \textit{appears above} $\ell'$ on the trail. We use the
-convention that $\trailIdx{\ell} < \trailIdx{\ell'}$ whenever $\ell$
-is in the trail and $\ell'$ is not; i.e., $i < \varnothing$ for any
-index $i$.
-
-The literals on $\trail$ have been assigned \true by the solver by
-either (a) having been set as a decisions or (b) having forced by
-unit-propagation. If $\ell\in \trail$ was was set by unit-propgation
-it will have an associated reason clause $\reason(\ell)$. This clause
-has two properties (1) $\ell\in \reason(\ell)$ and (2) for every
-$x\in \reason(\ell)$ we have that
-$\trailIdx{\lnot x} < \trailIdx{\ell}$. That is, $\reason(\ell)$ is a
-clause that implies that $\ell$ must assigned \true because every
-other literal in it has been assigned \false above $\ell$ in
-$\trail$. Literals $\ell$ set as decsions do not have reason clauses
-and we let $\reason(\ell) = \varnothing$.
-
-CDCL \sat solvers perform clause learning when the literals on the
-trail falsify a clause $c$. '
-
-\section{Find i-uip Clause}
+However, other clause learning schemes can be used in SAT solvers
+without changes to their main data structures. Furthermore, advances
+in our understanding allow us to better understand the potential
+advantages and disadvantages of these alternate schemes. In this paper
+we reexamine some of these previous clause learning schemes, with a
+focus on the schemes described in
+\cite{DBLP:conf/iccad/ZhangMMM01}. Improved understanding of \sat
+solvers, obtained from the last decade of research, allows us to see
+that in their original form these other clause learning schemes suffer
+significant disadvantages over 1-UIP clause learning.
+
+One of the previously proposed schemes was the $i$-UIP scheme
+\cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
+to exploit the main ideas of this scheme that avoids its main
+disadvantages. In particular, we propose to use a $i$-UIP like clause
+learning scheme to generate smaller learnt clauses which retain the
+good properties of standard 1-UIP clauses. Our method is related to,
+but not the same as various clause mininimization methods that aim to
+remove literals from the 1-UIP clause, e.g.,
+\cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
+Our method is orthogonal to clause minimization. In particular, 
+our approach can learn a clause that is completely different from the
+1-UIP clause but which still serves all of the same purposes as the 1-UIP
+clause. Like the 1-UIP clause, various minimization techniques can be
+applied on top of our method to further reduce the size of the clauses
+we learn.
+
+We present various versions of our method and show that these variants
+are often capable of learning shorter clauses that the 1-UIP scheme,
+and that this can lead to useful performance gains in state of the art
+\sat solvers.
+
+\section{Clause Learning Framework}
+We first provide some background and a framework for understanding
+clause learning as typically used in CDCL \sat solvers. A
+propositional formula $F$ expressed in Conjunctive Normal Form (CNF),
+contains a set of variables $V$. A literal is a variable $v\in V$ or
+its negation $\lnot v$. For a literal $\ell$ we let $\var(\ell)$
+denote its underlying variable. A CNF consists of a conjunction of
+clauses, each if which is a disjunction of literals. We often view a
+clause as being a set of literals and employ set notation, e.g.,
+$\ell\in C$ and $C'\subset C$.
+
+Two clauses $c_1$ and $c_2$ can be \textit{resolved} when they contain
+conflicting literals $\ell\in c_1$ and $\lnot \ell \in c_2$. Their
+resolvant $c_1 \resolve c_2$ is the new clause
+$(c_1 \cup c_2) - \{\ell, \lnot \ell\}$. The resolvant is tautological
+if $c_1$ and $c_2$ contain more than one pair of conflicting literals.
+
+We assume the reader is familiar with the operations of CDCL \sat
+solvers, and the main data structures used in such solvers. A good
+source for this background is \cite{DBLP:series/faia/SilvaLM09}.
+
+\paragraph{The Trail.}
+CDCL \sat solvers maintain a \textbf{trail}, $\trail$, which is a
+\textit{non-contradictory, non-redundant sequence of literals} that
+have been assigned \true by the solver; i.e.
+$\ell\in\trail \implies \lnot\ell \not\in\trail$, and $\trail$
+contains no duplicates. Newly assigned literals are added to the end
+of the trail, and on backtrack literals are removed from the end of
+the trail and unassigned.  If literal $\ell$ is on the trail let
+$\trailIdx{\ell}$ denote its index on the trail sequence
+($\trail[\trailIdx{\ell}] = \ell$). For convenience, we also let
+$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$,
+where $\var(\ell)$ is $\ell$'s variable, even though $\lnot \ell$ nor
+$\var(\ell)$ are actually on $\trail$. If $x$ and $y$ are both on the
+trail and $\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$
+  appears before $y$ on the trail}. In a slight abuse of notation we
+also use $\trailIdx{v}$ on a variable $v$ where
+$\trailIdx{v}=\trailIdx{\ell}$ and $\ell$ a literal of $v$ on the
+trail.
+
+Two types of true literals appear on the trail: decision literals that
+have been assumed to be true by the solver, and unit propagated
+literals that are forced to be true because they are the sole remining
+unfalsified literal of a clause. Each literal $\ell\in\trail$ has a
+decision level $\dlevel{\ell}$ which is equal to the number of
+decision literals appearing before $\ell$ on the trail plus one;
+hence, $\dlevel{d}=1$ for the first decision literal $d\in\trail$.
+The set of literals on $\trail$ that have the same decision level
+forms a contiguous subsequence of $\trail$ that starts with a decision
+literal $d_i$ and ends just before the next decision literal
+$d_{i+1}$. If $\dlevel{d_i} = i$ we call this subsequence of $\trail$
+the \textit{$i$-th decision level}.
+
+Each literal $\ell\in\trail$ also has a clausal reason
+$\reason{\ell}$. If $\ell$ is a unit propgated literal,
+$\reason{\ell}$ is a clause of the formula such that
+$\ell \in \reason{\ell}$ and
+$\forall x \in \reason{\ell}.\, x\neq \ell \implies (\lnot x \in
+\trail \land \trailIdx{\lnot x} < \trailIdx{\ell}$. That is,
+$\reason{\ell}$ is a clause that has become unit implying $\ell$ due
+to the literals on the trail above $\ell$. If $\ell$ is a decision
+literal then $\reason{\ell} = \varnothing$.
+
+In most \sat solver implementations clause learning is initiated as
+soon as a clause is falsified by $\trail$. In this paper we will be
+concerned with the subsequent clause learning process which uses
+$\trail$ to derive a new clause that can be added to the formula. In
+some implementations the trail might be altered during clause
+learning. Here, however, we will assume that $\trail$ remains intact
+during clause learning and is only changed after the new clause is
+derived. After the new clause is lerned the $\trail$ will be changed by backtracking.
+
+Say that $\trail$ falsifies a clause $c_I$, and that the last decision
+literal $d_k$ in $\trail$ has decision level $k$. Consider
+$\trail_{k-1}$ the prefix of $\trail$ above the last decision level,
+i.e., the sequence of literals
+$\trail[0]$---$\trail[\trailIdx{d_k}-1]$. We will assume that
+$\trail_{k-1}$ is \textbf{propagation complete}, although the full
+trail $\trail$ might not be. This means that (a) no clause was
+falsified by $\trail_{k-1}$. And (b) if $c_u$ is a clause containing
+the literal $x$ and all literals in $c_u$ except for $x$ are falsified
+by $\trail_{k-1}$, then $x\in \trail_{k-1}$ and
+$\dlevel{x} \leq\max\{\dlevel{y} | y\in c_u \land y\neq x\}$. This
+means that if $x$ appears in a clause made unit it must have been
+added to the trail, and added at the first decision level that this
+occurred.  Note, however, $x$ might appear in more than one clause.
+The less than equal condition above ensures that $x$ must appear in
+$\trail_{k-1}$ at the first decision level any of these clauses were
+made unit.
+
+Any clause falsified by $\trail$ is called a \textbf{conflict}. When a
+conflict is found, the final level of the trail, $k$, need not be
+propagation complete as the solver typically stops propagation as soon
+as it finds a conflict. This means that (a) other clauses might be
+falsified by $\trail$ besides the conflict found, and (b) other
+literals might be unit implied by $\trail$ but not added to $\trail$.
+
+\begin{defn}[Trail Resolvant]
+    A trail resolvant is a clause arising from resolving a conflict
+    against the reason clause of some literal $\ell \in \trail$ where
+    $\reason{\ell}\neq \varnothing$. Every trail resolvant is also a
+    conflict.
+\end{defn}
+
+The following things can be noted about trail resolvants: (1) trail
+resolvants are never tautological, as the polarity of all literals in
+$\reason{\ell}$ other than $\ell$ must agree with the polarity of all
+literals in the conflict (they are all falsified by $\trail$); (2) one
+polarity of the variable $\var(\ell)$ resolved on must be a unit
+propagated literal whose negation appears in the conflict; and (3) any
+literal in the conflict that is unit progagated in $\trail$ can be
+resolved upon (the corresponding variable must appear in different
+polarities in the conflict and in $\trail$).
+
+\begin{defn}[Trail Resolution]
+    A trail resolution is a sequence of trail resolvants applied to an
+    initial conflict $c_I$ yielding a new conflict $c_L$.  A trail
+    resolution is \textbf{ordered} if the sequence of variables $v_1$,
+    \dots, $v_m$ resolved have strictly decreasing trail indicies:
+    $\trailIdx{v_i} > \trailIdx{v_{i+1}}$ ($1\leq i < m$). (Note that
+    this implies that no variable is resolved on more than once).
+\end{defn}
+
+Ordered trail resolutions resolve unit propagated literals from the
+end of the trail to the beginning. Without lost of generality, we can
+require that all trail resolutions be ordered.
+
+\begin{Obs}
+    If the unordered trail resolution $U$ yields the conflict clause
+    $c_L$ from an initial conflict $c_I$, then there exists an ordered
+    trail resolution $O$ that yields a conflict clause $c'_L$ such
+    that $c'_L\subseteq c_L$.
+\end{Obs}
+\noindent
+\textbf{Proof.} Let $U$ be the sequence of clauses $c_I = c_0$, $c_1$,
+$\ldots$, $c_m= c_L$ obtained by resolving on the sequence of
+variables $v_1$, $\ldots$, $v_m$ whose corresponding literals on
+$\trail$ are $l_1$, $\ldots$, $l_m$. Reordering these resolution steps
+so that the variables are resolved in order of decreasing trail index
+and removing duplicates yields an ordered trail resolution $O$ with
+the desired properties. Since no reason clause contains literals with
+higher trail indicies, $O$ must be a valid trail resolution if $U$
+was, and furthermore $O$ yields the clause
+$c'_L = \bigcup_{i=1}^m \reason{l_i} - \{l_1, \lnot l_1, \ldots, l_m,
+\lnot l_m\}$. Since $U$ resolves on the same variables (in a different
+order) we must that that the clause it produces $c_L\subseteq
+c'_L$. But $c_L$ might retain one or more of the literals $\lnot l_1$,
+$\ldots$, $\lnot l_m$, since resolving on a literal $l_j$ with higher
+trail index might reintroduce a literal $\lnot l_i$ with lower trail
+index that was previously resolved away by an earlier resolution.
+\whitebox
+
+The relevance of trail resolutions is that all proposed clause
+learning schemes we are aware of use trail resolutions to produce
+learnt clauses. Furthermore, the commonly used technique for clause
+minimization \cite{DBLP:conf/sat/SorenssonB09} is also equivalent to a
+trail resolution that yields the minimized clause from the
+un-minimized clause. Interestingly, it is standard in \sat solver
+implementations to perform resolution going backwards along the
+trail. That is, these implementations are typically using ordered
+trail resolutions. The above observation shows that this in fact the
+right way to do this.
+
+Ordered trail resolutions are a special case of \textit{trivial
+  resolutions} \cite{DBLP:journals/jair/BeameKS04}. Trail resolutions
+are specific to the trail data structure typically used in \sat
+solvers. If $\trail$ falsifies a clause at its last decision level,
+then its associated implication graph
+\cite{DBLP:journals/tc/Marques-SilvaS99} contains a conflict
+node. Cuts in the implication graph that separate the conflict from
+the rest of the graph correspond to conflict clauses
+\cite{DBLP:journals/jair/BeameKS04}. It is not difficult to see that
+the proof Proposition~4 of \cite{DBLP:journals/jair/BeameKS04} applies
+also to trail resolutions. This means that \textit{any conflict clause
+  in the trail's implication graph can be derived using a trail
+  resolution.}
+
+\subsection{Some alternate Clause Learning Schemes}
+A number of different clause learning schemes for generating a new
+learnt clause from the initial conflict have been presented in prior
+work, e.g.,
+\cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
+Figure~\ref{fig:cl_schemes} gives a specification of some of these
+methods: (a) the all-decision scheme which resolves away all implied
+literals leaving a learnt clause over only decision literals; (c) the
+1-UIP scheme which resolves away literals from the deepest decision
+level leaving a learnt clause with a single literal at the deepest
+level; (d) the all-UIP scheme which resolves away literals from each
+decision level leaving a learnt clause with a single literal at each
+decision level; and (e) the $i$-UIP scheme which resolves away
+literals from the $i$ deepest decision levels leaving a learnt clause
+with a single literal at its $i$ deepest decision levels. It should be
+noted that when resolving away literals at decision level $i$ new
+literals at decision levels less than $i$ might be introduced into the
+clause. Hence, it is important in the $i$-UIP and all-UIP schemes to
+use ordered trail resolutions.
+
+Both the all-decision and all-UIP schemes yield a clause with only one
+literal at each decision level, and the all-UIP clause will be no
+larger that the all-decision clause. Furthermore, it is known
+\cite{DBLP:journals/tc/Marques-SilvaS99} that once we reduce the
+number of literals at a decision level $d$ to one, we could continue
+performing resolutions and later achieve a different single literal at
+the level $d$. In particular, a decision level might contain more than
+one unique implication point. The algorithms given in
+Figure~\ref{fig:cl_schemes} stop at the first UIP of a level, except
+for the all-decision schemes with stops at the last UIP of each level.
+
+\begin{figure}[t!]
+{\fss{8pt}{10pt}
+\begin{tabular}[t]{|l|c|l|}
+  \cline{1-1}\cline{3-3}
+  \begin{minipage}[t]{0.48\textwidth}
+    \textbf{(a) All Decision Clause\rule{0pt}{1.1\topskip}}
+    \begin{algorithmic}[0]
+        \State \hspace*{-1em}\textbf{all\_decision}($C_I$)
+        \State $C\gets C_I$
+        \While{$\{l\,|\, l\in C \land\reason{l} \neq\varnothing\} \neq \emptyset$}
+        \State \begin{tabular}{ll}
+                 $\ell \gets \mbox{}$ & literal with highest trail index in\\
+                                      & $\{l\,|\, l\in C \land\reason{l} \neq\varnothing\}$
+               \end{tabular}
+        \State $C \gets C\resolve \reason{\lnot \ell}$
+        \EndWhile
+        \State \textbf{return} $C$
+    \end{algorithmic}
+  \end{minipage}
+  & &
+  \begin{minipage}[t]{0.475\textwidth}
+      \textbf{(b) Make level $i$ contain a single literal\rule{0pt}{1.1\topskip}}
+      \begin{algorithmic}[0]
+          \State \hspace*{-1em} \textbf{UIP\_level}($C$, $i$)
+          \While{$\left|\left\{\ell \left| \begin{array}{l}\ell \in C\\
+                                           \mbox{}\land \reason{l} \neq \varnothing \\
+                                           \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}\right| > 1$}
+          \State \begin{tabular}{ll}
+                   $l \gets \mbox{}$ & literal with highest trail index in\\
+                                        & $\left\{\ell \left| \begin{array}{l}\ell \in C
+                                          \land \reason{l} \neq \varnothing \\
+                                          \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}$
+                 \end{tabular}
+          \State $C \gets C\resolve \reason{\lnot l}$
+          \EndWhile
+          \State \textbf{return} $C$\strut
+      \end{algorithmic}
+  \end{minipage}
+  \\\cline{1-1}\cline{3-3}
+   \multicolumn{1}{c}{\vspace*{-7pt}}\\\cline{1-1}\cline{3-3}
+  \begin{minipage}[t]{0.475\textwidth}
+    \begin{minipage}{\textwidth}
+        \textbf{(c) First UIP Clause \rule{0pt}{1.1\topskip}}
+        \begin{algorithmic}[0]
+            \State \hspace*{-1em}\textbf{1-UIP}($C_I$)
+            \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
+            \State \textbf{return} \textbf{UIP\_level}($C_I$, i)
+        \end{algorithmic}
+        \vspace*{5pt}
+    \end{minipage}\\[5pt]
+    \hspace*{-1.8pt}\rule{1.03\textwidth}{.4pt}\\[5pt]
+    \begin{minipage}{\textwidth}
+        \textbf{(d) All UIP Clause \rule{0pt}{1.1\topskip}}
+        \begin{algorithmic}[0]
+          \State \hspace*{-1em}\textbf{all-UIP}($C_I$, $i$)
+          \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
+          \State \textbf{return} \textbf{$i$-UIP}(C, i)
+      \end{algorithmic}
+      \end{minipage}
+  \end{minipage}
+ & &
+  \begin{minipage}[t]{0.48\textwidth}
+      \vspace*{-20pt}
+      \textbf{(e) $i$-UIP Clause}
+      \begin{algorithmic}[0]
+          \State \hspace*{-1em}\textbf{$i$-UIP}($C_I$, $i$)
+          \State $C\gets C_I$
+          \State $d\gets \max\{\dlevel{l}\,|\, l \in C\}$
+          \For{($j \gets 1$; $j\leq i$; $j\gets j+1$)}
+          \State \textbf{if} ($d = \varnothing$): \textbf{break}
+          \State $C\gets\mbox{}$ \textbf{UIP\_level(C, d)}
+          \State $d\gets \max\left\{\dlevel{l} \left|\begin{array}{l}l\in C \\ \mbox{} \land \dlevel{l}<d
+                                                  \end{array}\right.\right\}$
+          \EndFor
+          \State \textbf{return} $C$\strut
+      \end{algorithmic}
+      \vspace*{3pt}
+      \textit{Maximum of an empty set is $\varnothing$ \strut}
+  \end{minipage}
+\\\cline{1-1}\cline{3-3}
+\end{tabular}
+}
+\caption{Some different clause learning schemes. All use the
+      current trail $\trail$ and take as input an initial clause $C_I$
+      falsified by $\trail$ at its deepest level.\label{fig:cl_schemes}}
+\end{figure}
+
+ %   \floatstyle{boxed}
+ %   \restylefloat{figure}
+
+\subsection{Asserting Clauses and LBD---Reasons to prefer 1-UIP
+  clauses}
+An \textbf{asserting clause} \cite{DBLP:journals/ai/PipatsrisawatD11}
+is a conflict clause $C_L$ that has exactly one literal $\ell$ at its
+deepest level, i.e.,
+$\forall l\in C_L. \dlevel{l} \leq \dlevel{\ell} \land (\dlevel{l} =
+\ell \implies l = \ell$). All of the clause learning schemes in
+Figure~\ref{fig:cl_schemes} produced learnt clauses that are asserting.
+
+The main advantage of asserting clauses is that they are 1-Empowering
+\cite{DBLP:journals/ai/PipatsrisawatD11}, i.e., they allow unit
+propagation to derive a new forced literal. Hence, asserting clauses
+can be used to guide backtracking---the solver can backtrack from the
+current deepest level to the point the learnt clause first becomes
+unit and then use the learnt clause to add a new unit implicant to the
+trail. Since all but the deepest level was propagation complete, this
+means that the asserting clause must be a brand new clause; otherwise
+that unit implication would already have been made. On the other hand,
+if the learnt clause $C_L$ is not asserting then \emph{it could be
+  that it is a duplicate of another clause already in the formula}. In
+particular, $C_L$ contains at least two literals at the deepest level
+since it is not asserting. If two of its literals at the deepest level
+were were not completely unit propagated (unit propagation is aborted
+as soon as a conflict is found), then $C_L$ could already be in the
+formula: it's two watches might not be fully propagated and $C_L$
+could be a falsified clause not detected by the solver. (In general,
+more than one clause might be falsified by $\trail$ and the \sat
+solver will generally stop when it finds the first one).
+
+The LBD of the learnt clause $C_L$ is the number of different decision
+levels in it:
+$\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
+\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a empirically
+successful predictor of clause usefulness: clauses with lower LBD tend
+to be more useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09} from
+the initial falsified clause $C_I$ the 1-UIP scheme will produce a
+clause $C_L$ whose LBD is minimum amongst all asserting clauses that
+can be learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a
+reason clause $\reason{l}$ ($l\in C$), then $\LBD(C') \geq \LBD(C)$
+since $\reason{l}$ must contain at least one other literal with the
+same decision level as $l$. That is, each trail resolution step can
+only increase the LBD of the learnt clause. The 1-UIP scheme performs
+the minimum number of trail resolution steps required to generate an
+asserting clause.
+
+Putting these two observations together we see that the 1-UIP scheme
+produces asserting clauses with lowest possible LBD. This is a
+compelling reasons for using this scheme. Hence, it is not surprising
+that modern \sat sovlers almost exclusively use 1-UIP clause
+learning.\footnote{Knuth in his sat13 CDCL solver \cite{Knuth:Sat13}
+  uses an all-decision clause when the 1-UIP clause is too large. In
+  this context an all-UIP clause could also be used as it would be no
+  larger than the all decision clause.}
+
+\section{Using $i$-UIP Clause Learning}
 \label{sec:i-uip}
-Classical CDCL SAT solvers analyze each conflict with 1-UIP resolution
-learning scheme to derive an asserting clause $\oneUIPClause$. Even
-though prior studies shows that further resolutions on $\oneUIPClause$
-against the assertion trail $\assertionTrail$ cannot reduce the
-clause's literal block distance ($\LBD$), the resolutions could
-potentially reduce the clause's size. Clause size is an important
-quality measurement because a smaller clause 1) consumes less memory,
-2) requires less steps to force a literal and 3) decreases the size of
-future conflict clauses.
-
-Inspired by both $\LBD$ and clause size, i-UIP resolution learning
-attempts to resolve away literals in $\oneUIPClause$ against the
-assertion trail $\assertionTrail$ to minimize the clause size without
-increasing the clause's $\LBD$. The goal of i-UIP learning is to find
-a clause $\iUIPClause$ whose literals are either the unique
-implication points in their respective decision levels or are directly
-implied by literals from a foreign decision
-level. Alg. ~\ref{alg:i-uip} is a pseudo-code implementation of i-UIP
-learning.
+Although learning clauses with low LBD has been shown to be important
+in \sat solving \cite{DBLP:conf/ijcai/AudemardS09}, clause size is
+also important. Smaller clauses consume less memory and help to
+decrease the size of future learnt clauses. They are also semantically
+stronger than longer clauses.
+
+The $i$-UIP scheme will tend to produce small clauses since the
+clauses can contain at most one literal per decision level. However,
+since more trail resolution steps are required to generate them they
+will also tend to contain more decision levels and thus have higher
+LBD. They will, however, be asserting.
+
+Since, $\LBD$ has shown itself to be more important than size our
+approach is to use $i$-UIP learning when, and only when, it succeeds
+in reducing the size of the clause \emph{without increasing its
+  LBD}. It can be noted that the $i$-UIP scheme first computes the
+1-UIP clause when it reduces the deepest level to a single UIP
+literal. It then proceeds to reducing the shallower levels (see $i$-UIP's
+for loop in Figure~\ref{fig:cl_schemes}). So our approach will start
+with the normal 1-UIP clause and then try to apply $i$-UIP learning to
+reduce other levels to single literals. As noted above, clause
+minimization is orthogonal to our approach, so we also first apply
+standard clause minimization \cite{DBLP:conf/sat/SorenssonB09} to the
+1-UIP clause. That is, our algorithm will start with the clause that
+most \sat solvers learn from a conflict (a minimized 1-UIP
+clause). Our approach is specified in Alg. ~\ref{alg:i-uip}.
 
@@
 \begin{algorithmic}[1]
-\Require  $\oneUIPClause$ is a valid and minimized 1-UIP clause
-\Require  $\assertionTrail$ is a valid assertion trail
-\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$} 
-    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause} \label{ln:init}
-    \State \code{DecisionLvs} $\gets$ Decision levels in $\oneUIPClause$ in descending order
-\For{$i \in \code{DecisionLvs}$} \label{ln:iterDecisionLvls}
-    \State $L_{i} \gets \Set{l}{\code{level}(l) = i \wedge l \in \iUIPClause}$
-    \While{$|\code{unmarked}(L_i)| > 1$} \label{ln:uip}
-        \State $p \gets \text{lit with the highest trail position in } L_{i} $ \label{ln:trailaccess}
-        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs}} \label{ln:resolvable}
-            \If{$\IUIPPURE$}
-                \State Unresolve literals at level $i$  \label{ln:unresolve}
-                \State Go to the next decision level $i$ \label{ln:abondon}
-            \ElsIf{$\IUIPMIN$}
-                \State $\code{Mark}(p)$ \label{ln:mark}
+\Require $\oneUIPClause$ is minimized 1-UIP clause
+\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$, \type} 
+    \State $\iUIPClause \gets \oneUIPClause$
+    \State $\dlevels \gets\mbox{}$ decision levels in $C_i$ in descending order
+           \Comment{These never change}
+    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest level $\dlevels[0]$}
+       \State $C_i\gets\mbox{}$\tryuiplevel($C_i$, $\dlevels[i]$, \type)
+              \Comment{Try to reduce this level to UIP}
+       \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \leq i\}\bigr|
+                    + (|\dlevels| - (i+1)) \geq |\oneUIPClause|$}
+           \State \textbf{return} $\oneUIPClause$.\label{ln:earlyTerminate} \Comment{can't generate smaller clause}
        \EndIf
-        \Else 
-            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(\neg{p}, \assertionTrail)$
-           \State  $\code{Update}(L_{i})$
+    \EndFor
+    \State \textbf{if} ($\type = \IUIPPURE$) $C_i \gets \mathbf{miminize}(C_i)$\label{ln:puremin}
+    \State \textbf{return} if $|\iUIPClause| < |\oneUIPClause|$) then  $\iUIPClause$ else  $\oneUIPClause$\label{ln:returnmin}
+\EndProcedure
+\Statex
+\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{Do not add new decision levels}
+    \State $\ctry = C_i$
+    \State $L_i = \left\{\ell \left| \ell \in \ctry 
+                         \land \dlevel{l} = i\right.\right\}$\label{ln:litsati1}
+    \While{$|L_i| > 1$}
+        \State $p \gets\mbox{}$ \textbf{remove} lit with the highest trail index from $L_i$ 
+        \If{($\exists q \in \reason{\lnot p}.\, \dlevel{q} \not \in \mathit{decLvls}$)}\label{ln:newdlevel}
+        \Comment{Would add new decision levels}
+             \If{($\type = \IUIPPURE$)}
+                 \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
+             \ElsIf{($\type = \IUIPMIN$)}
+                 \State \textbf{continue} \Comment{Don't try to resolve away $p$}
+             \EndIf
+        \Else\label{ln:nonewdlevel}
+            \State $\ctry \gets \ctry \resolve \reason{\lnot p}$
+            \State $L_i = L_i \cup \{\ell\,|\, 
+                                     \ell\in\reason{\lnot p} \land \ell\neq \lnot p \land \dlevel{\ell} = i
+                                    \}$\label{ln:litsati2}
         \EndIf
      \EndWhile
-\EndFor \label{ln:endCompleteResolution}
-\\
-\If{$\IUIPPURE$}\label{ln: 2ndMinStart}   \Comment{$\iUIPClause$ from $\IUIPPURE$ can be minimized}
-    \State $\iUIPClause  \gets \textbf{Minimize}(\iUIPClause)$ 
-\EndIf\label{ln: 2ndMinEnd}
-\\
-\If{$|\iUIPClause| < |\oneUIPClause|$} \label{ln:checkSize}
-    \State \textbf{return} $\iUIPClause$
-\Else
-    \State \textbf{return} $\oneUIPClause$
-\EndIf  \label{ln:cSizeCheck}
+    \State \textbf{return} $\ctry$
 \EndProcedure
@@
 
-The algorithm $\IUIP$ computes $\iUIPClause$ from (line~\ref{ln:init}
-to \ref{ln:endCompleteResolution}), and then returns the smaller
-clause between $\oneUIPClause$ and $\iUIPClause$
-(line~\ref{ln:checkSize} to \ref{ln:cSizeCheck}).  The algorithm first
-initializes $\iUIPClause$ with the minimized\cite{}
-$\oneUIPClause$. Next, the algorithm iterate through the decision
-levels of $\iUIPClause$ in descending order
-(line~\ref{ln:iterDecisionLvls}) and tries to find the unique
-implication point in each level (line~\ref{ln:uip}). Since $\IUIP$
-needs to preserve $\iUIPClause$'s $\LBD$ during resolutions, before
-resolving away a target literal $p$, the algorithm preemptively checks
-whether the reasoning clause of $\neg{p}$ contains any literal $q$
-from an unseen decision level (line~\ref{ln:resolvable}). If such a
-literal $q$ exists, then the algorithm cannot find the unique
-implication point (UIP) at the current level $i$ because resolving
-away $p$ will introduce $q$ into the clause and increases $\LBD$. One
-solution is to abandon level $i$ by reverting back to the state before
-any literal at level $i$ is resolved away (line~\ref{ln:unresolve}),
-and then move on to the next level (line~\ref{ln:abondon}). We denote
-$\IUIP$ learning with this solution as $\IUIPPURE$. The final
-$\iUIPClause$ produced by $\IUIPPURE$ will contains exactly one
-literal for levels where the LBD-persevering UIP exist. However,
-$\IUIPPURE$ does not minimize the number of literals in the decision
-levels where the LBD-preserving UIP does not exist, and clause
-minimization techniques\cite{} can be used to further reduce the
-clause size (line~\ref{ln: 2ndMinStart} to \ref{ln: 2ndMinEnd}).
-
-When $\IUIP$ cannot resolve away a literal $p$ without increasing
-$\iUIPClause$'s LBD, instead of skipping the entire decision level
-$i$, another solution $\IUIPMIN$ simply marks literal $p$ and keep it
-in $\iUIPClause$ (line~\ref{ln:mark}). The learning scheme then
-continue the resolutions at level $i$. $\IUIPMIN$ will ignore all the
-unsolvable literals and find the "local" unique implication point at
-every decision level.  The algorithm terminates when there is exactly
-one unmarked literal left for each decision level of $\iUIPClause$,
-representing the local unique implication points. The clause
-$\iUIPClause$ learnt by $\IUIPMIN$ is minimized\footnote{clause size
-  cannot be reduced via minimization techniques\cite{}.}, we defer the
-proof of minimization to Appendix~\ref{}.
+The Algorithm $\IUIP$ attempts to reduce each level in the 1-UIP
+clause to a single literal (a UIP). The 1-UIP clause  has only
+one literal at its deepest level, so that level, $\dlevels[0]$, can be
+skipped. The subroutine \tryuiplevel is used to reduce each level. It
+uses trail resoluions to achieve this, subject to the constraint that
+no new decision levels can be introduced in the clause. In particular,
+\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
+decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
+(line~\ref{ln:litsati1}), in order of decreasing trail index, until only
+one literal at level $i$ remains. If the resolution step will not
+introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
+performed updating $\ctry$. In addition, all new literals at level $i$
+are added to $L_i$. These can only be lits at level $i$ in the reason
+clause.
+
+On the other hand, if the resolution step would introduce new decision
+levels (line~\ref{ln:newdlevel}) then there are two options.  The
+first option we call $\IUIPPURE$. With $\IUIPPURE$ we abort our
+attempt to UIP this level and return the clause with level $i$
+unchanged. In the second option, called $\IUIPMIN$, we continue
+without performing the resolution, \textit{keeping} the current
+literal $p$ in $\ctry$. $\IUIPMIN$ will continue to try to resolve
+away the other literals in $L_i$ (note that $p$ is no longer in $L_i$)
+until $L_i$ is reduced to a single literal. Hence, $\IUIPMIN$ can
+leave multiple literals at level $i$---all of those with reasons
+containing new levels along with one other. Note that since the sole
+remaining literal $u\in L_i$ is at a lower trail index than all of the
+kept literals, so there is no point in trying to resolve away
+$u$---either it will be the decision clause for level $i$ having no
+reason, or its reason will contain at least one other literal at level
+$i$.
+
+After trying to UIP each level the clause $C_i$ is obtained. If we
+were using $\IUIPPURE$ we can once again apply minimization
+(line~\ref{ln:puremin}). Recursive clause minimization
+\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\IUIPMIN$
+clause as all but one literal of each level introduces a new level and
+thus cannot be recursively removed.\footnote{Other more powerful
+  minimization techniques could still be applied.}
+
+Finally, the Algorithm returns the shorter of the initial 1-UIP clause
+and the newly computed clause (line~\ref{ln:returnmin}). Because we
+return the newly computed clause only when it is shorter an early
+termination test can be used (line~\ref{ln:earlyTerminate}). After the
+algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
+the literals at those levels cannot be further changed. Furthermore,
+we know that the best that can be done is to reduce the remaining
+$|\dlevels| - (i+1)$ levels down to a single literal each. These two
+observations give a lower bound on the size of the resulting clause,
+and if that lower bound is as large as the size of the initial 1-UIP
+clause we can terminate and return the initial 1-UIP clause.
 
 \begin{example}
-    Consider the assertion trail
-    $\assertionTrail =
-    {..a_2,b_2,c_2,d_2..e_5,f_5,g_5,h_6,i_6,j_6,k_6..}$ where literals
-    $a_2 ... k_{6}$ are asserted in order with decision level shown as
-    the underscript. Let $L_{ext}$ denote the set of literals on
-    $\assertionTrail$ outside of decision level 2,5,6 and 10, and
-    $l^{*}_{ext}$ be some literal in $L_{ext}$. The reasoning clauses
-    involved in building $\assertionTrail$ are:
-\begin{enumerate}
-  \item[] $C_k = \{\neg{f_5} \vee \neg{j_6} \vee k_6\}$,
-  $C_j = \{\neg{f_5} \vee \neg{i_6} \vee j_6\}$, 
-  $C_i = \{\neg{e_5} \vee \neg{h_6} \vee i_6\}$,
-  $C_g = \{\neg{a_2} \vee \neg{f_5} \vee g_5\}$
-  $C_f = \{\neg{l^{*}_{ext}} \vee \neg{e_5} \vee f_5\}$,
-  $C_d = \{\neg{b_2} \vee \neg{c_2} \vee d_2\}$,
-  $C_c = \{\neg{a_2} \vee \neg{b_2} \vee c_2 \}$,
-  $C_b = \{\neg{l^{*}_{ext}} \vee \neg{a_2} \vee b_2\}$
-\end{enumerate}
-Suppose 1-UIP learning produces clause
-$\oneUIPClause = \{\neg{m_{10}} \vee \neg{k_6} \vee \neg{j_6} \vee
-\neg{i_6} \vee \neg{h_6} \vee \neg{g_5} \vee \neg{d_2} \vee
-\neg{c_2}\}$ where $\neg{k_{10}}$ is the UIP from the conflicting
-level. $\IUIP$ first tries to find the UIP for level 6 by resolving
-$\oneUIPClause$ with $C_k$, $C_j$ and, $C_i$ in order and produces
-clause
-$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
-\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ where
-$\neg{h_6}$ is the UIP.
-
-$\IUIP$ attempts to find the UIP for level 5 by resolving $C_i$ with
-$C_g$ and $C_f$ in order. However, resolving with $C_f$ will introduce
-$\neg{l^{*}_{ext}}$ into $C_i$ and increase the clause's
-LBD. $\IUIPPURE$ undoes all resolutions at level 5 and skips the
-entire level, whereas $\IUIPMIN$ only skips the resolution with $C_f$
-and tries to find the UIP by ignoring $\neg{f_5}$.  After level 5,
-$\IUIPPURE$ preserves the same clause and $\IUIPMIN$ produces
-$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
-\neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \vee \neg{a_2}\}$ where
-$\neg{e_5}$ is the UIP when $\neg{f_5}$ is ignored. The same $\IUIP$
-procedure is then performed on level 2 to produce the final
-$\iUIPClause$.
-
-$\IUIPPURE$ learned
-$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{g_5} \vee
-\neg{f_5} \vee \neg{e_5} \vee \neg{d_2} \vee \neg{c_2} \}$ and reduces
-the learnt clause size by 1. $\IUIPMIN$ learned
-$\iUIPClause = \{\neg{m_{10}} \vee \neg{h_6} \vee \neg{f_5} \vee
-\neg{e_5} \vee \neg{b_2} \vee \neg{a_2}\}$ and reduces the learnt
-clause size by 2.
+    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
+    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
+    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
+    $\ldots$ where the subscript indicates the decision level of each
+    literal and the literals are in order of increasing trail
+    index. Let the clauses
+    $C_x$ denote the reason clause for literal $x_i$ with 
+    \[\begin{array}{|l|l|l|}\hline
+      C_{a} = \varnothing 
+      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
+      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
+      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
+      & C_{\ell} =\varnothing
+      & C_{e} =\varnothing \\\hline
+      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
+      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
+      & C_{h} =\varnothing \\\hline
+      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
+      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
+      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
+      \end{array}
+  \]
+  Suppose 1-UIP learning yields the clause
+  $\oneUIPClause = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
+  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
+  is the UIP from the conflicting level. $\IUIP$ first tries to find
+  the UIP for level 6 by resolving $\oneUIPClause$ with $C_k$, $C_j$
+  and then $C_i$ producing the clause
+  $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$,
+  $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where
+  $\neg{h_6}$ is the UIP for level 6.
+
+  $\IUIP$ then attempts to find the UIP for level 5 by resolving $C^*$
+  with $C_g$ and then $C_f$. However, resolving with $C_f$ would
+  introduce $\ell_1$ and a new decision level into $C^*$. $\IUIPPURE$
+  thus leaves level 5 unchanged . $\IUIPMIN$, on the other hand, skips
+  the resolution with $C_f$  leaving $f_5$ in $C^*$. Besides $f_5$
+  only one other literal at level 5 remains in the clause, $e_5$,
+  so $\IUIPMIN$ does not do any further resolutions at this level. Hence,
+  $\IUIPPURE$ yields $C^*$ unchanged, while $\IUIPMIN$ yeilds
+  $C^*_{\mathit{min}} = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$,
+  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2}, \lnot a_2)$.
+
+  Finally, $\IUIP$ processes level 2. Resolving away $d_2$ and then
+  $c_2$ will lead to an attempt to resolve away $b_2$. But again this
+  would introduce a new decision level with the literal $\ell_1$. So
+  $\IUIPPURE$ will leave level 2 unchanged and $\IUIPMIN$ will leave
+  $b_2$ unresolved. The final clauses produced by $\IUIPPURE$ would be
+  $(\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$,
+  $\neg{c_2}$, $\neg{a_2})$, a reduction of 1 over the 1-UIP clause,
+  and by $\IUIPMIN$ would be $(\neg{m_{10}}$, $\neg{h_6}$,
+  $\neg{f_5}$, $\neg{e_5}$, $\neg{b_2}$, $\neg{a_2})$, a reduction of
+  2 over the 1-UIP clause.
+\whitebox
 \end{example}
@@
 simple. However, to make the algorithm a practical learning scheme
-that is compatible with modern SAT solvers, a number of issues need to
+that is compatible with modern SAT solvers some further issues need to
 be addressed. The rest of the section details the key optimizations
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[b]
 \caption{Control-$\IUIP$}\label{alg:enableCondition}
@@
 
-\begin{figure}
+2\begin{figure}
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF.pdf}
-    \caption{ Relative clause size reduction distribution.
-X axis indicates the relative size of difference between i-uip and 1-uip clauses (calculated as $\dfrac{|\oneUIPClause|-|\iUIPClause|}{|\oneUIPClause|}$ ) for each instance, and Y axis shows the probability density.}
+    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF}
+    \caption{ Relative clause size reduction distribution. X axis
+      indicates the relative size of difference between i-uip and
+      1-uip clauses (calculated as
+      $\dfrac{|\oneUIPClause|-|\iUIPClause|}{|\oneUIPClause|}$ ) for
+      each instance, and Y axis shows the probability density.}
      \label{fig:len_pdf}
@@
     \includegraphics[width=0.8\textwidth]{figures/i-uip-sizes-compare.pdf}
-    \caption{Average clause size comparison plot.
-Each point in the plot represents an instance.
-X and Y axis shows the clause length from $\IUIP$ and 1-UIP, respectively.
-Each green (red) dot represents an compared instance between $\MapleBase$ and $\MapleIUIMIN$ ($\IUIPPURE$).
-}
+    \caption{Average clause size comparison plot. Each point in the
+      plot represents an instance. X and Y axis shows the clause
+      length from $\IUIP$ and 1-UIP, respectively. Each green (red)
+      dot represents an compared instance between $\MapleBase$ and
+      $\MapleIUIMIN$ ($\IUIPPURE$). }
     \label{fig:len_compare}
@@
 \bibliographystyle{splncs04}
-
 \end{document}

### 2020-02-16 paper.tex
fb 1st pass complete

@@
 %\PassOptionsToPackage{hypnens}{url}
-\title{Clause size reduction with $i$-UIP Learning}
+\title{Clause size reduction with all-UIP Learning}
 \author{Nick Feng \and Fahiem Bacchus}
@@
 \newcommand{\code}[1]{\text{#1}}
-\newcommand{\inputF}{\ensuremath{F}}
-\newcommand{\oneUIPClause}{\ensuremath{C_{1}}}
-\newcommand{\iUIPClause}{\ensuremath{C_{i}}}
 \newcommand{\assertionTrail}{\trail}
-\newcommand{\LBD}{\text{LBD}}
-\newcommand{\GAP}{\text{Gap}}
-\newcommand{\IUIP}{\textbf{LBDstable-$i$UIP}}
-\newcommand{\tryuiplevel}{\textbf{try-uip-level}\xspace}
-\newcommand{\IUIPPURE}{\textbf{Pure-$i$UIP}}
-\newcommand{\IUIPMIN}{\textbf{Min-$i$UIP}}
-\newcommand{\IUIPGreedy}{\textbf{i-UIP-Greedy}}
-\newcommand{\IUIPActive}{\textbf{i-UIP-Inclusive}}
-\newcommand{\IUIPDist}{\textbf{i-UIP-Exclusive}}
+\newcommand{\LBD}{\text{LBD}\xspace}
+\newcommand{\gap}{\text{Gap}}
+\newcommand{\tgap}{t_{\mathit{gap}}}
+\newcommand{\ntries}{n_{\mathit{tries}}}
+\newcommand{\nsuc}{n_{\mathit{succ}}}
+\newcommand{\allUip}{\textit{stable-allUIP}}
+\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
+\newcommand{\allUipPure}{\textit{pure-allUIP}\xspace}
+\newcommand{\allUipMin}{\textit{min-allUIP}\xspace}
+\newcommand{\allUipAct}{\textit{allUIP-Active}}
+\newcommand{\allUipIn}{\textit{allUIP-Inclusive}}
+\newcommand{\allUipEx}{\textit{allUIP-Exclusive}}
 \newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
@@
 \newcommand{\expSATShort}{\textit{expMaple} }
-\newcommand{\MapleIUIPPURE}{\text{Maple-\IUIPPURE}}
-\newcommand{\MapleIUIMIN}{\text{Maple-\IUIPMIN}}
+\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
+\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
 \newcommand{\MapleEightShort}{\textit{MapleCB}}
@@
 %Comments
-\newcommand{\nf}[1]{\textcolor{red}{#1}}
-\newcommand{\fb}[1]{\textcolor{blue}{#1}}
+\newcommand{\nf}[1]{{\color{red}{#1}}}
+\newcommand{\fb}[1]{{\color{blue}{#1}}}
 
@@
 
-One of the previously proposed schemes was the $i$-UIP scheme
+One of the previously proposed schemes was the all-UIP scheme
 \cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
 to exploit the main ideas of this scheme that avoids its main
-disadvantages. In particular, we propose to use a $i$-UIP like clause
+disadvantages. In particular, we propose to use a all-UIP like clause
 learning scheme to generate smaller learnt clauses which retain the
 good properties of standard 1-UIP clauses. Our method is related to,
-but not the same as various clause mininimization methods that aim to
-remove literals from the 1-UIP clause, e.g.,
+but not the same as, various clause mininimization methods that try to
+remove redundant literals from the 1-UIP clause yielding a clause that
+is a subset of the 1-UIP clause, e.g.,
 \cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
-Our method is orthogonal to clause minimization. In particular, 
-our approach can learn a clause that is completely different from the
-1-UIP clause but which still serves all of the same purposes as the 1-UIP
-clause. Like the 1-UIP clause, various minimization techniques can be
-applied on top of our method to further reduce the size of the clauses
-we learn.
+Our method is orthogonal to clause minimization. In particular, our
+approach can learn a clause that is not a subset of 1-UIP clause but
+which still serves all of the same purposes as the 1-UIP clause.
+Various minimization techniques can be applied on top of our method to
+remove redundant literals from the clauses we learn.
 
@@
 denote its underlying variable. A CNF consists of a conjunction of
-clauses, each if which is a disjunction of literals. We often view a
+clauses, each of which is a disjunction of literals. We often view a
 clause as being a set of literals and employ set notation, e.g.,
@@
 
-Two clauses $c_1$ and $c_2$ can be \textit{resolved} when they contain
-conflicting literals $\ell\in c_1$ and $\lnot \ell \in c_2$. Their
-resolvant $c_1 \resolve c_2$ is the new clause
-$(c_1 \cup c_2) - \{\ell, \lnot \ell\}$. The resolvant is tautological
-if $c_1$ and $c_2$ contain more than one pair of conflicting literals.
+Two clauses $C_1$ and $C_2$ can be \emph{resolved} when they contain
+\emph{conflicting} literals $\ell\in C_1$ and $\lnot \ell \in
+C_2$. Their resolvant $C_1 \resolve C_2$ is the new clause
+$(C_1 \cup C_2) - \{\ell, \lnot \ell\}$. The resolvant will be a
+tautology (i.e., a clause containing a literal $l$ and its negation
+$\lnot l$) if $C_1$ and $C_2$ contain more than one pair of
+conflicting literals.
 
@@
 the trail and unassigned.  If literal $\ell$ is on the trail let
-$\trailIdx{\ell}$ denote its index on the trail sequence
-($\trail[\trailIdx{\ell}] = \ell$). For convenience, we also let
-$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$,
-where $\var(\ell)$ is $\ell$'s variable, even though $\lnot \ell$ nor
-$\var(\ell)$ are actually on $\trail$. If $x$ and $y$ are both on the
-trail and $\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$
-  appears before $y$ on the trail}. In a slight abuse of notation we
-also use $\trailIdx{v}$ on a variable $v$ where
-$\trailIdx{v}=\trailIdx{\ell}$ and $\ell$ a literal of $v$ on the
-trail.
-
-Two types of true literals appear on the trail: decision literals that
-have been assumed to be true by the solver, and unit propagated
-literals that are forced to be true because they are the sole remining
-unfalsified literal of a clause. Each literal $\ell\in\trail$ has a
-decision level $\dlevel{\ell}$ which is equal to the number of
-decision literals appearing before $\ell$ on the trail plus one;
-hence, $\dlevel{d}=1$ for the first decision literal $d\in\trail$.
-The set of literals on $\trail$ that have the same decision level
-forms a contiguous subsequence of $\trail$ that starts with a decision
-literal $d_i$ and ends just before the next decision literal
-$d_{i+1}$. If $\dlevel{d_i} = i$ we call this subsequence of $\trail$
-the \textit{$i$-th decision level}.
+$\trailIdx{\ell}$ denote its index on the trail, i.e,
+$\trail[\trailIdx{\ell}] = \ell$. For convenience, we also let
+$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$ even
+though neither $\lnot \ell$ nor $\var(\ell)$ are actually on
+$\trail$. If $x$ and $y$ are both on the trail and
+$\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$ appears before
+  $y$ on the trail}.
+
+Two types of true literals appear on the trail: \emph{decision
+  literals} that have been assumed to be true by the solver, and
+\emph{unit propagated literals} that are forced to be true because
+they are the sole remaining unfalsified literal of a clause. Each
+literal $\ell\in\trail$ has a decision level $\dlevel{\ell}$ which is
+equal to the number of decision literals appearing before $\ell$ on
+the trail plus one; hence, $\dlevel{d}=1$ for the first decision
+literal $d\in\trail$.  The set of literals on $\trail$ that have the
+same decision level forms a contiguous subsequence of $\trail$ that
+starts with a decision literal $d_i$ and ends just before the next
+decision literal $d_{i+1}$. If $\dlevel{d_i} = i$ we call this
+subsequence of $\trail$ the \textit{$i$-th decision level}.
 
@@
 
-In most \sat solver implementations clause learning is initiated as
-soon as a clause is falsified by $\trail$. In this paper we will be
-concerned with the subsequent clause learning process which uses
-$\trail$ to derive a new clause that can be added to the formula. In
-some implementations the trail might be altered during clause
-learning. Here, however, we will assume that $\trail$ remains intact
-during clause learning and is only changed after the new clause is
-derived. After the new clause is lerned the $\trail$ will be changed by backtracking.
-
-Say that $\trail$ falsifies a clause $c_I$, and that the last decision
+In most \sat solvers clause learning is initiated as soon as a clause
+is falsified by $\trail$. In this paper we will be concerned with the
+subsequent clause learning process which uses $\trail$ to derive a new
+clause. In some implementations the trail might be altered during
+clause learning. Here, however, we will assume that $\trail$
+\emph{remains intact during clause learning} and is only changed after
+the new clause is learnt. After the new clause is learnt the $\trail$
+will be changed by backtracking.
+
+Say that $\trail$ falsifies a clause $C_I$, and that the last decision
 literal $d_k$ in $\trail$ has decision level $k$. Consider
@@
 trail $\trail$ might not be. This means that (a) no clause was
-falsified by $\trail_{k-1}$. And (b) if $c_u$ is a clause containing
-the literal $x$ and all literals in $c_u$ except for $x$ are falsified
+falsified by $\trail_{k-1}$. And (b) if $C_u$ is a clause containing
+the literal $x$ and all literals in $C_u$ except for $x$ are falsified
 by $\trail_{k-1}$, then $x\in \trail_{k-1}$ and
-$\dlevel{x} \leq\max\{\dlevel{y} | y\in c_u \land y\neq x\}$. This
+$\dlevel{x} \leq\max\{\dlevel{y} | y\in C_u \land y\neq x\}$. This
 means that if $x$ appears in a clause made unit it must have been
-added to the trail, and added at the first decision level that this
-occurred.  Note, however, $x$ might appear in more than one clause.
-The less than equal condition above ensures that $x$ must appear in
-$\trail_{k-1}$ at the first decision level any of these clauses were
-made unit.
+added to the trail, and added at or before decision level the clause
+became unit. Note that more than one clause might be made unit by
+$\trail$ forcing $x$, or $x$ might be set as a decision before being
+forced. This condition ensures that $x$ appears in $\trail$ at or
+before the first decision level it is forced by any clause.
 
@@
     A trail resolvant is a clause arising from resolving a conflict
-    against the reason clause of some literal $\ell \in \trail$ where
-    $\reason{\ell}\neq \varnothing$. Every trail resolvant is also a
-    conflict.
+    against the reason clause of some literal $\ell \in \trail$. Every
+    trail resolvant is also a conflict.
 \end{defn}
@@
 propagated literal whose negation appears in the conflict; and (3) any
-literal in the conflict that is unit progagated in $\trail$ can be
-resolved upon (the corresponding variable must appear in different
-polarities in the conflict and in $\trail$).
+variable in the conflict that is unit progagated in $\trail$ can be
+resolved upon (the variable must appear in different polarities in the
+conflict and in $\trail$).
 
@@
     A trail resolution is a sequence of trail resolvants applied to an
-    initial conflict $c_I$ yielding a new conflict $c_L$.  A trail
+    initial conflict $C_I$ yielding a new conflict $C_L$.  A trail
     resolution is \textbf{ordered} if the sequence of variables $v_1$,
     \dots, $v_m$ resolved have strictly decreasing trail indicies:
-    $\trailIdx{v_i} > \trailIdx{v_{i+1}}$ ($1\leq i < m$). (Note that
+    $\trailIdx{v_{i+1}} < \trailIdx{v_i}$ ($1\leq i < m$). (Note that
     this implies that no variable is resolved on more than once).
@@
     If the unordered trail resolution $U$ yields the conflict clause
-    $c_L$ from an initial conflict $c_I$, then there exists an ordered
-    trail resolution $O$ that yields a conflict clause $c'_L$ such
-    that $c'_L\subseteq c_L$.
+    $C_L$ from an initial conflict $C_I$, then there exists an ordered
+    trail resolution $O$ that yields a conflict clause $C'_L$ such
+    that $C'_L\subseteq C_L$.
 \end{Obs}
 \noindent
-\textbf{Proof.} Let $U$ be the sequence of clauses $c_I = c_0$, $c_1$,
-$\ldots$, $c_m= c_L$ obtained by resolving on the sequence of
+\textbf{Proof.} Let $U$ be the sequence of clauses $C_I = C_0$, $C_1$,
+$\ldots$, $C_m= C_L$ obtained by resolving on the sequence of
 variables $v_1$, $\ldots$, $v_m$ whose corresponding literals on
@@
 was, and furthermore $O$ yields the clause
-$c'_L = \bigcup_{i=1}^m \reason{l_i} - \{l_1, \lnot l_1, \ldots, l_m,
+$C'_L = \bigcup_{i=1}^m \reason{l_i} - \{l_1, \lnot l_1, \ldots, l_m,
 \lnot l_m\}$. Since $U$ resolves on the same variables (in a different
-order) we must that that the clause it produces $c_L\subseteq
-c'_L$. But $c_L$ might retain one or more of the literals $\lnot l_1$,
-$\ldots$, $\lnot l_m$, since resolving on a literal $l_j$ with higher
-trail index might reintroduce a literal $\lnot l_i$ with lower trail
-index that was previously resolved away by an earlier resolution.
+order) using the same reason clauses we must have $C'_L\subseteq
+C_L$. It can, however, be the case that $C'_L$ is proper subset of
+$C_L$: if $l_i$ is resolved away it might be reintroduced when
+resolving on $l_{i+1}$ if $\trailIdx{l_{i+1}}  > \trailIdx{l_i}$.
 \whitebox
@@
           \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
+          \State \Comment{$i$ is large enough to ensure all levels are UIP}
           \State \textbf{return} \textbf{$i$-UIP}(C, i)
+
       \end{algorithmic}
@@
 deepest level, i.e.,
-$\forall l\in C_L. \dlevel{l} \leq \dlevel{\ell} \land (\dlevel{l} =
-\ell \implies l = \ell$). All of the clause learning schemes in
+$\forall x\in C_L. \dlevel{x} \leq \dlevel{\ell} \land (\dlevel{x} =
+\ell \implies x = \ell$). All of the clause learning schemes in
 Figure~\ref{fig:cl_schemes} produced learnt clauses that are asserting.
@@
 particular, $C_L$ contains at least two literals at the deepest level
-since it is not asserting. If two of its literals at the deepest level
-were were not completely unit propagated (unit propagation is aborted
-as soon as a conflict is found), then $C_L$ could already be in the
-formula: it's two watches might not be fully propagated and $C_L$
-could be a falsified clause not detected by the solver. (In general,
-more than one clause might be falsified by $\trail$ and the \sat
-solver will generally stop when it finds the first one).
+If two of its literals at the deepest level were were not completely
+unit propagated (unit propagation is aborted as soon as a conflict is
+found), then $C_L$ could already be in the formula: it's two watches
+might not be fully propagated and $C_L$ could be a falsified clause
+not detected by the solver. (In general, more than one clause might be
+falsified by $\trail$ and the \sat solver will generally stop when it
+finds the first one).
 
@@
 $\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
-\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a empirically
-successful predictor of clause usefulness: clauses with lower LBD tend
-to be more useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09} from
-the initial falsified clause $C_I$ the 1-UIP scheme will produce a
-clause $C_L$ whose LBD is minimum amongst all asserting clauses that
-can be learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a
-reason clause $\reason{l}$ ($l\in C$), then $\LBD(C') \geq \LBD(C)$
-since $\reason{l}$ must contain at least one other literal with the
-same decision level as $l$. That is, each trail resolution step can
-only increase the LBD of the learnt clause. The 1-UIP scheme performs
-the minimum number of trail resolution steps required to generate an
-asserting clause.
+\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a successful
+predictor of clause usefulness: clauses with lower LBD tend to be more
+useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09}, from the
+initial falsified clause $C_I$ the 1-UIP scheme will produce a clause
+$C_L$ whose LBD is minimum among all asserting clauses that can be
+learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a reason
+clause $\reason{l}$, then $\LBD(C') \geq \LBD(C)$ since $\reason{l}$
+must contain at least one other literal with the same decision level
+as $l$ and might contain literals with decision levels not in
+$C$. That is, each trail resolution step can only increase the \LBD of
+the learnt clause. Hence, the 1-UIP scheme yields an asserting clause
+with minimum LBD as it performs the minimum number of trail resolutions
+required to generate an asserting clause.
+
+The other schemes must perform more trail resolutions. In fact, all of
+these schemes (all-decision, all-uip, i-uip) use trail resolutions in
+which the 1-UIP clause appears. That is, they all must first generate
+the 1-UIP clause and then continue with further trail resolution
+steps. These extra resolution steps can introduce many addition
+decision levels into the final clause. Hence, these schemes yield
+learnt clauses whose LBD can be larger, and never smaller, that the
+1-UIP clause.
 
@@
 
-\section{Using $i$-UIP Clause Learning}
+\section{Using all-UIP Clause Learning}
 \label{sec:i-uip}
-Although learning clauses with low LBD has been shown to be important
-in \sat solving \cite{DBLP:conf/ijcai/AudemardS09}, clause size is
-also important. Smaller clauses consume less memory and help to
-decrease the size of future learnt clauses. They are also semantically
-stronger than longer clauses.
-
-The $i$-UIP scheme will tend to produce small clauses since the
-clauses can contain at most one literal per decision level. However,
-since more trail resolution steps are required to generate them they
-will also tend to contain more decision levels and thus have higher
-LBD. They will, however, be asserting.
-
-Since, $\LBD$ has shown itself to be more important than size our
-approach is to use $i$-UIP learning when, and only when, it succeeds
-in reducing the size of the clause \emph{without increasing its
-  LBD}. It can be noted that the $i$-UIP scheme first computes the
-1-UIP clause when it reduces the deepest level to a single UIP
-literal. It then proceeds to reducing the shallower levels (see $i$-UIP's
-for loop in Figure~\ref{fig:cl_schemes}). So our approach will start
-with the normal 1-UIP clause and then try to apply $i$-UIP learning to
-reduce other levels to single literals. As noted above, clause
-minimization is orthogonal to our approach, so we also first apply
-standard clause minimization \cite{DBLP:conf/sat/SorenssonB09} to the
-1-UIP clause. That is, our algorithm will start with the clause that
-most \sat solvers learn from a conflict (a minimized 1-UIP
-clause). Our approach is specified in Alg. ~\ref{alg:i-uip}.
-
-\begin{algorithm}[t]
-\caption{\IUIP}\label{alg:i-uip}
+Although learning clauses with low LBD has been shown empirically to
+be important in \sat solving than learning short clauses
+\cite{DBLP:conf/ijcai/AudemardS09}, clause size is still
+important. Smaller clauses consume less memory and help to decrease
+the size of future learnt clauses. They are also semantically stronger
+than longer clauses.
+
+The all-UIP scheme will tend to produce small clauses since the
+clauses contain at most one literal per decision level. However, the
+all-UIP clause can have much higher LBD. Since, $\LBD$ is important
+than size our approach is to use all-UIP learning when, and only when,
+it succeeds in reducing the size of the clause \emph{without
+  increasing its LBD}. The all-UIP scheme first computes the 1-UIP
+clause when it reduces the deepest level to a single UIP literal. It
+then proceeds to reduce the shallower levels (see all-UIP's for loop
+in Figure~\ref{fig:cl_schemes}). So our approach will start with the
+1-UIP clause and then try to apply all-UIP learning to reduce other
+levels to single literals. As noted above, clause minimization is
+orthogonal to our approach, so we also first apply standard clause
+minimization \cite{DBLP:conf/sat/SorenssonB09} to the 1-UIP
+clause. That is, our algorithm will start with the clause that most
+\sat solvers learn from a conflict (a minimized 1-UIP clause). Our
+approach is specified in Alg. ~\ref{alg:i-uip}.
+
+\begin{algorithm}[!t]
+\caption{\allUip}\label{alg:i-uip}
 \begin{algorithmic}[1]
-\Require $\oneUIPClause$ is minimized 1-UIP clause
-\Procedure{\IUIP}{$\oneUIPClause, \assertionTrail$, \type} 
-    \State $\iUIPClause \gets \oneUIPClause$
+\Require $C_1$ is minimized 1-UIP clause
+\Require $\tgap\geq 0$ is global parameter, $\ntries$ and $\nsuc$
+         are globals used to configure $\tgap$
+\Procedure{\allUip}{$C_1, \assertionTrail$, \type} 
+    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return} $C_1$\label{ln:control}
+    \State $C_i \gets C_1$
     \State $\dlevels \gets\mbox{}$ decision levels in $C_i$ in descending order
@@
               \Comment{Try to reduce this level to UIP}
-       \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \leq i\}\bigr|
-                    + (|\dlevels| - (i+1)) \geq |\oneUIPClause|$}
-           \State \textbf{return} $\oneUIPClause$.\label{ln:earlyTerminate} \Comment{can't generate smaller clause}
+       \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq \dlevels[i]\}\bigr|
+                    + (|\dlevels| - (i+1)) \geq |C_1|$}\label{ln:earlyTerminate}
+           \State \textbf{return} $C_1$. \Comment{can't generate smaller clause}
        \EndIf
     \EndFor
-    \State \textbf{if} ($\type = \IUIPPURE$) $C_i \gets \mathbf{miminize}(C_i)$\label{ln:puremin}
-    \State \textbf{return} if $|\iUIPClause| < |\oneUIPClause|$) then  $\iUIPClause$ else  $\oneUIPClause$\label{ln:returnmin}
+    \State \textbf{if} ($\type = \allUipPure$) $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
+    \State $\ntries$++\label{ln:try}
+    \State \textbf{if} ($|C_i| < |C_1|$) $\nsuc$++, \textbf{return} $C_i$ 
+           \label{ln:returnmin}\label{ln:suc}
+    \State \textcolor{gray}{$\triangleright\mbox{}$ Use next line instead for} \allUipAct
+    \State \textcolor{gray}{
+           \textbf{if} ($|C_i| < |C_1| 
+                        \land (\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$)
+                       $\nsuc$++, \textbf{return} $C_i$} \label{ln:activity}
+    \State \textbf{return} $C_1$
+%    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
+%           \textbf{then}  $C_i$ \textbf{else}  $C_1$
 \EndProcedure
@@
         \Comment{Would add new decision levels}
-             \If{($\type = \IUIPPURE$)}
+             \If{($\type = \allUipPure$)}
                  \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
-             \ElsIf{($\type = \IUIPMIN$)}
+             \ElsIf{($\type = \allUipMin$)}
                  \State \textbf{continue} \Comment{Don't try to resolve away $p$}
@@
 
-The Algorithm $\IUIP$ attempts to reduce each level in the 1-UIP
-clause to a single literal (a UIP). The 1-UIP clause  has only
-one literal at its deepest level, so that level, $\dlevels[0]$, can be
+The Algorithm $\allUip$ attempts to reduce each level in the 1-UIP
+clause to a single literal (a UIP). The 1-UIP clause has only one
+literal at its deepest level, so that level, $\dlevels[0]$, can be
 skipped. The subroutine \tryuiplevel is used to reduce each level. It
@@
 decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
-(line~\ref{ln:litsati1}), in order of decreasing trail index, until only
-one literal at level $i$ remains. If the resolution step will not
+(line~\ref{ln:litsati1}), in order of decreasing trail index, until
+only one literal at level $i$ remains. If the resolution step will not
 introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
 performed updating $\ctry$. In addition, all new literals at level $i$
-are added to $L_i$. These can only be lits at level $i$ in the reason
+are added to $L_i$. These are the literals at level $i$ in the reason
 clause.
@@
 levels (line~\ref{ln:newdlevel}) then there are two options.  The
-first option we call $\IUIPPURE$. With $\IUIPPURE$ we abort our
+first option we call $\allUipPure$. With $\allUipPure$ we abort our
 attempt to UIP this level and return the clause with level $i$
-unchanged. In the second option, called $\IUIPMIN$, we continue
+unchanged. In the second option, called $\allUipMin$, we continue
 without performing the resolution, \textit{keeping} the current
-literal $p$ in $\ctry$. $\IUIPMIN$ will continue to try to resolve
+literal $p$ in $\ctry$. $\allUipMin$ then continues to try to resolve
 away the other literals in $L_i$ (note that $p$ is no longer in $L_i$)
-until $L_i$ is reduced to a single literal. Hence, $\IUIPMIN$ can
+until $L_i$ is reduced to a single literal. Hence, $\allUipMin$ can
 leave multiple literals at level $i$---all of those with reasons
-containing new levels along with one other. Note that since the sole
-remaining literal $u\in L_i$ is at a lower trail index than all of the
-kept literals, so there is no point in trying to resolve away
+containing new levels along with one other.\footnote{Since the sole
+  remaining literal $u\in L_i$ is at a lower trail index than all of
+  the other literals, so there is no point in trying to resolve away
   $u$---either it will be the decision clause for level $i$ having no
-reason, or its reason will contain at least one other literal at level
-$i$.
+  reason, or its reason will contain at least one other literal at
+  level $i$.} Observe that the number of literals at level $i$ can not
+be increased after processing it with \allUipPure. \allUipMin can,
+however, potentially increase the number of literals at a level. In
+resolving away a ltieral $l$ at level $i$ more literals might be
+introduced into level $i$, and some of these might not be removable by
+\allUipMin if their reasons contain new levels. However, both
+\allUipPure and \allUipMin can increase the number of literals at
+levels less that $i$ as new literals can be introduced into those
+levels when the literals at level $i$ are resolved away. These added
+literals at the lower levels might not be removable from the clause,
+and thus both methods might yield a longer clause than the inputted
+1-UIP clause.
 
 After trying to UIP each level the clause $C_i$ is obtained. If we
-were using $\IUIPPURE$ we can once again apply minimization
+were using $\allUipPure$ we can once again apply minimization
 (line~\ref{ln:puremin}). Recursive clause minimization
-\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\IUIPMIN$
+\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\allUipMin$
 clause as all but one literal of each level introduces a new level and
@@
 algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
-the literals at those levels cannot be further changed. Furthermore,
+the literals at those levels will not be further changed. Furthermore,
 we know that the best that can be done is to reduce the remaining
@@
 and if that lower bound is as large as the size of the initial 1-UIP
-clause we can terminate and return the initial 1-UIP clause.
+clause we can terminate and return the initial 1-UIP clause. As
+observed above, \allUipPure cannot trigger this early termination
+condition (it cannot increase the size of any processed level), but
+\allUipMin can.
+
+\paragraph{$\tgap$:}
+$\allUip$ can produce significantly smaller clause. However, when it
+does not yield a smaller clause, the cost of the additional resolution
+steps can hurt the solver's performance. Since resolution cannot
+reduce a clause's $\LBD$, The maximum size reduction obtainable from
+$\allUip$ is the difference between the 1-UIP clause's size and its
+$\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$ is small,
+applying $\allUip$ is unlikely to be cost effective. Our approach is
+to dynamically set a threshold on $\gap(C_1)$, $\tgap$, such that when
+$\gap(C_1) < \tgap$ we do not attempt to reduce the clause
+(line~\ref{ln:control}). Initially, $\tgap = 0$, and we count the
+number of times $\allUip$ is attempted ($\ntries$) and the number of
+times it successfully yields a shorter clause ($\nsuc$)
+(line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
+rate is less than 80\%, we decrease $\tgap$ by one (not allowing it to
+become negative), and if it is greater than 80\% we increase $\tgap$
+by one.
 
@@
   Suppose 1-UIP learning yields the clause
-  $\oneUIPClause = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
+  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
   $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
-  is the UIP from the conflicting level. $\IUIP$ first tries to find
-  the UIP for level 6 by resolving $\oneUIPClause$ with $C_k$, $C_j$
+  is the UIP from the conflicting level. $\allUip$ first tries to find
+  the UIP for level 6 by resolving $C_1$ with $C_k$, $C_j$
   and then $C_i$ producing the clause
@@
 
-  $\IUIP$ then attempts to find the UIP for level 5 by resolving $C^*$
+  $\allUip$ then attempts to find the UIP for level 5 by resolving $C^*$
   with $C_g$ and then $C_f$. However, resolving with $C_f$ would
-  introduce $\ell_1$ and a new decision level into $C^*$. $\IUIPPURE$
-  thus leaves level 5 unchanged . $\IUIPMIN$, on the other hand, skips
+  introduce $\ell_1$ and a new decision level into $C^*$. $\allUipPure$
+  thus leaves level 5 unchanged . $\allUipMin$, on the other hand, skips
   the resolution with $C_f$  leaving $f_5$ in $C^*$. Besides $f_5$
   only one other literal at level 5 remains in the clause, $e_5$,
-  so $\IUIPMIN$ does not do any further resolutions at this level. Hence,
-  $\IUIPPURE$ yields $C^*$ unchanged, while $\IUIPMIN$ yeilds
+  so $\allUipMin$ does not do any further resolutions at this level. Hence,
+  $\allUipPure$ yields $C^*$ unchanged, while $\allUipMin$ yeilds
   $C^*_{\mathit{min}} = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$,
@@
 
-  Finally, $\IUIP$ processes level 2. Resolving away $d_2$ and then
+  Finally, $\allUip$ processes level 2. Resolving away $d_2$ and then
   $c_2$ will lead to an attempt to resolve away $b_2$. But again this
   would introduce a new decision level with the literal $\ell_1$. So
-  $\IUIPPURE$ will leave level 2 unchanged and $\IUIPMIN$ will leave
-  $b_2$ unresolved. The final clauses produced by $\IUIPPURE$ would be
+  $\allUipPure$ will leave level 2 unchanged and $\allUipMin$ will leave
+  $b_2$ unresolved. The final clauses produced by $\allUipPure$ would be
   $(\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$,
   $\neg{c_2}$, $\neg{a_2})$, a reduction of 1 over the 1-UIP clause,
-  and by $\IUIPMIN$ would be $(\neg{m_{10}}$, $\neg{h_6}$,
+  and by $\allUipMin$ would be $(\neg{m_{10}}$, $\neg{h_6}$,
   $\neg{f_5}$, $\neg{e_5}$, $\neg{b_2}$, $\neg{a_2})$, a reduction of
@@
 
-The core algorithm of $\IUIP$ as a clause reduction technique is
-simple. However, to make the algorithm a practical learning scheme
-that is compatible with modern SAT solvers some further issues need to
-be addressed. The rest of the section details the key optimizations
-and augmentations of $\IUIP$ as a practical clause learning scheme.
-
-\subsection{Control i-UIP Learning}
-This simple and greedy $\IUIP$ learning scheme can produce
-significantly smaller clause. However, when $\IUIP$ does not reduce
-the clause size, the cost of the additional resolution steps will hurt
-the solver's performance. Since resolution cannot reduce a clause's
-$\LBD$, The maximum size reduction from $\IUIP$ is the difference
-between the clause's size and $\LBD$, denote as the clause's gap value
-($\GAP(\oneUIPClause) = |\oneUIPClause| - \LBD(\oneUIPClause)$). For
-an 1-UIP clause with a small $\GAP$, applying $\IUIP$ is unlikely to
-achieve cost effective results. Therefore, we propose a heuristic
-based approach to to enable and disable $\IUIP$ learning based on
-input clause's $\GAP$.
-
-\begin{algorithm}[b]
-\caption{Control-$\IUIP$}\label{alg:enableCondition}
-\begin{algorithmic}[1]
-\Require  $\oneUIPClause$ is a valid 1-UIP clause
-\Require  $ t_{gap} \ge 0$ is a dynamically calculated gap threshold
-\Procedure{Control-$\IUIP$}{$\oneUIPClause,  t_{gap}$} 
-    \State $Gap \gets |\oneUIPClause| - \textit{LBD}(\oneUIPClause)$
-     \If{$Gap > t_{gap}$} \label{ln:compareGap}
-        \State $\iUIPClause \gets \IUIP(\oneUIPClause, \phi)$
-        \State \textcolor{gray}{\textbf{I-UIP-Greedy}($\iUIPClause$ , $\oneUIPClause$)}
-               \Comment{additional clause selection policy, see sec~\ref{sec:greedy}}
-        \If {$|\iUIPClause| < |\oneUIPClause|$} 
-            \State $\text{Succeed} \gets \text{Succeed}+1$ \label{ln:updateS}
-        \EndIf
-        \State $\text{Attempted} \gets \text{Attempted}+1$ \label{ln:updateA}
-     \EndIf
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-Alg.~\ref{alg:enableCondition} compares the input $\oneUIPClause$'s
-$\GAP$ against a floating target threshold $t_{gap}$
-(line~\ref{ln:compareGap}). The threshold $t_{gap}$ represent the
-expected minimal $\GAP$ required for $\oneUIPClause$ to achieve a
-predetermined success rate (80\%) from performing $\IUIP$ learning.
-
-The gap threshold $t_{gap}$ is initialized to 0, and is updated at
-every restart based on $\IUIP$'s success rate from the previous
-restart interval. More specifically, the algorithm collects the
-statistics of the number of $\IUIP$ learning attempted
-(line~\ref{ln:updateA} in alg.~\ref{alg:enableCondition}) and the
-number of attempts succeed (line~\ref{ln:updateS}) for each restart
-interval, and use them to calculate the success rate. If the success
-rate is below 80, the threshold $t_{gap}$ is increased to restrict
-$\IUIP$for the next restart interval. On the other hand, the threshold
-is decreased to encourage more aggressive $\IUIP$ learning for the
-next restart interval.
-
+\subsection{Variants of $\allUip$}
+We also developed and experimented with a few variants of the
+$\allUip$ algorithm which describe below.
+
+\subsubsection{$\allUipAct$: Clauses with Active
+  Variables.} \label{sec:active} $\allUip$ learning might introduce
+literals with low variable activity into $C_i$. Low activity variables
+are variables that have had low recent participation in clause
+learning. Hence, clauses with variables of low activity might not be
+as currently useful to the solver. Our variant $\allUipAct$, shown as
+the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
+computes the average variable activity of the newly produced all-UIP
+clause $C_i$ and the original 1-UIP clause $C_1$. The new clause $C_i$
+will be returned only if it is both smaller and has higher average
+variable activity than the original 1-UIP clause. There are, of
+course, generalizations of this approach where one has a weighted
+tradeoff between these factors that allows preferring the new clause
+when it has large gains in one metric even though it has small losses
+in the other. We did not, however, experiment with such
+generalization.
+
+\subsubsection{Adjust Variable Activity.} \label{sec: varajust} An
+alternative to filtering clauses with low average variable
+($\allUipAct$) is to alter the way variable activities are updated to
+account for our new clause learning method. The popular branching
+heuristics VSIDS \cite{DBLP:conf/dac/MoskewiczMZZM01} and LBR
+\cite{DBLP:conf/sat/LiangGPC16} bump the variable activity for all
+literals appearing in the learnt clause $C_L$ and all literals
+resolved away during the conflict analysis that yielded $C_L$ from the
+initially detected conflict $C_I$ (all literals on the conflict side).
+
+We did not apply this approach to the $\allUip$ clause, as we did not
+want to bump the activity of the literals at above the deepest
+decision level that $\allUip$ resolves away. Intuitively, these
+literals did not directly contribute to generating the
+conflict. Instead, we tried two modifications to the variable activity
+bumping schemes.
+
+Let $C_1$ be the 1-UIP learnt clause and $C_i$ be the $\allUip$ learnt
+clause. First, we kept all of the variable activity bumps normally
+done by 1-UIP learning.\footnote{So extra techniques used by the
+  underlying solver, like reason side rate and locality
+  \cite{DBLP:conf/sat/LiangGPC16}, were kept intact.}  Then, when the
+$\allUip$ scheme was successful, i.e., $C_i$ was to be used as the new
+learnt clause, we perform further updates to the variable
+activities. In the $\allUipIn$ approach all variables variables
+appearing in $C_i$ that are not in $C_1$ have their activities
+bumped. Intuitively, since the clause $C_i$ is being added to the
+clause database we want to increase the activity of all of its
+variables. On the other hand, in the $\allUipEx$ approach in addition
+to bumping the activity of the new variables in $C_i$ we also remove
+the activity bumps of those variables in $C_1$ that are no longer in
+$C_i$.
+
+In sum, the two modified variable activity update schemes we
+experimented with were
 \[
-    t_{gap}=
-\begin{cases}
-    t_{gap} + 1& \text{if } \frac{\text{Succeed}}{\text{Attempted}} < 0.8\\
-    max(t_{gap} - 1, 0),              & \text{otherwise}
-\end{cases}
+    \begin{array}{l}
+    \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}l) \\
+    \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land 
+    \forall l\in C_1 - C_i.\,\mathit{unbumpActivity(l)}
+  \end{array}
 \]
 
-\subsection{Early stop i-UIP Learning}
-At any point of $\IUIPMIN$ learning, if the number of marked literals
-in $\iUIPClause$ (literals which are forced into the clause to
-preserve $\LBD$) exceeds the input clause $\oneUIPClause$'s $\GAP$, we
-can abort $\IUIP$ learning for $\oneUIPClause$ because the size of the
-final $\iUIPClause$ is at least the size of $\oneUIPClause$.  The
-early stopping rule prevents solver from wasting time on traversing a
-large implication graph when $\IUIP$ has already failed.
-
-The early stopping rule may prematurely stops $\IUIPPURE$ learning
-which would otherwise succeed because $\iUIPClause$ can be further
-minimized. However, the experiments indicates such false negative
-cases are rare, and the early stopping rule generally helps the
-learning scheme's performance.
-
-\subsection{Greedy Active Clause Selection} \label{sec:greedy} Even
-though $\IUIP$ learning can reduce the size of the learnt clause
-$\iUIPClause$, but it may introduce literals with low variable
-activity into $\iUIPClause$.  Inactive literals prevents the clause
-from being asserted to force literal implication. Therefore, a
-practical clause learning scheme should consider both size and
-variable activity. We propose an optional extension $\IUIPGreedy$ to
-filter out inactive $\oneUIPClause$.
-
-\begin{algorithm}[t]
-\caption{$\IUIPGreedy$}\label{alg:greedySelection}
-\begin{algorithmic}[1]
-\Require  $\iUIPClause$ is a valid i-UIP clause
-\Require  $\oneUIPClause$ is a valid 1-UIP clause
-\Procedure{$\IUIPGreedy$}{$\iUIPClause, \oneUIPClause$} 
-   \If{$|\iUIPClause| < |\oneUIPClause| \wedge (\code{AvgVarAct}(\iUIPClause) > \code{AvgVarAct}(\oneUIPClause)) $} \label{ln:checkSizeAndActivity}
-    \State \textbf{return} $\iUIPClause$
-\Else
-    \State \textbf{return} $\oneUIPClause$
-\EndIf  \label{ln:cSizeCheck}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-After computing the $\iUIPClause$, $\IUIPGreedy$ compares both the
-size and the average variable activity for $\oneUIPClause$ and
-$\iUIPClause$ (alg.~\ref{alg:greedySelection} at
-line~\ref{ln:checkSizeAndActivity}). The algorithm learns
-$\iUIPClause$ if the clause has smaller size and higher average
-variable activity.
-
-\subsection{Adjust Variable Activity} \label{sec: varajust} Two
-popular branching heuristics for modern SAT solvers are VSIDS and
-LBR. Both heuristics increase the variable activity for all literals
-involved in resolutions during 1-UIP learning. Since $\IUIP$ extends
-1-UIP with deeper resolutions against the trail, the variables
-activities for the fresh literals involved in the additional
-resolution steps need to be adjusted as well. We purpose two optional
-schemes for adjusting variable activities, $\IUIPActive$ and
-$\IUIPDist$.
-
-After learning $\oneUIPClause$ from 1-UIP scheme, $\IUIPActive$
-collects all literals appear in the the $\IUIP$ clause $\iUIPClause$,
-and increase their variable activity uniformly according to the
-current branching heuristic if the literals' variable activity have
-not yet been bumped during 1-UIP. The scheme does not bump variable
-activities for transient literals that are resolved away at
-non-conflicting level because, unlike transient literals at
-conflicting level, these literals cannot be re-asserted after the
-immediate backtracking. Notice that other post-analyze extensions such
-as Reason Side Rate (RSR) and Locality\cite{} can be applied after
-$\IUIPActive$.
-
-\[ \forall l \in  \iUIPClause \cdot NotBumped(var(l)) \implies  bumpActivity(var(l)) \]
-
-$\IUIPDist$ collects literals appear exclusively in $\iUIPClause$ and
-bump their variable activity uniformly. It also find all the literals
-in $\oneUIPClause$ that are resolved away during $\IUIP$, and unbump
-their variable activity if they have been bumped during 1-UIP. The
-unbumped literals are no longer in the learned clause $\iUIPClause$,
-keep their variable activity bumped does not help solver to use
-$\iUIPClause$.
-
-\[ \forall l_{i} \in  \iUIPClause \setminus \oneUIPClause \cdot bumpActivity(var(l_{i})) 
-\]
-\[ \forall l_{1} \in  \oneUIPClause \setminus \iUIPClause \cdot unbumpActivity(var(l_{1})) 
-\]
-
-\subsection{Integrate with Chronological Backtracking}
-In SAT solver with chronological backtracking, literals on the
-assertion trail $\assertionTrail$ are not always sorted by decision
-levels. This change imposes a challenge to $\IUIP$ learning since the
-previous implementation relies on solver's ability to efficiently
-access all literals from any decision level in descending trail order
-(alg.~\ref{alg:i-uip} line~\ref{ln:trailaccess}).
-
-To mitigate this challenge, we modify the solver to track the precise
-trail position of all asserted literals with a single vector. We then
-build a priority queue $lit\_Order$ to manage literal's resolution
-order. The queue $lit\_Order$ prioritizes literals with higher
-decision level, and it favors literal with higher trail position when
-decision levels are tied. The order of $lit\_Order$ represents the
-correct resolution order of $\IUIP$ because 1) an asserted literal's
-decision level is the maximum level among all of its reasoning
-literals, and 2) an asserted literal always appears higher in the
-trail than its reasoning literal.
-
-\begin{algorithm}[t]
-\caption{\IUIP-CB}\label{alg:i-uip-CB}
-\begin{algorithmic}[1]
-\Require  $\oneUIPClause$ is a valid 1-UIP clause
-\Require  $\assertionTrail$ is a valid assertion trail
-\Require  $lit\_Order$ is a priority queue
-\Procedure{\IUIP-CB}{$\oneUIPClause, \assertionTrail, lit\_Order $} 
-    \State ...
-    \State $\iUIPClause \gets \oneUIPClause $ \Comment{initialize i-UIP clause}
-    \State forall $l \in \oneUIPClause \cdot \code{Enqueu}(lit\_Order, l)$  \label{ln:enqueue}
-    \State \code{DecisionLvs} $\gets \Set{\code{level}(l)}{l \in \oneUIPClause}$
-    \While{$lit\_Order \neq \emptyset$} \label{ln:pop}
-        \State $p \gets \code{dequeu}(lit\_Order, l) $ \label{ln:dequeue}
-        \If{$\exists q \in \code{Reason}(\neg{p}, \assertionTrail) \cdot \code{level}(q) \not\in$ \code{DecisionLvs} \\ $\vee p$ is the last reminaing lit in its decision level } \label{ln:resolvable}
-          \State Pass
-        \Else 
-            \State  $\iUIPClause \gets (\oneUIPClause ) \bowtie \code{Reason}(q, \assertionTrail)$
-            \State   forall $l \in \code{Reason}(q, \assertionTrail) \cdot \code{Enqueu}(lit\_Order, l)$ \label{ln:reequeue}
-        \EndIf
-    \EndWhile
-\State ...
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-Alg.~\ref{alg:i-uip-CB} is the pesudo-code implementation of the
-augmented $\IUIP$ for chronological backtracking with the priority
-queue $lit\_Order$. The algorithm first populates $lit\_Order$ with
-all literals in $\oneUIPClause$ (line~\ref{ln:enqueue}), and then
-continuously pops literals until the queue is empty
-(line~\ref{ln:pop}). When a literal is resolved away, all of its
-reasoning literals are added into $lit\_Order$ if they are not already
-in the queue (line~\ref{ln:reequeue}). The algorithm will always
-terminate because a literal cannot entered the queue twice (guaranteed
-by the properties of the trail order) and there are finite amount of
-literals on the trail.
+\subsubsection{Chronological Backtracking.}
+We tested our new clause learning schemes on solvers that utilized
+Chronological Backtracking
+\cite{DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}. When
+chronological backtracking is used the literals on the trail might no
+longer be sorted by decision level. So resolving literals in the
+conflict by highest trail index first no longer works. However, we can
+define a new ordering on the literals to replace the trail index
+ordering. Let $l_1$ and $l_2$ be two literals on the trail
+$\trail$. We say that $l_1 >_{\mathit{chron}} l_2$ if
+$\dlevel{l_1} > \dlevel{l_2} \lor (\dlevel{l_1} = \dlevel{l_2} \land
+\trailIdx{l_1} > \trailIdx{l_2})$. That is, literals with higher
+decision level come first, and if that is equal then the literal with
+higher trail index comes first.
+
+Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19} it can be
+observed that all clause learning schemes continue to work as long
+literals are resolved away from the initial conflict in decreasing
+$>_{\mathit{chron}}$ order. In our implementation we used a heap
+(priority queue) to achieve this ordering of the literal resolutions
+in order to add our new schemes to those solvers using chronological
+backtracking.
 
 \section{Implementation and Experiments}
-
-\subsection{Clause Reduction with $\IUIP$}
-To evaluate $\IUIP$'s effectiveness as a clause reduction technique,
-we implement $\IUIPPURE$ and $\IUIPMIN$ with Control-I-UIP and Early
-Stopping rules in Sec~\ref{sec:i-uip} on top of \text{\MapleBase}
-\cite{}, the winner of SAT Race 2015 application track.  We than
-compare the performance of the baseline $\MapleBase$ with
-$\MapleIUIPPURE$ and $\MapleIUIMIN$ on the full set of benchmarks from
-SAT RACE 2019 main track.
-
-The benchmark contains 400 instances divided into two groups of 200,
-new and old, representing historical instances and fresh instances in
-the 2019 race, respectively. We randomly partition the old group
-instances into six partitions of size 30 and one partition of size
-20. Each partition is then assigned to a XeonE5-2 CPU node with 16
-cores (2 sockets 8 cores and 1 thread) and 96649 MB memory. The new
-group is partitioned based on their contributor (e.g. Heule
-contributed 22 matrix multiplication instances), and each partition is
-assigned to a aforementioned CPU node. To speed up the experiment, we
-allow a CPU node to solver at most seven instances concurrently.
+\subsubsection{Clause Reduction with $\allUip$.}
+We implemented our new clause learning schemes in the
+\text{\MapleBase} \cite{}, the winner of SAT Race 2015 application
+track. We then evaluated our schemes on the full set of benchmarks
+from SAT RACE 2019 main track which contains 400 instances. We ran our
+experiments on \fb{insert GHz of CPU}, allowing \fb{insert time bound}
+seconds per instance and a maximum of \fb{insert RAM bound}.
+
+\fb{Delete figure 2 and use only figure 8 with all of the
+  solvers. First briefly comment on the performance and make a brief
+  observation about clause sizes.
+
+  Then have a new subsubsection that looks at clause sizes wrt. to the
+  two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
+  them side by side (using a tabular environment as I did in figure 1).
+
+  Remove fig 5 and put those numbers in the text instead of in a table
+  as part of your discussion.
+
+  Then have a new subsubsection on DRAT trim and get rid of Fig 7.
+
+  Then a subsubsection on \allUip in modern solvers. Some of our
+  observations have no solid foundation, so check them to make sure
+  that you are just not stating your intuitions without reasonable
+  evidence. We need to to fit into 15 pages total along with a short
+  1/3 page for conclusions.
+
+  Finally add all citations.
+
+  If we add the more complex \allUipMin we also will need more space
+  to describe that, so prune some of your conclusions especially those
+  that are simply conjectures.
+}
 
@@
   the reduction is not cumulative. }  for each instances. For
-$\MapleIUIPPURE$ and $\MapleIUIMIN$, we also captures the $\IUIP$
+$\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
 learning attempted rate and success rate.
@@
 \hline
-$\MapleIUIPPURE$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
-\hline
-$\MapleIUIMIN$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
+$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
+\hline
+$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
 \hline
@@
 \end{center}
-\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPURE$  and $\MapleIUIMIN$ on SAT2019 race main track.
+\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 race main track.
 CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
@@
 
-Fig.~\ref{fig:t1} shows that $\IUIPPURE$ ($\IUIPPURE$) solved seven
+Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
 (five) more instances than the baseline solver with lower PAR-2
-scores. $\IUIPPURE$ has marginally lower PAR-2 score than
-$\IUIPPURE$. Both $\IUIPPURE$ and $\IUIPMIN$ produce clause with
+scores. $\allUipPure$ has marginally lower PAR-2 score than
+$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
 significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
@@
 distribution (PDF) of the relative clause size of both i-uip learning
-schemes ($\IUIPPURE$ in red and $\IUIPMIN$ in green) for each
-instance. $\IUIPPURE$ ($\IUIPMIN$) produces shorter clauses for 77.7\%
+schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
+instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
 (88.5\%) of instances, and average relative reduction from 1-UIP is
 16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
-clause size of $\IUIPMIN$, $\IUIPPURE$ and 1-UIP, and it shows that
+clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
 both i-UIP learning schemes in general produce the smaller clauses,
 and the size reduction is more significant for instances with large
-average 1-UIP clause size. Moreover, $\IUIPMIN$ clauses (indicated in
-green) is consistently smaller than $\IUIPPURE$ clauses.
+average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
+green) is consistently smaller than $\allUipPure$ clauses.
 
@@
       1-uip clauses (calculated as
-      $\dfrac{|\oneUIPClause|-|\iUIPClause|}{|\oneUIPClause|}$ ) for
+      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
       each instance, and Y axis shows the probability density.}
@@
       plot represents an instance. X and Y axis shows the clause
-      length from $\IUIP$ and 1-UIP, respectively. Each green (red)
+      length from $\allUip$ and 1-UIP, respectively. Each green (red)
       dot represents an compared instance between $\MapleBase$ and
-      $\MapleIUIMIN$ ($\IUIPPURE$). }
+      $\MapleIUIMin$ ($\allUipPure$). }
     \label{fig:len_compare}
@@
 \nf{ Do we need this?  We additionally looked at the 14 instances
-  solved by $\IUIPMIN$ but not by 1-UIP. $\IUIPMIN$ produces smaller
+  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
   clauses for all of them with average relative reduction of 22\% and
@@
   relative reduction over 30\%. For the 9 instances solved by 1-UIP
-  but not by $\IUIPMIN$, $\IUIPMIN$ only produce smaller clause for
+  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
   three instances and with average relative reduction of 3.3\%.}
 
-$\IUIPMIN$ outperformed $\IUIPPURE$ in clause size.
-This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.
+$\allUipMin$ outperformed $\allUipPure$ in clause size.
+This results agrees with our observation in Fig.~\ref{fig:t2}: $\allUipMin$ attempted $\allUip$ learning more frequently, and it is more likely to succeed.
 
@@
 \hline
-Solver & $\IUIP$ attempt rate & $\IUIP$ success rate  \\ 
-\hline
-$\MapleIUIPPURE$ & 16.1\% & 43.4\% \\ 
-\hline
-$\MapleIUIMIN$ & 28.8\% & 59.3\% \\ 
+Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
+\hline
+$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
+\hline
+$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
 \hline
@@
 \end{center}
-\caption{Compare $\IUIPPURE$ and $\IUIPMIN$ i-uip attempt rate and success rate.
-$\IUIPMIN$ scheme attempted $\IUIP$ more frequently, and it is more likely to successfully produce smaller $\iUIPClause$ clause .}
+\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate.
+$\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to successfully produce smaller $C_i$ clause .}
 \label{fig:t2}
@@
 Fig.~\ref{fig:t3} shows that the optimized proof construct by
-$\IUIPMIN$ and $\IUIPPURE$ are significantly smaller than 1-UIP
+$\allUipMin$ and $\allUipPure$ are significantly smaller than 1-UIP
 proofs. The relative proof size reduction roughly correlates to the
@@
 \hline
-$\MapleIUIPPURE$ & 487.2 & 6.90\% \\ 
-\hline
-$\MapleIUIMIN$ & 413.2 & 17.18\% \\ 
+$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
+\hline
+$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
 \hline
@@
 \end{center}
-\caption{Optimized UNSAT proof comparison for 1-UIP $\IUIPPURE$ and $\IUIPMIN$.
+\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$.
 Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
@@
     \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
-    \caption{Average optimized proof size between 1-uip and $\IUIPMIN$.}
+    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
     \label{fig:proof_compare}
@@
 \hline
-\nf{$\IUIPPURE$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
-\hline
-$\IUIPMIN$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
-\hline
-$\IUIPGreedy$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
-\hline
-$\IUIPActive$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
-\hline
-$\IUIPDist$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
+\nf{$\allUipPure$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
+\hline
+$\allUipMin$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
+\hline
+$\allUipAct$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
+\hline
+$\allUipIn$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
+\hline
+$\allUipEx$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$,
-$\IUIPActive$, and $\IUIPDist$ on SAT2019 race main track.}
+\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\allUipPure$, $\allUipMin$, $\allUipAct$,
+$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
 \label{fig:t4}
@@
 
-\subsection{$\IUIP$ as a Practical Learning Scheme}
-To evaluate $\IUIP$'s effectiveness as a clause learning scheme, we
-re-implement $\IUIPMIN$ on $\MapleBase$ with the extensions mentioned
+\subsection{$\allUip$ as a Practical Learning Scheme}
+To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
+re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
 in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
-$\IUIP$ configurations ( $\IUIPMIN$, $\IUIPPURE$, $\IUIPGreedy$,
-$\IUIPActive$, and $\IUIPDist$) on the SAT Race 2019 main track
+$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
+$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
 benchmark and reported solved instances, PAR-2 score and average
@@
 Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
-scheme $\IUIPPURE$ solved the most overall instances (228) and the
-most UNSAT instances (93). $\IUIPGreedy$ had the lowest PAR-2
-score. $\IUIPActive$ solved the most SAT instances (138). $\IUIPDist$
+scheme $\allUipPure$ solved the most overall instances (228) and the
+most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
+score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
 produced the shortest average clause size, but solved the second least
 instances, one more instance than the baseline 1-UIP learning. All
-configurations of $\IUIP$ outperformed the baseline 1-UIP scheme in
+configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
 solved instances, PAR-2 score and average clause size.
 
-\subsection{$\IUIP$ on Modern SAT solvers}
-To validate $\IUIP$ as a generalizable learning scheme on modern SAT
-solvers, we re-implement $\IUIP$ on the winners of 2017, 2018 and 2019
+\subsection{$\allUip$ on Modern SAT solvers}
+To validate $\allUip$ as a generalizable learning scheme on modern SAT
+solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
 SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$). $\expSATShort$
@@
 partially sidestep the problem of variable activity. For each solver,
-we compare the base 1-UIP learning scheme against $\IUIPPURE$,
-$\IUIPMIN$ and the top two $\IUIP$ variants, $\IUIPGreedy$ and
-$\IUIPActive$, on the SAT Race 2019 main track benchmark. We report
+we compare the base 1-UIP learning scheme against $\allUipPure$,
+$\allUipMin$ and the top two $\allUip$ variants, $\allUipAct$ and
+$\allUipIn$, on the SAT Race 2019 main track benchmark. We report
 solved instances, PAR-2 score and the average clause size.
@@
 \hline
-$\MapleSeven$-i-greedy & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
+$\MapleSeven$-i-Act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
 \hline
@@
 \hline
-$\MapleEightShort$-i-greedy & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
+$\MapleEightShort$-i-Act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
 \hline
@@
 \hline
-$\MapleNineShort$-i-greedy & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
+$\MapleNineShort$-i-Act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
 \hline
@@
 \hline
-$\expSATShort$-i-greedy & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
+$\expSATShort$-i-Act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of 1-UIP, $\IUIPPURE$. $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ on SAT2019 race main track.}
+\caption{Benchmark results of 1-UIP, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
 \label{fig:t5}
@@
 
-Table~\ref{fig:t5} shows the benchmark result of $\IUIP$
+Table~\ref{fig:t5} shows the benchmark result of $\allUip$
 configurations on different solvers. All four configurations of
-$\IUIP$ outperformed 1-UIP learning for the SAT 2017 race winner,
-$\MapleSeven$. More specifically, $\IUIPPURE$, $\IUIPMIN$,
-$\IUIPGreedy$ and $\IUIPActive$ solved 12, 8, 5 and 2 more instances,
-respectively, whiling producing smaller clauses. $\IUIPPURE$ solved
+$\allUip$ outperformed 1-UIP learning for the SAT 2017 race winner,
+$\MapleSeven$. More specifically, $\allUipPure$, $\allUipMin$,
+$\allUipAct$ and $\allUipIn$ solved 12, 8, 5 and 2 more instances,
+respectively, whiling producing smaller clauses. $\allUipPure$ solved
 more UNSAT and SAT instances while other configurations improved on
-solving SAT instances.  The improvement of $\IUIP$ is more significant
+solving SAT instances.  The improvement of $\allUip$ is more significant
 on $\MapleSeven$ than on $\MapleBase$ for both solved instances and
-clause size reduction. This may suggests that $\IUIP$ and the recent
+clause size reduction. This may suggests that $\allUip$ and the recent
 learnt clause minimization approach~\cite{} synergies well because
@@
 
-We observed significant improvement of $\IUIP$ for the SAT 2018
-winner, $\MapleEightShort$. Three out of four $\IUIP$ configurations
+We observed significant improvement of $\allUip$ for the SAT 2018
+winner, $\MapleEightShort$. Three out of four $\allUip$ configurations
 outperformed 1-UIP by 5, 4 an 4 instances,
@@
 chronological backtracking (CB) for long distance
-backtracks. Therefore, we used $\IUIP$-CB extension for all $\IUIP$
-configurations. The results indicates that the improvement of $\IUIP$
+backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
+configurations. The results indicates that the improvement of $\allUip$
 is slightly shadowed by the adoption of CB. More specifically, we
@@
 an interesting interaction effect that shows the limitations of both
-CB and $\IUIP$ learning for future research.
-
-$\IUIP$ learning schemes showed significant improvement for
+CB and $\allUip$ learning for future research.
+
+$\allUip$ learning schemes showed significant improvement for
 $\MapleNineShort$. Three out of four configurations improved solved
@@
 uses duplicate learning (DL) to prioritize clauses that are learned
-multiple times. $\IUIP$ and DL didn't synergies well possibly because
+multiple times. $\allUip$ and DL didn't synergies well possibly because
 i-UIP clauses are less likely to be duplicated. We observed that
-$\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the
+$\allUipMin$ ,in average, added 12\% less duplicated clauses into the
 core clause database than 1-UIP. One possible explanation is that
-$\IUIP$ learning can reduce an 1-UIP clause to different i-UIP clauses
+$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
 for different assertion trails. Therefore, the solver is less likely
@@
 
-We observed significant improvement of $\IUIP$ for $\expSATShort$,
-three out of four configurations of $\IUIP$ significantly outperformed
+We observed significant improvement of $\allUip$ for $\expSATShort$,
+three out of four configurations of $\allUip$ significantly outperformed
 1-UIP learning by 4, 7 and 8 more instances, respectively, while
 producing smaller clauses. We expect all three configurations of
-$\IUIP$ to solve the similar amount of instances with close PAR-2
+$\allUip$ to solve the similar amount of instances with close PAR-2
 scores because the additional random walk exploration allows the
@@
 impact on the solver's performance. However, we instead, observed that
-both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP
-learning scheme. One possible explanation is that $\IUIPGreedy$ and
-$\IUIPActive$ schemes could easily overcompensate variable activities
+both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
+learning scheme. One possible explanation is that $\allUipAct$ and
+$\allUipIn$ schemes could easily overcompensate variable activities
 for i-UIP clauses, and $\expSATShort$'s random walk exploration could

### 2020-02-18 paper.tex
address comments and pass over evaluation section, ongoing

@@
 \section{Implementation and Experiments}
-\subsubsection{Clause Reduction with $\allUip$.}
-We implemented our new clause learning schemes in the
-\text{\MapleBase} \cite{}, the winner of SAT Race 2015 application
-track. We then evaluated our schemes on the full set of benchmarks
-from SAT RACE 2019 main track which contains 400 instances. We ran our
-experiments on \fb{insert GHz of CPU}, allowing \fb{insert time bound}
-seconds per instance and a maximum of \fb{insert RAM bound}.
-
-\fb{Delete figure 2 and use only figure 8 with all of the
-  solvers. First briefly comment on the performance and make a brief
-  observation about clause sizes.
-
-  Then have a new subsubsection that looks at clause sizes wrt. to the
-  two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
-  them side by side (using a tabular environment as I did in figure 1).
-
-  Remove fig 5 and put those numbers in the text instead of in a table
-  as part of your discussion.
-
-  Then have a new subsubsection on DRAT trim and get rid of Fig 7.
-
-  Then a subsubsection on \allUip in modern solvers. Some of our
-  observations have no solid foundation, so check them to make sure
-  that you are just not stating your intuitions without reasonable
-  evidence. We need to to fit into 15 pages total along with a short
-  1/3 page for conclusions.
-
-  Finally add all citations.
-
-  If we add the more complex \allUipMin we also will need more space
-  to describe that, so prune some of your conclusions especially those
-  that are simply conjectures.
-}
-
-Beside solved instances count and PAR-2 score, we additionally measure
-the average clause length and clause reduction ratio (both cumulative
-and non-cumulative)\footnote{The cumulative reduction ratio is
-  obtained through learning all clauses with the target learning
-  scheme; Therefore, the reduction is cumulative. The non-cumulative
-  reduction ratio is obtained by running the target scheme for
-  measurement only (the minimized 1-UIP clause is learned); Therefore,
-  the reduction is not cumulative. }  for each instances. For
-$\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
-learning attempted rate and success rate.
-
-\begin{figure} 
-\begin{center}
-\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
-\hline
-Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
-\hline
-$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
-\hline
-$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
-\hline
-$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
-\hline
-\end{tabular}
-\end{center}
-\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 race main track.
-CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
-\label{fig:t1}
-\end{figure}
-
-Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
-(five) more instances than the baseline solver with lower PAR-2
-scores. $\allUipPure$ has marginally lower PAR-2 score than
-$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
-significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
-respectively. Fig.~\ref{fig:len_pdf} shows the probability density
-distribution (PDF) of the relative clause size of both i-uip learning
-schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
-instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
-(88.5\%) of instances, and average relative reduction from 1-UIP is
-16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
-clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
-both i-UIP learning schemes in general produce the smaller clauses,
-and the size reduction is more significant for instances with large
-average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
-green) is consistently smaller than $\allUipPure$ clauses.
-
-2\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF}
-    \caption{ Relative clause size reduction distribution. X axis
-      indicates the relative size of difference between i-uip and
-      1-uip clauses (calculated as
-      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
-      each instance, and Y axis shows the probability density.}
-     \label{fig:len_pdf}
-\end{figure}
-\begin{figure} \label{fig:len_compare}
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/i-uip-sizes-compare.pdf}
-    \caption{Average clause size comparison plot. Each point in the
-      plot represents an instance. X and Y axis shows the clause
-      length from $\allUip$ and 1-UIP, respectively. Each green (red)
-      dot represents an compared instance between $\MapleBase$ and
-      $\MapleIUIMin$ ($\allUipPure$). }
-    \label{fig:len_compare}
-\end{figure}
-\nf{ Do we need this?  We additionally looked at the 14 instances
-  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
-  clauses for all of them with average relative reduction of 22\% and
-  maximum 77\% (30 vs 135). Seven out of 14 instances has size
-  relative reduction over 30\%. For the 9 instances solved by 1-UIP
-  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
-  three instances and with average relative reduction of 3.3\%.}
-
-$\allUipMin$ outperformed $\allUipPure$ in clause size.
-This results agrees with our observation in Fig.~\ref{fig:t2}: $\allUipMin$ attempted $\allUip$ learning more frequently, and it is more likely to succeed.
-
-\begin{figure} 
-\begin{center} 
-\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
-\hline
-Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
-\hline
-$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
-\hline
-$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
-\hline
-\end{tabular}
-\end{center}
-\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate.
-$\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to successfully produce smaller $C_i$ clause .}
-\label{fig:t2}
-\end{figure}
-
-A solver produce smaller clauses can construct smaller proofs. For
-UNSAT instances, we measure their DRUP\cite{} proof checking time as
-well as the size of the optimized DRUP proof. We used DART-trim
-\cite{} with 5000 timeout to check and optimize DRUP proofs.
-
-Fig.~\ref{fig:t3} shows that the optimized proof construct by
-$\allUipMin$ and $\allUipPure$ are significantly smaller than 1-UIP
-proofs. The relative proof size reduction roughly correlates to the
-average clause size reduction. Fig.~\ref{fig:proof_compare} shows the
-absolute proof size comparison results.
-
-\begin{figure} 
-\begin{center} 
-\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
-\hline
-Solver & optimized proof size (MB) & relative reduction size  \\ 
-\hline
-$\MapleBase$ & 613.9 & 0  \\ 
-\hline
-$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
-\hline
-$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
-\hline
-\end{tabular}
-\end{center}
-\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$.
-Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
-\label{fig:t3}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
-    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
-    \label{fig:proof_compare}
-\end{figure}
+%\subsubsection{Clause Reduction with $\allUip$.}
+We implemented $\allUip$ learning schemes on
+\text{\MapleBase} \cite{liang2016maple}, the winner of SAT Race 2015 application track.
+We then evaluated these schemes and compare against 1-UIP baseline on the full set of benchmarks from SAT RACE 2019 main track which contains 400 instances.
+We ran our experiments on 2.70 GHz XeonE5-2680 CPUs
+%\fb{insert GHz of CPU}
+, allowing 
+%\fb{insert time bound}
+5000 seconds per instance and a maximum of 12 GB memory.
+%\fb{insert RAM bound}.
+
+%\fb{Delete figure 2 and use only figure 8 with all of the
+%  solvers. First briefly comment on the performance and make a brief observation about clause sizes.
+ %Then have a new subsubsection that looks at clause sizes wrt. to the
+  %two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
+  %them side by side (using a tabular environment as I did in figure 1).
+
+  %Remove fig 5 and put those numbers in the text instead of in a table
+  %as part of your discussion.
+
+  %Then have a new subsubsection on DRAT trim and get rid of Fig 7.
+
+  %Then a subsubsection on \allUip in modern solvers. Some of our
+  %observations have no solid foundation, so check them to make sure
+  %that you are just not stating your intuitions without reasonable
+  %evidence. We need to to fit into 15 pages total along with a short
+  %1/3 page for conclusions.
+
+  %Finally add all citations.
+
+  %If we add the more complex \allUipMin we also will need more space
+  %to describe that, so prune some of your conclusions especially those
+  %that are simply conjectures.
+%}
 
@@
 \hline
-\nf{$\allUipPure$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
-\hline
-$\allUipMin$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
-\hline
-$\allUipAct$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
-\hline
-$\allUipIn$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
-\hline
-$\allUipEx$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
+$\allUipPure$ &\textbf{228} (135, \bf{93}) +7 & 4867.37 & 49.88 \\
+\hline
+$\allUipMin$ & 226 (135, 91) +5 & 4890.67 & 45.2 \\ 
+\hline
+$\allUipAct$ & 226 (135, 91)  +5 & \textbf{4866.94} & 47.7 \\
+\hline
+$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52..12 \\
+\hline
+$\allUipEx$ & 223 (134, 89) +2& 5015.23& \textbf{43.2} \\
 \hline
@@
 
-\subsection{$\allUip$ as a Practical Learning Scheme}
-To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
-re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
-in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
-$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
-$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
-benchmark and reported solved instances, PAR-2 score and average
-clause size.
-
-Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
-scheme $\allUipPure$ solved the most overall instances (228) and the
-most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
-score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
-produced the shortest average clause size, but solved the second least
-instances, one more instance than the baseline 1-UIP learning. All
-configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
-solved instances, PAR-2 score and average clause size.
-
-\subsection{$\allUip$ on Modern SAT solvers}
-To validate $\allUip$ as a generalizable learning scheme on modern SAT
+Fig.~\ref{fig:t4} shows each learning scheme's solved instances count, PAR-2 score, and average clause size\footnote{The average clause size is measured by dividing learnt literals by learnt clauses.}.
+$\allUip$ learning schemes improved solved instances, and PAR-2 scores over 1-UIP.
+More specifically, $\allUipPure$ solved the most instances (+7 over 1-UIP) and the most UNSAT instances (+4);  $\allUipIn$ solved the most SAT instances (+6); and $\allUipAct$ yields the best PAR-2 score (-151 than 1-UIP).
+Besides the performance gain, $\allUip$ schemes yield significantly smaller average clause size.
+
+\subsubsection{Clause Reduction with $\allUip$}
+To precisely measure $\allUip$ clause reduction power, we compare each instance's learnt clause size from $\allUipMin$ and $\allUipPure$ against 1-UIP.
+Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the relative clause size of $\allUip$ learning
+schemes ($\allUipMin$ in green and $\allUipPure$ in red) for each instance.
+$\allUipMin$ ($\allUipPure$ resp.) produces shorter clauses for 88.5\% (77.7\%) of instances, and the average relative reduction rate from 1-UIP is 18.6\% (16.7\%).
+Fig.~\ref{fig:len_compare} compares the absolute clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
+Both $\allUip$ schemes generally yield smaller clauses,
+and the size reduction is more significant for instances with larger 1-UIP clause size.
+
+Horizontally, The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and \ref{fig:len_compare} indicate $\allUipMin$ often achieves higher clause reduction than $\allUipPure$.
+We also observed that $\allUipMin$
+attempted algorithm~\ref{alg:i-uip} more frequently than $\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed (59.3\% vs 43.4\%). This observation agrees with our experiment results.
+
+\begin{figure}[t!]
+{\fss{8pt}{10pt}
+\begin{tabular}[t]{l c l}
+
+  \begin{minipage}[t]{0.5\textwidth}
+   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF}
+   \caption{Relative clause size reduction distribution. X axis
+      indicates the relative size of difference between i-uip and
+      1-uip clauses (calculated as
+      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
+      each instance, and Y axis shows the probability density.}
+       \label{fig:len_pdf}
+  \end{minipage}
+  & &
+  \begin{minipage}[t]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{figures/i-uip-sizes-compare.pdf}
+    \caption{Average clause size comparison plot. Each point in the
+      plot represents an instance. X and Y axis shows the clause
+      length from $\allUip$ and 1-UIP, respectively. Each green (red)
+      dot represents an compared instance between $\MapleBase$ and
+      $\MapleIUIMin$ ($\allUipPure$).}
+      \label{fig:len_compare}
+  \end{minipage}
+
+\end{tabular}
+}
+\end{figure}
+
+%Beside solved instances count and PAR-2 score, we additionally measure
+%the average clause size. 
+%and clause reduction ratio (both cumulative
+%and non-cumulative)\footnote{The cumulative reduction ratio is
+%  obtained through learning all clauses with the target learning
+%  scheme; Therefore, the reduction is cumulative. The %non-cumulative
+%  reduction ratio is obtained by running the target scheme for
+%  measurement only (the minimized 1-UIP clause is learned); Therefore,
+%  the reduction is not cumulative. }  for each instances. 
+%For $\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
+%learning attempted rate and success rate.
+
+%\begin{figure} 
+%\begin{center}
+%\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
+%\hline
+%Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
+%\hline
+%$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
+%\hline
+%$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
+%\hline
+%$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 %race main track.
+%CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, %and the values in the brackets are the non-cumulative reduction ratio.}
+%\label{fig:t1}
+%\end{figure}
+
+%Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
+%(five) more instances than the baseline solver with lower PAR-2
+%scores. $\allUipPure$ has marginally lower PAR-2 score than
+%$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
+%significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
+%respectively. Fig.~\ref{fig:len_pdf} shows the probability density
+%distribution (PDF) of the relative clause size of both i-uip learning
+%schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
+%instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
+%(88.5\%) of instances, and average relative reduction from 1-UIP is
+%16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
+%clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
+%both i-UIP learning schemes in general produce the smaller clauses,
+%and the size reduction is more significant for instances with large
+%average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
+%5green) is consistently smaller than $\allUipPure$ clauses.
+
+%\nf{ Do we need this?  We additionally looked at the 14 instances
+%  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
+%  clauses for all of them with average relative reduction of 22\% and
+%  maximum 77\% (30 vs 135). Seven out of 14 instances has size
+%  relative reduction over 30\%. For the 9 instances solved by 1-UIP
+%  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
+%  three instances and with average relative reduction of 3.3\%.}
+
+%\begin{figure} 
+%\begin{center} 
+%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
+%\hline
+%Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
+%\hline
+%$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
+%\hline
+%$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate.
+%$\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to %successfully produce smaller $C_i$ clause .}
+%\label{fig:t2}
+%\end{figure}
+\subsubsection{Proof Reduction with $\allUip$}
+A learning scheme yields smaller clauses (lemmas) may construct smaller causal proofs. For
+UNSAT instances, we additionally compare the size of the optimized DRAT proof from $\allUipPure$, $\allUipMin$ and 1-UIP schemes.
+We used DART-trim \cite{wetzler2014drat} with 10000 second timeout to check and optimize every DRAT proof once \footnote{Applying DART-trim multiple times can further reduce the proof size until fix-point, but the full optimization is time consuming and impractical for our experiments.}.
+
+The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$ are 413.2MB and 487.2MB, respectively.
+Both sizes are significantly smaller than the average optimized proof size from 1-UIP, 613.9MB.
+The average relative proof size reduction per instance of $\allUipMin$ and $\allUipPure$ are 17.18\% and 6.9\% compare to 1-UIP, which roughly correlate to the observation from relative clause reduction plot in Fig~\ref{fig:len_pdf}.
+Moreover, comparing the absolute proof size between $\allUipMin$ and 1-UIP yields a similar trend as clause size comparison in Fig~\ref{fig:len_compare}.
+
+%Fig.~\ref{fig:proof_compare} shows the
+%absolute proof size comparison between $\allUipMin$ and 1-UIP, and displays a similar trend %as the clause size comparison plot in Fig.~\ref{fig:len_compare}.
+
+%\begin{figure} 
+%\begin{center} 
+%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
+%\hline
+%Solver & optimized proof size (MB) & relative reduction size  \\ 
+%\hline
+%$\MapleBase$ & 613.9 & 0  \\ 
+%\hline
+%$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
+%\hline
+%$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$.
+%Optimized proof size measures the average absolute proof size in MB, and relative reduction %size measures the average relative reduction for all UNSAT instances.}
+%\label{fig:t3}
+%\end{figure}
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
+%    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
+%    \label{fig:proof_compare}
+%\end{figure}
+
+%\subsection{$\allUip$ as a Practical Learning Scheme}
+%To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
+%re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
+%in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
+%$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
+%$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
+%benchmark and reported solved instances, PAR-2 score and average
+%clause size.
+
+%Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
+%scheme $\allUipPure$ solved the most overall instances (228) and the
+%most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
+%score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
+%produced the shortest average clause size, but solved the second least
+%instances, one more instance than the baseline 1-UIP learning. All
+%configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
+%solved instances, PAR-2 score and average clause size.
+
+\subsubsection{$\allUip$ on Modern SAT solvers}
+To validate $\allUip$ as practical learning schemes on modern SAT
 solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
-SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$). $\expSATShort$
+SAT Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT} and $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$). $\expSATShort$
 is a top ten solver from 2019 SAT race which uses random walk
@@
 a top solver in the 2019 SAT Race without using chronological
-backtracking; 2) the combination of random walk simulation and
-variable activity branching heuristic allows our learning schemes to
-partially sidestep the problem of variable activity. For each solver,
+backtracking; 2) the random walk simulation 
+branching heuristic is different from local branching heuristic (VSIDS and LRB) that we have considered,
+hence it is in the blind spot of $\allUip$.
 we compare the base 1-UIP learning scheme against $\allUipPure$,
@@
 Table~\ref{fig:t5} shows the benchmark result of $\allUip$
-configurations on different solvers. All four configurations of
-$\allUip$ outperformed 1-UIP learning for the SAT 2017 race winner,
-$\MapleSeven$. More specifically, $\allUipPure$, $\allUipMin$,
-$\allUipAct$ and $\allUipIn$ solved 12, 8, 5 and 2 more instances,
-respectively, whiling producing smaller clauses. $\allUipPure$ solved
+configurations on modern solvers.
+Overall, we observed similar performance gain on all modern solvers as we have seen on $\MapleBase$ in Fig.~\ref{fig:t4}.
+More specifically, almost all configurations improved on solved instance (+4.5 instances in average) and PAR-2 score (-83.6 in average).
+The average clause size reduction is consistent across all solvers.
+Each configuration also exhibits different strength: $\allUipPure$ solved the most instances with the best PAR-2 score on two solvers, $\allUipMin$ yields small clauses, $\allUipIn$ solved the most SAT instances, and $\allUipAct$ has stable performance.
+
+On the SAT 2017 race winner $\MapleSeven$, all four configurations of
+$\allUip$ outperformed 1-UIP learning by 12, 8, 5, and 2 instances, respectively. $\allUipPure$ solved
 more UNSAT and SAT instances while other configurations improved on
-solving SAT instances.  The improvement of $\allUip$ is more significant
-on $\MapleSeven$ than on $\MapleBase$ for both solved instances and
-clause size reduction. This may suggests that $\allUip$ and the recent
-learnt clause minimization approach~\cite{} synergies well because
-i-UIP clauses are shorter with more common literals which allows
-vivification~\cite{} to prune literals more aggressively through unit
-propagation.
-
-We observed significant improvement of $\allUip$ for the SAT 2018
-winner, $\MapleEightShort$. Three out of four $\allUip$ configurations
-outperformed 1-UIP by 5, 4 an 4 instances,
-respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using
-chronological backtracking (CB) for long distance
-backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
-configurations. The results indicates that the improvement of $\allUip$
-is slightly shadowed by the adoption of CB. More specifically, we
-believe CB prevents decision levels from being compressed through the
-process of backtracking and literal assertion. The shorter i-UIP
-clauses can bring related literals closer, which in turns compresses
-the assertion stack and the decision levels. However, since CB
-discourages long distance backtracking, the effect of learning shorter
-clauses is shadowed until a full restart.  We believe we have observed
-an interesting interaction effect that shows the limitations of both
-CB and $\allUip$ learning for future research.
-
-$\allUip$ learning schemes showed significant improvement for
-$\MapleNineShort$. Three out of four configurations improved solved
-instances by 6, 5 and 5 instances, respectively. $\MapleNineShort$
-uses duplicate learning (DL) to prioritize clauses that are learned
-multiple times. $\allUip$ and DL didn't synergies well possibly because
-i-UIP clauses are less likely to be duplicated. We observed that
-$\allUipMin$ ,in average, added 12\% less duplicated clauses into the
-core clause database than 1-UIP. One possible explanation is that
-$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
-for different assertion trails. Therefore, the solver is less likely
-to learn the same i-UIP clauses.
-
-We observed significant improvement of $\allUip$ for $\expSATShort$,
-three out of four configurations of $\allUip$ significantly outperformed
-1-UIP learning by 4, 7 and 8 more instances, respectively, while
-producing smaller clauses. We expect all three configurations of
-$\allUip$ to solve the similar amount of instances with close PAR-2
-scores because the additional random walk exploration allows the
-learning schemes to partially sidestep the activity problem; hence,
-the learning adjustment for variable activities should have less
-impact on the solver's performance. However, we instead, observed that
-both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
-learning scheme. One possible explanation is that $\allUipAct$ and
-$\allUipIn$ schemes could easily overcompensate variable activities
-for i-UIP clauses, and $\expSATShort$'s random walk exploration could
-use future search information to mitigate the negative effects of our
-overcompensation.
+solving SAT instances. The clause size reduction of $\allUip$ is more significant on $\MapleSeven$ compare to $\MapleBase$. 
+%This may suggests that $\allUip$ and the recent
+%learnt clause minimization approach~\cite{} synergies well because
+%i-UIP clauses are shorter with more common literals which allows
+%vivification~\cite{} to prune literals more aggressively through unit
+%propagation.
+
+The SAT 2018 race winner$\MapleEightShort$ uses chronological backtracking (CB) for long distance backtracks.
+Therefore, we applied $\allUip$-CB extension for all $\allUip$ configurations.
+Three out of four $\allUip$ configurations outperformed 1-UIP by 5, 4 an 4 instances,
+respectively. 
+%$\MapleEightShort$ improved from $\MapleSeven$ by using
+%chronological backtracking (CB) for long distance
+%backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
+%configurations.
+
+%The results indicates that the improvement of $\allUip$
+%is slightly shadowed by the adoption of CB. More specifically, we
+%believe CB prevents decision levels from being compressed through the
+%process of backtracking and literal assertion. The shorter i-UIP
+%clauses can bring related literals closer, which in turns compresses
+%the assertion stack and the decision levels. However, since CB
+%discourages long distance backtracking, the effect of learning shorter
+%clauses is shadowed until a full restart.  We believe we have observed
+%an interesting interaction effect that shows the limitations of both
+%CB and $\allUip$ learning for future research.
+
+On the SAT 2019 race winner $\MapleNineShort$,  three out of four$\allUip$ configurations solved instances 6, 5 and 5 more instances than 1-UIP, respectively.
+The solver uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
+We observed that $\allUip$ configurations
+i-UIP clauses are less likely to be duplicated. As an example, $\allUipMin$ in average, added 12\% less duplicated clauses into the
+core clause database than 1-UIP. This observation is surprising and the cause is unclear.
+
+%One possible explanation is that
+%$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
+%for different assertion trails. Therefore, the solver is less likely
+%to learn the same i-UIP clauses.
+
+On $\expSATShort$,
+three out of four $\allUip$ configurations solved 4, 7 and 8 more instances
+than 1-UIP learning.
+We noticed that $\allUipAct$ and $\allUipIn$ show significant better performance than $\allUipMin$ and $\allUipPure$ on this solver.
+The random walk simulation branching heuristic however didn't impact the performance of $\allUip$ schemes significantly.
+%$\allUip$ to solve the similar amount of instances with close PAR-2
+%scores because the additional random walk exploration allows the
+%learning schemes to partially sidestep the activity problem; hence,
+%the learning adjustment for variable activities should have less
+%impact on the solver's performance. However, we instead, observed that
+%both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
+%learning scheme. One possible explanation is that $\allUipAct$ and
+%$\allUipIn$ schemes could easily overcompensate variable activities
+%for i-UIP clauses, and $\expSATShort$'s random walk exploration could
+%use future search information to mitigate the negative effects of our
+%overcompensation.
 

### 2020-02-18 paper.tex
fix new figure link
add cites info
fix typos

@@
 schemes ($\allUipMin$ in green and $\allUipPure$ in red) for each instance.
-$\allUipMin$ ($\allUipPure$ resp.) produces shorter clauses for 88.5\% (77.7\%) of instances, and the average relative reduction rate from 1-UIP is 18.6\% (16.7\%).
+$\allUipMin$ ($\allUipPure$ resp.) produces shorter clauses for 88.5\% (77.7\%) of instances, and the average relative reduction ratio over 1-UIP is 18.6\% (16.7\%).
 Fig.~\ref{fig:len_compare} compares the absolute clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
 Both $\allUip$ schemes generally yield smaller clauses,
-and the size reduction is more significant for instances with larger 1-UIP clause size.
+and the size reduction is more significant for instances with larger 1-UIP clause.
 
@@
   \begin{minipage}[t]{0.5\textwidth}
-   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF}
+   \includegraphics[width=\textwidth]{figures/clause_reduction_PDF.pdf}
    \caption{Relative clause size reduction distribution. X axis
@@
   \begin{minipage}[t]{0.5\textwidth}
-    \includegraphics[width=\textwidth]{figures/i-uip-sizes-compare.pdf}
+    \includegraphics[width=\textwidth]{figures/clause_size_compare.pdf}
     \caption{Average clause size comparison plot. Each point in the
@@
 The solver uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
-We observed that $\allUip$ configurations
-i-UIP clauses are less likely to be duplicated. As an example, $\allUipMin$ in average, added 12\% less duplicated clauses into the
-core clause database than 1-UIP. This observation is surprising and the cause is unclear.
+We observed that clauses learnt from $\allUip$ configurations are less likely to be duplicated.
+As an example, $\allUipMin$ in average, added 12\% less duplicated clauses into the
+core clause database than 1-UIP. This observation is surprising, and the cause is unclear.
 

### 2020-02-21 paper.tex
added abstract

@@
 %\PassOptionsToPackage{hypnens}{url}
-\title{Clause size reduction with all-UIP Learning}
-\author{Nick Feng \and Fahiem Bacchus}
-\authorrunning{N. Feng and F. Bacchus}
-\institute{Department of Computer Science, University of Toronto, Canada, 
-  \email{\{nfeng,fbacchus\}@cs.toronto.edu}}
 
@@
 
+\title{Clause size reduction with all-UIP Learning}
+\author{Nick Feng \and Fahiem Bacchus}
+\authorrunning{N. Feng and F. Bacchus}
+\institute{Department of Computer Science, University of Toronto, Canada, 
+  \email{\{nfeng,fbacchus\}@cs.toronto.edu}}
+
 \begin{document}
@@
 \begin{abstract}
+    Almost all CDCL \sat solvers use the 1-UIP clause learning scheme
+    to learn new clauses from conflicts, and our current understanding
+    of \sat solving provides good reasons for using that scheme. In
+    particular, the 1-UIP scheme yields asserting clauses, and these
+    asserting clauses have minimum LBD among all possible asserting
+    clauses. As a result of these advantages other clause learning
+    schemes, like $i$-UIP and all-UIP, that had been proposed in early
+    work are not used in modern solvers. In this paper, we propose a
+    new technique for exploiting the all-UIP scheme in clause
+    learning. Our technique is to employ all-UIP learning under the
+    constraint that the clause's LBD does not increase (over the
+    minimum established by the 1-UIP clause). Our method can generate
+    clauses that are significantly smaller that the 1-UIP clause while
+    not increasing the LBD. Unlike previous clause minimization
+    methods, our technique is not limited to learning a sub-clause of
+    the 1-UIP clause. It can instead learn a completely different
+    clause. We show empirically that our method can improve the
+    performance of state of the art solvers.
 \end{abstract}

### 2020-02-22 paper.tex
slight change abstract

--- 2020-02-18 paper.tex
@@
 
+\title{Clause size reduction with all-UIP Learning}
+\author{Nick Feng \and Fahiem Bacchus}
+\authorrunning{N. Feng and F. Bacchus}
+\institute{Department of Computer Science, University of Toronto, Canada, 
+  \email{\{fengnick,fbacchus\}@cs.toronto.edu}}
+
 \begin{document}
@@
 \begin{abstract}
+    Clause learning is an essential technique in \sat solvers.
+    Almost all modern CDCL \sat solvers use the 1-UIP 
+    clause learning scheme to learn new clauses from conflicts, 
+    and our current understanding
+    of \sat solving provides good reasons for using that scheme. In
+    particular, the 1-UIP scheme yields asserting clauses, and these
+    asserting clauses have minimum LBD among all possible asserting
+    clauses. As a result of these advantages, other clause learning
+    schemes, like $i$-UIP and all-UIP, that had been proposed in prior
+    work are not used in modern solvers. In this paper, we propose a
+    new technique for exploiting the all-UIP clause learning scheme. 
+    Our technique $\allUip$ employs all-UIP learning under the
+    constraint that the learnt clause's LBD does not increase (over the
+    minimum established by the 1-UIP clause). $\allUip$ can generate
+    clauses that are significantly smaller than the 1-UIP clause while
+    preserving the minimum LBD. Unlike previous clause minimization
+    methods, our technique is not limited to learning a sub-clause of
+    the 1-UIP clause. It can instead learn a completely different
+    clause. We show empirically that $\allUip$ can improve the
+    performance of state of the art solvers.
 \end{abstract}

### 2020-02-22 paper.tex
small change

@@
     Clause learning is an essential technique in \sat solvers.
-    Almost all modern CDCL \sat solvers use the 1-UIP 
-    clause learning scheme to learn new clauses from conflicts, 
+    Most of modern CDCL \sat solvers use the 1-UIP 
+    clause learning scheme for learning new clauses from conflicts, 
     and our current understanding
@@
     constraint that the learnt clause's LBD does not increase (over the
-    minimum established by the 1-UIP clause). $\allUip$ can generate
+    minimum established by the 1-UIP clause). $\allUip$ can learn
     clauses that are significantly smaller than the 1-UIP clause while
@@
     the 1-UIP clause. It can instead learn a completely different
-    clause. We show empirically that $\allUip$ can improve the
-    performance of state of the art solvers.
+    clause with the same asserting literal. We show empirically 
+    that $\allUip$ can improve the performance of state of the art solvers.
 \end{abstract}

### 2020-02-22 paper.tex
revised the abstract

--- 2020-02-22 paper.tex
@@
 \begin{abstract}
-    Clause learning is an essential technique in \sat solvers.
-    Almost all modern CDCL \sat solvers use the 1-UIP 
-    clause learning scheme to learn new clauses from conflicts, 
+    Almost all CDCL \sat solvers use the 1-UIP 
+    clause learning scheme for learning new clauses from conflicts, 
     and our current understanding
@@
     clauses. As a result of these advantages, other clause learning
-    schemes, like $i$-UIP and all-UIP, that had been proposed in prior
+    schemes, like $i$-UIP and all-UIP, that were proposed in early
     work are not used in modern solvers. In this paper, we propose a
     new technique for exploiting the all-UIP clause learning scheme. 
-    Our technique $\allUip$ employs all-UIP learning under the
+    Our technique is to employ all-UIP learning under the
     constraint that the learnt clause's LBD does not increase (over the
-    minimum established by the 1-UIP clause). $\allUip$ can generate
+    minimum established by the 1-UIP clause). Our method can learn
     clauses that are significantly smaller than the 1-UIP clause while
@@
     the 1-UIP clause. It can instead learn a completely different
-    clause. We show empirically that $\allUip$ can improve the
-    performance of state of the art solvers.
+    clause. We show empirically 
+    that our method can improve the performance of state of the art solvers.
 \end{abstract}

### 2020-02-24 paper.tex
fix bib file

@@
 UNSAT instances, we additionally compare the size of the optimized DRAT proof from $\allUipPure$, $\allUipMin$ and 1-UIP schemes.
-We used DART-trim \cite{wetzler2014drat} with 10000 second timeout to check and optimize every DRAT proof once \footnote{Applying DART-trim multiple times can further reduce the proof size until fix-point, but the full optimization is time consuming and impractical for our experiments.}.
+We used DART-trim \cite{DBLP:conf/sat/WetzlerHH14} with 10000 second timeout to check and optimize every DRAT proof once \footnote{Applying DART-trim multiple times can further reduce the proof size until fix-point, but the full optimization is time consuming and impractical for our experiments.}.
 
@@
 solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
-SAT Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT} and $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$). $\expSATShort$
+SAT Race\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT} and $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$). $\expSATShort$
 is a top ten solver from 2019 SAT race which uses random walk

### 2020-02-24 paper.tex
some typos

--- 2020-02-22 paper.tex
@@
 advantages and disadvantages of these alternate schemes. In this paper
-we reexamine some of these previous clause learning schemes, with a
+we reexamine these previous clause learning schemes, with a
 focus on the schemes described in
@@
 Our method is orthogonal to clause minimization. In particular, our
-approach can learn a clause that is not a subset of 1-UIP clause but
+approach can learn a clause that is not a subset of the 1-UIP clause but
 which still serves all of the same purposes as the 1-UIP clause.
@@
 clause learning as typically used in CDCL \sat solvers. A
-propositional formula $F$ expressed in Conjunctive Normal Form (CNF),
+propositional formula $F$ expressed in Conjunctive Normal Form (CNF)
 contains a set of variables $V$. A literal is a variable $v\in V$ or
@@
 $(C_1 \cup C_2) - \{\ell, \lnot \ell\}$. The resolvant will be a
-tautology (i.e., a clause containing a literal $l$ and its negation
-$\lnot l$) if $C_1$ and $C_2$ contain more than one pair of
+tautology (i.e., a clause containing a literal $x$ and its negation
+$\lnot x$) if $C_1$ and $C_2$ contain more than one pair of
 conflicting literals.
@@
 Each literal $\ell\in\trail$ also has a clausal reason
-$\reason{\ell}$. If $\ell$ is a unit propgated literal,
+$\reason{\ell}$. If $\ell$ is a unit propagated literal,
 $\reason{\ell}$ is a clause of the formula such that
 $\ell \in \reason{\ell}$ and
-$\forall x \in \reason{\ell}.\, x\neq \ell \implies (\lnot x \in
-\trail \land \trailIdx{\lnot x} < \trailIdx{\ell}$. That is,
+$\forall x \in \reason{\ell}.\, x\neq \ell \implies \bigl(\lnot x \in
+\trail \land \trailIdx{\lnot x} < \trailIdx{\ell}\bigr)$. That is,
 $\reason{\ell}$ is a clause that has become unit implying $\ell$ due
@@
 \emph{remains intact during clause learning} and is only changed after
-the new clause is learnt. After the new clause is learnt the $\trail$
+the new clause is learnt. After the new clause is learnt $\trail$
 will be changed by backtracking.
@@
 propagated literal whose negation appears in the conflict; and (3) any
-variable in the conflict that is unit progagated in $\trail$ can be
+variable in the conflict that is unit propagated in $\trail$ can be
 resolved upon (the variable must appear in different polarities in the
@@
     resolution is \textbf{ordered} if the sequence of variables $v_1$,
-    \dots, $v_m$ resolved have strictly decreasing trail indicies:
+    \dots, $v_m$ resolved have strictly decreasing trail indices:
     $\trailIdx{v_{i+1}} < \trailIdx{v_i}$ ($1\leq i < m$). (Note that
@@
 the desired properties. Since no reason clause contains literals with
-higher trail indicies, $O$ must be a valid trail resolution if $U$
+higher trail indices, $O$ must be a valid trail resolution if $U$
 was, and furthermore $O$ yields the clause
@@
 $\forall x\in C_L. \dlevel{x} \leq \dlevel{\ell} \land (\dlevel{x} =
-\ell \implies x = \ell$). All of the clause learning schemes in
+\dlevel{\ell} \implies x = \ell$). All of the clause learning schemes in
 Figure~\ref{fig:cl_schemes} produced learnt clauses that are asserting.
@@
   that it is a duplicate of another clause already in the formula}. In
-particular, $C_L$ contains at least two literals at the deepest level
+particular, $C_L$ contains at least two literals at the deepest level.
 If two of its literals at the deepest level were were not completely
 unit propagated (unit propagation is aborted as soon as a conflict is
-found), then $C_L$ could already be in the formula: it's two watches
+found), then $C_L$ could already be in the formula: its two watches
 might not be fully propagated and $C_L$ could be a falsified clause
@@
 compelling reasons for using this scheme. Hence, it is not surprising
-that modern \sat sovlers almost exclusively use 1-UIP clause
+that modern \sat solvers almost exclusively use 1-UIP clause
 learning.\footnote{Knuth in his sat13 CDCL solver \cite{Knuth:Sat13}
@@
 Although learning clauses with low LBD has been shown empirically to
-be important in \sat solving than learning short clauses
+be more important in \sat solving than learning short clauses
 \cite{DBLP:conf/ijcai/AudemardS09}, clause size is still
@@
 clauses contain at most one literal per decision level. However, the
-all-UIP clause can have much higher LBD. Since, $\LBD$ is important
+all-UIP clause can have much higher LBD. Since, $\LBD$ is more important
 than size our approach is to use all-UIP learning when, and only when,
@@
 clause. That is, our algorithm will start with the clause that most
-\sat solvers learn from a conflict (a minimized 1-UIP clause). Our
+\sat solvers learn from a conflict, a minimized 1-UIP clause. Our
 approach is specified in Alg. ~\ref{alg:i-uip}.
@@
 skipped. The subroutine \tryuiplevel is used to reduce each level. It
-uses trail resoluions to achieve this, subject to the constraint that
+uses trail resolutions to achieve this, subject to the constraint that
 no new decision levels can be introduced in the clause. In particular,
@@
   remaining literal $u\in L_i$ is at a lower trail index than all of
-  the other literals, so there is no point in trying to resolve away
+  the other literals  there is no point in trying to resolve away
   $u$---either it will be the decision clause for level $i$ having no
@@
 be increased after processing it with \allUipPure. \allUipMin can,
-however, potentially increase the number of literals at a level. In
-resolving away a ltieral $l$ at level $i$ more literals might be
+however, potentially increase the number of literals at level $i$. In
+resolving away a literal $l$ at level $i$ more literals might be
 introduced into level $i$, and some of these might not be removable by
@@
 steps can hurt the solver's performance. Since resolution cannot
-reduce a clause's $\LBD$, The maximum size reduction obtainable from
+reduce a clause's $\LBD$, the maximum size reduction obtainable from
 $\allUip$ is the difference between the 1-UIP clause's size and its
@@
 (line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
-rate is less than 80\%, we decrease $\tgap$ by one (not allowing it to
-become negative), and if it is greater than 80\% we increase $\tgap$
+rate is greater  than 80\%, we decrease $\tgap$ by one (not allowing it to
+become negative), and if it is less than 80\% we increase $\tgap$
 by one.
@@
   so $\allUipMin$ does not do any further resolutions at this level. Hence,
-  $\allUipPure$ yields $C^*$ unchanged, while $\allUipMin$ yeilds
+  $\allUipPure$ yields $C^*$ unchanged, while $\allUipMin$ yields
   $C^*_{\mathit{min}} = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$,
@@
 We also developed and experimented with a few variants of the
-$\allUip$ algorithm which describe below.
+$\allUip$ algorithm which we describe below.
 
@@
 in the other. We did not, however, experiment with such
-generalization.
+generalizations.
 
@@
 We did not apply this approach to the $\allUip$ clause, as we did not
-want to bump the activity of the literals at above the deepest
-decision level that $\allUip$ resolves away. Intuitively, these
-literals did not directly contribute to generating the
-conflict. Instead, we tried two modifications to the variable activity
-bumping schemes.
+want to bump the activity of the literals above the deepest decision
+level that $\allUip$ resolves away. Intuitively, these literals did
+not directly contribute to generating the conflict. Instead, we tried
+two modifications to the variable activity bumping schemes.
 
@@
     \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}l) \\
-    \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land 
-    \forall l\in C_1 - C_i.\,\mathit{unbumpActivity(l)}
+    \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land\mbox{} \\
+    \qquad \forall l\in C_1 - C_i.\,\mathit{unbumpActivity(l)}
   \end{array}
@@
 Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19} it can be
-observed that all clause learning schemes continue to work as long
+observed that all clause learning schemes continue to work as long as
 literals are resolved away from the initial conflict in decreasing
@@
 %\subsubsection{Clause Reduction with $\allUip$.}
-We implemented $\allUip$ learning schemes on
-\text{\MapleBase} \cite{liang2016maple}, the winner of SAT Race 2015 application track.
-We then evaluated these schemes and compare against 1-UIP baseline on the full set of benchmarks from SAT RACE 2019 main track which contains 400 instances.
-We ran our experiments on 2.70 GHz XeonE5-2680 CPUs
-%\fb{insert GHz of CPU}
-, allowing 
+We implemented $\allUip$ learning schemes on \text{\MapleBase}
+\cite{liang2016maple}, the winner of SAT Race 2015 application
+track. We then evaluated these schemes and compare against the 1-UIP
+baseline on the full set of benchmarks from SAT RACE 2019 main track
+which contains 400 instances. We ran our experiments on 2.70 GHz
+XeonE5-2680 CPUs, allowing
 %\fb{insert time bound}
@@
 \hline
-$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52..12 \\
+$\allUipIn$ & 225 (\textbf{138}, 87) +4 & 4958.49 & 52.12 \\
 \hline
@@
 
-Fig.~\ref{fig:t4} shows each learning scheme's solved instances count, PAR-2 score, and average clause size\footnote{The average clause size is measured by dividing learnt literals by learnt clauses.}.
-$\allUip$ learning schemes improved solved instances, and PAR-2 scores over 1-UIP.
-More specifically, $\allUipPure$ solved the most instances (+7 over 1-UIP) and the most UNSAT instances (+4);  $\allUipIn$ solved the most SAT instances (+6); and $\allUipAct$ yields the best PAR-2 score (-151 than 1-UIP).
-Besides the performance gain, $\allUip$ schemes yield significantly smaller average clause size.
-
-\subsubsection{Clause Reduction with $\allUip$}
-To precisely measure $\allUip$ clause reduction power, we compare each instance's learnt clause size from $\allUipMin$ and $\allUipPure$ against 1-UIP.
-Fig.~\ref{fig:len_pdf} shows the probability density distribution (PDF) of the relative clause size of $\allUip$ learning
-schemes ($\allUipMin$ in green and $\allUipPure$ in red) for each instance.
-$\allUipMin$ ($\allUipPure$ resp.) produces shorter clauses for 88.5\% (77.7\%) of instances, and the average relative reduction ratio over 1-UIP is 18.6\% (16.7\%).
-Fig.~\ref{fig:len_compare} compares the absolute clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
-Both $\allUip$ schemes generally yield smaller clauses,
-and the size reduction is more significant for instances with larger 1-UIP clause.
-
-Horizontally, The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and \ref{fig:len_compare} indicate $\allUipMin$ often achieves higher clause reduction than $\allUipPure$.
-We also observed that $\allUipMin$
-attempted algorithm~\ref{alg:i-uip} more frequently than $\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed (59.3\% vs 43.4\%). This observation agrees with our experiment results.
+Fig.~\ref{fig:t4} shows each learning scheme's solved instances count,
+PAR-2 score, and average learnt clause size. We found that the
+$\allUip$ learning schemes improved solved instances, PAR-2 scores,
+and learnt clause size over 1-UIP. More specifically, $\allUipPure$
+solved the most instances (+7 over 1-UIP) and the most UNSAT instances
+(+4); $\allUipIn$ solved the most SAT instances (+6); and $\allUipAct$
+yields the best PAR-2 score (-151 than 1-UIP). In all cases the
+$\allUip$ schemes learnt significantly smaller clauses on average.
+
+\subsubsection{Clause Reduction with $\allUip$.}
+To precisely measure $\allUip$'s clause reduction power, we compare
+each instance's learnt clause size from $\allUipMin$ and $\allUipPure$
+against 1-UIP. Fig.~\ref{fig:len_pdf} shows the probability density
+distribution (PDF) of the relative clause size of the $\allUip$
+learning schemes ($\allUipMin$ in green and $\allUipPure$ in red) for
+each instance. $\allUipMin$ ($\allUipPure$ resp.) produces shorter
+clauses for 88.5\% (77.7\%) of instances, and the average relative
+reduction ratio over 1-UIP is 18.6\%
+(16.7\%). Fig.~\ref{fig:len_compare} compares the average learnt
+clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
+Both $\allUip$ schemes generally yield smaller clauses, and the size
+reduction is more significant for instances with larger 1-UIP clauses.
+
+The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and
+\ref{fig:len_compare} indicate $\allUipMin$ often achieves higher
+clause reduction than $\allUipPure$. We also observed that
+$\allUipMin$ attempted algorithm~\ref{alg:i-uip} more frequently than
+$\allUipPure$ (28.8\% vs 16.1\%), and is more likely to succeed
+(59.3\% vs 43.4\%). This observation agrees with our experiment
+results.
 
@@
    \includegraphics[width=\textwidth]{figures/clause_reduction_PDF.pdf}
-   \caption{Relative clause size reduction distribution. X axis
-      indicates the relative size of difference between i-uip and
+   \caption{Relative clause size reduction distribution. The $X$ axis
+      indicates the relative size of difference between all-uip and
       1-uip clauses (calculated as
       $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
-      each instance, and Y axis shows the probability density.}
+      each instance, and the $Y$ axis shows the probability density.}
        \label{fig:len_pdf}
@@
     \caption{Average clause size comparison plot. Each point in the
-      plot represents an instance. X and Y axis shows the clause
+      plot represents an instance. The $X$ and $Y$ axes shows the clause
       length from $\allUip$ and 1-UIP, respectively. Each green (red)
@@
 %\end{figure}
-\subsubsection{Proof Reduction with $\allUip$}
-A learning scheme yields smaller clauses (lemmas) may construct smaller causal proofs. For
-UNSAT instances, we additionally compare the size of the optimized DRAT proof from $\allUipPure$, $\allUipMin$ and 1-UIP schemes.
-We used DART-trim \cite{wetzler2014drat} with 10000 second timeout to check and optimize every DRAT proof once \footnote{Applying DART-trim multiple times can further reduce the proof size until fix-point, but the full optimization is time consuming and impractical for our experiments.}.
-
-The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$ are 413.2MB and 487.2MB, respectively.
-Both sizes are significantly smaller than the average optimized proof size from 1-UIP, 613.9MB.
-The average relative proof size reduction per instance of $\allUipMin$ and $\allUipPure$ are 17.18\% and 6.9\% compare to 1-UIP, which roughly correlate to the observation from relative clause reduction plot in Fig~\ref{fig:len_pdf}.
-Moreover, comparing the absolute proof size between $\allUipMin$ and 1-UIP yields a similar trend as clause size comparison in Fig~\ref{fig:len_compare}.
+\subsubsection{Reduced Proof sizes with $\allUip$.}
+A learning scheme that yields smaller clauses (lemmas) might also
+construct smaller causal proofs. For UNSAT instances, we additionally
+compare the size of the optimized DRAT proof from $\allUipPure$,
+$\allUipMin$ and 1-UIP schemes. We used the DART-trim tool
+\cite{wetzler2014drat} with a 10000 second timeout to check and
+optimize every DRAT proof once \footnote{Applying DART-trim multiple
+  times can further reduce the proof size until a fix-point. However,
+  the full optimization is time consuming and was impractical for our
+  experiments.}.
+
+The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
+are 413.2MB and 487.2MB, respectively. Both sizes are significantly
+smaller than the average optimized proof size from 1-UIP, 613.9MB.
+The average proof size reduction per instance for $\allUipMin$ and
+$\allUipPure$ was 17.18\% and 6.9\% compared with 1-UIP, which roughly
+correlate with our observed clause size observation in
+Fig~\ref{fig:len_pdf}.
 
@@
 \subsubsection{$\allUip$ on Modern SAT solvers}
-To validate $\allUip$ as practical learning schemes on modern SAT
-solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
-SAT Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT} and $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$). $\expSATShort$
-is a top ten solver from 2019 SAT race which uses random walk
-simulation to help branching. We chose $\expSATShort$ because 1) it is
-a top solver in the 2019 SAT Race without using chronological
-backtracking; 2) the random walk simulation 
-branching heuristic is different from local branching heuristic (VSIDS and LRB) that we have considered,
-hence it is in the blind spot of $\allUip$.
-we compare the base 1-UIP learning scheme against $\allUipPure$,
-$\allUipMin$ and the top two $\allUip$ variants, $\allUipAct$ and
-$\allUipIn$, on the SAT Race 2019 main track benchmark. We report
-solved instances, PAR-2 score and the average clause size.
+To validate $\allUip$ as a practical learning scheme for modern SAT
+solvers, we implemented $\allUip$ in the winners of 2017, 2018 and
+2019 SAT
+Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
+and in the $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$)
+solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
+uses random walk simulation to help branching. We chose $\expSATShort$
+because (1) it one of the top solvers in the 2019 SAT Race that does
+not use chronological backtracking; and (2) the random walk simulation
+branching heuristic is different from local branching heuristic (VSIDS
+and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
+$\allUipEx$. We compared these solver's base 1-UIP learning scheme
+with $\allUipPure$, $\allUipMin$ and the top two $\allUip$ variants,
+$\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main track
+benchmark. We report solved instances, PAR-2 score and the average
+clause size.
 
@@
 
-Table~\ref{fig:t5} shows the benchmark result of $\allUip$
-configurations on modern solvers.
-Overall, we observed similar performance gain on all modern solvers as we have seen on $\MapleBase$ in Fig.~\ref{fig:t4}.
-More specifically, almost all configurations improved on solved instance (+4.5 instances in average) and PAR-2 score (-83.6 in average).
-The average clause size reduction is consistent across all solvers.
-Each configuration also exhibits different strength: $\allUipPure$ solved the most instances with the best PAR-2 score on two solvers, $\allUipMin$ yields small clauses, $\allUipIn$ solved the most SAT instances, and $\allUipAct$ has stable performance.
+Table~\ref{fig:t5} shows the benchmark result of the $\allUip$
+configurations in our suite of modern solvers. Overall, we observed
+similar performance gain on all modern solvers as we have seen on
+$\MapleBase$ in Fig.~\ref{fig:t4}. More specifically, almost all
+configurations improved on solved instance (+4.5 instances in average)
+and PAR-2 score (-83.6 in average).  The average clause size reduction
+is consistent across all solvers. Each configuration also exhibits
+different strength: $\allUipPure$ solved the most instances with the
+best PAR-2 score on two solvers, $\allUipMin$ yields small clauses,
+$\allUipIn$ solved the most SAT instances, and $\allUipAct$ has stable
+performance.
 
 On the SAT 2017 race winner $\MapleSeven$, all four configurations of
-$\allUip$ outperformed 1-UIP learning by 12, 8, 5, and 2 instances, respectively. $\allUipPure$ solved
-more UNSAT and SAT instances while other configurations improved on
-solving SAT instances. The clause size reduction of $\allUip$ is more significant on $\MapleSeven$ compare to $\MapleBase$. 
+$\allUip$ outperformed 1-UIP learning by 12, 8, 5, and 2 instances,
+respectively. $\allUipPure$ solved more UNSAT and SAT instances while
+the other configurations improved on solving SAT instances. The clause
+size reduction of $\allUip$ is more significant on $\MapleSeven$
+than on to $\MapleBase$.
 %This may suggests that $\allUip$ and the recent
@@
 
-The SAT 2018 race winner$\MapleEightShort$ uses chronological backtracking (CB) for long distance backtracks.
-Therefore, we applied $\allUip$-CB extension for all $\allUip$ configurations.
-Three out of four $\allUip$ configurations outperformed 1-UIP by 5, 4 an 4 instances,
+The SAT 2018 race winner$\MapleEightShort$ uses chronological
+backtracking (CB) for long distance backtracks. Three out of four
+$\allUip$ configurations outperformed 1-UIP by 5, 4 an 4 instances,
 respectively.
@@
 
-On the SAT 2019 race winner $\MapleNineShort$,  three out of four$\allUip$ configurations solved instances 6, 5 and 5 more instances than 1-UIP, respectively.
-The solver uses duplicate learning (DL) to prioritize clauses that are learned multiple times.
-We observed that clauses learnt from $\allUip$ configurations are less likely to be duplicated.
-As an example, $\allUipMin$ in average, added 12\% less duplicated clauses into the
-core clause database than 1-UIP. This observation is surprising, and the cause is unclear.
+On the SAT 2019 race winner $\MapleNineShort$, three out of four
+$\allUip$ configurations solved instances 6, 5 and 5 more instances
+than 1-UIP, respectively. $\MapleNineShort$ prioritizes clauses that
+are learned multiple times. We observed that clauses learnt from
+$\allUip$ configurations are less likely to be duplicated. As an
+example, $\allUipMin$ in average, added 12\% less duplicated clauses
+into the core clause database than 1-UIP. This observation is
+surprising, and the cause is unclear.
 
@@
 
-On $\expSATShort$,
-three out of four $\allUip$ configurations solved 4, 7 and 8 more instances
-than 1-UIP learning.
-We noticed that $\allUipAct$ and $\allUipIn$ show significant better performance than $\allUipMin$ and $\allUipPure$ on this solver.
-The random walk simulation branching heuristic however didn't impact the performance of $\allUip$ schemes significantly.
+On $\expSATShort$, three out of four $\allUip$ configurations solved
+4, 7 and 8 more instances than 1-UIP learning. We noticed that
+$\allUipAct$ and $\allUipIn$ show significant better performance than
+$\allUipMin$ and $\allUipPure$ on this solver. The random walk
+simulation branching heuristic, however, didn't impact the performance
+of $\allUip$ schemes significantly.
 %$\allUip$ to solve the similar amount of instances with close PAR-2

### 2020-02-24 paper.tex

@@
 %\end{figure}
+
 \subsubsection{Reduced Proof sizes with $\allUip$.}
@@
 $\allUipMin$ and 1-UIP schemes. We used the DART-trim tool
-\cite{wetzler2014drat} with a 10000 second timeout to check and
+\cite{DBLP:conf/sat/WetzlerHH14} with a 10000 second timeout to check and
 optimize every DRAT proof once \footnote{Applying DART-trim multiple
@@
 2019 SAT
-Race\cite{luo2017effective,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
+Race\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
 and in the $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$)

### 2020-02-24 paper.tex
fix typo, update proof reduction results with 30000 proof checking time

@@
 $C_L$ whose LBD is minimum among all asserting clauses that can be
-learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a reason
+learnt from $C_I$. If $C'$ is a trail resolvent of $C$ and a reason
 clause $\reason{l}$, then $\LBD(C') \geq \LBD(C)$ since $\reason{l}$
@@
 course, generalizations of this approach where one has a weighted
-tradeoff between these factors that allows preferring the new clause
+trade-off between these factors that allows preferring the new clause
 when it has large gains in one metric even though it has small losses
@@
 A learning scheme that yields smaller clauses (lemmas) might also
-construct smaller causal proofs. For UNSAT instances, we additionally
-compare the size of the optimized DRAT proof from $\allUipPure$,
-$\allUipMin$ and 1-UIP schemes. We used the DART-trim tool
-\cite{DBLP:conf/sat/WetzlerHH14} with a 10000 second timeout to check and
+construct smaller causal proofs. For 88 UNSAT instances solved mutually
+by $\allUipPure$, $\allUipMin$ and 1-UIP schemes, we additionally
+compare the size of the optimized DRAT proof from the three learning schemes. 
+We used the DART-trim tool
+\cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
 optimize every DRAT proof once \footnote{Applying DART-trim multiple
@@
 The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
-are 413.2MB and 487.2MB, respectively. Both sizes are significantly
-smaller than the average optimized proof size from 1-UIP, 613.9MB.
+are 556.6MB and 654.4MB, respectively. Both sizes are significantly
+smaller than the average optimized proof size from 1-UIP, 824.9MB.
 The average proof size reduction per instance for $\allUipMin$ and
-$\allUipPure$ was 17.18\% and 6.9\% compared with 1-UIP, which roughly
+$\allUipPure$ was 16.5\% and 4.6\% compared with 1-UIP, which roughly
 correlate with our observed clause size observation in

### 2020-02-25 paper.tex
Nick's pass, added comments

@@
 used in modern \sat solvers \cite{DBLP:conf/sat/KatebiSS11}. In early
-\sat research a number of different clause learning techniques were
+\sat research \nf{,} a number of different clause learning techniques were
 proposed
@@
 However, other clause learning schemes can be used in SAT solvers
-without changes to their main data structures. Furthermore, advances
+without %changes%
+\nf{alternations} to %their%
+\nf{solvers'} main data structures. Furthermore, advances
 in our understanding allow us to better understand the potential
 advantages and disadvantages of these alternate schemes. In this paper
-we reexamine these previous clause learning schemes, with a
+we reexamine these %previous clause learning schemes,%
+\nf{previously proposed alternative schemes,} with a
 focus on the schemes described in
@@
 solvers, obtained from the last decade of research, allows us to see
-that in their original form these other clause learning schemes suffer
+that in their original form these %other%
+\nf{alternative} clause learning schemes suffer
 significant disadvantages over 1-UIP clause learning.
@@
 Various minimization techniques can be applied on top of our method to
-remove redundant literals from the clauses we learn.
+remove redundant literals from the clauses %we learn%
+\nf{we have learnt}.
 
 We present various versions of our method and show that these variants
-are often capable of learning shorter clauses that the 1-UIP scheme,
+are often capable of learning shorter clauses than the 1-UIP scheme,
 and that this can lead to useful performance gains in state of the art
@@
 equal to the number of decision literals appearing before $\ell$ on
-the trail plus one; hence, $\dlevel{d}=1$ for the first decision
+the trail and plus one \nf{if $\ell$ is a decision literal}; hence, $\dlevel{d}=1$ for the first decision
 literal $d\in\trail$.  The set of literals on $\trail$ that have the
@@
 
-In most \sat solvers clause learning is initiated as soon as a clause
+In most \sat solvers\nf{,} clause learning is initiated as soon as a clause
 is falsified by $\trail$. In this paper we will be concerned with the
 subsequent clause learning process which uses $\trail$ to derive a new
-clause. In some implementations the trail might be altered during
+clause. In some implementations\nf{,} the trail might be altered during
 clause learning. Here, however, we will assume that $\trail$
@@
 
-\begin{defn}[Trail Resolvant]
-    A trail resolvant is a clause arising from resolving a conflict
+\begin{defn}[Trail resolvent]
+    A trail resolvent is a clause arising from resolving a conflict
     against the reason clause of some literal $\ell \in \trail$. Every
-    trail resolvant is also a conflict.
+    trail resolvent is also a conflict.
 \end{defn}
 
-The following things can be noted about trail resolvants: (1) trail
-resolvants are never tautological, as the polarity of all literals in
+The following things can be noted about trail resolvents: (1) trail
+resolvents are never tautological, as the polarity of all literals in
 $\reason{\ell}$ other than $\ell$ must agree with the polarity of all
@@
 \begin{defn}[Trail Resolution]
-    A trail resolution is a sequence of trail resolvants applied to an
+    A trail resolution is a sequence of trail resolvents applied to an
     initial conflict $C_I$ yielding a new conflict $C_L$.  A trail
@@
 with a single literal at its $i$ deepest decision levels. It should be
-noted that when resolving away literals at decision level $i$ new
+noted that when resolving away literals at decision level $i$\nf{,} new
 literals at decision levels less than $i$ might be introduced into the
@@
 current deepest level to the point the learnt clause first becomes
-unit and then use the learnt clause to add a new unit implicant to the
+unit\nf{,} and then use the learnt clause to add a new unit implicant to the
 trail. Since all but the deepest level was propagation complete, this
@@
 however, potentially increase the number of literals at level $i$. In
-resolving away a literal $l$ at level $i$ more literals might be
+resolving away a literal $l$ at level $i$\nf{,} more literals might be
 introduced into level $i$, and some of these might not be removable by
@@
 
-Finally, the Algorithm returns the shorter of the initial 1-UIP clause
+Finally, the \nf{a}lgorithm returns the shorter of the initial 1-UIP clause
 and the newly computed clause (line~\ref{ln:returnmin}). Because we
-return the newly computed clause only when it is shorter an early
+return the newly computed clause only when it is shorter\nf{,} an early
 termination test can be used (line~\ref{ln:earlyTerminate}). After the
@@
 and if that lower bound is as large as the size of the initial 1-UIP
-clause we can terminate and return the initial 1-UIP clause. As
-observed above, \allUipPure cannot trigger this early termination
-condition (it cannot increase the size of any processed level), but
-\allUipMin can.
+clause\nf{,} we can terminate and return the initial 1-UIP clause. As
+observed above, \nf{this termination condition does not account for the 
+final clause minimization(line~\ref{ln:puremin}) for \allUipPure, and it could
+prematurely terminates the algorithm which would otherwise yield a shorter clause.
+However, we discovered such false negative is indeed rare, and applying early termination on \allUipPure generally improves performance.}
+%\allUipPure cannot trigger this early termination
+%condition (it cannot increase the size of any processed level), but
+%\allUipMin can.
 
@@
 \cite{DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}. When
-chronological backtracking is used the literals on the trail might no
+chronological backtracking is used\nf{,} the literals on the trail might no
 longer be sorted by decision level. So resolving literals in the
@@
 
-Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19} it can be
+Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19}\nf{,} it can be
 observed that all clause learning schemes continue to work as long as

### 2020-02-25 paper.tex
change statement location of the n_try update in algo
fix typo

@@
 particular, $C_L$ contains at least two literals at the deepest level.
-If two of its literals at the deepest level were were not completely
+If two of its literals at the deepest level were \nf{forced but} not completely
 unit propagated (unit propagation is aborted as soon as a conflict is
@@
 \Procedure{\allUip}{$C_1, \assertionTrail$, \type} 
-    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return} $C_1$\label{ln:control}
+    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return}
+    $C_1$\label{ln:control}
+    \nf{\State $\ntries$++\label{ln:try}} 
     \State $C_i \gets C_1$
@@
     \State \textbf{if} ($\type = \allUipPure$) $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
-    \State $\ntries$++\label{ln:try}
+    %\State $\ntries$++\label{ln:try}% 
     \State \textbf{if} ($|C_i| < |C_1|$) $\nsuc$++, \textbf{return} $C_i$ 

### 2020-02-27 paper.tex
update optimized proof checking size counting only mutally solved instances

@@
 clauses for 88.5\% (77.7\%) of instances, and the average relative
-reduction ratio over 1-UIP is 18.6\%
-(16.7\%). Fig.~\ref{fig:len_compare} compares the average learnt
+reduction ratio over 1-UIP is 18.5\%
+(9.6\%). Fig.~\ref{fig:len_compare} compares the average learnt
 clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
@@
 The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
-are 556.6MB and 654.4MB, respectively. Both sizes are significantly
+are 556.6MB and 698.5MB, respectively. Both sizes are significantly
 smaller than the average optimized proof size from 1-UIP, 824.9MB.
 The average proof size reduction per instance for $\allUipMin$ and
-$\allUipPure$ was 16.5\% and 4.6\% compared with 1-UIP, which roughly
+$\allUipPure$ was 16.5\% and 3.6\% compared with 1-UIP, which roughly
 correlate with our observed clause size observation in

### 2020-02-28 paper.tex
add conclusion

@@
 
+\section{Conclusion}We introduced $\allUip$, a new kind of clause learning schemes that combines the strengths of the existing all-UIP and 1-UIP learning scheme to learn shorter asserting clauses while preserving the minimum LBDs.
+As our experiments have shown, employing $\allUip$ and its variants in modern CDCL solvers achieves significant clause reduction and yields useful performance gains.
+
+Our scheme extends 1-UIP learning by performing further resolution beyond the deepest decision level to find the UIP for each involved level.
+Since resolutions may increase the clause's LBD through introducing literals from new decision level, we presented two methods to block off these literals.
+Although our learning scheme is conceptually simple, implementing it as an efficient clause learning scheme for modern solvers is not straight forward.
+Therefore, we presented optimizations to reduce and balance the learning cost.
+We additionally presented variants of our schemes to account for features (e.g.
+local branching heuristics and chronological backtracking) used in state of the art solvers.
+
+Although the field of SAT solving has long converged on using 1-UIP learning scheme, we've shown the possibility of developing an effective alternative through understanding 1-UIP and other existing schemes' strengths and weaknesses.
+Our learning scheme can be generalized and further improved by exploring more fine-grained trade-offs between different clause quality metrics beyond clause size and LBD.
+We also plan to study the interaction between clause learning and variable branching.
+Since most of the branching heuristics are tailored for 1-UIP scheme, the interactions with other learning schemes still need further exploration which could advance our understanding of SAT solving.
+
 \bibliography{sat}{}

### 2020-02-28 paper.tex
fixing typos

@@
 
-\section{Conclusion}We introduced $\allUip$, a new kind of clause learning schemes that combines the strengths of the existing all-UIP and 1-UIP learning scheme to learn shorter asserting clauses while preserving the minimum LBDs.
+\section{Conclusion}We introduced $\allUip$, a new kind of clause learning scheme that combines the strengths of the existing all-UIP and 1-UIP learning schemes to learn shorter asserting clauses while preserving the minimum LBDs.
 As our experiments have shown, employing $\allUip$ and its variants in modern CDCL solvers achieves significant clause reduction and yields useful performance gains.
@@
 We also plan to study the interaction between clause learning and variable branching.
-Since most of the branching heuristics are tailored for 1-UIP scheme, the interactions with other learning schemes still need further exploration which could advance our understanding of SAT solving.
+Since most of the branching heuristics are tailored for 1-UIP scheme, the interactions with other learning schemes still need further study which could advance our understanding of SAT solving.
 

### 2020-02-29 paper.tex
fb-final mods

--- 2020-02-27 paper.tex
@@
 \section{Introduction}
-Clause learning is an essential technique in \sat solvers. There is good
-evidence to indicate that it is, in fact, the most important technique
-used in modern \sat solvers \cite{DBLP:conf/sat/KatebiSS11}. In early
-\sat research \nf{,} a number of different clause learning techniques were
-proposed
+Clause learning is an essential technique in \sat solvers. There is
+good evidence to indicate that it is, in fact, the most important
+technique used in modern \sat solvers
+\cite{DBLP:conf/sat/KatebiSS11}. In early \sat research a number of
+different clause learning techniques were proposed
 \cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
@@
 However, other clause learning schemes can be used in SAT solvers
-without %changes%
-\nf{alternations} to %their%
-\nf{solvers'} main data structures. Furthermore, advances
-in our understanding allow us to better understand the potential
+without changes to the main data structures. Furthermore, advances in
+our understanding allow us to better understand the potential
 advantages and disadvantages of these alternate schemes. In this paper
-we reexamine these %previous clause learning schemes,%
-\nf{previously proposed alternative schemes,} with a
-focus on the schemes described in
-\cite{DBLP:conf/iccad/ZhangMMM01}. Improved understanding of \sat
-solvers, obtained from the last decade of research, allows us to see
-that in their original form these %other%
-\nf{alternative} clause learning schemes suffer
-significant disadvantages over 1-UIP clause learning.
+we reexamine these previously proposed schemes with a focus on the
+schemes described in \cite{DBLP:conf/iccad/ZhangMMM01}. Improved
+understanding of \sat solvers, obtained from the last decade of
+research, allows us to see that in their original form these
+alternative clause learning schemes suffer significant disadvantages
+over 1-UIP clause learning.
 
@@
 Our method is orthogonal to clause minimization. In particular, our
-approach can learn a clause that is not a subset of the 1-UIP clause but
-which still serves all of the same purposes as the 1-UIP clause.
+approach can learn a clause that is not a subset of the 1-UIP clause
+but which still serves all of the same purposes as the 1-UIP clause.
 Various minimization techniques can be applied on top of our method to
-remove redundant literals from the clauses %we learn%
-\nf{we have learnt}.
+remove redundant literals from the clauses we have learnt.
 
@@
 equal to the number of decision literals appearing before $\ell$ on
-the trail and plus one \nf{if $\ell$ is a decision literal}; hence, $\dlevel{d}=1$ for the first decision
+the trail plus one. Hence, $\dlevel{d}=1$ for the first decision
 literal $d\in\trail$.  The set of literals on $\trail$ that have the
@@
 
-In most \sat solvers\nf{,} clause learning is initiated as soon as a clause
+In most \sat solvers, clause learning is initiated as soon as a clause
 is falsified by $\trail$. In this paper we will be concerned with the
 subsequent clause learning process which uses $\trail$ to derive a new
-clause. In some implementations\nf{,} the trail might be altered during
+clause. In some implementations the trail might be altered during
 clause learning. Here, however, we will assume that $\trail$
 \emph{remains intact during clause learning} and is only changed after
-the new clause is learnt. After the new clause is learnt $\trail$
-will be changed by backtracking.
+the new clause is learnt. After the new clause is learnt $\trail$ will
+be changed by backtracking.
 
@@
 with a single literal at its $i$ deepest decision levels. It should be
-noted that when resolving away literals at decision level $i$\nf{,} new
+noted that when resolving away literals at decision level $i$, new
 literals at decision levels less than $i$ might be introduced into the
@@
 current deepest level to the point the learnt clause first becomes
-unit\nf{,} and then use the learnt clause to add a new unit implicant to the
-trail. Since all but the deepest level was propagation complete, this
-means that the asserting clause must be a brand new clause; otherwise
-that unit implication would already have been made. On the other hand,
-if the learnt clause $C_L$ is not asserting then \emph{it could be
-  that it is a duplicate of another clause already in the formula}. In
-particular, $C_L$ contains at least two literals at the deepest level.
-If two of its literals at the deepest level were \nf{forced but} not completely
-unit propagated (unit propagation is aborted as soon as a conflict is
-found), then $C_L$ could already be in the formula: its two watches
-might not be fully propagated and $C_L$ could be a falsified clause
-not detected by the solver. (In general, more than one clause might be
-falsified by $\trail$ and the \sat solver will generally stop when it
-finds the first one).
+unit, and then use the learnt clause to add a new unit implicant to
+the trail. Since all but the deepest level was propagation complete,
+this means that the asserting clause must be a brand new clause;
+otherwise that unit implication would already have been made. On the
+other hand, if the learnt clause $C_L$ is not asserting then \emph{it
+  could be that it is a duplicate of another clause already in the
+  formula}. In particular, $C_L$ contains at least two literals at the
+deepest level.  If two of its literals at the deepest level were not
+completely unit propagated (unit propagation is aborted as soon as a
+conflict is found), then $C_L$ could already be in the formula: its
+two watches might not be fully propagated and $C_L$ could be a
+falsified clause not detected by the solver. (In general, more than
+one clause might be falsified by $\trail$ and the \sat solver will
+generally stop when it finds the first one).
 
@@
     $C_1$\label{ln:control}
-    \nf{\State $\ntries$++\label{ln:try}} 
+    \State $\ntries$++\label{ln:try}
     \State $C_i \gets C_1$
@@
 however, potentially increase the number of literals at level $i$. In
-resolving away a literal $l$ at level $i$\nf{,} more literals might be
+resolving away a literal $l$ at level $i$, more literals might be
 introduced into level $i$, and some of these might not be removable by
@@
 
-Finally, the \nf{a}lgorithm returns the shorter of the initial 1-UIP clause
+Finally, the algorithm returns the shorter of the initial 1-UIP clause
 and the newly computed clause (line~\ref{ln:returnmin}). Because we
-return the newly computed clause only when it is shorter\nf{,} an early
+return the newly computed clause only when it is shorter, an early
 termination test can be used (line~\ref{ln:earlyTerminate}). After the
@@
 and if that lower bound is as large as the size of the initial 1-UIP
-clause\nf{,} we can terminate and return the initial 1-UIP clause. As
-observed above, \nf{this termination condition does not account for the 
-final clause minimization(line~\ref{ln:puremin}) for \allUipPure, and it could
-prematurely terminates the algorithm which would otherwise yield a shorter clause.
-However, we discovered such false negative is indeed rare, and applying early termination on \allUipPure generally improves performance.}
+clause we can terminate and return the initial 1-UIP clause.  This
+termination condition does not, however, account for the final clause
+minimization (line~\ref{ln:puremin}) of \allUipPure, and it could
+prematurely terminate processing that would otherwise yield a shorter
+clause. However, we discovered such false negative is rare, and
+applying early termination on \allUipPure generally improves
+performance.
 %\allUipPure cannot trigger this early termination
@@
 \cite{DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}. When
-chronological backtracking is used\nf{,} the literals on the trail might no
+chronological backtracking is used, the literals on the trail might no
 longer be sorted by decision level. So resolving literals in the
@@
 
-Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19}\nf{,} it can be
+Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19}, it can be
 observed that all clause learning schemes continue to work as long as
@@
 \begin{center}
-\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
+\begin{tabular}{ | m{3.2cm} | m{3.8cm}| m{2cm} | m{2.75cm} |  } 
 \hline
@@
 
+\section{Conclusion}
+In this paper we introduce a new clause learning scheme, $\allUip$,
+that combines the strengths of the existing all-UIP and 1-UIP learning
+schemes to learn shorter asserting clauses, while preserving the
+minimum LBDs of 1-UIP clauses. We provide empirical evidence that
+employing $\allUip$ and its variants in modern CDCL solvers achieves
+significant clause reduction and yields useful performance gains.
+
+Our scheme extends 1-UIP learning by performing further resolution
+beyond the deepest decision level in an attempt to find the UIP at
+each level in the learnt clause. Since resolutions may increase the
+clause's LBD by introducing literals from new decision level, we
+presented two methods to block such literals from entering the
+clause. Although our learning scheme is conceptually simple, and we
+have presented optimizations to reduce and balance the learning
+cost. We additionally presented variants of our schemes to account for
+features used in state of the art solvers, e.g., local branching
+heuristics and chronological backtracking.
+
+Although the field of SAT solving has long converged on using the
+1-UIP learning scheme, we've shown the possibility of developing an
+effective alternative through understanding 1-UIP and other existing
+schemes' strengths and weaknesses.  Our learning scheme can be
+generalized and further improved by exploring more fine-grained
+trade-offs between different clause quality metrics beyond clause size
+and LBD. We also plan to study the interaction between clause learning
+and variable branching. Since most of the branching heuristics are
+tailored for 1-UIP scheme, the interactions with other learning
+schemes still needs further study.
+
 \bibliography{sat}{}

### 2020-02-29 paper.tex
adjust plot's color scheme, a small wording change to specifiy how n_attempt and n_success is updated

@@
 (line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
-rate is greater  than 80\%, we decrease $\tgap$ by one (not allowing it to
+rate since the last restart is greater than 80\%, we decrease $\tgap$ by one (not allowing it to
 become negative), and if it is less than 80\% we increase $\tgap$

### 2020-05-03 paper.tex
fix bib and case inconsistency

@@
 
-\title{Clause size reduction with all-UIP Learning}
+\title{Clause Size Reduction with all-UIP Learning}
 \author{Nick Feng \and Fahiem Bacchus}
@@
 \emph{conflicting} literals $\ell\in C_1$ and $\lnot \ell \in
-C_2$. Their resolvant $C_1 \resolve C_2$ is the new clause
-$(C_1 \cup C_2) - \{\ell, \lnot \ell\}$. The resolvant will be a
+C_2$. Their resolvent $C_1 \resolve C_2$ is the new clause
+$(C_1 \cup C_2) - \{\ell, \lnot \ell\}$. The resolvent will be a
 tautology (i.e., a clause containing a literal $x$ and its negation
@@
 
-\begin{defn}[Trail resolvent]
+\begin{defn}[Trail Resolvent]
     A trail resolvent is a clause arising from resolving a conflict
@@
 
-\subsection{Some alternate Clause Learning Schemes}
+\subsection{Some Alternate Clause Learning Schemes}
 A number of different clause learning schemes for generating a new
@@
 
-\subsection{Asserting Clauses and LBD---Reasons to prefer 1-UIP
-  clauses}
+\subsection{Asserting Clauses and LBD---Reasons to Prefer 1-UIP
+  Clauses}
 An \textbf{asserting clause} \cite{DBLP:journals/ai/PipatsrisawatD11}
@@
 
-\subsubsection{Adjust Variable Activity.} \label{sec: varajust} An
+\subsubsection{Adjust Variable Activity} \label{sec: varajust} An
 alternative to filtering clauses with low average variable
@@
 
-\subsubsection{Chronological Backtracking.}
+\subsubsection{Chronological Backtracking}
 We tested our new clause learning schemes on solvers that utilized
@@
 
-\subsubsection{Reduced Proof sizes with $\allUip$.}
+\subsubsection{Reduced Proof Sizes with $\allUip$.}
 A learning scheme that yields smaller clauses (lemmas) might also
@@
 compare the size of the optimized DRAT proof from the three learning schemes. 
-We used the DART-trim tool
+We used the DRAT-trim tool
 \cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
-optimize every DRAT proof once \footnote{Applying DART-trim multiple
+optimize every DRAT proof once \footnote{Applying DRAT-trim multiple
   times can further reduce the proof size until a fix-point. However,

### 2020-05-07 paper.tex
change remaining i-uip to all-uip
fb camera ready revisions

@@
 \newcommand{\assertionTrail}{\trail}
+\newcommand{\oneuip}{1-UIP\xspace}
+\newcommand{\alluip}{all-UIP\xspace}
 \newcommand{\LBD}{\text{LBD}\xspace}
-\newcommand{\gap}{\text{Gap}}
+\newcommand{\gap}{\text{gap}}
 \newcommand{\tgap}{t_{\mathit{gap}}}
@@
 \newcommand{\nsuc}{n_{\mathit{succ}}}
-\newcommand{\allUip}{\textit{stable-allUIP}}
+\newcommand{\stablealluip}{\textit{stable-alluip}}
 \newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
-\newcommand{\allUipPure}{\textit{pure-allUIP}\xspace}
-\newcommand{\allUipMin}{\textit{min-allUIP}\xspace}
-\newcommand{\allUipAct}{\textit{allUIP-Active}}
-\newcommand{\allUipIn}{\textit{allUIP-Inclusive}}
-\newcommand{\allUipEx}{\textit{allUIP-Exclusive}}
+\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
+\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
+\newcommand{\allUipAct}{\textit{alluip-active}}
+\newcommand{\allUipIn}{\textit{alluip-inclusive}}
+\newcommand{\allUipEx}{\textit{alluip-exclusive}}
 \newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
@@
 \newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
-\newcommand{\expSAT}{\textit{expMaple\_CM\_GCBumpOnlyLRB} }
+\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
 \newcommand{\expSATShort}{\textit{expMaple} }
@@
 \newcommand{\Array}[2]{[\,#1\mid#2\,]}
-\newcommand{\target}{\textbf{Succ_t}}
 \newcommand{\trail}{\ensuremath{\mathcal{T}}}
@@
 
-\title{Clause Size Reduction with all-UIP Learning}
+\title{Clause Size Reduction with \alluip Learning}
 \author{Nick Feng \and Fahiem Bacchus}
@@
 \begin{abstract}
-    Almost all CDCL \sat solvers use the 1-UIP 
+    Almost all CDCL \sat solvers use the \oneuip 
     clause learning scheme for learning new clauses from conflicts, 
@@
     of \sat solving provides good reasons for using that scheme. In
-    particular, the 1-UIP scheme yields asserting clauses, and these
-    asserting clauses have minimum LBD among all possible asserting
+    particular, the \oneuip scheme yields asserting clauses, and these
+    asserting clauses have minimum \LBD among all possible asserting
     clauses. As a result of these advantages, other clause learning
-    schemes, like $i$-UIP and all-UIP, that were proposed in early
+    schemes, like $i$-UIP and \alluip, that were proposed in early
     work are not used in modern solvers. In this paper, we propose a
-    new technique for exploiting the all-UIP clause learning scheme. 
-    Our technique is to employ all-UIP learning under the
-    constraint that the learnt clause's LBD does not increase (over the
-    minimum established by the 1-UIP clause). Our method can learn
-    clauses that are significantly smaller than the 1-UIP clause while
-    preserving the minimum LBD. Unlike previous clause minimization
+    new technique for exploiting the \alluip clause learning scheme. 
+    Our technique is to employ \alluip learning under the
+    constraint that the learnt clause's \LBD does not increase (over the
+    minimum established by the \oneuip clause). Our method can learn
+    clauses that are significantly smaller than the \oneuip clause while
+    preserving the minimum \LBD. Unlike previous clause minimization
     methods, our technique is not limited to learning a sub-clause of
-    the 1-UIP clause. It can instead learn a completely different
+    the \oneuip clause. It can instead learn a completely different
     clause. We show empirically 
@@
 However, following the revolutionary performance improvements achieved
-by the Chaff \sat solver, the field has converged on using the 1-UIP
+by the Chaff \sat solver, the field has converged on using the \oneuip
 (first Unique Implication Point) scheme
@@
 pioneered in the Chaff solver).\footnote{The idea of UIP clauses was
-  first mentioned in \cite{DBLP:journals/tc/Marques-SilvaS99}, and
-  1-UIP clauses along with other UIP clauses were learnt and used in
+  first mentioned in \cite{DBLP:conf/iccad/SilvaS96}, and
+  \oneuip clauses along with other UIP clauses were learnt and used in
   the earlier GRASP \sat solver.} Since then almost all \sat solvers
-have employed the 1-UIP clause learning scheme, along with clause
+have employed the \oneuip clause learning scheme, along with clause
 minimization \cite{DBLP:conf/sat/SorenssonB09}, as their primary
@@
 alternative clause learning schemes suffer significant disadvantages
-over 1-UIP clause learning.
-
-One of the previously proposed schemes was the all-UIP scheme
+over \oneuip clause learning.
+
+One of the previously proposed schemes was the \alluip scheme
 \cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
 to exploit the main ideas of this scheme that avoids its main
-disadvantages. In particular, we propose to use a all-UIP like clause
+disadvantage which is that it can learn clauses with higher \LBD
+scores. In particular, we propose to use a \alluip like clause
 learning scheme to generate smaller learnt clauses which retain the
-good properties of standard 1-UIP clauses. Our method is related to,
+good properties of standard \oneuip clauses. Our method is related to,
 but not the same as, various clause mininimization methods that try to
-remove redundant literals from the 1-UIP clause yielding a clause that
-is a subset of the 1-UIP clause, e.g.,
+remove redundant literals from the \oneuip clause yielding a clause that
+is a subset of the \oneuip clause, e.g.,
 \cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
 Our method is orthogonal to clause minimization. In particular, our
-approach can learn a clause that is not a subset of the 1-UIP clause
-but which still serves all of the same purposes as the 1-UIP clause.
+approach can learn a clause that is not a subset of the \oneuip clause
+but which still serves all of the same purposes as the \oneuip clause.
 Various minimization techniques can be applied on top of our method to
@@
 We present various versions of our method and show that these variants
-are often capable of learning shorter clauses than the 1-UIP scheme,
+are often capable of learning shorter clauses than the \oneuip scheme,
 and that this can lead to useful performance gains in state of the art
@@
   literals} that have been assumed to be true by the solver, and
-\emph{unit propagated literals} that are forced to be true because
-they are the sole remaining unfalsified literal of a clause. Each
-literal $\ell\in\trail$ has a decision level $\dlevel{\ell}$ which is
-equal to the number of decision literals appearing before $\ell$ on
-the trail plus one. Hence, $\dlevel{d}=1$ for the first decision
-literal $d\in\trail$.  The set of literals on $\trail$ that have the
-same decision level forms a contiguous subsequence of $\trail$ that
-starts with a decision literal $d_i$ and ends just before the next
-decision literal $d_{i+1}$. If $\dlevel{d_i} = i$ we call this
-subsequence of $\trail$ the \textit{$i$-th decision level}.
+\emph{unit propagated literals} that are \emph{forced to be true}
+because they are the sole remaining unfalsified literal of a
+clause. Each literal $\ell\in\trail$ has a decision level
+$\dlevel{\ell}$. Let $k$ be the number of decision literals appearing
+before $\ell$ on the trail. When $\ell$ is a unit propagated literal
+$\dlevel{\ell}=k$, and when $\ell$ is a decision literal
+$\dlevel{\ell}=k+1$. For example, $\dlevel{d}=1$ for the first
+decision literal $d\in\trail$, and $\dlevel{\ell}=0$ for all literals
+$\ell$ appearing before $d$ on the trail.  The set of literals on
+$\trail$ that have the same decision level forms a contiguous
+subsequence of $\trail$ that starts with a decision literal $d_i$ and
+ends just before the next decision literal $d_{i+1}$. If
+$\dlevel{d_i} = i$ we call this subsequence of $\trail$ the
+\textit{$i$-th decision level}.
 
@@
 subsequent clause learning process which uses $\trail$ to derive a new
-clause. In some implementations the trail might be altered during
-clause learning. Here, however, we will assume that $\trail$
-\emph{remains intact during clause learning} and is only changed after
-the new clause is learnt. After the new clause is learnt $\trail$ will
-be changed by backtracking.
+clause. We will try to make as few assumptions about how $\trail$ is
+managed by the \sat solver as possible. One assumption we will make is
+that $\trail$ \emph{remains intact during clause learning} and is only
+changed after the new clause is learnt. After the new clause is learnt
+$\trail$ will be changed by backtracking.
 
@@
 $\trail[0]$---$\trail[\trailIdx{d_k}-1]$. We will assume that
-$\trail_{k-1}$ is \textbf{propagation complete}, although the full
-trail $\trail$ might not be. This means that (a) no clause was
+$\trail_{k-1}$ is \textbf{unit propagation complete}, although the
+full trail $\trail$ might not be. This means that (a) no clause was
 falsified by $\trail_{k-1}$. And (b) if $C_u$ is a clause containing
@@
 Any clause falsified by $\trail$ is called a \textbf{conflict}. When a
-conflict is found, the final level of the trail, $k$, need not be
+conflict is found, the final level of the trail, $k$, need not be unit
 propagation complete as the solver typically stops propagation as soon
@@
 \begin{Obs}
+    \label{obs:ordered}
     If the unordered trail resolution $U$ yields the conflict clause
@@
 trail. That is, these implementations are typically using ordered
-trail resolutions. The above observation shows that this in fact the
-right way to do this.
+trail resolutions. Observation~\ref{obs:ordered} shows that this in
+fact the right way to do this.
 
@@
 literals leaving a learnt clause over only decision literals; (c) the
-1-UIP scheme which resolves away literals from the deepest decision
+\oneuip scheme which resolves away literals from the deepest decision
 level leaving a learnt clause with a single literal at the deepest
-level; (d) the all-UIP scheme which resolves away literals from each
+level; (d) the \alluip scheme which resolves away literals from each
 decision level leaving a learnt clause with a single literal at each
@@
 literals at decision levels less than $i$ might be introduced into the
-clause. Hence, it is important in the $i$-UIP and all-UIP schemes to
+clause. Hence, it is important in the $i$-UIP and \alluip schemes to
 use ordered trail resolutions.
 
-Both the all-decision and all-UIP schemes yield a clause with only one
-literal at each decision level, and the all-UIP clause will be no
+Both the all-decision and \alluip schemes yield a clause with only one
+literal at each decision level, and the \alluip clause will be no
 larger that the all-decision clause. Furthermore, it is known
@@
 the level $d$. In particular, a decision level might contain more than
-one unique implication point. The algorithms given in
-Figure~\ref{fig:cl_schemes} stop at the first UIP of a level, except
-for the all-decision schemes with stops at the last UIP of each level.
+one unique implication point, and in some contexts the term all-UIP is
+used to refer to all the unique implication points that exist in a
+particular decision level \cite{DBLP:conf/sat/SabharwalSS12} rather
+than the \alluip clause learning scheme as is used here. The
+algorithms given in Figure~\ref{fig:cl_schemes} stop at the first UIP
+of a level, except for the all-decision schemes with stops at the last
+UIP of each level.
 
 \begin{figure}[t!]
-{\fss{8pt}{10pt}
+{\fss{9pt}{10pt}
 \begin{tabular}[t]{|l|c|l|}
@@
         \begin{algorithmic}[0]
-            \State \hspace*{-1em}\textbf{1-UIP}($C_I$)
+            \State \hspace*{-1em}\textbf{\oneuip}($C_I$)
             \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
@@
     \begin{minipage}{\textwidth}
-        \textbf{(d) All UIP Clause \rule{0pt}{1.1\topskip}}
+        \textbf{(d) \alluip Clause \rule{0pt}{1.1\topskip}}
         \begin{algorithmic}[0]
-          \State \hspace*{-1em}\textbf{all-UIP}($C_I$, $i$)
+          \State \hspace*{-1em}\textbf{\alluip}($C_I$, $i$)
           \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
@@
 
-\subsection{Asserting Clauses and LBD---Reasons to Prefer 1-UIP
+\subsection{Asserting Clauses and \LBD---Reasons to Prefer \oneuip
   Clauses}
@@
 unit, and then use the learnt clause to add a new unit implicant to
-the trail. Since all but the deepest level was propagation complete,
-this means that the asserting clause must be a brand new clause;
-otherwise that unit implication would already have been made. On the
-other hand, if the learnt clause $C_L$ is not asserting then \emph{it
-  could be that it is a duplicate of another clause already in the
-  formula}. In particular, $C_L$ contains at least two literals at the
-deepest level.  If two of its literals at the deepest level were not
-completely unit propagated (unit propagation is aborted as soon as a
-conflict is found), then $C_L$ could already be in the formula: its
-two watches might not be fully propagated and $C_L$ could be a
-falsified clause not detected by the solver. (In general, more than
-one clause might be falsified by $\trail$ and the \sat solver will
-generally stop when it finds the first one).
-
-The LBD of the learnt clause $C_L$ is the number of different decision
+the trail. Since all but the deepest level was unit propagation
+complete, this means that the asserting clause must be a brand new
+clause; otherwise that unit implication would already have been
+made. On the other hand, if the learnt clause $C_L$ is not asserting
+then it could be that it is a duplicate of another clause already in
+the formula.
+
+\begin{example}
+    Suppose that $a$ is a unit propagated literal and $d$ is a
+    decision literal with $\dlevel{d}>\dlevel{a}$. Let the sequence of
+    clauses watched by $\lnot d$ be $(\lnot d$, $x$, $\lnot a)$,
+    $(\lnot d$, $y$, $\lnot x$, $\lnot a)$, $(\lnot d$, $\lnot y$,
+    $\lnot x$, $\lnot a)$, $(\lnot d$, $\lnot x$, $\lnot a)$. When $d$
+    is unit propagated the clauses on $\lnot d$'s watch list will be
+    checked in this order.
+
+    Hence, unit propagation of $d$ will extend the trail by first
+    adding the unit propagated literal $x$ (with $\reason{x}=(x$,
+    $\lnot d$, $\lnot a)$) and then the unit propagated literal $y$
+    (with $\reason{y}=(y$, $\lnot x$, $\lnot a$, $\lnot d)$). Now the
+    third clause on $\lnot d$'s watch list, $(\lnot d$, $\lnot y$,
+    $\lnot x$, $\lnot a)$ is detected to be a conflict.
+
+    Clause learning can now be initiated from conflict $C_I=(\lnot d$,
+    $\lnot y$, $\lnot x$, $\lnot a)$. This clause has 3 literals at
+    level $\dlevel{d}=10$. If we stop clause learning before reaching
+    an asserting clause, then it is possible to simply resolve $C_I$
+    with $\reason{y}$ to obtain the learnt clause $C_L = (\lnot d$,
+    $\lnot x$, $\lnot a)$. However, this non-asserting learnt clause
+    is a duplicate of the fourth clause on $\lnot d$'s watch list
+    which is already in the formula.\footnote{In this example, the
+      fourth clause on $\lnot d$'s watch list subsumes the third
+      clause. But it is not difficult to construct more elaborate
+      examples where there are no subsumed clauses and we still obtain
+      learnt clauses that are duplicates of clauses already in the
+      formula.}  This issue can arise whenever $C_L$ contains two or
+    more literals at the deepest level (i.e., whenever $C_L$ is not
+    asserting). In such cases $C_L$ might be a clause already in the
+    formula with its two watches not yet fully unit propagated (and
+    thus $C_L$ is not detected by the \sat solver to be a conflict)
+    since propagation is typically stopped as soon as a conflict is
+    detected.
+
+\end{example}
+
+The \LBD of the learnt clause $C_L$ is the number of different decision
 levels in it:
 $\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
-\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a successful
-predictor of clause usefulness: clauses with lower LBD tend to be more
+\cite{DBLP:conf/ijcai/AudemardS09}. Empirically \LBD is a successful
+predictor of clause usefulness: clauses with lower \LBD tend to be more
 useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09}, from the
-initial falsified clause $C_I$ the 1-UIP scheme will produce a clause
-$C_L$ whose LBD is minimum among all asserting clauses that can be
+initial falsified clause $C_I$ the \oneuip scheme will produce a clause
+$C_L$ whose \LBD is minimum among all asserting clauses that can be
 learnt from $C_I$. If $C'$ is a trail resolvent of $C$ and a reason
@@
 as $l$ and might contain literals with decision levels not in
-$C$. That is, each trail resolution step can only increase the \LBD of
-the learnt clause. Hence, the 1-UIP scheme yields an asserting clause
-with minimum LBD as it performs the minimum number of trail resolutions
-required to generate an asserting clause.
+$C$. That is, the each trail resolution step might increase the \LBD
+of the learnt clause and can never decrease the \LBD. Hence, the \oneuip
+scheme yields an asserting clause with minimum \LBD as it performs the
+minimum number of trail resolutions required to generate an asserting
+clause.
 
 The other schemes must perform more trail resolutions. In fact, all of
-these schemes (all-decision, all-uip, i-uip) use trail resolutions in
-which the 1-UIP clause appears. That is, they all must first generate
-the 1-UIP clause and then continue with further trail resolution
+these schemes (all-decision, all-UIP, i-UIP) use trail resolutions in
+which the \oneuip clause appears. That is, they all must first generate
+the \oneuip clause and then continue with further trail resolution
 steps. These extra resolution steps can introduce many addition
 decision levels into the final clause. Hence, these schemes yield
-learnt clauses whose LBD can be larger, and never smaller, that the
-1-UIP clause.
-
-Putting these two observations together we see that the 1-UIP scheme
-produces asserting clauses with lowest possible LBD. This is a
+learnt clauses whose \LBD can be larger, and never smaller, that the
+\oneuip clause.
+
+Putting these two observations together we see that the \oneuip scheme
+produces asserting clauses with lowest possible \LBD. This is a
 compelling reasons for using this scheme. Hence, it is not surprising
-that modern \sat solvers almost exclusively use 1-UIP clause
+that modern \sat solvers almost exclusively use \oneuip clause
 learning.\footnote{Knuth in his sat13 CDCL solver \cite{Knuth:Sat13}
-  uses an all-decision clause when the 1-UIP clause is too large. In
-  this context an all-UIP clause could also be used as it would be no
+  uses an all-decision clause when the \oneuip clause is too large. In
+  this context an \alluip clause could also be used as it would be no
   larger than the all decision clause.}
 
-\section{Using all-UIP Clause Learning}
+\section{Using \alluip Clause Learning}
 \label{sec:i-uip}
-Although learning clauses with low LBD has been shown empirically to
+Although learning clauses with low \LBD has been shown empirically to
 be more important in \sat solving than learning short clauses
@@
 
-The all-UIP scheme will tend to produce small clauses since the
+The \alluip scheme will tend to produce small clauses since the
 clauses contain at most one literal per decision level. However, the
-all-UIP clause can have much higher LBD. Since, $\LBD$ is more important
-than size our approach is to use all-UIP learning when, and only when,
-it succeeds in reducing the size of the clause \emph{without
-  increasing its LBD}. The all-UIP scheme first computes the 1-UIP
-clause when it reduces the deepest level to a single UIP literal. It
-then proceeds to reduce the shallower levels (see all-UIP's for loop
-in Figure~\ref{fig:cl_schemes}). So our approach will start with the
-1-UIP clause and then try to apply all-UIP learning to reduce other
-levels to single literals. As noted above, clause minimization is
-orthogonal to our approach, so we also first apply standard clause
-minimization \cite{DBLP:conf/sat/SorenssonB09} to the 1-UIP
-clause. That is, our algorithm will start with the clause that most
-\sat solvers learn from a conflict, a minimized 1-UIP clause. Our
-approach is specified in Alg. ~\ref{alg:i-uip}.
+\alluip clause can have much higher \LBD. Since $\LBD$ is more
+important than size, our approach is to use \alluip learning when, and
+only when, it succeeds in reducing the size of the clause
+\emph{without increasing its \LBD}. The \alluip scheme first computes
+the \oneuip clause when it reduces the deepest level to a single UIP
+literal. It then proceeds to reduce the shallower levels (see
+\alluip's for loop in Figure~\ref{fig:cl_schemes}). So our approach
+will start with the \oneuip clause and then try to apply \alluip
+learning to reduce other levels to single literals. As noted above,
+clause minimization is orthogonal to our approach, so we also first
+apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
+to the \oneuip clause. That is, our algorithm will start with the
+clause that most \sat solvers learn from a conflict, a minimized
+\oneuip clause. Our approach is specified in Algorithm~\ref{alg:i-uip}
+($\stablealluip$).
 
 \begin{algorithm}[!t]
-\caption{\allUip}\label{alg:i-uip}
+\caption{\stablealluip}\label{alg:i-uip}
 \begin{algorithmic}[1]
-\Require $C_1$ is minimized 1-UIP clause
-\Require $\tgap\geq 0$ is global parameter, $\ntries$ and $\nsuc$
-         are globals used to configure $\tgap$
-\Procedure{\allUip}{$C_1, \assertionTrail$, \type} 
-    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return}
-    $C_1$\label{ln:control}
-    \State $\ntries$++\label{ln:try}
-    \State $C_i \gets C_1$
-    \State $\dlevels \gets\mbox{}$ decision levels in $C_i$ in descending order
-           \Comment{These never change}
-    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest level $\dlevels[0]$}
-       \State $C_i\gets\mbox{}$\tryuiplevel($C_i$, $\dlevels[i]$, \type)
-              \Comment{Try to reduce this level to UIP}
-       \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq \dlevels[i]\}\bigr|
-                    + (|\dlevels| - (i+1)) \geq |C_1|$}\label{ln:earlyTerminate}
-           \State \textbf{return} $C_1$. \Comment{can't generate smaller clause}
-       \EndIf
-    \EndFor
-    \State \textbf{if} ($\type = \allUipPure$) $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
+    \Require $C_1$ is minimized \oneuip clause
+    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are globals used to
+    configure $\tgap$,   \type{} a global determining how the
+    introduction of new decision levels is handled.
+
+    \Procedure{\stablealluip}{$C_1, \assertionTrail$, \type} \State
+    \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return}
+    $C_1$\label{ln:control} \State $\ntries$++\label{ln:try} \State
+    $C_i \gets C_1$ \State $\dlevels \gets\mbox{}$ decision levels in
+    $C_i$ in descending order \Comment{These never change}
+    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest
+      level $\dlevels[0]$} \State $C_i\gets\mbox{}$\tryuiplevel($C_i$,
+    $\dlevels[i]$, \type) \Comment{Try to reduce this level to UIP}
+    \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq
+      \dlevels[i]\}\bigr| + (|\dlevels| - (i+1)) \geq
+      |C_1|$}\label{ln:earlyTerminate} \State \textbf{return}
+    $C_1$. \Comment{can't generate smaller clause} \EndIf \EndFor
+    \State \textbf{if} ($\type = \allUipPure$)
+    $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
     %\State $\ntries$++\label{ln:try}% 
@@
 \Statex 
-\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{Do not add new decision levels}
+\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{
+  \parbox[t]{3in}{
+    Do not add new decision levels, \type{} determines
+    what to do if new decision levels would be introduced
+  }
+}
     \State $\ctry = C_i$
@@
 
-The Algorithm $\allUip$ attempts to reduce each level in the 1-UIP
-clause to a single literal (a UIP). The 1-UIP clause has only one
+Algorithm~\ref{alg:i-uip} attempts to reduce each level in the \oneuip
+clause to a single literal (a UIP). The \oneuip clause has only one
 literal at its deepest level, so that level, $\dlevels[0]$, can be
@@
   the other literals there is no point in trying to resolve away
-  $u$---either it will be the decision clause for level $i$ having no
+  $u$---either it will be the decision literal for level $i$ having no
   reason, or its reason will contain at least one other literal at
@@
 and thus both methods might yield a longer clause than the inputted
-1-UIP clause.
+\oneuip clause.
 
@@
 
-Finally, the algorithm returns the shorter of the initial 1-UIP clause
+Finally, the algorithm returns the shorter of the initial \oneuip clause
 and the newly computed clause (line~\ref{ln:returnmin}). Because we
@@
 observations give a lower bound on the size of the resulting clause,
-and if that lower bound is as large as the size of the initial 1-UIP
-clause we can terminate and return the initial 1-UIP clause.  This
+and if that lower bound is as large as the size of the initial \oneuip
+clause we can terminate and return the initial \oneuip clause.  This
 termination condition does not, however, account for the final clause
@@
 \paragraph{$\tgap$:}
-$\allUip$ can produce significantly smaller clause. However, when it
+$\stablealluip$ can produce significantly smaller clause. However, when it
 does not yield a smaller clause, the cost of the additional resolution
@@
 reduce a clause's $\LBD$, the maximum size reduction obtainable from
-$\allUip$ is the difference between the 1-UIP clause's size and its
+$\stablealluip$ is the difference between the \oneuip clause's size and its
 $\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$ is small,
-applying $\allUip$ is unlikely to be cost effective. Our approach is
+applying $\stablealluip$ is unlikely to be cost effective. Our approach is
 to dynamically set a threshold on $\gap(C_1)$, $\tgap$, such that when
@@
 (line~\ref{ln:control}). Initially, $\tgap = 0$, and we count the
-number of times $\allUip$ is attempted ($\ntries$) and the number of
+number of times $\stablealluip$ is attempted ($\ntries$) and the number of
 times it successfully yields a shorter clause ($\nsuc$)
@@
   \]
-  Suppose 1-UIP learning yields the clause
+  Suppose \oneuip learning yields the clause
   $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
   $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
-  is the UIP from the conflicting level. $\allUip$ first tries to find
+  is the UIP from the conflicting level. $\stablealluip$ first tries to find
   the UIP for level 6 by resolving $C_1$ with $C_k$, $C_j$
@@
 
-  $\allUip$ then attempts to find the UIP for level 5 by resolving $C^*$
+  $\stablealluip$ then attempts to find the UIP for level 5 by resolving $C^*$
   with $C_g$ and then $C_f$. However, resolving with $C_f$ would
@@
 
-  Finally, $\allUip$ processes level 2. Resolving away $d_2$ and then
+  Finally, $\stablealluip$ processes level 2. Resolving away $d_2$ and then
   $c_2$ will lead to an attempt to resolve away $b_2$. But again this
@@
   $(\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$,
-  $\neg{c_2}$, $\neg{a_2})$, a reduction of 1 over the 1-UIP clause,
+  $\neg{c_2}$, $\neg{a_2})$, a reduction of 1 over the \oneuip clause,
   and by $\allUipMin$ would be $(\neg{m_{10}}$, $\neg{h_6}$,
   $\neg{f_5}$, $\neg{e_5}$, $\neg{b_2}$, $\neg{a_2})$, a reduction of
-  2 over the 1-UIP clause.
+  2 over the \oneuip clause.
 \whitebox
@@
 
-\subsection{Variants of $\allUip$}
+\subsection{Variants of $\stablealluip$}
 We also developed and experimented with a few variants of the
-$\allUip$ algorithm which we describe below.
+$\stablealluip$ algorithm which we describe below.
 
 \subsubsection{$\allUipAct$: Clauses with Active
-  Variables.} \label{sec:active} $\allUip$ learning might introduce
+  Variables.} \label{sec:active} $\stablealluip$ learning might introduce
 literals with low variable activity into $C_i$. Low activity variables
@@
 the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
-computes the average variable activity of the newly produced all-UIP
-clause $C_i$ and the original 1-UIP clause $C_1$. The new clause $C_i$
+computes the average variable activity of the newly produced \alluip
+clause $C_i$ and the original \oneuip clause $C_1$. The new clause $C_i$
 will be returned only if it is both smaller and has higher average
-variable activity than the original 1-UIP clause. There are, of
+variable activity than the original \oneuip clause. There are, of
 course, generalizations of this approach where one has a weighted
@@
 
-We did not apply this approach to the $\allUip$ clause, as we did not
+We did not apply this approach to the $\stablealluip$ clause, as we did not
 want to bump the activity of the literals above the deepest decision
-level that $\allUip$ resolves away. Intuitively, these literals did
+level that $\stablealluip$ resolves away. Intuitively, these literals did
 not directly contribute to generating the conflict. Instead, we tried
@@
 
-Let $C_1$ be the 1-UIP learnt clause and $C_i$ be the $\allUip$ learnt
+Let $C_1$ be the \oneuip learnt clause and $C_i$ be the $\stablealluip$ learnt
 clause. First, we kept all of the variable activity bumps normally
-done by 1-UIP learning.\footnote{So extra techniques used by the
+done by \oneuip learning.\footnote{So extra techniques used by the
   underlying solver, like reason side rate and locality
   \cite{DBLP:conf/sat/LiangGPC16}, were kept intact.}  Then, when the
-$\allUip$ scheme was successful, i.e., $C_i$ was to be used as the new
+$\stablealluip$ scheme was successful, i.e., $C_i$ was to be used as the new
 learnt clause, we perform further updates to the variable
@@
 \section{Implementation and Experiments}
-%\subsubsection{Clause Reduction with $\allUip$.}
-We implemented $\allUip$ learning schemes on \text{\MapleBase}
+%\subsubsection{Clause Reduction with $\stablealluip$.}
+We implemented $\stablealluip$ learning schemes on \text{\MapleBase}
 \cite{liang2016maple}, the winner of SAT Race 2015 application
-track. We then evaluated these schemes and compare against the 1-UIP
+track. We then evaluated these schemes and compare against the \oneuip
 baseline on the full set of benchmarks from SAT RACE 2019 main track
@@
 
-  %Then a subsubsection on \allUip in modern solvers. Some of our
+  %Then a subsubsection on \stablealluip in modern solvers. Some of our
   %observations have no solid foundation, so check them to make sure
@@
 \hline
-Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
-\hline
-1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
+Solver & \# solved (SAT, UNSAT) & PAR-2 & avg. clause Size \\ 
+\hline
+\oneuip & 221 (132, 89)  & 5018.89 & 62.6  \\ 
 \hline
@@
 \end{center}
-\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\allUipPure$, $\allUipMin$, $\allUipAct$,
+\caption{Benchmark results of $\MapleBase$ with \oneuip, $\allUipPure$, $\allUipMin$, $\allUipAct$,
 $\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
@@
 PAR-2 score, and average learnt clause size. We found that the
-$\allUip$ learning schemes improved solved instances, PAR-2 scores,
-and learnt clause size over 1-UIP. More specifically, $\allUipPure$
-solved the most instances (+7 over 1-UIP) and the most UNSAT instances
+$\stablealluip$ learning schemes improved solved instances, PAR-2 scores,
+and learnt clause size over \oneuip. More specifically, $\allUipPure$
+solved the most instances (+7 over \oneuip) and the most UNSAT instances
 (+4); $\allUipIn$ solved the most SAT instances (+6); and $\allUipAct$
-yields the best PAR-2 score (-151 than 1-UIP). In all cases the
-$\allUip$ schemes learnt significantly smaller clauses on average.
-
-\subsubsection{Clause Reduction with $\allUip$.}
-To precisely measure $\allUip$'s clause reduction power, we compare
+yields the best PAR-2 score ($-$151 than \oneuip). In all cases the
+$\stablealluip$ schemes learnt significantly smaller clauses on average.
+
+\subsubsection{Clause Reduction with $\stablealluip$.}
+To precisely measure $\stablealluip$'s clause reduction power, we compare
 each instance's learnt clause size from $\allUipMin$ and $\allUipPure$
-against 1-UIP. Fig.~\ref{fig:len_pdf} shows the probability density
-distribution (PDF) of the relative clause size of the $\allUip$
+against \oneuip. Fig.~\ref{fig:len_pdf} shows the probability density
+distribution (PDF) of the relative clause size of the $\stablealluip$
 learning schemes ($\allUipMin$ in green and $\allUipPure$ in red) for
@@
 clauses for 88.5\% (77.7\%) of instances, and the average relative
-reduction ratio over 1-UIP is 18.5\%
+reduction ratio over \oneuip is 18.5\%
 (9.6\%). Fig.~\ref{fig:len_compare} compares the average learnt
-clause size of $\allUipMin$, $\allUipPure$ and 1-UIP per instance.
-Both $\allUip$ schemes generally yield smaller clauses, and the size
-reduction is more significant for instances with larger 1-UIP clauses.
+clause size of $\allUipMin$, $\allUipPure$ and \oneuip per instance.
+Both $\stablealluip$ schemes generally yield smaller clauses, and the size
+reduction is more significant for instances with larger \oneuip clauses.
 
@@
    \caption{Relative clause size reduction distribution. The $X$ axis
-      indicates the relative size of difference between all-uip and
-      1-uip clauses (calculated as
+      indicates the relative size of difference between all-UIP and
+      1-UIP clauses (calculated as
       $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
@@
       plot represents an instance. The $X$ and $Y$ axes shows the clause
-      length from $\allUip$ and 1-UIP, respectively. Each green (red)
+      length from $\stablealluip$ and \oneuip, respectively. Each green (red)
       dot represents an compared instance between $\MapleBase$ and
@@
 %  reduction ratio is obtained by running the target scheme for
-%  measurement only (the minimized 1-UIP clause is learned); Therefore,
+%  measurement only (the minimized \oneuip clause is learned); Therefore,
 %  the reduction is not cumulative. }  for each instances. 
-%For $\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
+%For $\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\stablealluip$
 %learning attempted rate and success rate.
@@
 %\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 %race main track.
-%CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, %and the values in the brackets are the non-cumulative reduction ratio.}
+%CL Reduction\% is the clause size reduction ratio comparing to non-minimized \oneuip clauses, %and the values in the brackets are the non-cumulative reduction ratio.}
 %\label{fig:t1}
@@
 %$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
-%significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
+%significantly smaller size than \oneuip by 20.4\% and 27.7\%,
 %respectively. Fig.~\ref{fig:len_pdf} shows the probability density
@@
 %instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
-%(88.5\%) of instances, and average relative reduction from 1-UIP is
+%(88.5\%) of instances, and average relative reduction from \oneuip is
 %16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
-%clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
+%clause size of $\allUipMin$, $\allUipPure$ and \oneuip, and it shows that
 %both i-UIP learning schemes in general produce the smaller clauses,
 %and the size reduction is more significant for instances with large
-%average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
+%average \oneuip clause size. Moreover, $\allUipMin$ clauses (indicated in
 %5green) is consistently smaller than $\allUipPure$ clauses.
@@
 %\nf{ Do we need this?  We additionally looked at the 14 instances
-%  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
+%  solved by $\allUipMin$ but not by \oneuip. $\allUipMin$ produces smaller
 %  clauses for all of them with average relative reduction of 22\% and
 %  maximum 77\% (30 vs 135). Seven out of 14 instances has size
-%  relative reduction over 30\%. For the 9 instances solved by 1-UIP
+%  relative reduction over 30\%. For the 9 instances solved by \oneuip
 %  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
@@
 %\hline
-%Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
+%Solver & $\stablealluip$ attempt rate & $\stablealluip$ success rate  \\ 
 %\hline
@@
 %\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate.
-%$\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to %successfully produce smaller $C_i$ clause .}
+%$\allUipMin$ scheme attempted $\stablealluip$ more frequently, and it is more likely to %successfully produce smaller $C_i$ clause .}
 %\label{fig:t2}
@@
 
-\subsubsection{Reduced Proof Sizes with $\allUip$.}
+\subsubsection{Reduced Proof Sizes with $\stablealluip$}
 A learning scheme that yields smaller clauses (lemmas) might also
 construct smaller causal proofs. For 88 UNSAT instances solved mutually
-by $\allUipPure$, $\allUipMin$ and 1-UIP schemes, we additionally
+by $\allUipPure$, $\allUipMin$ and \oneuip schemes, we additionally
 compare the size of the optimized DRAT proof from the three learning schemes. 
@@
 are 556.6MB and 698.5MB, respectively. Both sizes are significantly
-smaller than the average optimized proof size from 1-UIP, 824.9MB.
+smaller than the average optimized proof size from \oneuip, 824.9MB.
 The average proof size reduction per instance for $\allUipMin$ and
-$\allUipPure$ was 16.5\% and 3.6\% compared with 1-UIP, which roughly
+$\allUipPure$ was 16.5\% and 3.6\% compared with \oneuip, which roughly
 correlate with our observed clause size observation in
@@
 %Fig.~\ref{fig:proof_compare} shows the
-%absolute proof size comparison between $\allUipMin$ and 1-UIP, and displays a similar trend %as the clause size comparison plot in Fig.~\ref{fig:len_compare}.
+%absolute proof size comparison between $\allUipMin$ and \oneuip, and displays a similar trend %as the clause size comparison plot in Fig.~\ref{fig:len_compare}.
 
@@
 %\end{center}
-%\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$.
+%\caption{Optimized UNSAT proof comparison for \oneuip $\allUipPure$ and $\allUipMin$.
 %Optimized proof size measures the average absolute proof size in MB, and relative reduction %size measures the average relative reduction for all UNSAT instances.}
@@
 
-%\subsection{$\allUip$ as a Practical Learning Scheme}
-%To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
+%\subsection{$\stablealluip$ as a Practical Learning Scheme}
+%To evaluate $\stablealluip$'s effectiveness as a clause learning scheme, we
 %re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
-%in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
-%$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
+%in section~\ref{sec:i-uip}. We evaluated \oneuip learning and five
+%$\stablealluip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
 %$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
@@
 %produced the shortest average clause size, but solved the second least
-%instances, one more instance than the baseline 1-UIP learning. All
-%configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
+%instances, one more instance than the baseline \oneuip learning. All
+%configurations of $\stablealluip$ outperformed the baseline \oneuip scheme in
 %solved instances, PAR-2 score and average clause size.
 
-\subsubsection{$\allUip$ on Modern SAT solvers}
-To validate $\allUip$ as a practical learning scheme for modern SAT
-solvers, we implemented $\allUip$ in the winners of 2017, 2018 and
-2019 SAT
-Race\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
-and in the $\expSAT$\cite{MdSolimul2019expMalpe} ($\expSATShort$)
+\begin{figure} 
+\begin{center}
+\fss{8pt}{10pt}
+% \begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  }
+\begin{tabular}{|l|l|l|l|} 
+\hline
+Solver & \#solved (SAT, UNSAT) $\Delta$ & PAR-2 & avg. clause Size \\ 
+\hline
+SAT 2017 Winner & & & \\
+$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
+\hline
+$\MapleSeven$-all-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
+\hline
+$\MapleSeven$-all-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
+\hline
+$\MapleSeven$-all-act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
+\hline
+$\MapleSeven$-all-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
+\hline
+\hline
+SAT 2018 Winner & & & \\
+$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
+\hline
+$\MapleEightShort$-all-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
+\hline
+$\MapleEightShort$-all-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
+\hline
+$\MapleEightShort$-all-act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
+\hline
+$\MapleEightShort$-all-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
+\hline
+\hline
+SAT 2019 Winner & & & \\
+$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
+\hline
+$\MapleNineShort$-all-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
+\hline
+$\MapleNineShort$-all-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
+\hline
+$\MapleNineShort$-all-act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
+\hline
+$\MapleNineShort$-all-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
+\hline
+\hline
+SAT 2019 Competitor & & &\\
+$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
+\hline
+$\expSATShort$-all-pure & 235 (136, 99) $-$2  & 4668.96 & 48.26 \\
+\hline
+$\expSATShort$-all-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
+\hline
+$\expSATShort$-all-act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
+\hline
+$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
+\label{fig:t5}
+\end{figure}
+
+\subsubsection{$\stablealluip$ in Modern SAT solvers}
+\begin{sloppypar}
+To validate $\stablealluip$ as a practical learning scheme for modern
+SAT solvers, we implemented $\stablealluip$ in the winners of 2017,
+2018 and 2019 SAT Race
+\cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
+and in the $\expSAT$ \cite{MdSolimul2019expMalpe} ($\expSATShort$)
 solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
@@
 and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
-$\allUipEx$. We compared these solver's base 1-UIP learning scheme
-with $\allUipPure$, $\allUipMin$ and the top two $\allUip$ variants,
-$\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main track
-benchmark. We report solved instances, PAR-2 score and the average
-clause size.
-
-\begin{figure} 
-\begin{center}
-\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
-\hline
-Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
-\hline
-SAT 2017 Winner & & & \\
-$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
-\hline
-$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
-\hline
-$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
-\hline
-$\MapleSeven$-i-Act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
-\hline
-$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
-\hline
-\hline
-SAT 2018 Winner & & & \\
-$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
-\hline
-$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
-\hline
-$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
-\hline
-$\MapleEightShort$-i-Act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
-\hline
-$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
-\hline
-\hline
-SAT 2019 Winner & & & \\
-$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
-\hline
-$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
-\hline
-$\MapleNineShort$-i-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
-\hline
-$\MapleNineShort$-i-Act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
-\hline
-$\MapleNineShort$-i-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
-\hline
-\hline
-SAT 2019 Competitor & & &\\
-$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
-\hline
-$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
-\hline
-$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
-\hline
-$\expSATShort$-i-Act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
-\hline
-$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Benchmark results of 1-UIP, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
-\label{fig:t5}
-\end{figure}
-
-Table~\ref{fig:t5} shows the benchmark result of the $\allUip$
+$\allUipEx$. We compared these solver's base \oneuip learning scheme
+with $\allUipPure$, $\allUipMin$ and the top two $\stablealluip$
+variants, $\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main
+track benchmark. We report solved instances, PAR-2 score and the
+average clause size.
+\end{sloppypar}
+
+Table~\ref{fig:t5} shows the benchmark result of the $\stablealluip$
 configurations in our suite of modern solvers. Overall, we observed
@@
 configurations improved on solved instance (+4.5 instances in average)
-and PAR-2 score (-83.6 in average).  The average clause size reduction
-is consistent across all solvers. Each configuration also exhibits
-different strength: $\allUipPure$ solved the most instances with the
-best PAR-2 score on two solvers, $\allUipMin$ yields small clauses,
-$\allUipIn$ solved the most SAT instances, and $\allUipAct$ has stable
-performance.
+and PAR-2 score ($-$83.6 in average).  The average clause size
+reduction is consistent across all solvers. Each configuration also
+exhibits different strength: $\allUipPure$ solved the most instances
+with the best PAR-2 score on two solvers, $\allUipMin$ yields small
+clauses, $\allUipIn$ solved the most SAT instances, and $\allUipAct$
+has stable performance.
 
 On the SAT 2017 race winner $\MapleSeven$, all four configurations of
-$\allUip$ outperformed 1-UIP learning by 12, 8, 5, and 2 instances,
+$\stablealluip$ outperformed \oneuip learning by 12, 8, 5, and 2 instances,
 respectively. $\allUipPure$ solved more UNSAT and SAT instances while
 the other configurations improved on solving SAT instances. The clause
-size reduction of $\allUip$ is more significant on $\MapleSeven$
+size reduction of $\stablealluip$ is more significant on $\MapleSeven$
 than on to $\MapleBase$.
-%This may suggests that $\allUip$ and the recent
+%This may suggests that $\stablealluip$ and the recent
 %learnt clause minimization approach~\cite{} synergies well because
@@
 backtracking (CB) for long distance backtracks. Three out of four
-$\allUip$ configurations outperformed 1-UIP by 5, 4 an 4 instances,
+$\stablealluip$ configurations outperformed \oneuip by 5, 4 an 4 instances,
 respectively.
@@
 %chronological backtracking (CB) for long distance
-%backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
+%backtracks. Therefore, we used $\stablealluip$-CB extension for all $\stablealluip$
 %configurations.
 
-%The results indicates that the improvement of $\allUip$
+%The results indicates that the improvement of $\stablealluip$
 %is slightly shadowed by the adoption of CB. More specifically, we
@@
 %an interesting interaction effect that shows the limitations of both
-%CB and $\allUip$ learning for future research.
+%CB and $\stablealluip$ learning for future research.
 
 On the SAT 2019 race winner $\MapleNineShort$, three out of four
-$\allUip$ configurations solved instances 6, 5 and 5 more instances
-than 1-UIP, respectively. $\MapleNineShort$ prioritizes clauses that
+$\stablealluip$ configurations solved instances 6, 5 and 5 more instances
+than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
 are learned multiple times. We observed that clauses learnt from
-$\allUip$ configurations are less likely to be duplicated. As an
+$\stablealluip$ configurations are less likely to be duplicated. As an
 example, $\allUipMin$ in average, added 12\% less duplicated clauses
-into the core clause database than 1-UIP. This observation is
+into the core clause database than \oneuip. This observation is
 surprising, and the cause is unclear.
@@
 %One possible explanation is that
-%$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
+%$\stablealluip$ learning can reduce an \oneuip clause to different i-UIP clauses
 %for different assertion trails. Therefore, the solver is less likely
@@
 
-On $\expSATShort$, three out of four $\allUip$ configurations solved
-4, 7 and 8 more instances than 1-UIP learning. We noticed that
+On $\expSATShort$, three out of four $\stablealluip$ configurations solved
+4, 7 and 8 more instances than \oneuip learning. We noticed that
 $\allUipAct$ and $\allUipIn$ show significant better performance than
@@
 simulation branching heuristic, however, didn't impact the performance
-of $\allUip$ schemes significantly.
-%$\allUip$ to solve the similar amount of instances with close PAR-2
+of $\stablealluip$ schemes significantly.
+%$\stablealluip$ to solve the similar amount of instances with close PAR-2
 %scores because the additional random walk exploration allows the
@@
 %learning scheme. One possible explanation is that $\allUipAct$ and
-%$\allUipIn$ schemes could easily overcompensate variable activities
+%$\allUipIn$ schemes could easily over compensate variable activities
 %for i-UIP clauses, and $\expSATShort$'s random walk exploration could
 %use future search information to mitigate the negative effects of our
-%overcompensation.
+%over compensation.
 
 \section{Conclusion}
-In this paper we introduce a new clause learning scheme, $\allUip$,
-that combines the strengths of the existing all-UIP and 1-UIP learning
+In this paper we introduce a new clause learning scheme, $\stablealluip$,
+that combines the strengths of the existing \alluip and \oneuip learning
 schemes to learn shorter asserting clauses, while preserving the
-minimum LBDs of 1-UIP clauses. We provide empirical evidence that
-employing $\allUip$ and its variants in modern CDCL solvers achieves
+minimum \LBD of \oneuip clauses. We provide empirical evidence that
+employing $\stablealluip$ and its variants in modern CDCL solvers achieves
 significant clause reduction and yields useful performance gains.
 
-Our scheme extends 1-UIP learning by performing further resolution
+Our scheme extends \oneuip learning by performing further resolution
 beyond the deepest decision level in an attempt to find the UIP at
 each level in the learnt clause. Since resolutions may increase the
-clause's LBD by introducing literals from new decision level, we
+clause's \LBD by introducing literals from new decision level, we
 presented two methods to block such literals from entering the
@@
 Although the field of SAT solving has long converged on using the
-1-UIP learning scheme, we've shown the possibility of developing an
-effective alternative through understanding 1-UIP and other existing
+\oneuip learning scheme, we've shown the possibility of developing an
+effective alternative through understanding \oneuip and other existing
 schemes' strengths and weaknesses.  Our learning scheme can be
@@
 trade-offs between different clause quality metrics beyond clause size
-and LBD. We also plan to study the interaction between clause learning
+and \LBD. We also plan to study the interaction between clause learning
 and variable branching. Since most of the branching heuristics are
-tailored for 1-UIP scheme, the interactions with other learning
+tailored for \oneuip scheme, the interactions with other learning
 schemes still needs further study.

### 2020-05-07 paper.tex
fb camera ready revisions-more mods

@@
 
-\subsubsection{$\allUipAct$: Clauses with Active
-  Variables.} \label{sec:active} $\stablealluip$ learning might introduce
-literals with low variable activity into $C_i$. Low activity variables
-are variables that have had low recent participation in clause
-learning. Hence, clauses with variables of low activity might not be
-as currently useful to the solver. Our variant $\allUipAct$, shown as
-the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
-computes the average variable activity of the newly produced \alluip
-clause $C_i$ and the original \oneuip clause $C_1$. The new clause $C_i$
-will be returned only if it is both smaller and has higher average
-variable activity than the original \oneuip clause. There are, of
-course, generalizations of this approach where one has a weighted
-trade-off between these factors that allows preferring the new clause
-when it has large gains in one metric even though it has small losses
-in the other. We did not, however, experiment with such
-generalizations.
-
-\subsubsection{Adjust Variable Activity} \label{sec: varajust} An
-alternative to filtering clauses with low average variable
-($\allUipAct$) is to alter the way variable activities are updated to
-account for our new clause learning method. The popular branching
-heuristics VSIDS \cite{DBLP:conf/dac/MoskewiczMZZM01} and LBR
+\subsubsection{$\allUipAct$: Clauses with Active Variables.}
+\label{sec:active} $\stablealluip$ learning might
+introduce literals with low variable activity into $C_i$. Low activity
+variables are variables that have had low recent participation in
+clause learning. Hence, clauses with variables of low activity might
+not be as currently useful to the solver. Our variant $\allUipAct$,
+shown as the optional line~\ref{ln:activity} in
+Algorithm~\ref{alg:i-uip}, computes the average variable activity of
+the newly produced \alluip clause $C_i$ and the original \oneuip
+clause $C_1$. The new clause $C_i$ will be returned only if it is both
+smaller and has higher average variable activity than the original
+\oneuip clause. There are, of course, generalizations of this approach
+where one has a weighted trade-off between these factors that allows
+preferring the new clause when it has large gains in one metric even
+though it has small losses in the other. We did not, however,
+experiment with such generalizations.
+
+\subsubsection{Adjust Variable Activity.}
+\label{sec: varajust} An alternative to filtering clauses with low
+average variable ($\allUipAct$) is to alter the way variable
+activities are updated to account for our new clause learning
+method. The popular branching heuristics VSIDS
+\cite{DBLP:conf/dac/MoskewiczMZZM01} and LBR
 \cite{DBLP:conf/sat/LiangGPC16} bump the variable activity for all
@@
 
-\subsubsection{Chronological Backtracking}
+\subsubsection{Chronological Backtracking.}
 We tested our new clause learning schemes on solvers that utilized
@@
 
-\subsubsection{Reduced Proof Sizes with $\stablealluip$}
+\subsubsection{Reduced Proof Sizes with $\stablealluip$.}
 A learning scheme that yields smaller clauses (lemmas) might also
@@
 
-\subsubsection{$\stablealluip$ in Modern SAT solvers}
+\subsubsection{$\stablealluip$ in Modern SAT solvers.}
 \begin{sloppypar}

### 2020-05-08 paper.tex
improved alg description

@@
 \newcommand{\nsuc}{n_{\mathit{succ}}}
-\newcommand{\stablealluip}{\textit{stable-alluip}}
+\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
 \newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
@@
 \newcommand{\ctry}{C_{\mathit{try}}}
-\newcommand{\type}{\textit{type}}
+\newcommand{\config}{\textit{config}}
 \newcommand{\unmarked}{\textbf{unmarked}}
@@
 apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
-to the \oneuip clause. That is, our algorithm will start with the
-clause that most \sat solvers learn from a conflict, a minimized
-\oneuip clause. Our approach is specified in Algorithm~\ref{alg:i-uip}
-($\stablealluip$).
+to the \oneuip clause. That is, our algorithm $\stablealluip$,
+Algorithm~\ref{alg:i-uip}, starts with the clause that most \sat
+solvers learn from a conflict, a minimized \oneuip clause.
 
@@
     \Require $C_1$ is minimized \oneuip clause
-    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are globals used to
-    configure $\tgap$,   \type{} a global determining how the
-    introduction of new decision levels is handled.
-
-    \Procedure{\stablealluip}{$C_1, \assertionTrail$, \type} \State
-    \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return}
-    $C_1$\label{ln:control} \State $\ntries$++\label{ln:try} \State
-    $C_i \gets C_1$ \State $\dlevels \gets\mbox{}$ decision levels in
-    $C_i$ in descending order \Comment{These never change}
-    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest
-      level $\dlevels[0]$} \State $C_i\gets\mbox{}$\tryuiplevel($C_i$,
-    $\dlevels[i]$, \type) \Comment{Try to reduce this level to UIP}
+    \Require \config{} a set of configuration parameters to give different versions \stablealluip.
+    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are used to
+         dynamically adjust $\tgap$
+
+    \Procedure{\stablealluip}{$C_1, \assertionTrail$}
+    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return} $C_1$\label{ln:control} 
+    \State $\ntries$++\label{ln:try} 
+    \State $C_i \gets C_1$
+    \State $\dlevels \gets\mbox{}$ decision levels in $C_1$ in descending order\label{ln:init_loop}
+        \Comment{These never change}
+    \For{($i=1$; $i < |\dlevels|$; $i$++)}\label{ln:begin_level_loop}
+         \Comment{skip the deepest level $\dlevels[0]$}
+        \State $C_i\gets\mbox{}$\tryuiplevel($C_i$, $\dlevels[i]$) 
+            \Comment{Try to reduce this level to UIP}
         \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq
             \dlevels[i]\}\bigr| + (|\dlevels| - (i+1)) \geq
-      |C_1|$}\label{ln:earlyTerminate} \State \textbf{return}
-    $C_1$. \Comment{can't generate smaller clause} \EndIf \EndFor
-    \State \textbf{if} ($\type = \allUipPure$)
-    $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
-    %\State $\ntries$++\label{ln:try}% 
-    \State \textbf{if} ($|C_i| < |C_1|$) $\nsuc$++, \textbf{return} $C_i$ 
-           \label{ln:returnmin}\label{ln:suc}
-    \State \textcolor{gray}{$\triangleright\mbox{}$ Use next line instead for} \allUipAct
-    \State \textcolor{gray}{
-           \textbf{if} ($|C_i| < |C_1| 
-                        \land (\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$)
-                       $\nsuc$++, \textbf{return} $C_i$} \label{ln:activity}
+            |C_1|$}\label{ln:earlyTerminate}
             \State \textbf{return} $C_1$ 
+                \Comment{can't generate smaller clause}\label{ln:end_level_loop}
+        \EndIf
+   \EndFor
+   \If{$\mathbf{\allUipPure} \in \config$}
+       \State $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
+   \EndIf
+   \If{$\bigl(|C_i| < |C_1| \land 
+       \mathbf{\allUipAct}\in\config\implies(\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$
+      \label{ln:activity}}
+        \State $\nsuc$++, \textbf{return} $C_i$  \label{ln:returnmin}\label{ln:suc}
+        \Comment{$C_i$ is smaller than the input clause}
+   \Else \State \textbf{return} $C_1$
+   \EndIf
 %    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
@@
 \Statex 
-\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{
-  \parbox[t]{3in}{
-    Do not add new decision levels, \type{} determines
-    what to do if new decision levels would be introduced
-  }
-}
+\Procedure{\tryuiplevel}{$C_i$, $i$}
+    \Comment{Do not add new decision levels}
     \State $\ctry = C_i$
@@
         \Comment{Would add new decision levels}
-             \If{($\type = \allUipPure$)}
+             \If{($\mathbf{\allUipPure}\in \config$)}
                  \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
-             \ElsIf{($\type = \allUipMin$)}
+             \ElsIf{($\mathbf{\allUipMin}\in\config$)}
                  \State \textbf{continue} \Comment{Don't try to resolve away $p$}
@@
 
-Algorithm~\ref{alg:i-uip} attempts to reduce each level in the \oneuip
-clause to a single literal (a UIP). The \oneuip clause has only one
-literal at its deepest level, so that level, $\dlevels[0]$, can be
-skipped. The subroutine \tryuiplevel is used to reduce each level. It
-uses trail resolutions to achieve this, subject to the constraint that
-no new decision levels can be introduced in the clause. In particular,
+Algorithm~\ref{alg:i-uip} tries to compute a clause shorter than the
+inputted \oneuip clause $C_1$. If a clause shorter than $C_1$ cannot
+be computed the routine returns $C_1$ unchanged. Line~\ref{ln:control}
+uses the parameter $\tgap$ to predict if Alg.~\ref{alg:i-uip} will be
+successful in producing a shorter clause. This predication is
+described below. If the prediction is negative $C_1$ is immediately
+returned and Alg.~\ref{alg:i-uip} is not attempted. Otherwise, a copy
+of $C_1$ is made in $C_i$ and $\ntries$, which counts the number of
+times Alg.~\ref{alg:i-uip} is attempted, is incremented.
+
+Then the decision levels of $C_1$ are computed and stored in
+$\dlevels$ in order from largest to lowest. The for loop of
+lines~\ref{ln:begin_level_loop}--\ref{ln:end_level_loop} is then
+executed for each decision level $\dlevels[i]$. In the loop the
+subroutine \tryuiplevel tries to reduce the set of literals at
+$\dlevels[i]$ down to a single UIP literal using a sequence of trail
+resolutions. Since $C_1$ is a \oneuip clause $\dlevels[0]$ (the
+deepest level) already contains only one literal, so we can start at
+$i=1$.
+
+After the call to \tryuiplevel a check (line~\ref{ln:earlyTerminate})
+is made to see if we can abort further processing. At this point the
+algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
+so the literals at those levels will change. Furthermore, we know that
+the best that can be done from this point on is to reduce the
+remaining $|\dlevels| - (i+1)$ levels down to a single literal
+each. Hence, adding these two numbers gives a lower bound on the size
+of the final computed clause. If that lower bound is as large as the
+size of the initial \oneuip clause we can terminate and return the initial
+\oneuip clause.
+
+After processing all decision levels, if \tryuiplevel is using the
+\allUipPure configuration, additional reduction in the clause size
+might be achieved by an another round of clause minimization
+(line~\ref{ln:puremin}). Finally, if the newly computed clause $C_i$
+is smaller that the input clause $C_1$ it is returned. Otherwise the
+original clause $C_1$ is returned. Additionally, if the configuration
+$\allUipAct$, described in Section~\ref{sec:active}, is being used,
+then we also require that the average activity level of the new clause
+$C_i$ be larger than $C_1$ before we can return the new clause $C_i$.
+
 \tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
@@
 introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
-performed updating $\ctry$. In addition, all new literals at level $i$
-are added to $L_i$. These are the literals at level $i$ in the reason
-clause.
+performed updating $\ctry$. In addition, all new literals added to
+$\ctry$ at level $i$ are added to $L_i$. These are the literals at
+level $i$ in the reason clause.
 
@@
 unchanged. In the second option, called $\allUipMin$, we continue
-without performing the resolution, \textit{keeping} the current
-literal $p$ in $\ctry$. $\allUipMin$ then continues to try to resolve
-away the other literals in $L_i$ (note that $p$ is no longer in $L_i$)
-until $L_i$ is reduced to a single literal. Hence, $\allUipMin$ can
-leave multiple literals at level $i$---all of those with reasons
-containing new levels along with one other.\footnote{Since the sole
-  remaining literal $u\in L_i$ is at a lower trail index than all of
-  the other literals there is no point in trying to resolve away
-  $u$---either it will be the decision literal for level $i$ having no
-  reason, or its reason will contain at least one other literal at
-  level $i$.} Observe that the number of literals at level $i$ can not
-be increased after processing it with \allUipPure. \allUipMin can,
-however, potentially increase the number of literals at level $i$. In
-resolving away a literal $l$ at level $i$, more literals might be
-introduced into level $i$, and some of these might not be removable by
-\allUipMin if their reasons contain new levels. However, both
-\allUipPure and \allUipMin can increase the number of literals at
-levels less that $i$ as new literals can be introduced into those
-levels when the literals at level $i$ are resolved away. These added
-literals at the lower levels might not be removable from the clause,
-and thus both methods might yield a longer clause than the inputted
-\oneuip clause.
+without performing the resolution, keeping the current literal $p$ in
+$\ctry$. $\allUipMin$ then continues to try to resolve away the other
+literals in $L_i$ (note that $p$ is no longer in $L_i$) until $L_i$ is
+reduced to a single literal. Hence, $\allUipMin$ can leave multiple
+literals at level $i$---all of those with reasons containing new
+levels along with one other.\footnote{Since the sole remaining literal
+  $u\in L_i$ is at a lower trail index than all of the other literals
+  there is no point in trying to resolve away $u$---either it will be
+  the decision literal for level $i$ having no reason, or its reason
+  will contain at least one other literal at level $i$.} Observe that
+the number of literals at level $i$ can not be increased after
+processing it with \allUipPure. \allUipMin can, however, potentially
+increase the number of literals at level $i$. In resolving away a
+literal $l$ at level $i$, more literals might be introduced into level
+$i$, and some of these might not be removable by \allUipMin if their
+reasons contain new levels. However, both \allUipPure and \allUipMin
+can increase the number of literals at levels less that $i$ as new
+literals can be introduced into those levels when the literals at
+level $i$ are resolved away. These added literals at the lower levels
+might not be removable from the clause, and thus both methods might
+yield a longer clause than the inputted \oneuip clause.
 
 After trying to UIP each level the clause $C_i$ is obtained. If we
-were using $\allUipPure$ we can once again apply minimization
-(line~\ref{ln:puremin}). Recursive clause minimization
-\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\allUipMin$
-clause as all but one literal of each level introduces a new level and
-thus cannot be recursively removed.\footnote{Other more powerful
-  minimization techniques could still be applied.}
-
-Finally, the algorithm returns the shorter of the initial \oneuip clause
-and the newly computed clause (line~\ref{ln:returnmin}). Because we
-return the newly computed clause only when it is shorter, an early
-termination test can be used (line~\ref{ln:earlyTerminate}). After the
-algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
-the literals at those levels will not be further changed. Furthermore,
-we know that the best that can be done is to reduce the remaining
-$|\dlevels| - (i+1)$ levels down to a single literal each. These two
-observations give a lower bound on the size of the resulting clause,
-and if that lower bound is as large as the size of the initial \oneuip
-clause we can terminate and return the initial \oneuip clause.  This
-termination condition does not, however, account for the final clause
-minimization (line~\ref{ln:puremin}) of \allUipPure, and it could
-prematurely terminate processing that would otherwise yield a shorter
-clause. However, we discovered such false negative is rare, and
-applying early termination on \allUipPure generally improves
-performance.
+were using $\allUipPure$ we can once again apply recursive clause
+minimization (line~\ref{ln:puremin})
+\cite{DBLP:conf/sat/SorenssonB09}, but this would be useless when
+using $\allUipMin$ as all but one literal of each level introduces a
+new level and thus cannot be recursively removed.\footnote{Other more
+  powerful minimization techniques could still be applied.}
+
 %\allUipPure cannot trigger this early termination
@@
 
-\paragraph{$\tgap$:}
-$\stablealluip$ can produce significantly smaller clause. However, when it
-does not yield a smaller clause, the cost of the additional resolution
-steps can hurt the solver's performance. Since resolution cannot
-reduce a clause's $\LBD$, the maximum size reduction obtainable from
-$\stablealluip$ is the difference between the \oneuip clause's size and its
-$\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$ is small,
-applying $\stablealluip$ is unlikely to be cost effective. Our approach is
-to dynamically set a threshold on $\gap(C_1)$, $\tgap$, such that when
-$\gap(C_1) < \tgap$ we do not attempt to reduce the clause
-(line~\ref{ln:control}). Initially, $\tgap = 0$, and we count the
-number of times $\stablealluip$ is attempted ($\ntries$) and the number of
-times it successfully yields a shorter clause ($\nsuc$)
-(line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
-rate since the last restart is greater than 80\%, we decrease $\tgap$ by one (not allowing it to
-become negative), and if it is less than 80\% we increase $\tgap$
-by one.
-
+\noindent\textbf{$\tgap$}:\,\,
+$\stablealluip$ can produce significantly smaller clauses. However,
+when it does not yield a smaller clause, the cost of the additional
+resolution steps can hurt the solver's performance. Since resolution
+cannot reduce a clause's $\LBD$, the maximum size reduction obtainable
+from $\stablealluip$ is the difference between the \oneuip clause's
+size and its $\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$
+is small, applying $\stablealluip$ is unlikely to be cost
+effective. Our approach is to dynamically set a threshold on
+$\gap(C_1)$, $\tgap$, such that when $\gap(C_1) < \tgap$ we do not
+attempt to reduce the clause (line~\ref{ln:control}). Initially,
+$\tgap = 0$, and we count the number of times $\stablealluip$ is
+attempted ($\ntries$) and the number of times it successfully yields a
+shorter clause ($\nsuc$) (line~\ref{ln:try} and \ref{ln:suc}). On
+every restart if the success rate since the last restart is greater
+than 80\% (less than 80\%), we decrease (increase) $\tgap$ by one
+not allowing it to become negative.
+
+\vspace*{-3pt}
 \begin{example}
@@
     literal and the literals are in order of increasing trail
-    index. Let the clauses
-    $C_x$ denote the reason clause for literal $x_i$ with 
+    index. 
     \[\begin{array}{|l|l|l|}\hline
@@
   \]
-  Suppose \oneuip learning yields the clause
+  Let the clauses $C_x$, show above, denote the reason clause for
+  literal $x_i$.  Suppose \oneuip learning yields the clause
   $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
-  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
-  is the UIP from the conflicting level. $\stablealluip$ first tries to find
-  the UIP for level 6 by resolving $C_1$ with $C_k$, $C_j$
-  and then $C_i$ producing the clause
-  $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$,
-  $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where
-  $\neg{h_6}$ is the UIP for level 6.
+  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where
+  $\neg{m_{10}}$ is the UIP from the conflicting
+  level. $\stablealluip$ first tries to find the UIP for level 6 by
+  resolving $C_1$ with $C_k$, $C_j$ and then $C_i$ producing the
+  clause $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$, $\neg{f_5}$,
+  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{h_6}$ is the UIP
+  for level 6.
 
@@
 \subsubsection{$\allUipAct$: Clauses with Active Variables.}
-\label{sec:active} $\stablealluip$ learning might
-introduce literals with low variable activity into $C_i$. Low activity
-variables are variables that have had low recent participation in
-clause learning. Hence, clauses with variables of low activity might
-not be as currently useful to the solver. Our variant $\allUipAct$,
-shown as the optional line~\ref{ln:activity} in
-Algorithm~\ref{alg:i-uip}, computes the average variable activity of
-the newly produced \alluip clause $C_i$ and the original \oneuip
-clause $C_1$. The new clause $C_i$ will be returned only if it is both
-smaller and has higher average variable activity than the original
-\oneuip clause. There are, of course, generalizations of this approach
-where one has a weighted trade-off between these factors that allows
-preferring the new clause when it has large gains in one metric even
-though it has small losses in the other. We did not, however,
-experiment with such generalizations.
+\label{sec:active} $\stablealluip$ learning might introduce literals
+with low variable activity into $C_i$. Low activity variables are
+variables that have had low recent participation in clause
+learning. Hence, clauses with variables of low activity might not be
+as currently useful to the solver. Our variant $\allUipAct$
+(line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip}) computes the
+average variable activity of the newly produced \alluip clause $C_i$
+and the original \oneuip clause $C_1$. The new clause $C_i$ will be
+returned only if it is both smaller and has higher average variable
+activity than the original \oneuip clause. There are, of course,
+generalizations of this approach where one has a weighted trade-off
+between these factors that allows preferring the new clause when it
+has large gains in one metric even though it has small losses in the
+other. We did not, however, experiment with such generalizations.
 

### 2020-05-10 paper.tex
add cadical results

@@
 \newcommand{\expSATShort}{\textit{expMaple} }
+\newcommand{\cadical}{\textit{CaDiCaL}}
+\newcommand{\defaultcadical}{\textit{CaDiCaL-default}}
 \newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
@@
 We implemented $\stablealluip$ learning schemes on \text{\MapleBase}
-\cite{liang2016maple}, the winner of SAT Race 2015 application
+\cite{liang2016maple}, the winner of SAT Race 2016 application
 track. We then evaluated these schemes and compare against the \oneuip
@@
 \hline
-$\MapleNineShort$-all-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
+$\MapleNineShort$-all-pure & 240 (142, \textbf{98}) +2 & 4519.08 &  43.32\\
 \hline
@@
 \hline
-$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
+$\expSATShort$-all-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33}
+\\
+\hline
+\hline
+ & & &\\
+$\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
+\hline
+$\defaultcadical$-all-pure & 248 (151, 97) $-$1  & 4373.38 & 82.44  \\
+\hline
+$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 46.29 \\ 
+\hline
+$\defaultcadical$-all-act & \textbf{252} (\textbf{152}, \textbf{100}) +3 & 4331.56 & 47.34 \\
+\hline
+$\defaultcadical$-all-inclusive & 251 (\textbf{153}, 98) +2 & 4335.88 & \textbf{42.61}
+\\
 \hline
@@
 and in the $\expSAT$ \cite{MdSolimul2019expMalpe} ($\expSATShort$)
-solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
+and $\cadical$~\cite{cadical} solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
 uses random walk simulation to help branching. We chose $\expSATShort$
@@
 and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
-$\allUipEx$. We compared these solver's base \oneuip learning scheme
+$\allUipEx$.
+We chose $\cadical$ because its default configuration\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) solved 244 total instances in 2019 SAT Race, the highest among all participants.
+For this experimented, we used the latest version of $\defaultcadical$ instead of the SAT Race participant~\cite{Armin2019Cadical}.
+We compared these solver's base \oneuip learning scheme
 with $\allUipPure$, $\allUipMin$ and the top two $\stablealluip$
@@
 
-On the SAT 2019 race winner $\MapleNineShort$, three out of four
-$\stablealluip$ configurations solved instances 6, 5 and 5 more instances
+On the SAT 2019 race winner $\MapleNineShort$, all four
+$\stablealluip$ configurations solved instances 2, 6, 5 and 5 more instances
 than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
@@
 of $\stablealluip$ schemes significantly.
-%$\stablealluip$ to solve the similar amount of instances with close PAR-2
-%scores because the additional random walk exploration allows the
-%learning schemes to partially sidestep the activity problem; hence,
-%the learning adjustment for variable activities should have less
-%impact on the solver's performance. However, we instead, observed that
-%both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
-%learning scheme. One possible explanation is that $\allUipAct$ and
-%$\allUipIn$ schemes could easily over compensate variable activities
-%for i-UIP clauses, and $\expSATShort$'s random walk exploration could
-%use future search information to mitigate the negative effects of our
-%over compensation.
+
+$\defaultcadical$ with \oneuip solved 249 total instances, 5 more than the result from the SAT 2019 Race.
+Applying $\allUipAct$ and $\allUipIn$ further improved the solved instance by 3 an 2, respectively.
+The average learnt clause size of \oneuip on $\defaultcadical$ is much larger than other solvers' (101 vs 60) while the average clause size of $\stablealluip$ configurations are similar for all solvers.
 

### 2020-05-10 paper.tex
final pass, fix some typos and out-of-date stats

@@
 \usepackage[noend]{algpseudocode}
-%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
 \usepackage{xspace}
@@
 \usepackage[T1]{fontenc}
-% \usepackage[T1,hyphens]{url}
 \usepackage{hyperref}
-%\PassOptionsToPackage{hypnens}{url}
 
@@
    \EndIf
-%    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
-%           \textbf{then}  $C_i$ \textbf{else}  $C_1$
 \EndProcedure
@@
   powerful minimization techniques could still be applied.}
-
-%\allUipPure cannot trigger this early termination
-%condition (it cannot increase the size of any processed level), but
-%\allUipMin can.
 
@@
     \begin{array}{l}
-    \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}l) \\
+    \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \\
     \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land\mbox{} \\
@@
 \section{Implementation and Experiments}
-%\subsubsection{Clause Reduction with $\stablealluip$.}
 We implemented $\stablealluip$ learning schemes on \text{\MapleBase}
@@
 XeonE5-2680 CPUs, allowing
-%\fb{insert time bound}
 5000 seconds per instance and a maximum of 12 GB memory.
-%\fb{insert RAM bound}.
-
-%\fb{Delete figure 2 and use only figure 8 with all of the
-%  solvers. First briefly comment on the performance and make a brief observation about clause sizes.
- %Then have a new subsubsection that looks at clause sizes wrt. to the
-  %two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
-  %them side by side (using a tabular environment as I did in figure 1).
-
-  %Remove fig 5 and put those numbers in the text instead of in a table
-  %as part of your discussion.
-
-  %Then have a new subsubsection on DRAT trim and get rid of Fig 7.
-
-  %Then a subsubsection on \stablealluip in modern solvers. Some of our
-  %observations have no solid foundation, so check them to make sure
-  %that you are just not stating your intuitions without reasonable
-  %evidence. We need to to fit into 15 pages total along with a short
-  %1/3 page for conclusions.
-
-  %Finally add all citations.
-
-  %If we add the more complex \allUipMin we also will need more space
-  %to describe that, so prune some of your conclusions especially those
-  %that are simply conjectures.
-%}
 
@@
 
-%Beside solved instances count and PAR-2 score, we additionally measure
-%the average clause size. 
-%and clause reduction ratio (both cumulative
-%and non-cumulative)\footnote{The cumulative reduction ratio is
-%  obtained through learning all clauses with the target learning
-%  scheme; Therefore, the reduction is cumulative. The %non-cumulative
-%  reduction ratio is obtained by running the target scheme for
-%  measurement only (the minimized \oneuip clause is learned); Therefore,
-%  the reduction is not cumulative. }  for each instances. 
-%For $\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\stablealluip$
-%learning attempted rate and success rate.
-
-%\begin{figure} 
-%\begin{center}
-%\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
-%\hline
-%Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
-%\hline
-%$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
-%\hline
-%$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
-%\hline
-%$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 %race main track.
-%CL Reduction\% is the clause size reduction ratio comparing to non-minimized \oneuip clauses, %and the values in the brackets are the non-cumulative reduction ratio.}
-%\label{fig:t1}
-%\end{figure}
-
-%Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
-%(five) more instances than the baseline solver with lower PAR-2
-%scores. $\allUipPure$ has marginally lower PAR-2 score than
-%$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
-%significantly smaller size than \oneuip by 20.4\% and 27.7\%,
-%respectively. Fig.~\ref{fig:len_pdf} shows the probability density
-%distribution (PDF) of the relative clause size of both i-uip learning
-%schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
-%instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
-%(88.5\%) of instances, and average relative reduction from \oneuip is
-%16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
-%clause size of $\allUipMin$, $\allUipPure$ and \oneuip, and it shows that
-%both i-UIP learning schemes in general produce the smaller clauses,
-%and the size reduction is more significant for instances with large
-%average \oneuip clause size. Moreover, $\allUipMin$ clauses (indicated in
-%5green) is consistently smaller than $\allUipPure$ clauses.
-
-%\nf{ Do we need this?  We additionally looked at the 14 instances
-%  solved by $\allUipMin$ but not by \oneuip. $\allUipMin$ produces smaller
-%  clauses for all of them with average relative reduction of 22\% and
-%  maximum 77\% (30 vs 135). Seven out of 14 instances has size
-%  relative reduction over 30\%. For the 9 instances solved by \oneuip
-%  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
-%  three instances and with average relative reduction of 3.3\%.}
-
-%\begin{figure} 
-%\begin{center} 
-%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
-%\hline
-%Solver & $\stablealluip$ attempt rate & $\stablealluip$ success rate  \\ 
-%\hline
-%$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
-%\hline
-%$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate.
-%$\allUipMin$ scheme attempted $\stablealluip$ more frequently, and it is more likely to %successfully produce smaller $C_i$ clause .}
-%\label{fig:t2}
-%\end{figure}
-
 \subsubsection{Reduced Proof Sizes with $\stablealluip$.}
@@
 
-%Fig.~\ref{fig:proof_compare} shows the
-%absolute proof size comparison between $\allUipMin$ and \oneuip, and displays a similar trend %as the clause size comparison plot in Fig.~\ref{fig:len_compare}.
-
-%\begin{figure} 
-%\begin{center} 
-%\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
-%\hline
-%Solver & optimized proof size (MB) & relative reduction size  \\ 
-%\hline
-%$\MapleBase$ & 613.9 & 0  \\ 
-%\hline
-%$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
-%\hline
-%$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Optimized UNSAT proof comparison for \oneuip $\allUipPure$ and $\allUipMin$.
-%Optimized proof size measures the average absolute proof size in MB, and relative reduction %size measures the average relative reduction for all UNSAT instances.}
-%\label{fig:t3}
-%\end{figure}
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
-%    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
-%    \label{fig:proof_compare}
-%\end{figure}
-
-%\subsection{$\stablealluip$ as a Practical Learning Scheme}
-%To evaluate $\stablealluip$'s effectiveness as a clause learning scheme, we
-%re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
-%in section~\ref{sec:i-uip}. We evaluated \oneuip learning and five
-%$\stablealluip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
-%$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
-%benchmark and reported solved instances, PAR-2 score and average
-%clause size.
-
-%Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
-%scheme $\allUipPure$ solved the most overall instances (228) and the
-%most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
-%score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
-%produced the shortest average clause size, but solved the second least
-%instances, one more instance than the baseline \oneuip learning. All
-%configurations of $\stablealluip$ outperformed the baseline \oneuip scheme in
-%solved instances, PAR-2 score and average clause size.
-
 \begin{figure} 
@@
 \fss{8pt}{10pt}
-% \begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  }
 \begin{tabular}{|l|l|l|l|} 
@@
 \hline
-$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 46.29 \\ 
+$\defaultcadical$-all-min & 248 (149, 99) $-$1 & 4398.31 & 43.93 \\ 
 \hline
@@
 uses random walk simulation to help branching. We chose $\expSATShort$
-because (1) it one of the top solvers in the 2019 SAT Race that does
+because (1) it is one of the top solvers in the 2019 SAT Race that does
 not use chronological backtracking; and (2) the random walk simulation
@@
 We chose $\cadical$ because its default configuration\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) solved 244 total instances in 2019 SAT Race, the highest among all participants.
-For this experimented, we used the latest version of $\defaultcadical$ instead of the SAT Race participant~\cite{Armin2019Cadical}.
+For this experimented, we used the latest available version of $\defaultcadical$ instead of the 2019 SAT Race participant~\cite{Armin2019Cadical}.
 We compared these solver's base \oneuip learning scheme
@@
 $\MapleBase$ in Fig.~\ref{fig:t4}. More specifically, almost all
-configurations improved on solved instance (+4.5 instances in average)
-and PAR-2 score ($-$83.6 in average).  The average clause size
+configurations improved on solved instance (+3.9 instances in average)
+and PAR-2 score ($-$57.7 in average).  The average clause size
 reduction is consistent across all solvers. Each configuration also
@@
 than on to $\MapleBase$.
-%This may suggests that $\stablealluip$ and the recent
-%learnt clause minimization approach~\cite{} synergies well because
-%i-UIP clauses are shorter with more common literals which allows
-%vivification~\cite{} to prune literals more aggressively through unit
-%propagation.
 
@@
 respectively.
-%$\MapleEightShort$ improved from $\MapleSeven$ by using
-%chronological backtracking (CB) for long distance
-%backtracks. Therefore, we used $\stablealluip$-CB extension for all $\stablealluip$
-%configurations.
-
-%The results indicates that the improvement of $\stablealluip$
-%is slightly shadowed by the adoption of CB. More specifically, we
-%believe CB prevents decision levels from being compressed through the
-%process of backtracking and literal assertion. The shorter i-UIP
-%clauses can bring related literals closer, which in turns compresses
-%the assertion stack and the decision levels. However, since CB
-%discourages long distance backtracking, the effect of learning shorter
-%clauses is shadowed until a full restart.  We believe we have observed
-%an interesting interaction effect that shows the limitations of both
-%CB and $\stablealluip$ learning for future research.
 
@@
 
-%One possible explanation is that
-%$\stablealluip$ learning can reduce an \oneuip clause to different i-UIP clauses
-%for different assertion trails. Therefore, the solver is less likely
-%to learn the same i-UIP clauses.
-
 On $\expSATShort$, three out of four $\stablealluip$ configurations solved
@@
 
-$\defaultcadical$ with \oneuip solved 249 total instances, 5 more than the result from the SAT 2019 Race.
+$\defaultcadical$ with \oneuip solved 249 instances, 5 more than the result from the SAT 2019 Race.
 Applying $\allUipAct$ and $\allUipIn$ further improved the solved instance by 3 an 2, respectively.
-The average learnt clause size of \oneuip on $\defaultcadical$ is much larger than other solvers' (101 vs 60) while the average clause size of $\stablealluip$ configurations are similar for all solvers.
+The average learnt clause size of \oneuip on $\defaultcadical$ is much larger than other solvers' (101 vs 60) while the average clause size of $\stablealluip$ configurations are similar on all solvers.
 

### 2020-05-11 paper.tex
fix a typo
fix a typo, simplify the description of cadical

@@
 algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
-so the literals at those levels will change. Furthermore, we know that
+so the literals at those levels will not change. Furthermore, we know that
 the best that can be done from this point on is to reduce the
@@
 
-$\defaultcadical$ with \oneuip solved 249 instances, 5 more than the result from the SAT 2019 Race.
-Applying $\allUipAct$ and $\allUipIn$ further improved the solved instance by 3 an 2, respectively.
-The average learnt clause size of \oneuip on $\defaultcadical$ is much larger than other solvers' (101 vs 60) while the average clause size of $\stablealluip$ configurations are similar on all solvers.
+$\defaultcadical$ with \oneuip solved 249 instances,   
+and $\allUipAct$ and $\allUipIn$ solved 2 and 3 more instance, respectively.  
+The average learnt clause size of \oneuip on 
+$\defaultcadical$ is much larger than other solvers' (101 vs 60) 
+while the average clause size of $\stablealluip$ configurations are similar on all solvers.
 

### 2020-05-11 paper.tex
fix typos

@@
 and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
-$\allUipEx$.
-We chose $\cadical$ because its default configuration\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) solved 244 total instances in 2019 SAT Race, the highest among all participants.
-For this experimented, we used the latest available version of $\defaultcadical$ instead of the 2019 SAT Race participant~\cite{Armin2019Cadical}.
+$\allUipEx$. We chose $\cadical$ because its default configuration\footnote{$\cadical$'s 
+default configuration is different from the specialized 'sat' configuration, 
+which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) 
+solved 244 instances in the 2019 SAT Race, the highest among 
+all participants. For this experimented, we used the latest available version of
+ $\defaultcadical$ instead of the 2019 SAT Race participant~\cite{Armin2019Cadical}. 
   We compared these solver's base \oneuip learning scheme

### 2020-05-11 paper.tex
change wording for better format

@@
 default configuration is different from the specialized 'sat' configuration, 
-which solved the most SAT instances in the 2019 SAT Race.} (denote as $\defaultcadical$) 
+which solved the most satisfiable instances in the 2019 SAT Race.} (denote as $\defaultcadical$) 
 solved 244 instances in the 2019 SAT Race, the highest among 
@@
 
-$\defaultcadical$ with \oneuip solved 249 instances,   
-and $\allUipAct$ and $\allUipIn$ solved 2 and 3 more instance, respectively.  
+$\defaultcadical$ with \oneuip solved 249 total instances. 
+Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 2 and 3 more instance, respectively.  
 The average learnt clause size of \oneuip on 

### 2020-05-15 paper.tex
fix the last typo, hopefully
last typo
fix font, fit into the page limit

@@
 \documentclass[runningheads]{llncs}
-\usepackage{times}
 \usepackage{amsmath}
@@
 \end{center}
-\caption{Benchmark results of $\MapleBase$ with \oneuip, $\allUipPure$, $\allUipMin$, $\allUipAct$,
+\caption{Results of $\MapleBase$ with \oneuip, $\allUipPure$, $\allUipMin$, $\allUipAct$,
 $\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
@@
 
-The results in Fig.~\ref{fig:t4}, ~\ref{fig:len_pdf} and
+The results in Fig.~\ref{fig:t4}, \ref{fig:len_pdf} and
 \ref{fig:len_compare} indicate $\allUipMin$ often achieves higher
@@
   \end{minipage}
-
 \end{tabular}
@@
 by $\allUipPure$, $\allUipMin$ and \oneuip schemes, we additionally
-compare the size of the optimized DRAT proof from the three learning schemes. 
+compared the size of the optimized DRAT proof from the three learning schemes. 
 We used the DRAT-trim tool
 \cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
-optimize every DRAT proof once \footnote{Applying DRAT-trim multiple
+optimize every DRAT proof once\footnote{Applying DRAT-trim multiple
   times can further reduce the proof size until a fix-point. However,
-  the full optimization is time consuming and was impractical for our
+  the full optimization is too time consuming for our
   experiments.}.
@@
 The average proof size reduction per instance for $\allUipMin$ and
-$\allUipPure$ was 16.5\% and 3.6\% compared with \oneuip, which roughly
-correlate with our observed clause size observation in
+$\allUipPure$ is 16.5\% and 3.6\% against \oneuip, which roughly
+correlate with our clause size observation in
 Fig~\ref{fig:len_pdf}.
@@
 \hline
- & & &\\
 $\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
@@
 the other configurations improved on solving SAT instances. The clause
-size reduction of $\stablealluip$ is more significant on $\MapleSeven$
+size reduction of $\stablealluip$ is more significant on this solver
 than on to $\MapleBase$.
@@
 than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
-are learned multiple times. We observed that clauses learnt from
-$\stablealluip$ configurations are less likely to be duplicated. As an
+are learned multiple times. We observed that
+$\stablealluip$ clauses are less likely to be duplicated. As an
 example, $\allUipMin$ in average, added 12\% less duplicated clauses
@@
 4, 7 and 8 more instances than \oneuip learning. We noticed that
-$\allUipAct$ and $\allUipIn$ show significant better performance than
+both $\allUipAct$ and $\allUipIn$ show better performance than
 $\allUipMin$ and $\allUipPure$ on this solver. The random walk
@@
 $\defaultcadical$ with \oneuip solved 249 total instances. 
-Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 2 and 3 more instance, respectively.  
+Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 3 and 2 more instances, respectively.  
 The average learnt clause size of \oneuip on 

### 2020-05-15 paper.tex
stablize Fig. 5

@@
 \hline
+\cadical version 1.2.1 & & &\\
 $\defaultcadical$ & 249 (150, 99)  & \textbf{4311.76} & 101.96 \\
@@
 and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
-$\allUipEx$. We chose $\cadical$ because its default configuration\footnote{$\cadical$'s 
-default configuration is different from the specialized 'sat' configuration, 
-which solved the most satisfiable instances in the 2019 SAT Race.} (denote as $\defaultcadical$) 
+$\allUipEx$. We chose $\cadical$ because its default configuration
+%\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most satisfiable instances in the 2019 SAT Race.} 
+(denote as $\defaultcadical$) 
 solved 244 instances in the 2019 SAT Race, the highest among 

### 2020-05-15 paper.tex
fix a typo

@@
 On the SAT 2019 race winner $\MapleNineShort$, all four
-$\stablealluip$ configurations solved instances 2, 6, 5 and 5 more instances
+$\stablealluip$ configurations solved 2, 6, 5 and 5 more instances
 than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
@@
 
-$\defaultcadical$ with \oneuip solved 249 total instances. 
+$\defaultcadical$ with \oneuip solved 249  instances. 
 Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 3 and 2 more instances, respectively.  

### 2020-05-15 CR/paper.tex
fb changes to CR

--- 2020-05-15 paper.tex
@@
 \newcommand{\allUipEx}{\textit{alluip-exclusive}}
-\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
+\newcommand{\MapleBase}{\textit{MapleCOMSPS-LRB}}
 \newcommand{\MapleSeven}{\textit{MapleLCMDist}}
@@
 \begin{abstract}
-    Almost all CDCL \sat solvers use the \oneuip 
-    clause learning scheme for learning new clauses from conflicts, 
-    and our current understanding
-    of \sat solving provides good reasons for using that scheme. In
-    particular, the \oneuip scheme yields asserting clauses, and these
-    asserting clauses have minimum \LBD among all possible asserting
-    clauses. As a result of these advantages, other clause learning
-    schemes, like $i$-UIP and \alluip, that were proposed in early
-    work are not used in modern solvers. In this paper, we propose a
-    new technique for exploiting the \alluip clause learning scheme. 
-    Our technique is to employ \alluip learning under the
-    constraint that the learnt clause's \LBD does not increase (over the
-    minimum established by the \oneuip clause). Our method can learn
-    clauses that are significantly smaller than the \oneuip clause while
-    preserving the minimum \LBD. Unlike previous clause minimization
-    methods, our technique is not limited to learning a sub-clause of
-    the \oneuip clause. It can instead learn a completely different
-    clause. We show empirically 
-    that our method can improve the performance of state of the art solvers.
+    Almost all CDCL \sat solvers use the \oneuip clause learning
+    scheme for learning new clauses from conflicts, and our current
+    understanding of \sat solving provides good reasons for using that
+    scheme. In particular, the \oneuip scheme yields asserting
+    clauses, and these asserting clauses have minimum \LBD among all
+    possible asserting clauses. As a result of these advantages, other
+    clause learning schemes, like $i$-UIP and \alluip, that were
+    proposed in early work are not used in modern solvers. In this
+    paper, we propose a new technique for exploiting the \alluip
+    clause learning scheme.  Our technique is to employ \alluip
+    learning under the constraint that the learnt clause's \LBD does
+    not increase (over the minimum established by the \oneuip
+    clause). Our method can learn clauses that are significantly
+    smaller than the \oneuip clause while preserving the minimum
+    \LBD. Unlike previous clause minimization methods, our technique
+    is not limited to learning a sub-clause of the \oneuip clause. We
+    show empirically that our method can improve the performance of
+    state of the art solvers.
 \end{abstract}
@@
 good properties of standard \oneuip clauses. Our method is related to,
-but not the same as, various clause mininimization methods that try to
+but not the same as, various clause minimization methods that try to
 remove redundant literals from the \oneuip clause yielding a clause that
@@
 but which still serves all of the same purposes as the \oneuip clause.
-Various minimization techniques can be applied on top of our method to
-remove redundant literals from the clauses we have learnt.
+Clause minimization techniques can be applied on top of our method 
+to remove redundant literals.
 
@@
 
-\paragraph{The Trail.}
+\vspace*{3pt}\noindent\textit{The Trail.}
 CDCL \sat solvers maintain a \textbf{trail}, $\trail$, which is a
@@
 that $\trail$ \emph{remains intact during clause learning} and is only
-changed after the new clause is learnt. After the new clause is learnt
-$\trail$ will be changed by backtracking.
+changed after the new clause is learnt (by backtracking).
 
@@
 Ordered trail resolutions resolve unit propagated literals from the
-end of the trail to the beginning. Without lost of generality, we can
-require that all trail resolutions be ordered.
+end of the trail to the beginning. W.l.o.g we can require that all trail
+resolutions be ordered.
 
@@
 trail. That is, these implementations are typically using ordered
-trail resolutions. Observation~\ref{obs:ordered} shows that this in
-fact the right way to do this.
+trail resolutions. Observation~\ref{obs:ordered} shows that this is
+correct.
 
@@
 
+{\setlength{\textfloatsep}{-10pt plus 1.0pt minus 2.0pt}
 \begin{figure}[t!]
-{\fss{9pt}{10pt}
-\begin{tabular}[t]{|l|c|l|}
+{\fss{8pt}{9pt}
+\setlength{\tabcolsep}{0.02em}
+\begin{tabular}[t]{|c|c|c|}
   \cline{1-1}\cline{3-3}
-  \begin{minipage}[t]{0.48\textwidth}
+  \begin{minipage}[t]{0.5\textwidth}
     \textbf{(a) All Decision Clause\rule{0pt}{1.1\topskip}}
     \begin{algorithmic}[0]
-        \State \hspace*{-1em}\textbf{all\_decision}($C_I$)
+        \State \hspace*{-.7em}\textbf{all-decision}($C_I$)
         \State $C\gets C_I$
@@
   & &
-  \begin{minipage}[t]{0.475\textwidth}
+  \begin{minipage}[t]{0.5\textwidth}
       \textbf{(b) Make level $i$ contain a single literal\rule{0pt}{1.1\topskip}}
       \begin{algorithmic}[0]
-          \State \hspace*{-1em} \textbf{UIP\_level}($C$, $i$)
+          \State \hspace*{-.7em} \textbf{UIP-level}($C$, $i$)
           \While{$\left|\left\{\ell \left| \begin{array}{l}\ell \in C\\
@@
    \multicolumn{1}{c}{\vspace*{-7pt}}\\\cline{1-1}\cline{3-3}
-  \begin{minipage}[t]{0.475\textwidth}
+  \begin{minipage}[t]{0.5\textwidth}
     \begin{minipage}{\textwidth}
@@
         \begin{algorithmic}[0]
-            \State \hspace*{-1em}\textbf{\oneuip}($C_I$)
+            \State \hspace*{-.7em}\textbf{\oneuip}($C_I$)
             \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
-            \State \textbf{return} \textbf{UIP\_level}($C_I$, i)
+            \State \textbf{return} \textbf{UIP-level}($C_I$, i)
         \end{algorithmic}
@@
     \end{minipage}\\[5pt]
-    \hspace*{-1.8pt}\rule{1.03\textwidth}{.4pt}\\[5pt]
+    \hspace*{-.6pt}\rule{1.005\textwidth}{.4pt}\\[5pt]
     \begin{minipage}{\textwidth}
@@
         \begin{algorithmic}[0]
-          \State \hspace*{-1em}\textbf{\alluip}($C_I$, $i$)
+          \State \hspace*{-.7em}\textbf{\alluip}($C_I$, $i$)
           \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
-          \State \Comment{$i$ is large enough to ensure all levels are UIP}
+          \State \Comment{{\fs{7pt}$i$ is large enough to ensure all levels are UIP}}
           \State \textbf{return} \textbf{$i$-UIP}(C, i)
@@
  & &
-  \begin{minipage}[t]{0.48\textwidth}
+  \begin{minipage}[t]{0.5\textwidth}
       \vspace*{-20pt}
@@
       \begin{algorithmic}[0]
-          \State \hspace*{-1em}\textbf{$i$-UIP}($C_I$, $i$)
+          \State \hspace*{-.7em}\textbf{$i$-UIP}($C_I$, $i$)
           \State $C\gets C_I$
@@
           \State \textbf{if} ($d = \varnothing$): \textbf{break}
-          \State $C\gets\mbox{}$ \textbf{UIP\_level(C, d)}
+          \State $C\gets\mbox{}$ \textbf{UIP-level(C, d)}
           \State $d\gets \max\left\{\dlevel{l} \left|\begin{array}{l}l\in C \\ \mbox{} \land \dlevel{l}<d
@@
 }
+\vspace*{-5pt}
 \caption{Some different clause learning schemes. All use the
@@
 \end{figure}
-
+}
  %   \floatstyle{boxed}
@@
 these schemes (all-decision, all-UIP, i-UIP) use trail resolutions in
-which the \oneuip clause appears. That is, they all must first generate
-the \oneuip clause and then continue with further trail resolution
-steps. These extra resolution steps can introduce many addition
-decision levels into the final clause. Hence, these schemes yield
-learnt clauses whose \LBD can be larger, and never smaller, that the
-\oneuip clause.
+which the \oneuip clause appears. That is, they all must first
+generate the \oneuip clause and then continue with further trail
+resolution steps. These extra resolution steps can introduce many
+addition decision levels into the final clause. Hence, these schemes
+learn clauses with \LBD at least as large as the \oneuip clauses.
 
@@
 apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
-to the \oneuip clause. That is, our algorithm $\stablealluip$,
-Algorithm~\ref{alg:i-uip}, starts with the clause that most \sat
+to the \oneuip clause. That is, our algorithm $\stablealluip$
+(Algorithm~\ref{alg:i-uip}), starts with the clause that most \sat
 solvers learn from a conflict, a minimized \oneuip clause.
@@
 \caption{\stablealluip}\label{alg:i-uip}
+\fss{9pt}{10pt}
 \begin{algorithmic}[1]
@@
 performed updating $\ctry$. In addition, all new literals added to
-$\ctry$ at level $i$ are added to $L_i$. These are the literals at
-level $i$ in the reason clause.
+$\ctry$ at level $i$ are added to $L_i$.
 
@@
 level $i$ are resolved away. These added literals at the lower levels
-might not be removable from the clause, and thus both methods might
-yield a longer clause than the inputted \oneuip clause.
+might not be removable from the clause, and thus both methods 
+can yield a longer clause than the input \oneuip clause.
 
@@
 
-\vspace*{-3pt}
+%\vspace*{-3pt}
 \begin{example}
@@
     index. 
-    \[\begin{array}{|l|l|l|}\hline
+    \[\fss{9pt}{9pt}
+        \begin{array}{|l|l|l|}\hline
       C_{a} = \varnothing 
@@
 
-\subsubsection{$\allUipAct$: Clauses with Active Variables.}
+\vspace*{3pt}\noindent\textbf{$\allUipAct$: Clauses with Active Variables.}
 \label{sec:active} $\stablealluip$ learning might introduce literals
@@
 
+\begin{sloppypar}
 In sum, the two modified variable activity update schemes we
-experimented with were
-\[
-    \begin{array}{l}
-    \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \\
-    \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land\mbox{} \\
-    \qquad \forall l\in C_1 - C_i.\,\mathit{unbumpActivity(l)}
-  \end{array}
-\]
-
-\subsubsection{Chronological Backtracking.}
+experimented with were (1)
+$\textbf{\allUipIn} \equiv \forall l\in C_i -
+C_1.\,\mathit{bumpActivity}(l)$ and (2)
+$\textbf{\allUipEx} \equiv \forall l\in C_i -
+C_1.\,\mathit{bumpActivity}(l) \land \bigl(\forall l\in C_1 -
+C_i.\,\mathit{unbumpActivity(l)}\bigr)$
+\end{sloppypar}
+
+\vspace*{3pt}\noindent\textbf{Chronological Backtracking.}
 We tested our new clause learning schemes on solvers that utilized
@@
 
+{
+  \setlength{\intextsep}{5pt}
 \begin{figure}
+    \fss{9pt}{10pt}
 \begin{center}
@@
 \end{figure}
+}
 
@@
 
-\subsubsection{Clause Reduction with $\stablealluip$.}
+\vspace*{3pt}\noindent\textbf{Clause Reduction with $\stablealluip$.}
 To precisely measure $\stablealluip$'s clause reduction power, we compare
@@
 \begin{figure}[t!]
-{\fss{8pt}{10pt}
 \begin{tabular}[t]{l c l}
-
   \begin{minipage}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/clause_reduction_PDF.pdf}
-   \caption{Relative clause size reduction distribution. The $X$ axis
+   \caption{\fss{9pt}{10pt}
+      Relative clause size reduction distribution. The $X$ axis
       indicates the relative size of difference between all-UIP and
@@
     \includegraphics[width=\textwidth]{figures/clause_size_compare.pdf}
-    \caption{Average clause size comparison plot. Each point in the
+    \caption{\fss{9pt}{10pt}
+      Average clause size comparison plot. Each point in the
       plot represents an instance. The $X$ and $Y$ axes shows the clause
@@
 \end{tabular}
-}
 \end{figure}
 
-\subsubsection{Reduced Proof Sizes with $\stablealluip$.}
+\vspace*{3pt}\noindent\textbf{Reduced Proof Sizes with $\stablealluip$.}
 A learning scheme that yields smaller clauses (lemmas) might also
@@
 
+{  \setlength{\intextsep}{10pt}
 \begin{figure} 
 \begin{center}
-\fss{8pt}{10pt}
+\fss{9pt}{10pt}
 \begin{tabular}{|l|l|l|l|} 
@@
 \hline
+\end{tabular}
+\end{center}
+\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track instances.}
+\label{fig:t5a}
+\end{figure}
+}
+
+{ \setlength{\intextsep}{10pt}
+  \setlength{\textfloatsep}{10pt}
+\begin{figure} 
+\begin{center}
+\fss{9pt}{10pt}
+\begin{tabular}{|l|l|l|l|} 
 \hline
@@
 \hline
-$\defaultcadical$-all-act & \textbf{252} (\textbf{152}, \textbf{100}) +3 & 4331.56 & 47.34 \\
+$\defaultcadical$-all-act & \textbf{252} (152, \textbf{100}) +3 & 4331.56 & 47.34 \\
 \hline
@@
 \end{center}
-\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
-\label{fig:t5}
+\caption{Benchmark results of \oneuip, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track instances.}
+\label{fig:t5b}
 \end{figure}
-
-\subsubsection{$\stablealluip$ in Modern SAT solvers.}
+}
+
 \begin{sloppypar}
-To validate $\stablealluip$ as a practical learning scheme for modern
-SAT solvers, we implemented $\stablealluip$ in the winners of 2017,
-2018 and 2019 SAT Race
+    \vspace*{3pt}\noindent\textbf{$\stablealluip$ in Modern SAT
+      solvers.}  To validate $\stablealluip$ in modern SAT solvers, we
+    implemented $\stablealluip$ in the winners of 2017, 2018 and 2019
+    SAT Race
     \cite{DBLP:conf/ijcai/LuoLXML17,ryvchin2018maple,Stepan2019MapleLCMDistChronoBT}
     and in the $\expSAT$ \cite{MdSolimul2019expMalpe} ($\expSATShort$)
-and $\cadical$~\cite{cadical} solver. $\expSATShort$ is a top ten solver from 2019 SAT race which
-uses random walk simulation to help branching. We chose $\expSATShort$
-because (1) it is one of the top solvers in the 2019 SAT Race that does
-not use chronological backtracking; and (2) the random walk simulation
-branching heuristic is different from local branching heuristic (VSIDS
-and LRB) that we have considered in $\allUipAct$, $\allUipIn$, and
-$\allUipEx$. We chose $\cadical$ because its default configuration
+    and $\cadical$~\cite{cadical} solvers. $\expSATShort$ is a top ten
+    solver from 2019 SAT race which uses random walk simulation to
+    help branching. We chose $\expSATShort$ because the random walk
+    simulation branching heuristic is different from local branching
+    heuristic (VSIDS and LRB) that we have considered in $\allUipAct$,
+    $\allUipIn$, and $\allUipEx$. We chose $\cadical$ because its
+    default configuration
 %\footnote{$\cadical$'s default configuration is different from the specialized 'sat' configuration, which solved the most satisfiable instances in the 2019 SAT Race.} 
-(denote as $\defaultcadical$) 
-solved 244 instances in the 2019 SAT Race, the highest among 
-all participants. For this experimented, we used the latest available version of
- $\defaultcadical$ instead of the 2019 SAT Race participant~\cite{Armin2019Cadical}. 
-  We compared these solver's base \oneuip learning scheme
-with $\allUipPure$, $\allUipMin$ and the top two $\stablealluip$
-variants, $\allUipAct$ and $\allUipIn$, on the SAT Race 2019 main
-track benchmark. We report solved instances, PAR-2 score and the
-average clause size.
+    ($\defaultcadical$) solved the most instances in the 2019 SAT Race
+    (244). For this experiment, we used the latest available version
+    of $\defaultcadical$ instead of the 2019 SAT Race
+    version~\cite{Armin2019Cadical}.  We compared these solvers'
+    base \oneuip learning scheme with $\allUipPure$, $\allUipMin$ and
+    the top two $\stablealluip$ variants, $\allUipAct$ and
+    $\allUipIn$, on the SAT Race 2019 main track benchmarks. We report
+    solved instances, PAR-2 score and the average clause size.
 \end{sloppypar}
 
-Table~\ref{fig:t5} shows the benchmark result of the $\stablealluip$
-configurations in our suite of modern solvers. Overall, we observed
-similar performance gain on all modern solvers as we have seen on
-$\MapleBase$ in Fig.~\ref{fig:t4}. More specifically, almost all
-configurations improved on solved instance (+3.9 instances in average)
-and PAR-2 score ($-$57.7 in average).  The average clause size
-reduction is consistent across all solvers. Each configuration also
-exhibits different strength: $\allUipPure$ solved the most instances
-with the best PAR-2 score on two solvers, $\allUipMin$ yields small
-clauses, $\allUipIn$ solved the most SAT instances, and $\allUipAct$
-has stable performance.
+Tables~\ref{fig:t5a} and \ref{fig:t5b} show the results of the
+$\stablealluip$ configurations in our suite of modern
+solvers. Overall, we observed similar performance gain on all modern
+solvers as we have seen on $\MapleBase$ in Fig.~\ref{fig:t4}. More
+specifically, almost all configurations improved on solved instance
+(+3.9 instances in average) and PAR-2 score ($-$57.7 in average).  The
+average clause size reduction is consistent across all solvers. Each
+configuration also exhibits different strengths: $\allUipPure$ solved
+the most instances with the best PAR-2 score on two solvers,
+$\allUipMin$ yields small clauses, $\allUipIn$ solved the most SAT
+instances, and $\allUipAct$ has stable performance.
 
 On the SAT 2017 race winner $\MapleSeven$, all four configurations of
-$\stablealluip$ outperformed \oneuip learning by 12, 8, 5, and 2 instances,
-respectively. $\allUipPure$ solved more UNSAT and SAT instances while
-the other configurations improved on solving SAT instances. The clause
+$\stablealluip$ solved more instances than \oneuip
+learning. $\allUipPure$ solved more UNSAT and SAT instances while the
+other configurations improved on solving SAT instances. The clause
 size reduction of $\stablealluip$ is more significant on this solver
-than on to $\MapleBase$.
-
-The SAT 2018 race winner$\MapleEightShort$ uses chronological
-backtracking (CB) for long distance backtracks. Three out of four
-$\stablealluip$ configurations outperformed \oneuip by 5, 4 an 4 instances,
-respectively.
-
-On the SAT 2019 race winner $\MapleNineShort$, all four
-$\stablealluip$ configurations solved 2, 6, 5 and 5 more instances
-than \oneuip, respectively. $\MapleNineShort$ prioritizes clauses that
-are learned multiple times. We observed that
-$\stablealluip$ clauses are less likely to be duplicated. As an
-example, $\allUipMin$ in average, added 12\% less duplicated clauses
-into the core clause database than \oneuip. This observation is
-surprising, and the cause is unclear.
-
-On $\expSATShort$, three out of four $\stablealluip$ configurations solved
-4, 7 and 8 more instances than \oneuip learning. We noticed that
-both $\allUipAct$ and $\allUipIn$ show better performance than
-$\allUipMin$ and $\allUipPure$ on this solver. The random walk
-simulation branching heuristic, however, didn't impact the performance
-of $\stablealluip$ schemes significantly.
-
-$\defaultcadical$ with \oneuip solved 249  instances. 
-Applying $\allUipAct$ and $\allUipIn$ helped the solver solve 3 and 2 more instances, respectively.  
-The average learnt clause size of \oneuip on 
-$\defaultcadical$ is much larger than other solvers' (101 vs 60) 
-while the average clause size of $\stablealluip$ configurations are similar on all solvers.
+than on $\MapleBase$. The SAT 2018 race winner $\MapleEightShort$ uses
+chronological backtracking (CB); three out of four configurations
+outperformed \oneuip. On the SAT 2019 race winner $\MapleNineShort$,
+all four $\stablealluip$ configurations solved more instances than
+\oneuip. $\MapleNineShort$ prioritizes clauses that are learned
+multiple times. We observed that $\stablealluip$ clauses are less
+likely to be duplicated. As an example, $\allUipMin$ on average, added
+12\% less duplicated clauses into the core clause database than
+\oneuip. This observation is surprising, and the cause is unclear.
+
+On $\expSATShort$, three out of four $\stablealluip$ configurations
+solved more instances than \oneuip learning. We noticed that both
+$\allUipAct$ and $\allUipIn$ show better performance than $\allUipMin$
+and $\allUipPure$ on this solver. The random walk simulation branching
+heuristic, however, didn't impact the performance of $\stablealluip$
+schemes significantly.
+
+$\defaultcadical$ with \oneuip solved 249 instances.  Applying
+$\allUipAct$ and $\allUipIn$ helped the solver solve 3 and 2 more
+instances, respectively.  The \oneuip clauses in $\defaultcadical$
+were much larger than other solvers on average (101 vs 60) but the
+$\stablealluip$ configurations yielded similar clause sizes.
 
 \section{Conclusion}
-In this paper we introduce a new clause learning scheme, $\stablealluip$,
-that combines the strengths of the existing \alluip and \oneuip learning
-schemes to learn shorter asserting clauses, while preserving the
-minimum \LBD of \oneuip clauses. We provide empirical evidence that
-employing $\stablealluip$ and its variants in modern CDCL solvers achieves
+In this paper we introduced a new clause learning scheme,
+$\stablealluip$, that preserves the strengths \oneuip learning while
+learning shorter clauses. We provided empirical evidence that using
+$\stablealluip$ and its variants in modern CDCL solvers achieves
 significant clause reduction and yields useful performance gains.
@@
 each level in the learnt clause. Since resolutions may increase the
-clause's \LBD by introducing literals from new decision level, we
+clause's \LBD by introducing literals from new decision levels, we
 presented two methods to block such literals from entering the
 clause. Although our learning scheme is conceptually simple, and we
-have presented optimizations to reduce and balance the learning
-cost. We additionally presented variants of our schemes to account for
-features used in state of the art solvers, e.g., local branching
-heuristics and chronological backtracking.
-
-Although the field of SAT solving has long converged on using the
-\oneuip learning scheme, we've shown the possibility of developing an
-effective alternative through understanding \oneuip and other existing
-schemes' strengths and weaknesses.  Our learning scheme can be
-generalized and further improved by exploring more fine-grained
+presented optimizations to reduce and balance the learning cost. We
+additionally presented variants of our schemes to account for features
+used in state of the art solvers, e.g., local branching heuristics and
+chronological backtracking.
+
+Although the field of SAT solving has converged on using the \oneuip
+learning scheme, we have shown the possibility of developing an
+effective alternative through understanding the strengths and
+weaknesses of \oneuip and clause learning schemes. Our learning scheme
+can be generalized and further improved by exploring more fine-grained
 trade-offs between different clause quality metrics beyond clause size
-and \LBD. We also plan to study the interaction between clause learning
-and variable branching. Since most of the branching heuristics are
-tailored for \oneuip scheme, the interactions with other learning
-schemes still needs further study.
+and \LBD. We also plan to study the interaction between clause
+learning and variable branching. Since most of the branching
+heuristics are tailored for \oneuip scheme, their interactions with
+other learning schemes requires further study.
 
